### 2016-03-05 manuscript/manuscript.Rnw
Draft manuscript

\documentclass{nature}
\bibliographystyle{naturemag.bst}
\usepackage{lineno}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage[hidelinks]{hyperref}

\title{Climatic warming strengthens a positive feedback between alpine shrubs and fire}
\author{James S. Camac\textasteriskcentered$^{1,2,3}$, Richard J. Williams$^4$, Carl-Henrik Wahren$^3$, Ary A. Hoffmann$^5$ \\
\& Peter A. Vesk$^2$}

\begin{document}
\linenumbers
\maketitle

\begin{affiliations}
\item Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia.
\item The Centre of Excellence for Environmental Decisions, School of BioSciences, The University of Melbourne, Parkville 3010, Victoria, Australia.
\item Research Centre for Applied Alpine Ecology, La Trobe University, Melbourne 3086, Victoria, Australia.
\item CSIRO Land and Water, tropical ecosystems research centre, PMB 44 Winnellie 0821, NT, Australia.
\item Bio21 Institute, School of BioSciences, The University of Melbourne, Vic. 3010, Australia.
\textasteriskcentered Email for correspondence: \texttt{james.camac@gmail.com}
\end{affiliations}

\pagebreak

\begin{abstract}
Climate change is expected to increase fire activity and exacerbate woody plant encroachment into arctic and alpine grasslands. What is unknown is whether increases in woody plants and fire will interact to exacerbate changes in structure, function and composition of these ecosystems. Here we use field surveys and experimental manipulations to examine warming and fire effects on recruitment, growth and survival of seedlings of evergreen obligate seeding alpine shrubs. We find that fire substantially increased shrub seedling establishment (up to 36 fold) and that warming doubled tall shrub seedling growth rates and may increase survival. Warming had weak or no effect on shrub recruitment, rates of grass post-fire gap infilling or grass competitive effects on shrub seedling growth and survival. These findings indicate that rising temperatures coupled with more frequent or severe fires will exacerbate increases in tall evergreen shrubs by increasing recruitment, doubling growth rates, and potentially increase survival.
\end{abstract}

Accurately forecasting the effects of climatic warming on vegetation dynamics requires an understanding of the mechanisms by which climate and vegetation interact. Most forecasting models include the direct effects of climatic conditions on species distributions, but largely ignore biotic interactions and the type, frequency and severity of disturbances that are also likely to change\cite{Guisan:2005gz,Tucker:2012dr,Giannini:2013gi}. Because disturbance strongly influences recruitment opportunities, and thus, composition and structure of plant communities and biomes worldwide\cite{Bond:2005eg, Turner:2010wo}, it has the capacity to exacerbate or diminish how climate affects vegetation. Consequently, it is imperative to understand how seedlings, the life stage most vulnerable to climate and disturbance and the life stage which determines the long-term persistence of a species and its capacity to establish in new areas, responds to both changing climatic and disturbance regimes\cite{Walck:2011dm, Briceno:2015iv}.

Field manipulative experiments have been widely used to investigate the effects of climate change on plant communities. These experiments have provided invaluable information on how climate change directly influences plant phenology\cite{Hoffmann:2010kf, Dorji:2013tt}, reproduction\cite{Klady:2011dw}, morphology\cite{Hudson:2009cn}, growth\cite{Hollister:2005vf}, floristic composition\cite{Elmendorf:2012kz} and biotic interactions\cite{Klanderud:2005uw}. However, because disturbances vary both spatially and temporally, most have focused on mature plant responses in relatively undisturbed vegetation\cite{Briceno:2015iv}. As such, few field climate experiments have been conducted on seedlings or vegetation in post-disturbance conditions (but see\cite{Munier:2010cg,Graae:2010ir,Camac:2015hja}). Consequently, little is known about how climate change influences seedling demographic rates\cite{Briceno:2015iv} and thus, vegetation recovery in post-disturbance conditions. We consider these issues in an Australian alpine heathland.

In alpine and arctic ecosystems, warming experiments and long-term monitoring have documented significant increases in the growth and cover of woody species\cite{Sturm:2001vd,Walker:2006ht,MyersSmith:2011fp}. The frequency and extent of wildfires have also increased in these environments over recent decades, a trend expected to continue\cite{Westerling:2006ve, Flannigan:2009kl, Qiu:2009th, Bradstock:2014jp}. In alpine and tundra ecosystems, current evidence from long-term and experimental studies indicate that shrub recruitment and encroachment is highest in disturbed areas\cite{Williams:1987ua, Batllori:2009fe, Frost:2013tm}, that climatic warming is likely to improve woody species growth rates\cite{Arft:1999iz, Elmendorf:2012ba, MyersSmith:2011fp}, and that shrubs are the most flammable vegetation component of these ecosystems\cite{Williams:2006tl, Higuera:2008dw, Higuera:2009ga}. The combination of this evidence indicates that more frequent or severe fire will increase recruitment opportunities for woody species, which when coupled with higher growth and survival rates caused by rising temperatures, will exacerbate shrub expansion\cite{Racine:2004tn,Lantz:2013fh} and ultimately increase the likelihood of more fires\cite{Higuera:2008dw, Higuera:2009ga,MyersSmith:2011fp}. In effect, warming will strengthen an existing feedback loop that not only has the potential to cause rapid changes in the composition and structure of alpine and arctic vegetation, but also have serious social, biodiversity and carbon sequestration consequences\cite{Mack:2011ew}.

While paleoecological studies have indicated that such a feedback may exist\cite{Higuera:2008dw, Higuera:2009ga}, there is a paucity of information on multiple mechanisms that may strengthen, mitigate or break this feedback loop (Fig. \ref{fig1}a). For example, we have little information on whether fire will stimulate shrub seedling recruitment in alpine and tundra ecosystems, nor do we know whether rates of seedling growth and survival will be improved under warmer post-fire conditions. Furthermore, we do not know how tussock grasslands (the other dominant community in these ecosystems) will respond to warmer post-fire conditions, and whether grasses have competitive, facilitative or no effect on shrub seedling demographic rates, or whether such effects are altered by warmer post-fire conditions.

We examine these unknowns and their impacts on the feedback between climatic warming, shrubs and fire in Australian alpine vegetation. As with tundra ecosystems, the Australian Alps have experienced rapid changes in climate. Since 1979, mean growing season temperatures have risen by approximately 0.4\textdegree C and annual precipitation has fallen by 6\%\cite{Wahren:2013ji}, with a consequent decline in snow pack\cite{SanchezBayo:2013ci}. These climatic changes have been correlated with a 10 to 20\% increase in shrub cover and a 25\% decline in graminoids\cite{Wahren:2013ji}. Much of the Australian Alps has also been burnt by recent wildfires, the frequency and severity of which is expected to increase\cite{Bradstock:2014jp,Williams:2014ud}.

We took advantage of recent fires in alpine open heathland, a plant community that occupies ca. 25\% of the Australian alpine landscape. Under global warming, this plant community is likely to encroach grasslands and is itself susceptible to shrub thickening\cite{McDougall:2003gb}. To determine how warmer and drier post-fire conditions affects shrub cover, we combined field observations with a warming experiment that used seedlings of four dominant Australian evergreen obligate seeding alpine shrubs \textemdash \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) and \textit{Phebalium squamulosum} (Rutaceae; a tall shrub). This approach allowed us to examine: 1) how warmer post-fire conditions influence rates of both grass recovery and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affect shrub seedling demographic rates, and 3) whether such affects change under warmer conditions post-fire conditions.

\subsection{Drivers of shrub seedling establishment.}

We first investigated how altitude, topographic wetness (TWI), adult density, fire and, if burnt, fire severity, influenced the density of \textit{Grevillea} and \textit{Asterolasia} seedlings (two dominant shrubs of open heathland). In 2012-13, we estimated seedling density for both species across 40 alpine sites and found that the abundance of \textit{Grevillea} (Fig. \ref{fig2}) and \textit{Asterolasia} (Supplementary Fig. S1) seedlings were strongly influenced by the occurrence of fire. Sites burnt by fire in 2003 had seedling densities between 13 and 36 times higher than unburnt sites. In \textit{Grevillea}, mean seedling densities were 1.44/m$^2$ at burnt sites and 0.04/m$^2$ at unburnt sites. \textit{Asterolasia} had similar mean densities: 1.71 and 0.13 seedlings/m$^2$ at burnt and unburnt sites, respectively. For both species, seedling density was also higher in severely burnt sites. As expected, pre-fire adult density positively influenced \textit{Grevillea} seedling density, with sites containing higher abundances of pre-fire \textit{Grevillea} adults containing more seedlings. For both shrub species, we detected no significant change in seedling density along a 200 m altitudinal range (equivalent to a 1.5\textdegree C change in mean temperature). The field warming experiment (see below) indicates that this lack of elevation response in \textit{Grevillea} is due to both rates of recruitment (Supplementary Fig. S2) and mortality (see below) largely being insensitive to a 1\textdegree C change in temperature. We detected no significant effect of topographic wetness for either species.

\subsection{Experimental warming effects on shrub seedling growth.}

Knowing shrub seedlings were more abundant in burnt vegetation, we investigated whether warmer post-fire conditions affected seedling growth and mortality rates. Seedlings of \textit{Grevillea}, \textit{Asterolasia}, \textit{Prostanthera} and \textit{Phebalium} emerging after a wildfire were transplanted into experimentally burnt plots that were either passively warmed by 1\textdegree C using open top chambers (OTCs) or subject to current ambient conditions (controls). After 1812 days (1087 growing season days where OTC treatment was applied) or 5 years growth, mean seedling heights of the tall shrubs (\textit{Grevillea}, \textit{Prostanthera} and \textit{Phebalium}) growing in post-fire bare ground were taller in warmed plots by 9.7, 13.5 and 3.8 cm respectively (Fig. \ref{fig3}). Warming increased heights of both \textit{Grevillea} and \textit{Prostanthera} seedlings in all years, while \textit{Phebalium} did not respond to warming treatment until the second growing season. By contrast, seedlings of the prostrate shrub, \textit{Asterolasia}, showed no significant difference in growth rate between warmed and control plots in any year. For each species, we also observed similar treatment trends with stem diameter growth (Supplementary Fig. S3). Accounting for initial height and assuming logistic growth, mean annual predicted height growth rates of \textit{Grevillea}, \textit{Prostanthera} and \textit{Phebalium} were 2.4, 1.9 and 1.5 times that observed in control plots, respectively. Projecting these differences equates to an average 6 cm tall seedling (the mean initial size used in this experiment) growing in warmed conditions reaching maximum mature height in 39 fewer years for \textit{Grevillea} and 17 fewer years for both \textit{Prostanthera} and \textit{Phebalium} (Fig. \ref{fig3}). \textit{Asterolasia} was predicted to reach a mature height in approximately 22 years regardless of warming treatment.

We validated the \textit{Grevillea} and \textit{Asterolasia} growth responses observed in experimentally warmed plots by using an elevation gradient in open heathland and maximum height data for 2003 post-fire recruits (i.e. suspected 9 year old seedlings). Here, a 200 m altitudinal range is equivalent to an ambient temperature difference of approximately 1.5\textdegree C\cite{Brown:1989vx}, which is comparable to that observed between experimental warmed and control plots (1\textdegree C). In response to shifts in temperature, maximum heights of Grevillea post-fire recruits were expected to decrease with elevation, whereas \textit{Asterolasia} seedlings were not expected to show this pattern. Our experimental predictions were verified (Supplementary Fig. S4). Mean maximum heights of \textit{Grevillea} seedlings in burnt open heathland were 7.9 cm taller at 1660 m a.s.l compared to seedlings at 1860 m a.s.l. (19.4 cm vs 11.5 cm; a difference comparable to our experimental findings). In contrast, mean maximum heights of \textit{Asterolasia} seedlings did not vary significantly with elevation. Topographic Wetness Index (TWI) and fire severity had no detectable influence on maximum seedling heights in either species.

\subsection{Experimental warming effects on shrub seedling mortality.}

After five years and across all plots, 34\% (172 out of 512) of all seedlings growing in burnt bare ground patches had died. Most deaths occurred in \textit{Asterolasia} (65) followed by \textit{Phebalium} (63), \textit{Grevillea} (31) and \textit{Prostanthera} (13). \textit{Prostanthera} showed the largest treatment effect (Fig. \ref{fig4}), with annual mortality rates estimated to be near 0 in warmed plots and 5\% in control plots. This significant increase in survival may be a consequence of OTCs reducing the severity of spring frosts by rising minimum ambient and soil temperatures by 0.86\textdegree C and 1.6\textdegree C, respectively (see Supplementary Fig S5-7). Warming also reduced mean seedling mortality in \textit{Grevillea} and \textit{Phebalium} (Fig. \ref{fig4}); however, for both species, the effect was not statistically significant. By contrast, annual mortality rates in the prostrate shrub, \textit{Asterolasia}, were marginally higher in warmed plots, but again this effect was not significant (Fig. \ref{fig4}).

\subsection{Grass proximity effects on shrub seedling growth and mortality rates.}

Climate change is expected to alter biotic interactions because their strength and direction depend strongly on climatic conditions, particularly in alpine and arctic ecosystems\cite{Callaway:2002vb, Klanderud:2005uw}. Here, we assess the interactive effects of warming and inter-tussock gap size (i.e. the proximity of shrub seedlings to grass tussocks) on the growth and survival of \textit{Grevillea} seedlings (Fig. \ref{fig5}).  As with the seedlings growing in bare ground, we detected a strong positive effect of warming on growth rates and a marginal, yet not significant, decrease in mortality. We also found no significant effect of inter-tussock gap radius on growth or mortality rates. However, for growth, we did detect a marginally significant interaction between warming and gap size, such that in warmed plots growth rates slightly increased with gap size. This suggests that warmer conditions may increase competition between shrubs and grasses, a finding consistent with the abiotic stress hypothesis\cite{Callaway:2002vb}, which suggests that temperature stress in alpine and tundra ecosystems could increase the strength of competitive interactions. However, this effect was small, and likely to have little biological consequence.

\subsection{Experimental warming and rates of tussock grass gap infilling.}

Despite having little impact on shrub seedling growth and mortality rates, tussock grass may still reduce shrub recruitment opportunities by infilling post-fire bare ground gaps at a faster rate. Using five years of post-fire inter-tussock gap size changes in warmed and unwarmed plots we found that gaps were being infilled by grasses. However, the rate at which this occurred was not significantly influenced by a 1\textdegree C increase in temperature (Fig. \ref{fig6}).

\subsection{Strengthening of the warming-shrub-fire feedback.}

Here we quantified several unknown interactions between fire, shrub-grass relationships and climate to extend a conceptual model of alpine shrub dynamics under climate change (Fig \ref{fig1}b). We showed how these interactions strengthen a hypothesized feedback loop that can rapidly increase shrub cover in alpine and tundra ecosystems\cite{Higuera:2008dw}. The combination of rising temperatures and more frequent or severe fire creates conditions that allow shrub seedlings to establish in greater densities, and for tall shrubs, double their growth rates, and potentially increase their survival. These demographic effects will ultimately result in shrub thickening and expansion into grasslands. These effects, coupled with field\cite{Williams:2006tl}, experimental\cite{Fraser:NHT5Kx_W} and paleoecological\cite{Higuera:2008dw} evidence indicate that shrubs are the most flammable vegetation component of alpine and tundra ecosystems, and suggest that flammable fuel loads will accumulate faster and cover a larger proportion of the alpine landscape under a warmer environment. This further strengthens the feedback\cite{Higuera:2008dw, Higuera:2009ga} by potentially increasing the frequency or severity of fires, which then create further recruitment opportunities for shrubs with little demographic impact of neighbouring tussock grass.

While we have addressed several unknowns associated with this feedback between climatic warming, shrubs and fire, there are others we have not addressed that may also strengthen, weaken or break this feedback. The most obvious mechanism that will break this cycle involves shorter fire intervals that prevent obligate seeding shrubs reaching reproductive age and thereby exhausting the seedbank. However, this scenario is unlikely for the majority of alpine (or tundra) landscapes including in Australia, where it would require current fire intervals of 50 to 100 years to decrease to less than 20 years \textemdash the time observed for the species in this study to reach reproductive maturity\cite{Williams:2008ba}. Furthermore, if reproductive output is related to plant size\cite{Wenk:2015ko}, then climatic warming may allow obligate seeding species to reach reproductive maturity sooner, and consequently, increase species resilience to short fire intervals. Nevertheless, these factors and others such as changes in snow pack\cite{Wipf:2009jh}, soil moisture\cite{Engler:2011vq} or herbivory\cite{Post:2008wz} require further research because they are likely to be altered by the interactive effects of climate and disturbance in unpredictable ways\cite{MyersSmith:2011fp}.

Our findings provide mechanistic understanding as to why shrub cover has increased, often at the expense of grasslands, in many alpine and arctic ecosystems\cite{Racine:2004tn,MyersSmith:2011fp,Frost:2013tm,Lantz:2013fh}. But more importantly, our results provide evidence for underlying processes that could result in a warming-fire-shrub feedback that has been hypothesized in paleoecological studies\cite{Higuera:2008dw, Higuera:2009ga}. Based on current IPCC observations and predictions, average global temperature has already increased by 0.85\textdegree C since 1880 and is expected to rise by as much as 4.8\textdegree C by 2100\cite{IPCC:2013vv}. In alpine and tundra environments, temperatures\cite{ChapinIII:2005js}, shrub cover\cite{MyersSmith:2011fp} and the frequency and severity of fire\cite{Westerling:2006ve, Flannigan:2009kl, Qiu:2009th, Bradstock:2014jp} have all increased in the last few decades. These changes mean that the warming-shrub-fire feedback loop identified here has already strengthened, which could cause grasslands and other non-woody communities to transition to an alternative state of more shrubs and more fire, both of which are likely to have consequences for carbon sequestration, water supply and biodiversity.

\begin{methods}

We investigated shrub dynamics in open heathland because it is a common and highly flammable plant community in the Australian Alps above 1600 m a.s.l\cite{Williams:2006tl}. It is also an ecotone between closed heathland (>70\% shrub cover) and tussock grassland consisting of \textit{Grevillea australis} shrubs interspersed among a \textit{Poa hiemata} sward. Aerial photography has revealed that plant community has acted as an \textit{invasion front} of shrub expansion into grassland. It is also experiencing shrub thickening by closed heathland dominants such as \textit{Prostanthera cuneata}.

\subsection{Open Top Chamber Experiment.}

In March 2010, at 1750 m a.s.l, we burnt 32 randomly selected mature (60 cm tall and 1.5m$^2$) \textit{Grevillea australis} shrubs in open heathland to create patches of bare ground approximately 0.7 m$^2$ surrounded by burnt tussock grass, simulating disturbance in open heathland burnt by wildfire. After creating the bare ground patches we collected seedlings of dominant alpine shrub species from a nearby (< 2 km) site of similar altitude burnt by a late 2006 wildfire. We collected shrub seedlings of two dominant open heathland species (\textit{Grevillea australis} (Proteaceae; a tall shrub) and \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), a dominant closed heathland species (\textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) that typically grows on warmer aspects and a species common to both open and closed heathland (\textit{Phebalium squamulosum} (Rutaceae; a tall shrub). All four species are obligate seeders post-fire (i.e. fire killed)\cite{Camac:2013eb}.

A total of 640 seedlings, 256 \textit{Grevillea} (half used in \textit{Poa} inter-tussock experiment \textemdash see below) and 128 for each of \textit{Asterolasia}, \textit{Prostanthera} and \textit{Phebalium} were used. Four seedlings per species were randomly selected and transplanted into a 4\texttimes 4 square grid in the center of each burnt patch, with 14 cm between individuals and the edge of the patch dominated by resprouting tussock grass \textit{Poa hiemata}. To examine interactions between tussock grass and shrub seedlings we also randomly transplanted four additional \textit{Grevillea australis} into various sized inter-tussock gaps between burnt \textit{Poa hiemata} immediately surrounding the bare ground (Supplementary Fig. S6). The experimental site was fenced to prevent grazing by deer and horses. Rabbits or invertebrates also did not graze on these seedlings.

To simulate near-term warmer conditions predicted under IPCC\cite{IPCC:2013vv}, we randomly assigned Open Top Chambers (OTCs) to half (16) the plots, with the remainder treated as unwarmed controls. OTCs were placed over plots, ensuring all seedlings (including inter-tussock shrub seedlings) occurred within the 1.1 m$^2$ open top to minimise edge effects. OTCs were placed out at the start of the growing season (October) where they remained until snowfall (early June). This procedure was repeated for five growing seasons from 2010 to 2015.

Microclimatic changes were measured hourly using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) at four control and four OTC plots. Across five growing seasons (1087 days), OTCs simulated warmer conditions at the lower end of IPCC predictions\cite{IPCC:2013vv}. OTCs passively increased mean daily ambient and soil temperatures by 1\textdegree C and maximums and minimums by up to 2.4\textdegree C (Supplementary Methods and Supplementary Fig. S5-7). Chambers only marginally decreased soil moisture by 0.21\%.

Seedling survival, maximum height and stem diameter (nearest mm measured with Vernier calipers) were initially recorded in May 2010 and then subsequently re-measured at the end of each growing season (May-June). At the same time, we recorded the distance to nearest tussock or grass seedling in each of four cardinal directions for seedlings growing in inter-tussock gaps. Thus changes in inter-tussock gap size could be due to either vegetative growth or seedling establishment. We did not measure individual characteristics (e.g. height and basal diameter) of surrounding tussock grass because we could not distinguish individuals, and because height varied throughout the season. Numbers of natural \textit{Grevillea australis} recruits establishing within the plots were also recorded for the first two seasons.

\subsection{Seedling gradient study.}

We used 40 long-term open heathland sites\cite{Camac:2013eb} established after the 2003 fires consisting of 30 burnt sites and 10 sites thought to be unburnt for over 70 years. In the summer of 2012-13, at each site, seedling density/m$^2$ was estimated using 40 1 m$^2$ quadrats, evenly distributed in groups of 10 along four 50 m transects, with 10 m between transect lines, subsampling an area of 2000 m$^2$. Within plots we recorded the number and maximum height of \textit{Grevillea} and \textit{Asterolasia} seedlings. For unburnt sites we counted the number of mature \textit{Grevillea} plants (> 0.5 m$^2$) within 5 m either side of each transect. In burnt sites, this required counting the number of skeletons (there were no living adults at any burnt site) that still persist post-fire. We were unable estimate numbers of adult \textit{Asterolasia} because this species does not have a persistent woody skeleton post-fire. Site level data, elevation and Topographic Wetness Index (TWI), were obtained from a 30 m resolution digital elevation model. Lastly, for burnt sites, fire severity was estimated using minimum twig diameters, collected immediately after the 2003 fires.

\subsection{Data analysis.}

In total we built 11 hierarchical models to examine factors influencing tussock grass post-fire recovery and shrub seedling establishment, growth and mortality. For each model (Supplementary Methods) we used Bayesian inference and fitted models with Markov chain Monte Carlo (MCMC) sampling in R 3.2.2 using package Rstan 2.8.1. Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.
\end{methods}

\section*{Acknowledgements} 
This research was funded through Australian Research Council Linkage Grants, partnered through the Department of Sustainability and Environment, Parks Victoria, Commonwealth Scientific and Industrial Research Organisation (CSIRO). Funding was also received from the Centre of Excellence for Environmental Decisions (CEED) and Holsworth Research Committee. J.S.C was a recipient of an Australian Postgraduate Award. Monica Camac, Brad Farmilo, Monica Hursburgh, Paulius Kviecinskas, Karen Stott, Imogen Fraser, Sarah Bartlett, Paul Smart, Steven Rumbold, Lachlan Yourn and Chris Jones provided assistance with field measurements. William Morris and John Baumgartner provided insightful chats about modelling approaches and Warwick Papst provided field logistic advice. Daniel Falster and Richard Fitzjohn for enabled these analyses to be readily reproducible. Special thanks to David Bowman, Anaïs Gibert, Brad Farmilo, John Morgan, Kimberley Millers, Dale Nimmo, Michaela Plein, Libby Rumpff and Inka Veltheim for their comments on this manuscript.

\section*{Author contributions}
J.C. conceived, designed, performed the experiments, field surveys and analysis; C-H.W, R.W, A.H and P.V. supervised the development of this work, aided in data collection and provided statistical advice. All authors contributed to the writing of this manuscript.

\section*{Competing financial interests} 
The authors declare that they have no competing financial interests.

\bibliography{refs.bib}

\section*{Figures}

\begin{figure}
\label{fig1}
\begin{minipage}{\textwidth}
\includegraphics{figures/fig1.pdf}
\caption{\textbf{Conceptual diagram illustrating positive warming-shrub-fire feedback loop.} Top panel highlights the known positive feedback loop between fire, climatic warming and landscape flammability (solid blue lines). It also highlights several interactions (dashed \& black) which currently have a paucity of information but which may exacerbate or diminish this feedback loop. Bottom panel either assigns the direction or removes these interactions based on new evidence obtained from experimental and field surveys conducted in this study. It also includes an example of a potential, but unlikely, effect (red dotted line) that could break this feedback.}
\end{minipage}
\end{figure}

\begin{figure}
\label{fig2}
\begin{minipage}{\textwidth}
\includegraphics{figures/fig2.pdf}
\caption{\textbf{Mean \textit{Grevillea australis} seedling density with burning, fire severity, adult density, altitude and Topographic Wetness Index (TWI) gradients.} Top left panel = centered and standardized model coefficients (on the log scale). Top right panel is the effect of burning. All other panels are partial dependence plots for each covariate along their observed gradients in areas burnt by the 2003 fires. All errors bars are 95\% Bayesian Credible Intervals.}
\end{minipage}
\end{figure}

\begin{figure}
\label{fig3}
\begin{minipage}{\textwidth}
\includegraphics{figures/fig3.pdf}
\caption{\textbf{Observed and projected growth trajectories of four dominant Australian alpine shrubs.} Top panels: Thick lines represent mean ($\pm$ 95\% confidence intervals) observed heights at each May census. Thin lines represent individual growth trajectories. Bottom panels are the mean ($\pm$ 95\% Bayesian Credible Interval) projected height growth for each species over 70 years. In all cases, red or orange lines = seedlings growing in warmed (OTC) conditions and blue or light blue lines = seedlings growing in control conditions. Projections were based on an logistic growth model using initial seedling size of 6 cm (the mean initial height observed in the OTC experiment) and mean maximum heights observed in long-unburnt sites (i.e. 34, 65, 65 \& 70 cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Prostanthera} \& \textit{Phebalium}, respectively).}
\end{minipage}
\end{figure}

\begin{figure}
\label{fig4}
\begin{minipage}{\textwidth}
\includegraphics{figures/fig4.pdf}
\caption{\textbf{Mean ($\pm$ 95\% Bayesian Credible Interval) seedling probability of mortality for each species growing in control (blue) and warmed (red) conditions.}}
\end{minipage}
\end{figure}
\pagebreak

\begin{figure}
\label{fig5}
\begin{minipage}{\textwidth}
\includegraphics{figures/fig5.pdf}
\caption{\textbf{Experimental warming and inter-tussock gap size effects on \textit{Grevillea australis} seedlings.}First row = Growth rate effects, Second row = Mortality effects. Left panels are the centered and standardized model coefficients; right panels are the predicted growth and mortality rates along an inter-tussock gap size gradient warmed (red) and control (blue) conditions. All errors bars are 95\% Bayesian Credible Intervals.} 
\end{minipage}
\end{figure}
\pagebreak

\begin{figure}
\label{fig6}
\begin{minipage}{\textwidth}
\includegraphics{figures/fig6.pdf}
\caption{\textbf{Effects of warming on rates of tussock grass gap infilling.} Left panel = Model coefficients, Right panel = Predicted temporal change in size for an average 10 cm inter-tussock gap in warmed (red) and unwarmed (blue) conditions.} 
\end{minipage}
\end{figure}

\end{document}

### 2016-03-05 manuscript/supplementary.Rnw
Draft supplementary

\documentclass{nature}
\bibliographystyle{naturemag.bst}
\usepackage{lineno}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage[hidelinks]{hyperref}

\title{Climatic warming strengthens a positive feedback between alpine shrubs and fire (Supplementary Material)}
\author{James S. Camac\textasteriskcentered$^{1,2,3}$, Richard J. Williams$^4$, Carl-Henrik Wahren$^3$, Ary A. Hoffmann$^5$ \\
\& Peter A. Vesk$^2$}

\begin{document}
\linenumbers
\maketitle

\begin{affiliations}
\item Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia.
\item The Centre of Excellence for Environmental Decisions, School of BioSciences, The University of Melbourne, Parkville 3010, Victoria, Australia.
\item Research Centre for Applied Alpine Ecology, La Trobe University, Melbourne 3086, Victoria, Australia.
\item CSIRO Land and Water, tropical ecosystems research centre, PMB 44 Winnellie 0821, NT, Australia.
\item Bio21 Institute, School of BioSciences, The University of Melbourne, Vic. 3010, Australia.
\textasteriskcentered Email for correspondence: \texttt{james.camac@gmail.com}
\end{affiliations}

\pagebreak

\section*{Study site.}
Research sites were on the Bogong High Plains, a series of alpine and high subalpine plateaux in the Alpine National Park, approximately 250 km northeast of Melbourne, Australia. The undulating terrain ranges in altitude from 1400 to 2000 m a.s.l and contains a well-developed mantle of soil. Mean annual temperature is approximately 6\textdegree C, with persistent snow cover from June to September. Annual precipitation is 1200 mm, yet droughts occur at roughly decadal intervals\cite{Wahren:2013ji}. The vegetation is a mosaic of Eucalypt-woodland and treeless vegetation, with the latter largely consisting of heathlands and tussock grasslands, but also herbfields and wetlands. The treeline is irregular (due to cold air drainage), occurring at altitudes ranging from 1600 to 1700 m a.s.l. Approximately 50\% of the alpine ‘treeless’ landscape consists of heathlands, most of which were extensively burned by a 400,000 ha wildfire in 2003\cite{Williams:2006tl,Camac:2013eb}. Large wildfires in Australian alpine landscapes have occurred once or twice per century and typically occur during periods of extended regional drought coupled with severe fire weather and multiple ignitions in the surrounding montane forests\cite{Williams:2008ba}.

\section*{Open Top Chamber Experiment.}
Open Top Chamber Experiment
We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of grass recovery and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affect shrub seedling demographic rates, and 3) whether such affects change under warmer conditions post-fire conditions.

In alpine and tundra regions, shrub seedlings typically establish and grow in bare ground created by disturbance\cite{Williams:1987ua, Batllori:2009fe, Frost:2013tm}. Here, we used fire to create bare ground because 1) Australian alpine shrub seedlings occur in greatest abundance after fire (see seedling survey results) and 2) in many parts of the world, including Australia, wildfires are predicted to increase in frequency and severity\cite{Flannigan:2009kl, Westerling:2011jp}. Thus, to forecast likely changes in shrub abundance we need to know how warming affects growth and mortality rates of shrub seedlings growing in burnt sites.

We collected seedlings from a site burnt by a late 2006 fire because it allowed us to collect sufficient seedling numbers from species occurring in both open and closed heathland. We focused on \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) and \textit{Phebalium squamulosum} (Rutaceae; a tall shrub because they are all: obligate seeders post-fire (i.e. fire killed), common in the Australian Alps, and under global warming, likely candidates to invade non-shrubby plant communities such as alpine grasslands and herbfields. For each species, we collected a range of seedling heights (2 – 12 cm) using a forestry planter (12.5 cm deep by 5.5 cm wide). The size of the planter was selected to minimize root damage across the range of seedling sizes and to allow for the extraction of a constant volume (378 cm$^3$) of intact soil and roots. The planter also minimized soil disturbance during seedling extraction and replanting.

Four seedlings of each species were then transplanted into 70 cm$^2$ of burnt bare ground in the center of each experimental plot. An additional four \textit{Grevillea seedlings} were transplanted into various sized inter-tussock gaps surrounding this bare ground. Only \textit{Grevillea} seedlings were transplanted into inter-tussock gaps because of space constraints but also because it is the dominant open heathland shrub species (i.e. occurs closest to the shrub/grassland boundary) and is therefore the most likely to invade grasslands. The combination of bare ground and inter-tussock seedlings meant that each experimental plot had a total of 20 seedlings per 1 m$^2$ (Fig. S8). However, the two sets of seedlings (i.e. bare ground vs inter-tussock) were treated separately because they were spatially confounded (e.g. center vs edge of the 1 m$^2$ opening) and addressed different questions. Seedlings in bare ground patches (i.e. the center of the plot) were used to assess whether warmer post-fire conditions altered growth and mortality rates in four dominant alpine shrub species. By comparison, \textit{Grevillea} seedlings growing in inter-tussock gaps were used to determine whether inter-tussock gap size altered growth and mortality rates, and whether such effects were modified by warmer conditions.

To simulate near-term warmer conditions predicted under IPCC\cite{IPCC:2013tu}, we used Open Top Chambers (OTCs)\cite{Molau:1996vc}. OTCs were hexagonal, 58 cm tall with an open-top diameter of 110 cm and base diameter of 168 cm. The chambers were constructed following the International Tundra Experiment (ITEX) protocols\cite{Molau:1996vc} using sun-lite: a lightweight shatter-resistant fiberglass with high light transmittance (86\%) and low infrared transmittance (< 5\%; Solar Components Corporation, USA).

We measured changes in microclimate using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) randomly installed into four control plots and four OTC plots. Each Micro Station measured ambient temperature, relative humidity (both at 10 cm above ground), soil temperature (3 cm below ground) and volumetric soil moisture (between 3 and 10 cm) at hourly intervals for the entire duration of the experiment. Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions\cite{IPCC:2013tu} (Supplementary Fig. S5-7). OTCs passively raised mean daily ambient and soil temperatures by approximately 1 \textdegree C and 0.9 \textdegree C ($\pm$ 0.05 95\% CI) and marginally lowered both soil moisture (- 0.21\% $\pm$ 0.11) and relative humidity (- 1.2\% $\pm$ 0.17). OTCs also increased minimum daily ambient and soil temperatures by 0.86 \textdegree C ($\pm$ 0.06 95\% CI) and 1.6 ($\pm$ 0.12 95\% CI) respectively. Maximum temperatures were also strongly influenced by OTCs with ambient and soil maximum temperatures increasing by 2.4 \textdegree C ($\pm$  0.16 95\% CI) and 0.95 \textdegree C ($\pm$  0.42 95\% CI), respectively. Because OTCs strongly increase temperatures relative to other measures, throughout we refer to it as a ‘warming treatment’.

Natural \textit{Grevillea} recruits, likely from the soil seedbank, began establishing within the experimental plots by the end of the growing season in 2011. This provided us with the opportunity to examine how warming directly affected recruitment. We marked (with paint on stems) and recorded the number of recruits that germinated in both 2011 and 2012 seasons.

\section*{Seedling survey.}
We measured seedling density and maximum heights in open heathland across the Bogong High Plains. The primary aim of this study was to evaluate how variables such as altitude, Topographic Wetness Index (a measure of plant available water), adult density, fire, and, if burnt, fire severity, influence the density of \textit{Asterolasia} and \textit{Grevillea} seedlings in open heathland. A secondary aim was to validate growth rates observed in the OTC experiment by assessing maximum seedling heights of \textit{Asterolasia} and \textit{Grevillea} along altitude, soil moisture and fire severity gradients in areas burnt by the 2003 wildfires.

The study used 40 open heathland sites\cite{Camac:2013eb} established immediately post-2003 fires \cite{Williams:2006tl}. All sites have not been grazed by cattle since the 2003 fires. Thirty sites were burnt, while the remaining ten were unburnt, likely for more than 70 years\cite{Camac:2013eb}. This disparity in sampling effort between burnt and unburnt was due to very few open heathland remaining unburnt after both the 2003 and 2007 fires. These sites encompassed most of the Bogong High Plains (120 km$^2$) and ranged in altitudinal from 1660 m to 1860 m a.s.l. Sites were at least 0.25 ha and more than 500 m apart. In the summer of 2012-13, 9-years post-fire, seedling density in each site was estimated using 40 1 m$^2$ quadrats, evenly distributed in groups of 10 along four 50 m transects. Transects were positioned in the center of each site and were parallel to each other with exactly 10 m between transect lines. This effectively created a grid of 40 plots (10 plots \texttimes 4 transects) per site, subsampling an area of 2000 m$^2$.

Different variables were recorded at the three sampling levels (site, transect and plot). For each site, altitude and TWI were derived from a 30 m resolution digital elevation model of the Bogong High Plains. TWI, which approximates the relative plant-available water based on a site’s position in the landscape\cite{Moore:1993fk}, was derived for each site using $TWI= ln(\frac{a\alpha}{tan\theta})$, where, $\alpha$ is the upslope catchment area per unit width orthogonal to the flow direction and $\theta$ is the slope (in degrees). Across the 40 sites, TWI varied twofold (5 – 10 units). Altitude ranged by 200 m (from 1660 to 1860 m a.s.l), or by approximately 1.5\textdegree C from the lowest to the highest site (based on a lapse-rate of 0.77\textdegree C/100 m elevation\cite{Brown:1989vx}. This altitudinal temperature range is comparable to that observed in the OTC experiment (1.5\textdegree C vs 1.2\textdegree C, respectively).

The occurrence and severity of fire was initially determined immediately after the 2003 fire\cite{Williams:2006tl}. The severity of the fires in each burnt site was estimated using twig diameters, a proxy measure of fire severity\cite{Whight:1999vw}. These twig diameters were recorded between 1 and 10 months post-fire from 10 replicate \textit{Grevillea australis} individuals. Across the 30 burnt sites, the average twig diameter ranged from 2 to 13 mm, with smaller diameters assumed to have experienced lower fire intensity. This encompassed the full range of fire severity from lightly scorched sites containing shrubs with only the tips of the outer branches burnt, to severely burnt sites where the majority of large shrub branches were consumed.

\section*{Data analysis.}
In total we built 11 hierarchical models to examine factors influencing tussock grass post-fire recovery and shrub seedling establishment, growth and mortality. These models are well suited to datasets that have a hierarchical structure (e.g. plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset\cite{Gelman:2007te, Kery:2012wl}. For each model we used Bayesian inference and fitted models with Markov chain Monte Carlo (MCMC) sampling in R 3.2.2 using package Rstan 2.8.1.  For each model, three independent chains were executed. In all cases modelled parameters converged within 1000 iterations. Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic\cite{Gelman:1992ts, Brooks:1998ju}. After discarding the first 1000 iterations as ‘warm up’, a further 1000 iterations were taken from the joint posterior. Below we describe the model structure, parameters and priors used in these models. The residuals of all models were checked graphically against predictions with none showing any systematic pattern or severe heteroscedasticity. To ensure all covariate effects were identifiable we selectively removed each covariate and visually inspected whether the direction and magnitude of effects were severely altered; none were. Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine-Shrub-Experiment}.

\subsection{\textit{Modelling shrub seedling growth rates.}}
For the purpose of modelling seedling annual growth rates we used heights measured in May of each year from seedlings that survived throughout the passive warming experiment. This meant that our observations were of maximum height for each individual $i$ at each May census $t$. We modelled these heights, $height_{it}$, as a random realization from a normal distribution truncated at zero:

\begin{equation}\label{eq:growth_obs} 
height_{it} \sim \mathcal{N^+}(\widehat{height}_{it}, \sigma_{obs}),
\end{equation}

where $\widehat{height}_{i,t}$ is the estimated maximum height for each individual $i$ at each May census $t$ and $\sigma_obs$ is the standard deviation of the observation error. The parameter $\widehat{height}_{i,t}$ was estimated using a logistic parameterisation of the unified Richards equation\cite{Tjorve:2010bk}:

\begin{equation}\label{eq:logistic} 
\widehat{height}_{i,t} = \frac{Hmax_{j[i]}}{\left( 1 + \left(\frac{Hmax_{j[i]}}{Hinit_i} - 1\right)e^{R_i \frac{Days_t}{365.25}}\right)},
\end{equation}

We used a logistic equation because it can simulate sigmoidal growth. That is, a curve whereby initial seedling growth is slow before increasing in speed and finally levelling off when adult height is reached. We used this particular parameterisation because it accounts for seedling size at the beginning of an experiment and contained biologically interpretable parameters. Here, $Hmax_{j[i]}$ is the asymptotic height: the average mature height an individual $i$ belonging to species $j$ can achieves. We treated Hmax as a constant, not to be estimated by the model, owing to our experimental observations being restricted to the first five years of growth. Based on maximum height data from 5 to 22 long-unburnt sites (depending on the species) we estimated the average mature height (and thus Hmax) to be approximately 34, 67, 66 and 70 cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera}, respectively. $Hinit_i$ is the height of individual $i$ at the start of the experiment (i.e. the first May measurement). $Days_t$ is the number of days from the beginning of the experiment to each May census $t$.  The experiment ran for a total of 1812 days (or approximately 5 years). We divide the number of days at each census by 365.25 in order to estimate annual growth rates. The parameter of interest, $R_i$, is the mean growth rate constant for each individual $i$. This parameter controls the slope at inflection (i.e. the maximum annual relative growth rate). When multiplied by the asymptotic height, $Hmax$, it yields the maximum annual growth rate at the inflection point . We modelled $R_i$ separately for seedlings growing in bare ground (eqn. 1.3.1) compared to those growing in inter-tussock gaps (eqn. 1.3.2).

For individual seedlings of the four shrub species growing in the bare ground patch located in the centre of each plot we modelled $R_i$ using a varying intercept linear model:

\begin{equation}\label{eq:r_bg} 
\begin{aligned}
R_i &=  \alpha_{j[i]} + \beta_{j[i]} \times OTC_{k[i]} + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\alpha_j &\sim \mathcal{N}(\mu_{spp}^{CTL}, \sigma_{spp}^{CTL}) \\
\beta_j &\sim \mathcal{N}(\mu_{spp}^{OTC}, \sigma_{spp}^{OTC}) \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{obs_{ind}}).
\end{aligned}
\end{equation}

Here, the intercept term, $alpha_{j[i]}$, represents the expected control plot growth constant of species $j$ to which individual $i$ belongs. It is estimated from a normal distribution centred on $\mu_{spp}^{CTL}$ (the expected control plot growth rate for an average hypothetical species) with a standard deviation,$\sigma_{spp}^{CTL}$ which defines the variation between species’ growth rates at control plots. The parameter $\beta_{j[i]}$ is the OTC effect for each species $j$ to which individual $i$ belongs. $\beta_j$ is also estimated from a normal distribution, but unlike $\alpha_j$, it is centred on $\mu_{spp}^{OTC}$, (the expected OTC growth constant for an average hypothetical species) with a standard deviation, $\sigma_{spp}^{OTC}$ which defines the variation between species’ growth constants within OTCs. The terms $\epsilon_{plot_{k[i]}}$ and $\epsilon_i$ represent residual variation at the plot level (for each plot, $k$, within which individual $i$ is located), and the individual level (for each individual $i$), respectively. Both error terms were drawn from normal distributions centred on zero with standard deviations that defined the variation amongst plots, $\sigma_{plot}$, or amongst individuals, $\sigma_{obs_{ind}}$, respectively.

In contrast, for \textit{Grevillea australis} seedlings growing in inter-tussock gaps, the growth constant was modelled using a fixed intercept linear model:

\begin{equation}\label{eq:r_tussock} 
\begin{aligned}
R_{i,t} &=  \alpha + \beta_1 \times OTC_{k[i]} + \beta_2 \times Gap_{i,t} + \beta_3 \times (OTC_{k[i]} \times Gap_{i,t}) + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{obs_{ind}}).
\end{aligned}
\end{equation}

The structure and parameter definitions differ slightly to that in eqn 1.3.1. Here, the model has two additional parameters, $\beta_2$ and $\beta_3$. The parameter $\beta_2$ is the effect of inter-tussock gap radius, which varied over time $t$, while $\beta_3$ is the interactive effect of OTC and time-varying inter-tussock gap radius. The inclusion of these parameters, coupled $Gap_{i,t}$ being centered and standardized (see Data inputs into growth and mortality models), alters the definition of the intercept, $\alpha$, which now represents the mean control plot growth rate at the average inter-tussock gap size (12 cm). The parameter  $\beta_1$ is the OTC effect and $\epsilon_{plot_k}$ and $\epsilon_i$ are residual errors at the plot and individual level, respectively. Note that growth constant, $R$, now varies through time as a function of inter-tussock gap radius.

We used Bayesian inference and hence needed to specify prior distributions for all model parameters. In both growth models standard deviation parameters (i.e. $\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{ind}$) were all estimated from weakly informative positive half Cauchy priors with scale parameters of 25. We used these Cauchy priors in order to prevent over-inflated estimates of variances due to the small number of plots and species\cite{Gelman:2006uw}. In both sets of models the intercept terms ($\mu_{spp}^{CTL}$, $\alpha$), coefficients (i.e. inter-tusscok model: $\beta_1$, $\beta_2$ and $\beta_3$) and hyperparameter (i.e. multispecies model) for the mean OTC effect $\mu_{spp}^{OTC}$ were estimated from normal priors centered on zero with a scale parameter of 0.5. We chose this prior as it spanned the range of likely growth trajectories from very slow to very fast growing.

\subsection{\textit{Modelling shrub seedling mortality rates.}}
We were also interested in whether seedling instantaneous mortality rates varied amongst shrub species, OTC treatment and inter-tussock gap size. Our observations, $y_{i,t}$, were binary (0 = alive or 1 = dead) for each individual $i$, in each census, $t$. We modelled these binary observations through time as random realizations from a Bernoulli distribution based on the model-estimated cumulative hazard rate for each individual at each census:

\begin{equation}\label{eq:mort_obs} 
y_{i,t} \sim Bern(Cumulative\;hazard_{i,t}).
\end{equation}

As with growth rates, we modelled each individual's cumulative hazard per census, $Cumulative\;hazard_{i,t}$, separately for seedlings growing in bare ground patches to those in inter-tussock gaps. In both cases, $Cumulative\;hazard_{i,t}$, was modelled on the complementary log-log scale as a function of an instantaneous hazard rate for each individual, $\lambda_i$, multiplied by the length (in years) between each census, $t$. By doing this our regression model becomes equivalent to a parametric survival analysis\cite{Kleinbaum:2012tn}.

\begin{equation}\label{eq:cum_hazard} 
\log_e(-\log_e(1 - Cumulative\;hazard_{i,t}) = \lambda_i \times census\;length_t.
\end{equation}

We modelled individual instantaneous hazard rate, $\lambda_i$, in the same way we modelled $R_i$ in equations 1.3.1 \& 1.3.2. 
For individual seedlings of the four shrub species located in the bare ground patch of each plot, we modelled $\lambda_i$ using a varying intercept model:

\begin{equation}\label{eq:mort_bg} 
\begin{aligned}
\lambda_i &=  \alpha_{j[i]} + \beta_{j[i]} \times OTC_{k[i]} + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\alpha_j &\sim \mathcal{N}(\mu_{spp}^{CTL}, \sigma_{spp}^{CTL}) \\
\beta_j &\sim \mathcal{N}(\mu_{spp}^{OTC}, \sigma_{spp}^{OTC}) \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{obs_{ind}}).
\end{aligned}
\end{equation}

Here, the intercept term, $\alpha_{j[i]}$, represents the expected instantaneous hazard rate in a control plot for species $j$ within which individual $i$ belongs. It is estimated from a normal distribution centred on $\mu_{spp}^{CTL}$  (the control plot expected instantaneous hazard rate for the average hypothetical species) with a standard deviation, $\sigma_{spp}^{CTL}$, which defines the variation between species’ instantaneous hazard rates in control plots. The parameter $\beta_{j[i]}$ is the OTC effect for each species $j$ to which individual $i$ belongs. $\beta_{j[i]}$ is also estimated from a normal distribution, but unlike $\alpha_j$, it is centered on $\mu_{spp}^{OTC}$, (the OTC expected probability of death for the average hypothetical species) with a standard deviation, $\sigma_{spp}^{OTC}$, which defines the variation between species’ instantaneous hazard rates in OTCs. The parameters $\epsilon_{plot_k}$ and $\epsilon_i$ are residual (unexplained) variation at the plot and individual level, respectively.

In contrast, $\lambda_{i,t}$ for \textit{Grevillea} seedlings growing in various sized inter-tussock gaps was modelled using a fixed intercept linear model:

\begin{equation}\label{eq:mort_tussock} 
\begin{aligned}
\lambda_{i,t} &=  \alpha + \beta_1 \times OTC_{k[i]} + \beta_2 \times Gap_{i,t} + \beta_3 \times (OTC_{k[i]} \times Gap_{i,t}) + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{obs_{ind}}).
\end{aligned}
\end{equation}

Here, the structure and parameter definitions are similar to eqn 2.3.1. The intercept, $\alpha$, represents the expected instantaneous hazard rate in control plots in an average inter-tussock gap radius, $\beta_1$ is the effect of OTC, $\beta_2$ is the effect of inter-tussock gap radius, $\beta_3$ is the interactive effect of OTC and inter-tussock gap radius and $\epsilon_{plot_k}$ and $\epsilon_i$ are the residual variation at the plot and individual scales.

Like the growth models, the standard deviation parameters in both mortality models (i.e. $\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{ind}$) were each estimated from weakly informative positive half Cauchy priors with scale parameters of 25. The intercept term $\alpha$ for inter-tussock seedling model and the mean hyper-parameter $\mu_{spp}^{CTL}$ in the multispecies models were estimated from cauchy priors with a scale parameter of 2.5. We also used these Cauchy priors for inter-tussock model coefficients (i.e. $\beta_1$, $\beta_2$ and $\beta_3$) and the multi-species hyperparameter for the mean OTC effect $\mu_{spp}^{OTC}$. These Cauchy priors are the recommended default choice for logistic regression models\cite{Gelman:2008ua}.

\subsection{\textit{Modelling Grevillea recruitment rates.}}
We examined whether warming influenced natural recruitment rates of \textit{Grevillea} by modelling the number of recruits per 1 m$^2$ plot $k$ in each census, $t$ as a random realization from a negative binomial distribution:

\begin{equation}\label{eq:recruit_obs} 
recruits_{k,t} \sim NegBinomial(\mu_{k,t}, \phi),
\end{equation}

where $\mu_{k,t}$ is the log expected recruit count for each plot, and $\phi$ is the overdispersion parameter. We used a negative binomial distribution because count data was significantly overdispersed with $\phi$ being estimated as 3.95.

We modeled $\mu_{k,t}$ on the log scale using a hierarchical model:

\begin{equation}\label{eq:recuit_mod} 
\begin{aligned}
\log_e(\mu_{k,t}) &=  \alpha + \beta_1 \times OTC_k + \beta_2 \times Census_t + \epsilon_{plot_k} \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}),
\end{aligned}
\end{equation}

where, the intercept, $\alpha$, represents the average log recruitment in control plots in 2011, assuming all other parameters are at their mean. The parameter $\beta_1$ is the effects of OTC and $\beta_2$ describes the change in recruit numbers from 2011 to 2012. $\epsilon_{plot_k}$ is the residual variation at the plot scale.

Half Cauchy priors with a scale of 25 were used for both the overdispersion parameter, $\phi$ and the plot level standard deviation $\sigma_{plot}$. Coefficients $\beta_1$ and $\beta_2$ were estimated using normal priors centered on zero with a standard deviation of 2.5, encapsulating our prior knowledge that plots were unlikely to have more than 150 recruits/m$^2$.

\subsection{\textit{Data inputs for growth, mortality and recruitment models.}}
In each model, OTC treatment, $OTC_k$, was a binary (0 = Control, 1 = OTC) plot level covariate. By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census t. However, because some cardinal distances were recorded as censored data (i.e.  > 25 cm), these data were incomplete. We estimated these censored data using a censored model (see below). Once censored data were estimated, we then estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each censuses, for each inter-tussock seedling. Median inter-tussock gap radius ranged from 0.5 cm to 43 cm in both control and OTC plots. In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero (minus grand mean) and divided by two standard deviations\cite{Gelman:2007te}. The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as average responses and slope terms as partial dependencies conditional on other continuous variables at their mean. Furthermore, standardizing by two standard deviations allows the magnitude of effects to be directly compared between binary and continuous covariates\cite{Gelman:2007te}.

\subsection{\textit{Modelling censored Poa distances.}}
Approximately 15\% of the Poa cardinal distance data used to calculate the average inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, ≥ 26 cm.  We estimated these censored values by first estimating the distribution of distances in observed data, coupled with estimating covariate effects and then using these distributions to estimate censored observations.

Specifically, we first modelled the observed (non-censored) cardinal distances as random realizations from a lognormal distribution:

\begin{equation}\label{eq:distance_obs} 
distance_i \sim LN(\mu_i, \sigma_{obs}),
\end{equation}

where $\mu_i$ is the expected log estimated distance for cardinal distance $i$, and  $\sigma_obs$ is the standard deviation of the observation error.

We modelled the log estimated distance, $\mu_i$, using hierarchical linear regression:

\begin{equation}\label{eq:censored_mod} 
\begin{aligned}
\log_e(\mu_{k,t}) &=  \alpha + \beta \times OTC_{k[i]} + \epsilon_{census_{t[i]}} + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\epsilon_{census_t} &\sim \mathcal{N}(0, \sigma_{census}) \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{ind}).
\end{aligned}
\end{equation}

This regression model contained the effect of OTC ($\beta$) and estimated unexplained variation at the census level ($\epsilon_{census_t}$), plot level ($\epsilon_{plot_k}$) and inter-tussock gap level ($\epsilon_i$).

Again, like other models, half Cauchy priors with a scale of 25 were used for all standard deviation parameters. Both the intercept, $\alpha$, and the effect of OTC, $\beta$ (which are on the log scale), were estimated from weakly informed normal distributions centered on zero with a standard deviation of 2.5. Specifically, this prior encapsulated our knowledge that distances were very unlikely to be > 150 cm.

Using this model, coupled with our prior knowledge that the distances could not be less than 26 cm or larger than 100 cm (all plots were surrounded by grass) we could estimate the censored distances using a lognormal distribution truncated at both 26 cm and 100 cm:

\begin{equation}\label{eq:censored_obs} 
censored\;distance_i \sim LN(\mu_i, \sigma_{obs})[26,100],
\end{equation}

The mean posteriors from the censored data were then extracted from the model and used as our estimates for each censored observation. This complete dataset was then used subsequent analyses (e.g. inter-tussock seedling growth and mortality, changes in gap size \textemdash see below).

\subsection{\textit{Modelling rates of inter-tussock gap infilling by Poa.}}
As well as modelling how OTC affected shrub seedling growth and mortality rates, we were also interested in examining how warming influenced rates of grass infilling of inter-tussock gaps. To examine this we first modelled the median gap radius,$y_{i,t}$, for each inter-tussock gap $i$ in each census $t$, as a random realisation from a lognormal distribution:

\begin{equation}\label{eq:gap_obs} 
y_{i,t} \sim LN(\mu_{i,t}, \sigma_{obs}),
\end{equation}

where $\mu_{i,t}$ is the log estimated median radius for each inter-tussock gap, at each census, and  $\sigma_{obs}$ is the standard deviation of the observation error.

We then modelled $\mu_{i,t}$, as:

\begin{equation}\label{eq:log_gap_mod} 
\mu_{i,t} = \log_e(\alpha_i) + \beta_i \times \frac{Days_t}{365.25},
\end{equation}

which on the raw (non-logged scale) is equivalent to a standard exponential curve:

\begin{equation}\label{eq:gap_mod} 
\mu_{i,t} = \alpha_i\,e^{\beta_i\frac{Days_t}{365.25}}.
\end{equation}

This functional form allows gap radius to either increase or decrease exponentially. We did not force this curve to be a negative exponential (i.e. gaps decreased over time) because it is theoretically possible for inter-tussock gaps to increase over time. Here $\alpha_i$ is a known constant that represents the initial inter-tussock gap radius for each gap $i$ at the beginning (i.e. the intercept). $\beta_i$ represents the exponential rate for each gap and $Days_t$ is the number of days from the beginning of the experiment.

We estimated $\beta_i$ using a hierarchical linear regression:

\begin{equation}\label{eq:gap_rate} 
\begin{aligned}
\beta_i &= \alpha_0 + \beta \times OTC_{k[i]} + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{ind}).
\end{aligned}
\end{equation}

Here, $\alpha_0$ is the average rate of gap radius change in control plots, $\beta$ is the effect of OTC and $\epsilon_{plot_k}$ and $\epsilon_i$ represent the plot and gap level residual errors. 
Like all other models used in this study, we used half Cauchy priors with a scale of 25 for all standard deviation parameters. The intercept, $\alpha_0$, and the effect of OTC, $\beta$ were estimated from normal priors centered on zero with a standard deviation of 5. These priors encapsulated the full range of potential gap trajectories from increases or decreases or no change.

\subsection{\textit{Modelling drivers of seedling density.}}
We examined drivers of seedling density by modelling the number of seedlings per 1 m$^2$ as a random realization from a negative binomial distribution:

\begin{equation}\label{eq:density_obs} 
count_i \sim NegBinomial(\mu_i, \phi),
\end{equation}

where $\mu_i$ is the log expected seedling count for each plot, and $\phi$ is the overdispersion parameter. We used a negative binomial distribution because count data was significantly overdispersed for both species with $\phi$ being estimated as 0.54 ($\pm$ 0.1 95\% BCI) and 0.82 ($\pm$ 0.1 95\% BCI), for \textit{Grevillea} and \textit{Asterolasia}, respectively.

$\mu_i$ was modelled separately for \textit{Grevillea australis} and \textit{Asterolasia trymalioides} because the latter contained no measurements of adult density. The two models are structurally similar and only differ in that parameter $\beta_5$ is not included for \textit{Asterolasia}. Below we outline the full model structure used to estimate $\mu_i$.

We modeled $\mu_i$ on the log scale as a function of both site and transect covariates:

\begin{equation}\label{eq:density_mu_mod} 
\mu_i = \delta_{j[i]} + \gamma_{k[i]}
\end{equation}

where, $\delta_{j[i]}$ and $\gamma_{k[i]}$ represents the average log count for each site $j$ and transect $k$ within which plot $i$ is located, respectively.

We modelled $\delta_j$ using a varying intercept linear model:

\begin{equation}\label{eq:density_site} 
\begin{aligned}
\delta_j &= \alpha_j + \beta_1 \times Alt_j + \beta_2 \times TWI_j + \beta_3 \times (1 - Burnt_j) + \beta_4 \times (Severity_j \times Burnt_j) \\
\alpha_j &\sim \mathcal{N}(\mu_{site}, \sigma_{site}),
\end{aligned}
\end{equation}

where the intercept, $\alpha_j$, varies by site and represents the average log count per plot at each site, assuming all other parameters are at their mean. We estimated $\alpha_j$ from a normal distribution centred on $\mu_{site}$, the average seedling count per 1 m$^2$ plot for the average site. The parameters $\beta_1$ and $\beta_2$ are the effects of altitude and wetness, respectively, while $\beta_3$ is the effect of being an unburnt site and $\beta_4$ is the effect of fire severity in burnt sites.

We modelled $\gamma_k$ as:

\begin{equation}\label{eq:density_transect} 
\begin{aligned}
\gamma_k &= \beta_5 \times ADen_k + \epsilon_{tran_k}  \\
\epsilon_{tran_k} &\sim \mathcal{N}(0, \sigma_{tran}) \\.
\end{aligned}
\end{equation}

where $\beta_5$ is the effect of mature \textit{Grevillea} counts and $\epsilon_{tran_k}$ is the random error associated with transect $k$. Parameter $\beta_5$ was not included when modelling \textit{Asterolasia} seedling counts because of a lack of mature individual counts. For this species $\gamma_k = \epsilon_{tran_k}$. Interactions between covariates were also examined, however, we detected no significant interactions for either species.

Half Cauchy priors with a scale of 25 were used for all standard deviation parameters and the overdispersion parameter $\phi$. Hyperparameter $\mu_site$ and coefficients $\beta_1$ to $\beta_5$ were estimated using normal priors centered on zero with a standard deviation of 2.5, encapsulating our prior knowledge that plots were unlikely to have more than 150 seedlings/m$^2$.

\subsection{\textit{Modelling drivers of seedling maximum heights.}}
In order to validate growth rate findings in the OTC experiment, we assessed how an elevation range of 200 m in elevation (equivalent to a temperature change of 1.5\textdegree C \cite{Brown:1989vx}) influenced \textit{Grevillea} and \textit{Asterolasia} seedlings heights in areas burnt by the 2003 fires. We also assessed whether TWI and fire severity influenced these heights. Here, our observations were of maximum seedling heights for both species when at least one individual was present in a plot. We used maximum heights because these were likely to represent individuals that established immediately post-fire, and would have had 9 years of post-fire growth. We modelled these heights, $maxht_i$, as a random realization from a log-normal distribution:

\begin{equation}\label{eq:maxht_obs} 
maxht_i \sim LN(\widehat{\mu}_i, \sigma_{obs}),
\end{equation}

where $\widehat{maxht}_i$ is the model estimated log maximum height for each individual $i$ and $\sigma_{obs}$ is the standard deviation of the observation error. We modelled $\widehat{maxht}_i$ separately for \textit{Grevillea} and \textit{Asterolasia} using a varying intercept linear model:

\begin{equation}\label{eq:maxht_mod} 
\begin{aligned}
\widehat{maxht}_i &= \alpha_j + \beta_1 \times Alt_j + \beta_2 \times TWI_j + \beta_3 \times Severity_j \\
\alpha_j &\sim \mathcal{N}(\mu_{site}, \sigma_{site}),
\end{aligned}
\end{equation}

where the intercept, $\alpha_j$, varies by site and represents the average log maximum seedling height of each site, assuming all other parameters are at their mean. We estimated $\alpha_j$ from a normal distribution centred on $\mu_{site}$, the average log maximum height for the average site and a standard deviation of $sigma_{site}$ which defines site variability. The parameters $\beta_1$ and $\beta_2$ are the effects of altitude and topographic wetness, respectively and $\beta_3$ is the effect of fire severity.

Here, we estimated $\mu_{site}$ and coefficient parameter ($\beta_1$ to $\beta_3$) using weakly informative normal priors with a mean of 0 and a standard deviation of 2.5. These priors encapsulated our prior knowledge that seedling maximum heights would be less than 150 cm. All standard deviation parameters were modelled using positive half Cauchy priors with a scale of 25.

\subsection{\textit{Data inputs into density and maximum height models.}}
In all seedling survey models, $Burnt_j$ is a binary (0 = unburnt, 1 = burnt) site level covariate, which only affects seedling counts at unburnt sites. By contrast, fire severity $Severity_j$, was a continuous site level covariate that only affected seedling counts and maximum heights at burnt sites. Altitude ($Alt_j$), Topographic Wetness Index ($TWI_j$) and mature \textit{Grevillea} count ($ADen_k$) were also continuous covariates that affected all sites in both count and maximum height models. Collinearity among covariates was checked and all correlations were r < 0.16. The data from each continuous covariate was then centered on zero and standardized by two standard deviations\cite{Gelman:2007te}. 
\pagebreak
\section*{Figures}
\begin{suppfigure}
\label{figs1}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS1.pdf}
\caption{\textbf{Mean \textit{Asterolasia trymalioides} seedling density with burning, fire severity, adult density, altitude and Topographic Wetness Index (TWI) gradients.} Top left panel = centered and standardized model coefficients (on the log scale). Top right panel is the effect of burning. All other panels are partial dependence plots for each covariate along their observed gradients in areas burnt by the 2003 fires. All errors bars are 95\% Bayesian Credible Intervals. Phi = overdispersion parameter for negative binomial distribution.}
\end{minipage}
\end{suppfigure}
\pagebreak

\begin{suppfigure}
\label{figs2}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS2.pdf}
\caption{\textbf{Number of \textit{Grevillea australis} recruits in warmed (red) and control (blue) conditions in both 2011 and 2012. Errors bars are 95\% Bayesian Credible Intervals.}}
\end{minipage}
\end{suppfigure}
\pagebreak

\begin{suppfigure}
\label{figs3}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS3.pdf}
\caption{\textbf{Observed and projected diameter growth trajectories of four dominant Australian alpine shrubs.} Thick lines represent mean ($\pm$ 95\% CI) observed heights at each May census. Thin lines represent individual growth trajectories. Red or orange lines = seedlings growing in warmed (OTC) conditions and blue or light blue lines = seedlings growing in control conditions.} 
\end{minipage}
\end{suppfigure}
\pagebreak

\begin{suppfigure}
\label{figs4}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS4.pdf}
\caption{\textbf{Mean maximum seedling height along fire severity (TD; twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires.} Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses. The first row shows the centered and standardized model coefficients (on the log scale) for both species. Rows 2 to 4 are partial dependence plots for each covariate along their observed gradients. All errors bars are 95\% Bayesian Credible Intervals.} 
\end{minipage}
\end{suppfigure}
\pagebreak

\begin{suppfigure}
\label{figs5}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS5.pdf}
\caption{\textbf{Mean daily difference (OTC-Control) for ambient temperature, soil temperature, volumetric water content and relative humidity over experimental period.} Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls. Shaded areas are the winter snow periods when OTCs were not over plots. Note: amplitude changes in later years are due to shading effects caused by increased shrub growth.}
\end{minipage}
\end{suppfigure}
\pagebreak

\begin{suppfigure}
\label{figs6}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS6.pdf}
\caption{\textbf{Mean minimum daily difference (OTC-Control) for ambient temperature, soil temperature, volumetric water content and relative humidity over experimental period.} Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls. Shaded areas are the winter snow periods when OTCs were not over plots. Note: amplitude changes in later years are due to shading effects caused by increased shrub growth.}
\end{minipage}
\end{suppfigure}
\pagebreak

\begin{suppfigure}
\label{figs7}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS7.pdf}
\caption{\textbf{Mean maximum daily difference (OTC-Control) for ambient temperature, soil temperature, volumetric water content and relative humidity over experimental period.} Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls. Shaded areas are the winter snow periods when OTCs were not over plots. Note: amplitude changes in later years are due to shading effects caused by increased shrub growth.}
\end{minipage}
\end{suppfigure}
\pagebreak

\begin{suppfigure}
\label{figs8}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS8.pdf}
\caption{\textbf{Diagram outlining positions of transplanted seedling in an OTC plot.} The inner white square = the burnt bare ground patch used to assess heights of four dominant alpine shrub species. The grey area contains \textit{Grevillea australis} seedlings randomly transplanted into burnt inter-tussock gaps. The area beyond the grey section was subject to OTC panel sheltering (i.e. it was not within the 110 cm opening of the chamber) and therefore was not used in the experiment.} 
\end{minipage}
\end{suppfigure}
\pagebreak
\section*{References}
\bibliography{refs.bib}
\end{document}
\pagebreak

### 2016-03-07 manuscript/manuscript.Rnw
Filled in most key values. Still more to do

@@
 Here we use field surveys and experimental manipulations to examine warming and fire effects on recruitment, growth and survival of seedlings of evergreen obligate seeding alpine shrubs.
-We find that fire substantially increased shrub seedling establishment (up to 36 fold) and that warming doubled tall shrub seedling growth rates and may increase survival.
+We find that fire substantially increased shrub seedling establishment (up to \Sexpr{density_ratios(greaus_seedling_density_model, 'Grevillea')}) and that warming doubled tall shrub seedling growth rates and may increase survival.
 Warming had weak or no effect on shrub recruitment, rates of grass post-fire gap infilling or grass competitive effects on shrub seedling growth and survival.
@@
 \subsection{Drivers of shrub seedling establishment.}
-
 We first investigated how altitude, topographic wetness (TWI), adult density, fire and, if burnt, fire severity, influenced the density of \textit{Grevillea} and \textit{Asterolasia} seedlings (two dominant shrubs of open heathland).
@@
 S1) seedlings were strongly influenced by the occurrence of fire.
-Sites burnt by fire in 2003 had seedling densities between 13 and 36 times higher than unburnt sites.
-In \textit{Grevillea}, mean seedling densities were 1.44/m$^2$ at burnt sites and 0.04/m$^2$ at unburnt sites.
-\textit{Asterolasia} had similar mean densities: 1.71 and 0.13 seedlings/m$^2$ at burnt and unburnt sites, respectively.
+Sites burnt by fire in 2003 had seedling densities between \Sexpr{density_ratios(asttry_seedling_density_model,'Asterolasia')} and \Sexpr{density_ratios(greaus_seedling_density_model,'Grevillea')} times higher than unburnt sites.
+In \textit{Grevillea}, mean seedling densities were \Sexpr{density_means(greaus_seedling_density_model,'Grevillea', TRUE)}/m$^2$ at burnt sites and \Sexpr{density_means(greaus_seedling_density_model,'Grevillea', FALSE)}/m$^2$ at unburnt sites.
+\textit{Asterolasia} had similar mean densities: \Sexpr{density_means(asttry_seedling_density_model,'Asterolasia', TRUE)} and \Sexpr{density_means(asttry_seedling_density_model, 'Asterolasia', FALSE)} seedlings/m$^2$ at burnt and unburnt sites, respectively.
 For both species, seedling density was also higher in severely burnt sites.
@@
 The field warming experiment (see below) indicates that this lack of elevation response in \textit{Grevillea} is due to both rates of recruitment (Supplementary Fig.
-S2) and mortality (see below) largely being insensitive to a 1\textdegree C change in temperature.
+S2) and mortality (see below) largely being insensitive to a \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C change in temperature.
 We detected no significant effect of topographic wetness for either species.
@@
 Knowing shrub seedlings were more abundant in burnt vegetation, we investigated whether warmer post-fire conditions affected seedling growth and mortality rates.
-Seedlings of \textit{Grevillea}, \textit{Asterolasia}, \textit{Prostanthera} and \textit{Phebalium} emerging after a wildfire were transplanted into experimentally burnt plots that were either passively warmed by 1\textdegree C using open top chambers (OTCs) or subject to current ambient conditions (controls).
-After 1812 days (1087 growing season days where OTC treatment was applied) or 5 years growth, mean seedling heights of the tall shrubs (\textit{Grevillea}, \textit{Prostanthera} and \textit{Phebalium}) growing in post-fire bare ground were taller in warmed plots by 9.7, 13.5 and 3.8 cm respectively (Fig.
+Seedlings of \textit{Grevillea}, \textit{Asterolasia}, \textit{Prostanthera} and \textit{Phebalium} emerging after a wildfire were transplanted into experimentally burnt plots that were either passively warmed by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C using open top chambers (OTCs) or subject to current ambient conditions (controls).
+After \Sexpr{n_days(chamber_dates, FALSE)} days \Sexpr{n_days(chamber_dates, TRUE)} growing season days where OTC treatment was applied) or \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} years growth, mean seedling heights of the tall shrubs (\textit{Grevillea}, \textit{Prostanthera} and \textit{Phebalium}) growing in post-fire bare ground were taller in warmed plots by \Sexpr{ht_diff(non_tussock_otc_growth_data,'Grevillea')}, \Sexpr{ht_diff(non_tussock_otc_growth_data,'Prostanthera')} and \Sexpr{ht_diff(non_tussock_otc_growth_data,'Phebalium')} cm respectively (Fig.
 \ref{fig3}).
@@
 S3).
-Accounting for initial height and assuming logistic growth, mean annual predicted height growth rates of \textit{Grevillea}, \textit{Prostanthera} and \textit{Phebalium} were 2.4, 1.9 and 1.5 times that observed in control plots, respectively.
-Projecting these differences equates to an average 6 cm tall seedling (the mean initial size used in this experiment) growing in warmed conditions reaching maximum mature height in 39 fewer years for \textit{Grevillea} and 17 fewer years for both \textit{Prostanthera} and \textit{Phebalium} (Fig.
+Accounting for initial height and assuming logistic growth, mean annual predicted height growth rates of \textit{Grevillea}, \textit{Prostanthera} and \textit{Phebalium} were \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Grevillea')}, \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Prostanthera')} and \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Phebalium')} times that observed in control plots, respectively.
+Projecting these differences equates to an average 6 cm tall seedling (the mean initial size used in this experiment) growing in warmed conditions reaching maximum mature height in \Sexpr{hmax(non_tussock_otc_growth_model,'Grevillea')} fewer years for \textit{Grevillea} and \Sexpr{hmax(non_tussock_otc_growth_model,'Prostanthera')} fewer years for both \textit{Prostanthera} and \textit{Phebalium} (Fig.
 \ref{fig3}).
-\textit{Asterolasia} was predicted to reach a mature height in approximately 22 years regardless of warming treatment.
+\textit{Asterolasia} was predicted to reach a mature height in approximately \Sexpr{round(hmax(non_tussock_otc_growth_model,'Asterolasia', FALSE)[2])} to \Sexpr{round(hmax(non_tussock_otc_growth_model,'Asterolasia', FALSE)[1])} years regardless of warming treatment.
 
@@
 suspected 9 year old seedlings).
-Here, a 200 m altitudinal range is equivalent to an ambient temperature difference of approximately 1.5\textdegree C\cite{Brown:1989vx}, which is comparable to that observed between experimental warmed and control plots (1\textdegree C).
+Here, a 200 m altitudinal range is equivalent to an ambient temperature difference of approximately \Sexpr{0.77 * (200/100)}\textdegree C\cite{Brown:1989vx}, which is comparable to that observed between experimental warmed and control plots (\Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C).
 In response to shifts in temperature, maximum heights of Grevillea post-fire recruits were expected to decrease with elevation, whereas \textit{Asterolasia} seedlings were not expected to show this pattern.
@@
 S4).
-Mean maximum heights of \textit{Grevillea} seedlings in burnt open heathland were 7.9 cm taller at 1660 m a.s.l compared to seedlings at 1860 m a.s.l.
-(19.4 cm vs 11.5 cm; a difference comparable to our experimental findings).
+Mean maximum heights of \textit{Grevillea} seedlings in burnt open heathland were \Sexpr{max_ht_alt(greaus_seedling_max_ht_model)} cm taller at 1670 m a.s.l compared to seedlings at 1860 m a.s.l.
+(\Sexpr{max_ht_alt(greaus_seedling_max_ht_model, FALSE)[1]} cm vs \Sexpr{max_ht_alt(greaus_seedling_max_ht_model, FALSE)[2]} cm; a difference comparable to our experimental findings).
 In contrast, mean maximum heights of \textit{Asterolasia} seedlings did not vary significantly with elevation.
@@
 
-After five years and across all plots, 34\% (172 out of 512) of all seedlings growing in burnt bare ground patches had died.
-Most deaths occurred in \textit{Asterolasia} (65) followed by \textit{Phebalium} (63), \textit{Grevillea} (31) and \textit{Prostanthera} (13).
+After five years and across all plots, \Sexpr{total_deaths(non_tussock_otc_mortality_data,TRUE)}\% (\Sexpr{total_deaths(non_tussock_otc_mortality_data,FALSE)} out of 511) of all seedlings growing in burnt bare ground patches had died.
+Most deaths occurred in \textit{Asterolasia} (\Sexpr{death_counts(non_tussock_otc_mortality_data,'Asterolasia')}) followed by \textit{Phebalium} (\Sexpr{death_counts(non_tussock_otc_mortality_data, 'Phebalium')}), \textit{Grevillea} (\Sexpr{death_counts(non_tussock_otc_mortality_data,'Grevillea')}) and \textit{Prostanthera} (\Sexpr{death_counts(non_tussock_otc_mortality_data,'Prostanthera')}).
 \textit{Prostanthera} showed the largest treatment effect (Fig.
-\ref{fig4}), with annual mortality rates estimated to be near 0 in warmed plots and 5\% in control plots.
-This significant increase in survival may be a consequence of OTCs reducing the severity of spring frosts by rising minimum ambient and soil temperatures by 0.86\textdegree C and 1.6\textdegree C, respectively (see Supplementary Fig S5-7).
+\ref{fig4}), with annual mortality rates estimated to be near \Sexpr{annual_mort(non_tussock_otc_mortality_model, 'Prostanthera','otc')}\% in warmed plots and \Sexpr{annual_mort(non_tussock_otc_mortality_model, 'Prostanthera','ctl')}\% in control plots.
+This significant increase in survival may be a consequence of OTCs reducing the severity of spring frosts by rising minimum ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary,'CAmbient_Temp','mean_min','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary,'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C, respectively (see Supplementary Fig S5-7).
 Warming also reduced mean seedling mortality in \textit{Grevillea} and \textit{Phebalium} (Fig.
@@
 Using five years of post-fire inter-tussock gap size changes in warmed and unwarmed plots we found that gaps were being infilled by grasses.
-However, the rate at which this occurred was not significantly influenced by a 1\textdegree C increase in temperature (Fig.
+However, the rate at which this occurred was not significantly influenced by a \Sexpr{micro_clim(overall_microclimate_summary,'CAmbient_Temp','mean','grand_mean')}\textdegree C increase in temperature (Fig.
 \ref{fig6}).
@@
 \begin{figure}
-\label{fig1}
 \begin{minipage}{\textwidth}
@@
 It also includes an example of a potential, but unlikely, effect (red dotted line) that could break this feedback.}
+\label{fig1}
 \end{minipage}
@@
 \begin{figure}
-\label{fig2}
 \begin{minipage}{\textwidth}
@@
 All errors bars are 95\% Bayesian Credible Intervals.}
+\label{fig2}
 \end{minipage}
@@
 \begin{figure}
-\label{fig3}
 \begin{minipage}{\textwidth}
@@
 34, 65, 65 \& 70 cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Prostanthera} \& \textit{Phebalium}, respectively).}
+\label{fig3}
 \end{minipage}
@@
 \begin{figure}
-\label{fig4}
 \begin{minipage}{\textwidth}
@@
 \caption{\textbf{Mean ($\pm$ 95\% Bayesian Credible Interval) seedling probability of mortality for each species growing in control (blue) and warmed (red) conditions.}}
+\label{fig4}
 \end{minipage}
@@
 \begin{figure}
-\label{fig5}
 \begin{minipage}{\textwidth}
@@
 All errors bars are 95\% Bayesian Credible Intervals.}
+\label{fig5}
 \end{minipage}
@@
 \begin{figure}
-\label{fig6}
 \begin{minipage}{\textwidth}
@@
 \caption{\textbf{Effects of warming on rates of tussock grass gap infilling.} Left panel = Model coefficients, Right panel = Predicted temporal change in size for an average 10 cm inter-tussock gap in warmed (red) and unwarmed (blue) conditions.}
+\label{fig6}
 \end{minipage}

### 2016-03-08 R/data_processing.R
Updated hmax_prior function

@@ -127,21 +127,20 @@ get_mortality_data <- function(complete_otc_seedling_data, inter_tussock=TRUE) {
 }
 
 # Calculates average maximum height from multiple field studies
-hmax_priors <- function(csv_file) {
- data <- read.csv(csv_file)
- 
+hmax_priors <- function(data) {
  data <- data %>%
    filter(!is.na(species)) %>%
    droplevels() %>%
    group_by(site,species) %>%
-   summarise(max_ht = max(height_cm*10, na.rm=TRUE)) %>%
+   summarise(max_ht = max(height_cm, na.rm=TRUE)) %>%
    ungroup() %>%
    group_by(species) %>%
    summarise(n = n(),
-             mn_max_ht = mean(max_ht),
+             mn_max_ht = round(mean(max_ht)),
              sd_max_ht = sd(max_ht),
-             se_max_ht = sd_max_ht/sqrt(n))
-
+             se_max_ht = sd_max_ht/sqrt(n)) %>%
+   mutate(species = factor(species, labels= c('Asterolasia','Grevillea','Phebalium','Prostanthera')))
+ return(data)
 }
 
 # Cleans unburnt field data       

### 2016-03-08 manuscript/manuscript.Rnw
More edits

@@
 \begin{abstract}
-Climate change is expected to increase fire activity and exacerbate woody plant encroachment into arctic and alpine grasslands.
+Climate change is expected to increase fire activity and exacerbate woody plant encroachment in arctic and alpine grasslands.
 What is unknown is whether increases in woody plants and fire will interact to exacerbate changes in structure, function and composition of these ecosystems.
 Here we use field surveys and experimental manipulations to examine warming and fire effects on recruitment, growth and survival of seedlings of evergreen obligate seeding alpine shrubs.
-We find that fire substantially increased shrub seedling establishment (up to \Sexpr{density_ratios(greaus_seedling_density_model, 'Grevillea')}) and that warming doubled tall shrub seedling growth rates and may increase survival.
-Warming had weak or no effect on shrub recruitment, rates of grass post-fire gap infilling or grass competitive effects on shrub seedling growth and survival.
-These findings indicate that rising temperatures coupled with more frequent or severe fires will exacerbate increases in tall evergreen shrubs by increasing recruitment, doubling growth rates, and potentially increase survival.
+We find that fire substantially increased shrub seedling establishment (up to \Sexpr{density_ratios(greaus_seedling_density_model, 'Grevillea')}) and that warming doubled tall shrub seedling growth rates and can increase survival.
+Warming also had weak or no effect on shrub recruitment, rates of grass post-fire gap infilling or grass competitive effects on shrub seedling growth and survival.
+These findings indicate that rising temperatures coupled with more frequent or severe fires will likely exacerbate increases in tall evergreen shrubs by increasing recruitment, doubling growth rates, and potentially increasing seedling survival.
 \end{abstract}
@@
 Consequently, little is known about how climate change influences seedling demographic rates\cite{Briceno:2015iv} and thus, vegetation recovery in post-disturbance conditions.
-We consider these issues in an Australian alpine heathland.
+We consider these issues in Australian alpine heathland.
 
@@
 \ref{fig1}a).
-For example, we have little information on whether fire will stimulate shrub seedling recruitment in alpine and tundra ecosystems, nor do we know whether rates of seedling growth and survival will be improved under warmer post-fire conditions.
+For example, we have little information on whether fire will stimulate shrub seedling recruitment in alpine or tundra ecosystems, nor do we know how rates of seedling growth and survival will be affected under warmer post-fire conditions.
 Furthermore, we do not know how tussock grasslands (the other dominant community in these ecosystems) will respond to warmer post-fire conditions, and whether grasses have competitive, facilitative or no effect on shrub seedling demographic rates, or whether such effects are altered by warmer post-fire conditions.
@@
 Under global warming, this plant community is likely to encroach grasslands and is itself susceptible to shrub thickening\cite{McDougall:2003gb}.
-To determine how warmer and drier post-fire conditions affects shrub cover, we combined field observations with a warming experiment that used seedlings of four dominant Australian evergreen obligate seeding alpine shrubs \textemdash \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) and \textit{Phebalium squamulosum} (Rutaceae; a tall shrub).
-This approach allowed us to examine: 1) how warmer post-fire conditions influence rates of both grass recovery and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affect shrub seedling demographic rates, and 3) whether such affects change under warmer conditions post-fire conditions.
+To determine how warmer post-fire conditions affects shrub cover, we combined field observations with a warming experiment that used seedlings of four dominant Australian evergreen obligate seeding alpine shrubs \textemdash \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub),  \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub).
+This combination of data sources allowed us to examine: 1) landscape scale drivers of shrub seedling establishment; 2) how warmer post-fire conditions influence rates of both grass recovery and shrub seedling recruitment, growth and mortality; 3) how proximity to grass affect shrub seedling demographic rates, and 4) whether such grass effects change under warmer conditions post-fire conditions.
 
 \subsection{Drivers of shrub seedling establishment.}
-We first investigated how altitude, topographic wetness (TWI), adult density, fire and, if burnt, fire severity, influenced the density of \textit{Grevillea} and \textit{Asterolasia} seedlings (two dominant shrubs of open heathland).
+We first investigated how altitude, topographic wetness (TWI), adult density, fire and, if burnt, fire severity, influenced the density of \textit{Grevillea} and \textit{Asterolasia} seedlings (the two dominant shrubs of open heathland).
 In 2012-13, we estimated seedling density for both species across 40 alpine sites and found that the abundance of \textit{Grevillea} (Fig.
@@
 As expected, pre-fire adult density positively influenced \textit{Grevillea} seedling density, with sites containing higher abundances of pre-fire \textit{Grevillea} adults containing more seedlings.
-For both shrub species, we detected no significant change in seedling density along a 200 m altitudinal range (equivalent to a 1.5\textdegree C change in mean temperature).
+For both shrub species, we detected no significant change in seedling density along a 200 m altitudinal range (equivalent to a \Sexpr{round(0.77 * (200/100),1)}\textdegree C change in mean temperature\cite{Brown:1989vx}).
 The field warming experiment (see below) indicates that this lack of elevation response in \textit{Grevillea} is due to both rates of recruitment (Supplementary Fig.
@@
 
-Knowing shrub seedlings were more abundant in burnt vegetation, we investigated whether warmer post-fire conditions affected seedling growth and mortality rates.
-Seedlings of \textit{Grevillea}, \textit{Asterolasia}, \textit{Prostanthera} and \textit{Phebalium} emerging after a wildfire were transplanted into experimentally burnt plots that were either passively warmed by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C using open top chambers (OTCs) or subject to current ambient conditions (controls).
-After \Sexpr{n_days(chamber_dates, FALSE)} days \Sexpr{n_days(chamber_dates, TRUE)} growing season days where OTC treatment was applied) or \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} years growth, mean seedling heights of the tall shrubs (\textit{Grevillea}, \textit{Prostanthera} and \textit{Phebalium}) growing in post-fire bare ground were taller in warmed plots by \Sexpr{ht_diff(non_tussock_otc_growth_data,'Grevillea')}, \Sexpr{ht_diff(non_tussock_otc_growth_data,'Prostanthera')} and \Sexpr{ht_diff(non_tussock_otc_growth_data,'Phebalium')} cm respectively (Fig.
+As shrub seedlings were more abundant in burnt vegetation, we investigated how warmer post-fire conditions affected seedling growth and mortality rates.
+Seedlings of \textit{Grevillea}, \textit{Asterolasia}, \textit{Phebalium} and \textit{Prostanthera} emerging after a wildfire were transplanted into experimentally burnt plots that were either passively warmed by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C using open top chambers (OTCs) or subject to current ambient conditions (controls).
+After \Sexpr{n_days(chamber_dates, FALSE)} days (\Sexpr{n_days(chamber_dates, TRUE)} growing season days where OTC treatment was applied) or \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} years growth, mean seedling heights of the tall shrubs (\textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera}) growing in post-fire bare ground were taller in warmed plots by \Sexpr{ht_diff(non_tussock_otc_growth_data,'Grevillea')}, \Sexpr{ht_diff(non_tussock_otc_growth_data,'Phebalium')} and \Sexpr{ht_diff(non_tussock_otc_growth_data,'Prostanthera')} cm, respectively (Fig.
 \ref{fig3}).
@@
 S3).
-Accounting for initial height and assuming logistic growth, mean annual predicted height growth rates of \textit{Grevillea}, \textit{Prostanthera} and \textit{Phebalium} were \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Grevillea')}, \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Prostanthera')} and \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Phebalium')} times that observed in control plots, respectively.
-Projecting these differences equates to an average 6 cm tall seedling (the mean initial size used in this experiment) growing in warmed conditions reaching maximum mature height in \Sexpr{hmax(non_tussock_otc_growth_model,'Grevillea')} fewer years for \textit{Grevillea} and \Sexpr{hmax(non_tussock_otc_growth_model,'Prostanthera')} fewer years for both \textit{Prostanthera} and \textit{Phebalium} (Fig.
+Accounting for initial height and assuming logistic growth, mean annual predicted height growth rates of \textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera} were \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Grevillea')}, \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Phebalium')} and \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Prostanthera')} times that observed in control plots, respectively.
+Projecting these differences equates to an average 6 cm tall seedling (the mean initial size used in this experiment) growing in warmed conditions reaching maximum mature height in \Sexpr{hmax(non_tussock_otc_growth_model,'Grevillea')} fewer years for \textit{Grevillea} and \Sexpr{hmax(non_tussock_otc_growth_model,'Prostanthera')} fewer years for both \textit{Phebalium} and \textit{Prostanthera} (Fig.
 \ref{fig3}).
-\textit{Asterolasia} was predicted to reach a mature height in approximately \Sexpr{round(hmax(non_tussock_otc_growth_model,'Asterolasia', FALSE)[2])} to \Sexpr{round(hmax(non_tussock_otc_growth_model,'Asterolasia', FALSE)[1])} years regardless of warming treatment.
+\textit{Asterolasia} was predicted to reach a mature height in approximately \Sexpr{round(hmax(non_tussock_otc_growth_model,'Asterolasia', FALSE)[2])} to \Sexpr{round(hmax(non_tussock_otc_growth_model,'Asterolasia', FALSE)[1])} years.
 
-We validated the \textit{Grevillea} and \textit{Asterolasia} growth responses observed in experimentally warmed plots by using an elevation gradient in open heathland and maximum height data for 2003 post-fire recruits (i.e.
-suspected 9 year old seedlings).
-Here, a 200 m altitudinal range is equivalent to an ambient temperature difference of approximately \Sexpr{0.77 * (200/100)}\textdegree C\cite{Brown:1989vx}, which is comparable to that observed between experimental warmed and control plots (\Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C).
-In response to shifts in temperature, maximum heights of Grevillea post-fire recruits were expected to decrease with elevation, whereas \textit{Asterolasia} seedlings were not expected to show this pattern.
+We validated the \textit{Grevillea} and \textit{Asterolasia} growth responses observed in experimentally warmed plots by examining the effect of elevation on maximum height data collected from 2003 post-fire recruits across 40 open heathland sites.
+Here, a 200 m altitudinal range is equivalent to a mean ambient temperature difference of approximately \Sexpr{round(0.77 * (200/100),1)}\textdegree C\cite{Brown:1989vx}, which is comparable to that observed between experimental warmed and control plots (\Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C).
+In response to shifts in temperature, maximum heights of \textit{Grevillea} post-fire recruits were expected to decrease with elevation, whereas \textit{Asterolasia} seedlings were not expected to show this pattern.
 Our experimental predictions were verified (Supplementary Fig.
@@
 \ref{fig4}), with annual mortality rates estimated to be near \Sexpr{annual_mort(non_tussock_otc_mortality_model, 'Prostanthera','otc')}\% in warmed plots and \Sexpr{annual_mort(non_tussock_otc_mortality_model, 'Prostanthera','ctl')}\% in control plots.
-This significant increase in survival may be a consequence of OTCs reducing the severity of spring frosts by rising minimum ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary,'CAmbient_Temp','mean_min','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary,'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C, respectively (see Supplementary Fig S5-7).
+This significant decrease in mortality may be a consequence of OTCs reducing the severity of spring frosts by rising minimum ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary,'CAmbient_Temp','mean_min','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary,'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C, respectively (see Supplementary Fig S5-7).
 Warming also reduced mean seedling mortality in \textit{Grevillea} and \textit{Phebalium} (Fig.
@@
 \ref{fig5}).
-As with the seedlings growing in bare ground, we detected a strong positive effect of warming on growth rates and a marginal, yet not significant, decrease in mortality.
+As with the seedlings growing with no nearby \textit{poa tussock}, we detected a strong positive effect of warming on growth rates and a marginal, yet not significant, decrease in mortality.
 We also found no significant effect of inter-tussock gap radius on growth or mortality rates.
 However, for growth, we did detect a marginally significant interaction between warming and gap size, such that in warmed plots growth rates slightly increased with gap size.
-This suggests that warmer conditions may increase competition between shrubs and grasses, a finding consistent with the abiotic stress hypothesis\cite{Callaway:2002vb}, which suggests that temperature stress in alpine and tundra ecosystems could increase the strength of competitive interactions.
-However, this effect was small, and likely to have little biological consequence.
+This suggests that warmer conditions may increase competition between shrubs and grasses, a finding consistent with the abiotic stress hypothesis\cite{Callaway:2002vb}.
+However, this effect was small and unlikely to have significant biological consequence on shrub growth rates.
 
@@
 
-Despite having little impact on shrub seedling growth and mortality rates, tussock grass may still reduce shrub recruitment opportunities by infilling post-fire bare ground gaps at a faster rate.
-Using five years of post-fire inter-tussock gap size changes in warmed and unwarmed plots we found that gaps were being infilled by grasses.
-However, the rate at which this occurred was not significantly influenced by a \Sexpr{micro_clim(overall_microclimate_summary,'CAmbient_Temp','mean','grand_mean')}\textdegree C increase in temperature (Fig.
+Despite having little impact on shrub seedling growth and mortality rates, tussock grass may still limit shrub recruitment, and thus, shrub expansion, by infilling post-fire bare ground gaps at a faster rate under warmer conditions.
+Using five years of post-fire inter-tussock gap size changes in warmed and unwarmed plots we found that gaps were being infilled by grasses (Fig.
 \ref{fig6}).
+However, the rate at which this occurred was very slow, with a 10 cm radius gap decreasing by approximately \Sexpr{gap_size(gap_dynamic_model, 'otc')} cm over a ten year period.
+We also detected no significant effect of a \Sexpr{micro_clim(overall_microclimate_summary,'CAmbient_Temp','mean','grand_mean')}\textdegree C temperature rise on the rate of grass infilling.
 
@@
 
-Here we quantified several unknown interactions between fire, shrub-grass relationships and climate to extend a conceptual model of alpine shrub dynamics under climate change (Fig \ref{fig1}b).
+Here we quantified several unknown interactions between fire, shrub-grass relationships and climate to extend a conceptual model of alpine shrub dynamics under climate change (Fig.
+\ref{fig1}b).
 We showed how these interactions strengthen a hypothesized feedback loop that can rapidly increase shrub cover in alpine and tundra ecosystems\cite{Higuera:2008dw}.
-The combination of rising temperatures and more frequent or severe fire creates conditions that allow shrub seedlings to establish in greater densities, and for tall shrubs, double their growth rates, and potentially increase their survival.
+The combination of rising temperatures and more frequent or severe fire creates conditions that allow shrub seedlings to establish in greater densities, and for tall shrubs, double their growth rates and potentially increase their survival.
 These demographic effects will ultimately result in shrub thickening and expansion into grasslands.
@@
 The most obvious mechanism that will break this cycle involves shorter fire intervals that prevent obligate seeding shrubs reaching reproductive age and thereby exhausting the seedbank.
-However, this scenario is unlikely for the majority of alpine (or tundra) landscapes including in Australia, where it would require current fire intervals of 50 to 100 years to decrease to less than 20 years \textemdash the time observed for the species in this study to reach reproductive maturity\cite{Williams:2008ba}.
+However, this scenario is unlikely for the majority of alpine (or tundra) landscapes including in Australia, where it would require current fire intervals of 50 to 100 years to decrease to less than 20 years \textemdash the time estimated for the species in this study to reach reproductive maturity\cite{Williams:2008ba}.
 Furthermore, if reproductive output is related to plant size\cite{Wenk:2015ko}, then climatic warming may allow obligate seeding species to reach reproductive maturity sooner, and consequently, increase species resilience to short fire intervals.
@@
 We investigated shrub dynamics in open heathland because it is a common and highly flammable plant community in the Australian Alps above 1600 m a.s.l\cite{Williams:2006tl}.
-It is also an ecotone between closed heathland (>70\% shrub cover) and tussock grassland consisting of \textit{Grevillea australis} shrubs interspersed among a \textit{Poa hiemata} sward.
-Aerial photography has revealed that plant community has acted as an \textit{invasion front} of shrub expansion into grassland.
-It is also experiencing shrub thickening by closed heathland dominants such as \textit{Prostanthera cuneata}.
+It is also an ecotone between closed heathland (\textgreater 70\% shrub cover) and tussock grassland consisting of \textit{Grevillea australis} shrubs interspersed among a \textit{Poa hiemata} sward.
+Aerial photography has revealed that the plant community has acted as an \textit{invasion front} of shrub expansion into grassland and has itself experienced shrub thickening by closed heathland dominants such as \textit{Prostanthera cuneata}.
 
@@
 In March 2010, at 1750 m a.s.l, we burnt 32 randomly selected mature (60 cm tall and 1.5m$^2$) \textit{Grevillea australis} shrubs in open heathland to create patches of bare ground approximately 0.7 m$^2$ surrounded by burnt tussock grass, simulating disturbance in open heathland burnt by wildfire.
-After creating the bare ground patches we collected seedlings of dominant alpine shrub species from a nearby (< 2 km) site of similar altitude burnt by a late 2006 wildfire.
+After creating the bare ground patches we collected seedlings of dominant alpine shrub species from a nearby (\textless 2 km) site of similar altitude burnt by a late 2006 wildfire.
 We collected shrub seedlings of two dominant open heathland species (\textit{Grevillea australis} (Proteaceae; a tall shrub) and \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), a dominant closed heathland species (\textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) that typically grows on warmer aspects and a species common to both open and closed heathland (\textit{Phebalium squamulosum} (Rutaceae; a tall shrub).
@@
 To examine interactions between tussock grass and shrub seedlings we also randomly transplanted four additional \textit{Grevillea australis} into various sized inter-tussock gaps between burnt \textit{Poa hiemata} immediately surrounding the bare ground (Supplementary Fig.
-S6).
+S8).
 The experimental site was fenced to prevent grazing by deer and horses.
@@
 OTCs were placed out at the start of the growing season (October) where they remained until snowfall (early June).
-This procedure was repeated for five growing seasons from 2010 to 2015.
+This procedure was repeated for five growing seasons from May 2010 to May 2015.
 
 Microclimatic changes were measured hourly using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) at four control and four OTC plots.
-Across five growing seasons (1087 days), OTCs simulated warmer conditions at the lower end of IPCC predictions\cite{IPCC:2013vv}.
-OTCs passively increased mean daily ambient and soil temperatures by 1\textdegree C and maximums and minimums by up to 2.4\textdegree C (Supplementary Methods and Supplementary Fig.
+Across \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} growing seasons (\Sexpr{n_days(chamber_dates, TRUE)} chamber days), OTCs simulated warmer conditions at the lower end of IPCC predictions\cite{IPCC:2013vv} (Supplementary Methods and Supplementary Fig.
 S5-7).
-Chambers only marginally decreased soil moisture by 0.21\%.
+OTCs passively increased average ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C, and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively.
+Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C).
+Chambers only marginally decreased soil moisture by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CMoisture3to10cmBG','mean','grand_mean'))}\% and relative humidity by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CRH','mean','grand_mean'))}\%.
 
@@
 Within plots we recorded the number and maximum height of \textit{Grevillea} and \textit{Asterolasia} seedlings.
-For unburnt sites we counted the number of mature \textit{Grevillea} plants (> 0.5 m$^2$) within 5 m either side of each transect.
+For unburnt sites we counted the number of mature \textit{Grevillea} plants (\textgreater 0.5 m$^2$) within 5 m either side of each transect.
 In burnt sites, this required counting the number of skeletons (there were no living adults at any burnt site) that still persist post-fire.
@@
 In total we built 11 hierarchical models to examine factors influencing tussock grass post-fire recovery and shrub seedling establishment, growth and mortality.
-For each model (Supplementary Methods) we used Bayesian inference and fitted models with Markov chain Monte Carlo (MCMC) sampling in R 3.2.2 using package Rstan 2.8.1.
+For each model (Supplementary Methods) we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package Rstan \Sexpr{packageVersion("rstan")}.
 Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.
@@
 \includegraphics{figures/fig1.pdf}
-\caption{\textbf{Conceptual diagram illustrating positive warming-shrub-fire feedback loop.} Top panel highlights the known positive feedback loop between fire, climatic warming and landscape flammability (solid blue lines).
-It also highlights several interactions (dashed \& black) which currently have a paucity of information but which may exacerbate or diminish this feedback loop.
-Bottom panel either assigns the direction or removes these interactions based on new evidence obtained from experimental and field surveys conducted in this study.
-It also includes an example of a potential, but unlikely, effect (red dotted line) that could break this feedback.}
+\caption{\textbf{Conceptual diagram illustrating positive warming-shrub-fire feedback loop.} Hypothesised positive feedback loop between fire, climatic warming and landscape flammability (\textbf{a}).
+Solid blue lines = known mechansims.
+dahsed black lines = mechansims that may exacerbate or diminish this feedback loop but which we have a paucity of information on.
+Feedback loop based on new evidence obtained from experimental and field surveys conducted in this study (\textbf{b}).
+Also includes an example of a potential, but unlikely, effect (red dotted line) that could break this feedback.}
 \label{fig1}
@@
 \includegraphics{figures/fig2.pdf}
-\caption{\textbf{Mean \textit{Grevillea australis} seedling density with burning, fire severity, adult density, altitude and Topographic Wetness Index (TWI) gradients.} Top left panel = centered and standardized model coefficients (on the log scale).
-Top right panel is the effect of burning.
-All other panels are partial dependence plots for each covariate along their observed gradients in areas burnt by the 2003 fires.
+\caption{\textbf{Mean \textit{Grevillea australis} seedling density with burning, fire severity (minimum twig diameter), adult density, altitude and Topographic Wetness Index (TWI) gradients.} Centered and standardized model coefficients (on the log scale) (\textbf{a}) and effects of burning (\textbf{b}), fire severity  (\textbf{c}), altitude (\textbf{d}), topographic wetness (\textbf{e}) \& adult density (\textbf{f}), in areas burnt by the 2003 fires.
 All errors bars are 95\% Bayesian Credible Intervals.}
@@
 \includegraphics{figures/fig3.pdf}
-\caption{\textbf{Observed and projected growth trajectories of four dominant Australian alpine shrubs.} Top panels: Thick lines represent mean ($\pm$ 95\% confidence intervals) observed heights at each May census.
+\caption{\textbf{Observed and projected growth trajectories of four dominant Australian alpine shrubs.} Observed height growth (\textbf{a}): Thick lines represent mean ($\pm$ 95\% confidence intervals) observed heights at each May census.
 Thin lines represent individual growth trajectories.
-Bottom panels are the mean ($\pm$ 95\% Bayesian Credible Interval) projected height growth for each species over 70 years.
-In all cases, red or orange lines = seedlings growing in warmed (OTC) conditions and blue or light blue lines = seedlings growing in control conditions.
+Mean ($\pm$ 95\% Bayesian Credible Interval) projected height growth for each species over 70 years (\textbf{b}).
+In all cases, red \& orange lines = seedlings growing in warmed (OTC) conditions and blue \& light blue lines = seedlings growing in control conditions.
 Projections were based on an logistic growth model using initial seedling size of 6 cm (the mean initial height observed in the OTC experiment) and mean maximum heights observed in long-unburnt sites (i.e.
-34, 65, 65 \& 70 cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Prostanthera} \& \textit{Phebalium}, respectively).}
+\Sexpr{hmax_values(hmax,'Asterolasia')}, \Sexpr{hmax_values(hmax,'Grevillea')}, \Sexpr{hmax_values(hmax,'Phebalium')} \& \Sexpr{hmax_values(hmax,'Prostanthera')} cm for \textit{Asterolasia}, \textit{Grevillea},\textit{Phebalium}, \& \textit{Prostanthera}, respectively).}
 \label{fig3}
@@
 \includegraphics{figures/fig5.pdf}
-\caption{\textbf{Experimental warming and inter-tussock gap size effects on \textit{Grevillea australis} seedlings.}First row = Growth rate effects, Second row = Mortality effects.
-Left panels are the centered and standardized model coefficients; right panels are the predicted growth and mortality rates along an inter-tussock gap size gradient warmed (red) and control (blue) conditions.
+\caption{\textbf{Experimental warming and inter-tussock gap size effects on \textit{Grevillea australis} seedlings.} Top rows = Growth rate effects, Second row = Mortality effects.
+\textbf{a} \& \textbf{c} are the centered and standardized model coefficients; \textbf{b} \& \textbf{d} are growth and mortality rate response curves along an inter-tussock gap size gradient under warmed (red) and control (blue) conditions.
 All errors bars are 95\% Bayesian Credible Intervals.}
@@
 \includegraphics{figures/fig6.pdf}
-\caption{\textbf{Effects of warming on rates of tussock grass gap infilling.} Left panel = Model coefficients, Right panel = Predicted temporal change in size for an average 10 cm inter-tussock gap in warmed (red) and unwarmed (blue) conditions.}
+\caption{\textbf{Effects of warming on rates of tussock grass gap infilling.}
+Model coefficients (\textbf{a}) and projected temporal change in size for an average 10 cm inter-tussock gap in warmed (red) and unwarmed (blue) conditions (\textbf{b}).} 
 \label{fig6}

### 2016-03-08 manuscript/manuscript.Rnw
More updates

@@
 Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.
+Additional information of experimental design and analysis can be found in Supplementary methods.
 \end{methods}
@@
 \includegraphics{figures/fig2.pdf}
-\caption{\textbf{Mean \textit{Grevillea australis} seedling density with burning, fire severity (minimum twig diameter), adult density, altitude and Topographic Wetness Index (TWI) gradients.} Centered and standardized model coefficients (on the log scale) (\textbf{a}) and effects of burning (\textbf{b}), fire severity  (\textbf{c}), altitude (\textbf{d}), topographic wetness (\textbf{e}) \& adult density (\textbf{f}), in areas burnt by the 2003 fires.
-All errors bars are 95\% Bayesian Credible Intervals.}
+\caption{\textbf{Mean \textit{Grevillea australis} seedling density along burning, fire severity (minimum twig diameter), adult density, altitude and Topographic Wetness Index (TWI) gradients.} Centered and standardized model coefficients (on the log scale) (\textbf{a}) and effects of burning (\textbf{b}), fire severity  (\textbf{c}), altitude (\textbf{d}), topographic wetness (\textbf{e}) \& adult density (\textbf{f}), in areas burnt by the 2003 fires.
+All errors bars are 95\% Bayesian Credible Intervals.
+Phi = overdispersion parameter for negative binomial distribution.}
 \label{fig2}

### 2016-03-08 manuscript/supplementary.Rnw
Put \label under the caption
More edits to supp mat

@@
 
-\section*{Study site.}
+\section*{Study site}
 Research sites were on the Bogong High Plains, a series of alpine and high subalpine plateaux in the Alpine National Park, approximately 250 km northeast of Melbourne, Australia.
@@
 
-\section*{Open Top Chamber Experiment.}
-Open Top Chamber Experiment
+\section*{Open Top Chamber Experiment}
 We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of grass recovery and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affect shrub seedling demographic rates, and 3) whether such affects change under warmer conditions post-fire conditions.
@@
 In alpine and tundra regions, shrub seedlings typically establish and grow in bare ground created by disturbance\cite{Williams:1987ua, Batllori:2009fe, Frost:2013tm}.
-Here, we used fire to create bare ground because 1) Australian alpine shrub seedlings occur in greatest abundance after fire (see seedling survey results) and 2) in many parts of the world, including Australia, wildfires are predicted to increase in frequency and severity\cite{Flannigan:2009kl, Westerling:2011jp}.
+Here, we used fire to create bare ground because 1) Australian alpine shrub seedlings occur in greatest abundance after fire (see seedling survey results) and 2) in many parts of the world, including Australia, wildfires are predicted to increase in frequency and severity\cite{Westerling:2006ve,Flannigan:2009kl}.
 Thus, to forecast likely changes in shrub abundance we need to know how warming affects growth and mortality rates of shrub seedlings growing in burnt sites.
@@
 We collected seedlings from a site burnt by a late 2006 fire because it allowed us to collect sufficient seedling numbers from species occurring in both open and closed heathland.
-We focused on \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) and \textit{Phebalium squamulosum} (Rutaceae; a tall shrub because they are all: obligate seeders post-fire (i.e.
-fire killed), common in the Australian Alps, and under global warming, likely candidates to invade non-shrubby plant communities such as alpine grasslands and herbfields.
-For each species, we collected a range of seedling heights (2 – 12 cm) using a forestry planter (12.5 cm deep by 5.5 cm wide).
+We focused on \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) because they are all: obligate seeders post-fire (i.e.
+fire killed), common in the Australian Alps, and under climatic warming, likely candidates to invade non-shrubby plant communities such as alpine grasslands and herbfields.
+For each species, we collected a range of seedling heights (\Sexpr{range_init_size(complete_otc_seedling_data)[1]} to \Sexpr{range_init_size(complete_otc_seedling_data)[2]} cm) using a forestry planter (12.5 cm deep by 5.5 cm wide).
 The size of the planter was selected to minimize root damage across the range of seedling sizes and to allow for the extraction of a constant volume (378 cm$^3$) of intact soil and roots.
@@
 Four seedlings of each species were then transplanted into 70 cm$^2$ of burnt bare ground in the center of each experimental plot.
-An additional four \textit{Grevillea seedlings} were transplanted into various sized inter-tussock gaps surrounding this bare ground.
+An additional four \textit{Grevillea} seedlings were transplanted into various sized inter-tussock gaps surrounding this bare ground.
 Only \textit{Grevillea} seedlings were transplanted into inter-tussock gaps because of space constraints but also because it is the dominant open heathland shrub species (i.e.
@@
 
-To simulate near-term warmer conditions predicted under IPCC\cite{IPCC:2013tu}, we used Open Top Chambers (OTCs)\cite{Molau:1996vc}.
+To simulate near-term warmer conditions predicted under IPCC\cite{IPCC:2013vv}, we used Open Top Chambers (OTCs)\cite{Molau:1996vc}.
 OTCs were hexagonal, 58 cm tall with an open-top diameter of 110 cm and base diameter of 168 cm.
-The chambers were constructed following the International Tundra Experiment (ITEX) protocols\cite{Molau:1996vc} using sun-lite: a lightweight shatter-resistant fiberglass with high light transmittance (86\%) and low infrared transmittance (< 5\%; Solar Components Corporation, USA).
+The chambers were constructed following the International Tundra Experiment (ITEX) protocols\cite{Molau:1996vc} using sun-lite: a lightweight shatter-resistant fiberglass with high light transmittance (86\%) and low infrared transmittance (\textless 5\%; Solar Components Corporation, USA).
 
@@
 Each Micro Station measured ambient temperature, relative humidity (both at 10 cm above ground), soil temperature (3 cm below ground) and volumetric soil moisture (between 3 and 10 cm) at hourly intervals for the entire duration of the experiment.
-Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions\cite{IPCC:2013tu} (Supplementary Fig.
+Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions\cite{IPCC:2013vv} (Supplementary Fig.
 S5-7).
-OTCs passively raised mean daily ambient and soil temperatures by approximately 1 \textdegree C and 0.9 \textdegree C ($\pm$ 0.05 95\% CI) and marginally lowered both soil moisture (- 0.21\% $\pm$ 0.11) and relative humidity (- 1.2\% $\pm$ 0.17).
-OTCs also increased minimum daily ambient and soil temperatures by 0.86 \textdegree C ($\pm$ 0.06 95\% CI) and 1.6 ($\pm$ 0.12 95\% CI) respectively.
-Maximum temperatures were also strongly influenced by OTCs with ambient and soil maximum temperatures increasing by 2.4 \textdegree C ($\pm$  0.16 95\% CI) and 0.95 \textdegree C ($\pm$  0.42 95\% CI), respectively.
+OTCs passively increased average ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C, and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively.
+Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C).
+Chambers only marginally decreased soil moisture by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CMoisture3to10cmBG','mean','grand_mean'))}\% and relative humidity by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CRH','mean','grand_mean'))}\%.
 Because OTCs strongly increase temperatures relative to other measures, throughout we refer to it as a ‘warming treatment’.
@@
 
-\section*{Seedling survey.}
+\section*{Seedling survey}
 We measured seedling density and maximum heights in open heathland across the Bogong High Plains.
@@
 
-The study used 40 open heathland sites\cite{Camac:2013eb} established immediately post-2003 fires \cite{Williams:2006tl}.
+The study used 40 open heathland sites\cite{Camac:2013eb} established immediately after the 2003 fires\cite{Williams:2006tl}.
 All sites have not been grazed by cattle since the 2003 fires.
@@
 Sites were at least 0.25 ha and more than 500 m apart.
-In the summer of 2012-13, 9-years post-fire, seedling density in each site was estimated using 40 1 m$^2$ quadrats, evenly distributed in groups of 10 along four 50 m transects.
+In the summer of 2011-12, 9-years post-fire, seedling density in each site was estimated using forty 1 m$^2$ quadrats, evenly distributed in groups of 10 along four 50 m transects.
 Transects were positioned in the center of each site and were parallel to each other with exactly 10 m between transect lines.
-This effectively created a grid of 40 plots (10 plots \texttimes 4 transects) per site, subsampling an area of 2000 m$^2$.
+This effectively created a grid of 40 plots (10 plots\texttimes4 transects) per site, subsampling an area of 2000 m$^2$.
 
@@
 TWI, which approximates the relative plant-available water based on a site’s position in the landscape\cite{Moore:1993fk}, was derived for each site using $TWI= ln(\frac{a\alpha}{tan\theta})$, where, $\alpha$ is the upslope catchment area per unit width orthogonal to the flow direction and $\theta$ is the slope (in degrees).
-Across the 40 sites, TWI varied twofold (5 – 10 units).
-Altitude ranged by 200 m (from 1660 to 1860 m a.s.l), or by approximately 1.5\textdegree C from the lowest to the highest site (based on a lapse-rate of 0.77\textdegree C/100 m elevation\cite{Brown:1989vx}.
-This altitudinal temperature range is comparable to that observed in the OTC experiment (1.5\textdegree C vs 1.2\textdegree C, respectively).
+Across the 40 sites, TWI varied twofold (5 to 10 units; higher units = wetter).
+Altitude ranged by \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m (i.e.
+from \Sexpr{round(min(complete_field_survey_data$altitude_m))} to \Sexpr{round(max(complete_field_survey_data$altitude_m))} m a.s.l), or by approximately \Sexpr{round(0.77 * (190/100),1)}\textdegree C from the lowest to the highest site (based on a lapse-rate of 0.77\textdegree C/100 m elevation\cite{Brown:1989vx}.
+This altitudinal temperature range is comparable to that observed in the OTC experiment.
 
@@
 plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset\cite{Gelman:2007te, Kery:2012wl}.
-For each model we used Bayesian inference and fitted models with Markov chain Monte Carlo (MCMC) sampling in R 3.2.2 using package Rstan 2.8.1.
+For each model we used Bayesian inference and fitted models with sampling in R \Sexpr{paste(version$major, version$minor, sep='.')} using package Rstan \Sexpr{packageVersion("rstan")}.
 For each model, three independent chains were executed.
@@
 To ensure all covariate effects were identifiable we selectively removed each covariate and visually inspected whether the direction and magnitude of effects were severely altered; none were.
-Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine-Shrub-Experiment}.
+Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.
 
@@
 This meant that our observations were of maximum height for each individual $i$ at each May census $t$.
-We modelled these heights, $height_{it}$, as a random realization from a normal distribution truncated at zero:
+We modelled these heights, $height_{i,t}$, as a random realization from a normal distribution truncated at zero:
 
 \begin{equation}\label{eq:growth_obs} 
-height_{it} \sim \mathcal{N^+}(\widehat{height}_{it}, \sigma_{obs}),
+height_{i,t} \sim \mathcal{N}(\widehat{height}_{i,t}, \sigma_{obs}), 0 \leq height_{i,t} \leq \infty,
 \end{equation}
 
-where $\widehat{height}_{i,t}$ is the estimated maximum height for each individual $i$ at each May census $t$ and $\sigma_obs$ is the standard deviation of the observation error.
+where $\widehat{height}_{i,t}$ is the estimated maximum height for each individual $i$ at each May census $t$ and $\sigma_{obs}$ is the standard deviation of the observation error.
 The parameter $\widehat{height}_{i,t}$ was estimated using a logistic parameterisation of the unified Richards equation\cite{Tjorve:2010bk}:
@@
 We treated Hmax as a constant, not to be estimated by the model, owing to our experimental observations being restricted to the first five years of growth.
-Based on maximum height data from 5 to 22 long-unburnt sites (depending on the species) we estimated the average mature height (and thus Hmax) to be approximately 34, 67, 66 and 70 cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera}, respectively.
+Based on maximum height data from 5 to 22 long-unburnt sites (depending on the species) we estimated the average mature height (and thus Hmax) to be approximately \Sexpr{hmax_values(hmax,'Asterolasia')}, \Sexpr{hmax_values(hmax,'Grevillea')}, \Sexpr{hmax_values(hmax,'Phebalium')} and \Sexpr{hmax_values(hmax,'Prostanthera')} cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera}, respectively.
 $Hinit_i$ is the height of individual $i$ at the start of the experiment (i.e.
@@
 $Days_t$ is the number of days from the beginning of the experiment to each May census $t$.
-The experiment ran for a total of 1812 days (or approximately 5 years).
+The experiment ran for a total of \Sexpr{n_days(chamber_dates, FALSE)} days (or approximately \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} years).
 We divide the number of days at each census by 365.25 in order to estimate annual growth rates.
@@
 the maximum annual relative growth rate).
-When multiplied by the asymptotic height, $Hmax$, it yields the maximum annual growth rate at the inflection point .
-We modelled $R_i$ separately for seedlings growing in bare ground (eqn.
-1.3.1) compared to those growing in inter-tussock gaps (eqn.
-1.3.2).
+When multiplied by the asymptotic height, $Hmax$, it yields the maximum annual growth rate at the inflection point.
+We modelled $R_i$ separately for seedlings growing in bare ground (equation \ref{eq:r_bg}) compared to those growing in inter-tussock gaps (equation \ref{eq:r_tussock}).
 
@@
 
-The structure and parameter definitions differ slightly to that in eqn 1.3.1.
+The structure and parameter definitions differ slightly to that in equation \ref{eq:r_bg}.
 Here, the model has two additional parameters, $\beta_2$ and $\beta_3$.
 The parameter $\beta_2$ is the effect of inter-tussock gap radius, which varied over time $t$, while $\beta_3$ is the interactive effect of OTC and time-varying inter-tussock gap radius.
-The inclusion of these parameters, coupled $Gap_{i,t}$ being centered and standardized (see Data inputs into growth and mortality models), alters the definition of the intercept, $\alpha$, which now represents the mean control plot growth rate at the average inter-tussock gap size (12 cm).
+The inclusion of these parameters, coupled $Gap_{i,t}$ being centered and standardized (see Data inputs into growth and mortality models), alters the definition of the intercept, $\alpha$, which now represents the mean control plot growth rate at the average inter-tussock gap size (\Sexpr{mean_init_gap(tussock_otc_growth_data)} cm).
 The parameter  $\beta_1$ is the OTC effect and $\epsilon_{plot_k}$ and $\epsilon_i$ are residual errors at the plot and individual level, respectively.
@@
 
-We modelled individual instantaneous hazard rate, $\lambda_i$, in the same way we modelled $R_i$ in equations 1.3.1 \& 1.3.2. 
+We modelled individual instantaneous hazard rate, $\lambda_i$, in the same way we modelled $R_i$ in equations \ref{eq:r_bg} \& \ref{eq:r_tussock}. 
 For individual seedlings of the four shrub species located in the bare ground patch of each plot, we modelled $\lambda_i$ using a varying intercept model:
@@
 
-Here, the structure and parameter definitions are similar to eqn 2.3.1.
+Here, the structure and parameter definitions are similar to equation \ref{eq:r_tussock}.
 The intercept, $\alpha$, represents the expected instantaneous hazard rate in control plots in an average inter-tussock gap radius, $\beta_1$ is the effect of OTC, $\beta_2$ is the effect of inter-tussock gap radius, $\beta_3$ is the interactive effect of OTC and inter-tussock gap radius and $\epsilon_{plot_k}$ and $\epsilon_i$ are the residual variation at the plot and individual scales.
@@
 where $\mu_{k,t}$ is the log expected recruit count for each plot, and $\phi$ is the overdispersion parameter.
-We used a negative binomial distribution because count data was significantly overdispersed with $\phi$ being estimated as 3.95.
+We used a negative binomial distribution because count data was significantly overdispersed with $\phi$ being estimated as \Sexpr{phi_value(recruit_model,'mean')} (95\% BCI: \Sexpr{phi_value(recruit_model,'l_bci')}\textendash \Sexpr{phi_value(recruit_model,'u_bci')}).
 
@@
 However, because some cardinal distances were recorded as censored data (i.e.
-> 25 cm), these data were incomplete.
+\textgreater 25 cm), these data were incomplete.
 We estimated these censored data using a censored model (see below).
 Once censored data were estimated, we then estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each censuses, for each inter-tussock seedling.
-Median inter-tussock gap radius ranged from 0.5 cm to 43 cm in both control and OTC plots.
-In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero (minus grand mean) and divided by two standard deviations\cite{Gelman:2007te}.
+Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots.
+In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations\cite{Gelman:2007te}.
 The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as average responses and slope terms as partial dependencies conditional on other continuous variables at their mean.
@@
 \subsection{\textit{Modelling censored Poa distances.}}
-Approximately 15\% of the Poa cardinal distance data used to calculate the average inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, ≥ 26 cm.
+Approximately \Sexpr{prop_censored(cleaned_poa_distances)}\% of the Poa cardinal distance data used to calculate the average inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, ≥ 26 cm.
 We estimated these censored values by first estimating the distribution of distances in observed data, coupled with estimating covariate effects and then using these distributions to estimate censored observations.
@@
 
-where $\mu_i$ is the expected log estimated distance for cardinal distance $i$, and  $\sigma_obs$ is the standard deviation of the observation error.
+where $\mu_i$ is the expected log estimated distance for cardinal distance $i$, and  $\sigma_{obs}$ is the standard deviation of the observation error.
 
@@
 Both the intercept, $\alpha$, and the effect of OTC, $\beta$ (which are on the log scale), were estimated from weakly informed normal distributions centered on zero with a standard deviation of 2.5.
-Specifically, this prior encapsulated our knowledge that distances were very unlikely to be > 150 cm.
+Specifically, this prior encapsulated our knowledge that distances were very unlikely to be \textgreater 150 cm.
 
@@
 \begin{equation}\label{eq:censored_obs} 
-censored\;distance_i \sim LN(\mu_i, \sigma_{obs})[26,100],
+censored\;distance_i \sim LN(\mu_i, \sigma_{obs}), 26 \leq censored\;distance_i \leq 100.
 \end{equation}
@@
 As well as modelling how OTC affected shrub seedling growth and mortality rates, we were also interested in examining how warming influenced rates of grass infilling of inter-tussock gaps.
-To examine this we first modelled the median gap radius,$y_{i,t}$, for each inter-tussock gap $i$ in each census $t$, as a random realisation from a lognormal distribution:
+To examine this we first modelled the median gap radius, $y_{i,t}$, for each inter-tussock gap $i$ in each census $t$, as a random realisation from a lognormal distribution:
 
@@
 where $\mu_i$ is the log expected seedling count for each plot, and $\phi$ is the overdispersion parameter.
-We used a negative binomial distribution because count data was significantly overdispersed for both species with $\phi$ being estimated as 0.54 ($\pm$ 0.1 95\% BCI) and 0.82 ($\pm$ 0.1 95\% BCI), for \textit{Grevillea} and \textit{Asterolasia}, respectively.
+We used a negative binomial distribution because count data was significantly overdispersed for both species with $\phi$ being estimated as \Sexpr{phi_value(greaus_seedling_density_model$fit,'mean')} (95\% BCI: \Sexpr{phi_value(greaus_seedling_density_model$fit,'l_bci')}\textendash \Sexpr{phi_value(greaus_seedling_density_model$fit,'u_bci')}) and \Sexpr{phi_value(asttry_seedling_density_model$fit,'mean')} (95\% BCI: \Sexpr{phi_value(asttry_seedling_density_model$fit,'l_bci')}\textendash \Sexpr{phi_value(asttry_seedling_density_model$fit,'u_bci')}), for \textit{Grevillea} and \textit{Asterolasia}, respectively.
 
@@
 \subsection{\textit{Modelling drivers of seedling maximum heights.}}
-In order to validate growth rate findings in the OTC experiment, we assessed how an elevation range of 200 m in elevation (equivalent to a temperature change of 1.5\textdegree C \cite{Brown:1989vx}) influenced \textit{Grevillea} and \textit{Asterolasia} seedlings heights in areas burnt by the 2003 fires.
+In order to validate growth rate findings in the OTC experiment, we assessed how an elevation range of 190 m in elevation (equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C\cite{Brown:1989vx}) influenced \textit{Grevillea} and \textit{Asterolasia} seedlings heights in areas burnt by the 2003 fires.
 We also assessed whether TWI and fire severity influenced these heights.
@@
 Altitude ($Alt_j$), Topographic Wetness Index ($TWI_j$) and mature \textit{Grevillea} count ($ADen_k$) were also continuous covariates that affected all sites in both count and maximum height models.
-Collinearity among covariates was checked and all correlations were r < 0.16.
+Collinearity among covariates was checked and all correlations were r \textless 0.16.
 The data from each continuous covariate was then centered on zero and standardized by two standard deviations\cite{Gelman:2007te}.
 \pagebreak
-\section*{Figures}
+\section*{Supplementary figures}
 \begin{suppfigure}
-\label{figs1}
 \begin{minipage}{\textwidth}
 \includegraphics{figures/figS1.pdf}
-\caption{\textbf{Mean \textit{Asterolasia trymalioides} seedling density with burning, fire severity, adult density, altitude and Topographic Wetness Index (TWI) gradients.} Top left panel = centered and standardized model coefficients (on the log scale).
-Top right panel is the effect of burning.
-All other panels are partial dependence plots for each covariate along their observed gradients in areas burnt by the 2003 fires.
+\caption{\textbf{Mean \textit{Asterolasia trymalioides} seedling density along burning, fire severity (minimum twig diameter), altitude and Topographic Wetness Index (TWI) gradients.} Centered and standardized model coefficients (on the log scale) (\textbf{a}) and effects of burning (\textbf{b}), fire severity  (\textbf{c}), altitude (\textbf{d}) \& topographic wetness (\textbf{e}), in areas burnt by the 2003 fires.
 All errors bars are 95\% Bayesian Credible Intervals.
 Phi = overdispersion parameter for negative binomial distribution.}
+\label{figs1}
 \end{minipage}
@@
 \begin{suppfigure}
-\label{figs2}
 \begin{minipage}{\textwidth}
@@
 \caption{\textbf{Number of \textit{Grevillea australis} recruits in warmed (red) and control (blue) conditions in both 2011 and 2012. Errors bars are 95\% Bayesian Credible Intervals.}}
+\label{figs2}
 \end{minipage}
@@
 \begin{suppfigure}
-\label{figs3}
 \begin{minipage}{\textwidth}
@@
 Red or orange lines = seedlings growing in warmed (OTC) conditions and blue or light blue lines = seedlings growing in control conditions.}
+\label{figs3}
 \end{minipage}
@@
 \begin{suppfigure}
-\label{figs4}
 \begin{minipage}{\textwidth}
 \includegraphics{figures/figS4.pdf}
-\caption{\textbf{Mean maximum seedling height along fire severity (TD; twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires.} Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses.
-The first row shows the centered and standardized model coefficients (on the log scale) for both species.
-Rows 2 to 4 are partial dependence plots for each covariate along their observed gradients.
+\caption{\textbf{Mean maximum seedling height along fire severity (minimum twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires.} Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses.
+Centered and standardized model coefficients (on the log scale) (\textbf{a} \& \textbf{b}) and effects of fire severity  (\textbf{c} \& \textbf{d}), altitude (\textbf{e} \& \textbf{f}) \& topographic wetness (\textbf{g} \& \textbf{h}), in areas burnt by the 2003 fires.
 All errors bars are 95\% Bayesian Credible Intervals.}
+\label{figs4}
 \end{minipage}
@@
 \begin{suppfigure}
-\label{figs5}
 \begin{minipage}{\textwidth}
 \includegraphics{figures/figS5.pdf}
-\caption{\textbf{Mean daily difference (OTC-Control) for ambient temperature, soil temperature, volumetric water content and relative humidity over experimental period.} Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls.
+\caption{\textbf{Mean daily microclimate difference (OTC-Control)}.
+Ambient temperature (\textbf{a}), soil temperature (\textbf{b}), volumetric water content (\textbf{c}) and relative humidity over experimental period (\textbf{d}).
+Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls.
 Shaded areas are the winter snow periods when OTCs were not over plots.
-Note: amplitude changes in later years are due to shading effects caused by increased shrub growth.}
+Note: amplitude declines in later years are due to shading effects caused by increased shrub growth.}
+\label{figs5}
 \end{minipage}
@@
 \begin{suppfigure}
-\label{figs6}
 \begin{minipage}{\textwidth}
 \includegraphics{figures/figS6.pdf}
-\caption{\textbf{Mean minimum daily difference (OTC-Control) for ambient temperature, soil temperature, volumetric water content and relative humidity over experimental period.} Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls.
+\caption{\textbf{Mean minimum daily microclimate difference (OTC-Control)}.
+Ambient temperature (\textbf{a}), soil temperature (\textbf{b}), volumetric water content (\textbf{c}) and relative humidity over experimental period (\textbf{d}).
+Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls.
 Shaded areas are the winter snow periods when OTCs were not over plots.
-Note: amplitude changes in later years are due to shading effects caused by increased shrub growth.}
+Note: amplitude declines in later years are due to shading effects caused by increased shrub growth.}
+\label{figs6}
 \end{minipage}
@@
 \begin{suppfigure}
-\label{figs7}
 \begin{minipage}{\textwidth}
 \includegraphics{figures/figS7.pdf}
-\caption{\textbf{Mean maximum daily difference (OTC-Control) for ambient temperature, soil temperature, volumetric water content and relative humidity over experimental period.} Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls.
+\caption{\textbf{Mean maximum daily microclimate difference (OTC-Control)}.
+Ambient temperature (\textbf{a}), soil temperature (\textbf{b}), volumetric water content (\textbf{c}) and relative humidity over experimental period (\textbf{d}).
+Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls.
 Shaded areas are the winter snow periods when OTCs were not over plots.
-Note: amplitude changes in later years are due to shading effects caused by increased shrub growth.}
+Note: amplitude declines in later years are due to shading effects caused by increased shrub growth.}
+\label{figs7}
 \end{minipage}
@@
 \begin{suppfigure}
-\label{figs8}
 \begin{minipage}{\textwidth}
@@
 it was not within the 110 cm opening of the chamber) and therefore was not used in the experiment.}
+\label{figs8}
 \end{minipage}
 \end{suppfigure}
+
 \pagebreak
+
 \section*{References}

### 2016-03-08 R/manuscript_functions.R
Functions to extract key values for manuscript

@@ -1,0 +1,165 @@
+# These functions are for extracting key values for use in manuscript
+
+# Extract mean density estimates
+density_means <- function(model,species, burnt=TRUE) {
+  if(species == 'Grevillea') {
+  density <- round(summarise_density_predictions(model, 'Grevillea')$burnt_unburnt$mean,2)
+  } else {
+  density <- round(summarise_density_predictions(model, 'Asterolasia')$burnt_unburnt$mean,2)
+  }
+  if(isTRUE(burnt)) {
+    return(density[2])
+  } else
+    return(density[1])
+}
+
+# Calculate density ratios
+density_ratios <- function(model, species) {
+  if(species == 'Grevillea') {
+  density <- round(summarise_density_predictions(model, 'Grevillea')$burnt_unburnt$mean,2)
+  } else {
+  density <- round(summarise_density_predictions(model, 'Asterolasia')$burnt_unburnt$mean,2)
+  }
+  round(density[2]/density[1])
+}
+
+# Extract climate statistics
+micro_clim <- function(data, clim_var, summary_var, statistic) {
+  round(subset(data, sensor==clim_var & clim_stat == summary_var)[[statistic]],1)
+}
+
+# Extract height differences 
+ht_diff <- function(data,species) {
+  max_date <- max(data$date)
+  ctl_ht <- subset(summarise_otc_growth_obs(data, 'ht'),date== max_date & otc=='ctl' & spp == species)
+  otc_ht <- subset(summarise_otc_growth_obs(data, 'ht'),date== max_date & otc=='otc' & spp == species)
+  round(otc_ht$mean - ctl_ht$mean,1)
+}
+
+# Calculate growth rate ratios between otc and ctl
+growth_rate_ratio <- function(model,species) {
+  mod <- summarise_coefficients(model,c('alpha','b_otc'))
+  if(species =='Asterolasia') {
+    n = 1
+  }
+  if(species =='Grevillea') {
+    n = 2
+  }
+  if(species =='Phebalium') {
+    n = 3
+  }
+  if(species =='Prostanthera') {
+    n = 4
+  }
+  pars <- c(paste0('alpha_',n), paste0('b_otc_',n))
+  res <- subset(mod, parameter %in% pars)
+  round((res$mean[1] + res$mean[2])/res$mean[1],1)
+}
+
+# Extract year max height is predicted to be reached
+hmax <- function(model, species, diff=TRUE) {
[…]

### 2016-03-08 manuscript/supplementary.Rnw
removed period in one section

@@
 
-\section*{Data analysis.}
+\section*{Data analysis}
 In total we built 11 hierarchical models to examine factors influencing tussock grass post-fire recovery and shrub seedling establishment, growth and mortality.

### 2016-03-09 R/figures.R
Fixed ylim on asttry density figure

@@ -179,7 +179,7 @@ tussock_plots <- function(tussock_growth_model, tussock_mortality_model) {
   plot_grid(p1,p2,p3,p4, labels=letters[1:4], ncol = 2, label_size = 11)
 }
 
-density_count_plots <- function(density_model, species, ylim=c(0,32)) {
+density_count_plots <- function(density_model, species, ylim=c(0,35)) {
   if(species=='Grevillea') {
     params <- c('alpha_mu','b_unburnt','b_severity','b_altitude','b_twi','b_adult_density','phi')
     y_axis_labels <- c('Intercept','unburnt','severity','altitude','twi','adult density','phi')

### 2016-03-09 R/models.R
Upped adapt_delta for non tussock mort model

@@ -111,7 +111,7 @@ run_censored_model <- function(data) {
                         data = stan_data,
                         pars = c("full_data", "sigma_log_census","sigma_log_plot","sigma_log_ind","sigma_log_obs","log_alpha","y_censored","log_b_otc","log_census_error","log_plot_error","log_ind_error"), 
                         chains = 3,
-                        control=list(adapt_delta=0.995, stepsize=0.01, max_treedepth =15),
+                        control=list(adapt_delta=0.999, stepsize=0.01, max_treedepth =15),
                         iter = 2000))
   return(fit)
 }
@@ -560,7 +560,7 @@ run_non_tussock_mortalilty_model <- function(data) {
                          pars = c("mu_alpha","mu_b_otc",
                                   "sigma_alpha","sigma_b_otc","sigma_plot","sigma_ind",
                                   "alpha","b_otc", "p_death_otc","p_death_ctl"), 
-                         control=list(adapt_delta=0.999999,stepsize=0.1, max_treedepth =15),
+                         control=list(adapt_delta=0.99999999,stepsize=0.1, max_treedepth =15),
                          chains = 3,
                          iter = 2000))
   return(fit)

### 2016-03-09 manuscript/manuscript.Rnw
More minor edits

@@
 \begin{abstract}
-Climate change is expected to increase fire activity and exacerbate woody plant encroachment in arctic and alpine grasslands.
+Climate change is expected to increase fire activity and exacerbate woody plant encroachment in arctic and alpine landscapes.
 What is unknown is whether increases in woody plants and fire will interact to exacerbate changes in structure, function and composition of these ecosystems.
@@
 In alpine and tundra ecosystems, current evidence from long-term and experimental studies indicate that shrub recruitment and encroachment is highest in disturbed areas\cite{Williams:1987ua, Batllori:2009fe, Frost:2013tm}, that climatic warming is likely to improve woody species growth rates\cite{Arft:1999iz, Elmendorf:2012ba, MyersSmith:2011fp}, and that shrubs are the most flammable vegetation component of these ecosystems\cite{Williams:2006tl, Higuera:2008dw, Higuera:2009ga}.
-The combination of this evidence indicates that more frequent or severe fire will increase recruitment opportunities for woody species, which when coupled with higher growth and survival rates caused by rising temperatures, will exacerbate shrub expansion\cite{Racine:2004tn,Lantz:2013fh} and ultimately increase the likelihood of more fires\cite{Higuera:2008dw, Higuera:2009ga,MyersSmith:2011fp}.
-In effect, warming will strengthen an existing feedback loop that not only has the potential to cause rapid changes in the composition and structure of alpine and arctic vegetation, but also have serious social, biodiversity and carbon sequestration consequences\cite{Mack:2011ew}.
+The combination of this evidence indicates that more frequent or severe fire may increase recruitment opportunities for woody species, which when coupled with higher growth and survival rates caused by rising temperatures, will exacerbate shrub expansion\cite{Racine:2004tn,Lantz:2013fh} and ultimately increase the likelihood of more fires\cite{Higuera:2008dw, Higuera:2009ga,MyersSmith:2011fp}.
+In effect, warming could strengthen an existing feedback loop that not only has the potential to cause rapid changes in the composition and structure of alpine and arctic vegetation, but also have serious social, biodiversity and carbon sequestration consequences\cite{Mack:2011ew}.
 
@@
 For example, we have little information on whether fire will stimulate shrub seedling recruitment in alpine or tundra ecosystems, nor do we know how rates of seedling growth and survival will be affected under warmer post-fire conditions.
-Furthermore, we do not know how tussock grasslands (the other dominant community in these ecosystems) will respond to warmer post-fire conditions, and whether grasses have competitive, facilitative or no effect on shrub seedling demographic rates, or whether such effects are altered by warmer post-fire conditions.
+Furthermore, we do not know how tussock grasslands (the other dominant community in these ecosystems) will respond to warmer post-fire conditions, and whether grasses have competitive, facilitative or no effect on shrub seedling vital rates, or whether such effects are altered by warmer post-fire conditions.
 
@@
 We first investigated how altitude, topographic wetness (TWI), adult density, fire and, if burnt, fire severity, influenced the density of \textit{Grevillea} and \textit{Asterolasia} seedlings (the two dominant shrubs of open heathland).
-In 2012-13, we estimated seedling density for both species across 40 alpine sites and found that the abundance of \textit{Grevillea} (Fig.
+In 2011-12, we estimated seedling density for both species across 40 alpine sites and found that the abundance of \textit{Grevillea} (Fig.
 \ref{fig2}) and \textit{Asterolasia} (Supplementary Fig.
@@
 As expected, pre-fire adult density positively influenced \textit{Grevillea} seedling density, with sites containing higher abundances of pre-fire \textit{Grevillea} adults containing more seedlings.
-For both shrub species, we detected no significant change in seedling density along a 200 m altitudinal range (equivalent to a \Sexpr{round(0.77 * (200/100),1)}\textdegree C change in mean temperature\cite{Brown:1989vx}).
+For both shrub species, we detected no significant change in seedling density along a \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m altitudinal range (equivalent to a \Sexpr{round(0.77 * (200/100),1)}\textdegree C change in mean temperature\cite{Brown:1989vx}).
 The field warming experiment (see below) indicates that this lack of elevation response in \textit{Grevillea} is due to both rates of recruitment (Supplementary Fig.
@@
 \ref{fig5}).
-As with the seedlings growing with no nearby \textit{poa tussock}, we detected a strong positive effect of warming on growth rates and a marginal, yet not significant, decrease in mortality.
+As with the seedlings growing with no nearby \textit{Poa} tussock, we detected a strong positive effect of warming on growth rates and a marginal, yet not significant, decrease in mortality.
 We also found no significant effect of inter-tussock gap radius on growth or mortality rates.
@@
 
-Despite having little impact on shrub seedling growth and mortality rates, tussock grass may still limit shrub recruitment, and thus, shrub expansion, by infilling post-fire bare ground gaps at a faster rate under warmer conditions.
+Despite having little impact on shrub seedling growth and mortality rates, tussock grass may still limit shrub recruitment, and thus, shrub expansion, by infilling post-fire bare ground gaps (whether by vegatative growth or seedlings) at a faster rate under warmer conditions.
 Using five years of post-fire inter-tussock gap size changes in warmed and unwarmed plots we found that gaps were being infilled by grasses (Fig.
 \ref{fig6}).
-However, the rate at which this occurred was very slow, with a 10 cm radius gap decreasing by approximately \Sexpr{gap_size(gap_dynamic_model, 'otc')} cm over a ten year period.
+However, the rate at which this occurred was very slow, with a 10 cm radius gap predicted to decrease by approximately \Sexpr{gap_size(gap_dynamic_model, 'otc')} cm over a ten year period.
 We also detected no significant effect of a \Sexpr{micro_clim(overall_microclimate_summary,'CAmbient_Temp','mean','grand_mean')}\textdegree C temperature rise on the rate of grass infilling.
@@
 We used 40 long-term open heathland sites\cite{Camac:2013eb} established after the 2003 fires consisting of 30 burnt sites and 10 sites thought to be unburnt for over 70 years.
-In the summer of 2012-13, at each site, seedling density/m$^2$ was estimated using 40 1 m$^2$ quadrats, evenly distributed in groups of 10 along four 50 m transects, with 10 m between transect lines, subsampling an area of 2000 m$^2$.
+In the summer of 2011-12, at each site, seedling density/m$^2$ was estimated using 40 1 m$^2$ quadrats, evenly distributed in groups of 10 along four 50 m transects, with 10 m between transect lines, subsampling an area of 2000 m$^2$.
 Within plots we recorded the number and maximum height of \textit{Grevillea} and \textit{Asterolasia} seedlings.
@@
 For each model (Supplementary Methods) we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package Rstan \Sexpr{packageVersion("rstan")}.
-Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.
+Data and source code for reproducing analysis and figures are available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.
 Additional information of experimental design and analysis can be found in Supplementary methods.

### 2016-03-09 manuscript/supplementary.Rnw
more minor edits

@@
 \section*{Study site}
-Research sites were on the Bogong High Plains, a series of alpine and high subalpine plateaux in the Alpine National Park, approximately 250 km northeast of Melbourne, Australia.
+Our research sites were on the Bogong High Plains, a series of alpine and high subalpine plateaux in the Alpine National Park, approximately 250 km northeast of Melbourne, Australia.
 The undulating terrain ranges in altitude from 1400 to 2000 m a.s.l and contains a well-developed mantle of soil.
@@
 \section*{Open Top Chamber Experiment}
-We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of grass recovery and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affect shrub seedling demographic rates, and 3) whether such affects change under warmer conditions post-fire conditions.
+We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of grass bare ground gap infilling and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affect shrub seedling demographic rates, and 3) whether such affects change under warmer conditions post-fire conditions.
 
@@
 Below we describe the model structure, parameters and priors used in these models.
-The residuals of all models were checked graphically against predictions with none showing any systematic pattern or severe heteroscedasticity.
 To ensure all covariate effects were identifiable we selectively removed each covariate and visually inspected whether the direction and magnitude of effects were severely altered; none were.

### 2016-03-09 R/data_processing.R
Fixed changes to tidyr complete package which now requires nesting instead of c

@@ -151,7 +151,7 @@ clean_unburnt_data <- function(csv_file) {
     select(site, year, transect=tran, greaus_tran_adultden=greaus_adultden,
            plot, poa, bare_ground=bg, rock, species, adult, height, 
            diameter=stemdiam, canopy_length_1 = clength1, canopy_length_2 = clength2) %>%
-    complete(c(site,year,transect,greaus_tran_adultden,plot,poa,bare_ground,rock),species) %>%
+    tidyr::complete(nesting(site,year,transect,greaus_tran_adultden,plot,poa,bare_ground,rock),species) %>%
     filter(!is.na(species)) %>%
     mutate(greaus_tran_adultden = greaus_tran_adultden/500) %>% # converts to density/m2
     droplevels()

### 2016-03-09 R/figures.R
Bug in latest ggplot release results in custom_theme having no axis lines. This commit fixes this

@@ -7,6 +7,8 @@ partial_plot_theme <- function(legend.position = "none", strips = FALSE,...) {
                           legend.title = element_blank(),
                           strip.background = sb,
                           legend.position = legend.position,
+                          axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
+                          axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'),
                           plot.margin = unit(c(5,5,5,5), "mm"))
 }
 
@@ -14,6 +16,8 @@ partial_plot_theme <- function(legend.position = "none", strips = FALSE,...) {
 coefficent_plot_theme <- function() {
   theme_classic(base_size = 10) + theme(
                           axis.title.y = element_blank(),
+                          axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
+                          axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'),
                           plot.margin = unit(c(5,5,5,5), "mm"))
 }
 

### 2016-03-09 R/models.R
tightened sigma priors on mortality models to reduce divergence. Where too vague before

@@ -536,10 +536,10 @@ run_non_tussock_mortalilty_model <- function(data) {
     raw_ind ~ normal(0,1);
     mu_alpha ~ cauchy(0,2.5);
     mu_b_otc ~ cauchy(0,2.5);
-    sigma_alpha ~ cauchy(0,25);
-    sigma_b_otc ~ cauchy(0,25);
-    sigma_plot ~ cauchy(0,25);
-    sigma_ind ~ cauchy(0,25);
+    sigma_alpha ~ cauchy(0,2.5);
+    sigma_b_otc ~ cauchy(0,2.5);
+    sigma_plot ~ cauchy(0,2.5);
+    sigma_ind ~ cauchy(0,2.5);
   }
   generated quantities {
     real p_death_ctl[n_spp];
@@ -560,7 +560,7 @@ run_non_tussock_mortalilty_model <- function(data) {
                          pars = c("mu_alpha","mu_b_otc",
                                   "sigma_alpha","sigma_b_otc","sigma_plot","sigma_ind",
                                   "alpha","b_otc", "p_death_otc","p_death_ctl"), 
-                         control=list(adapt_delta=0.99999999,stepsize=0.1, max_treedepth =15),
+                         control=list(adapt_delta=0.9999,stepsize=0.001, max_treedepth =15),
                          chains = 3,
                          iter = 2000))
   return(fit)
@@ -654,8 +654,8 @@ run_tussock_mortality_model <- function(data, pred_poadist_range = NULL) {
     b_otc ~ cauchy(0,2.5);
     b_poadist ~ cauchy(0,2.5);
     b_otc_x_poadist ~ cauchy(0,2.5);
-    sigma_plot ~ cauchy(0,25);
-    sigma_ind ~ cauchy(0,25);
+    sigma_plot ~ cauchy(0,2.5);
+    sigma_ind ~ cauchy(0,2.5);
   }
   generated quantities{
   real p_death_ctl[n_preds];

### 2016-03-09 manuscript/manuscript.Rnw
More edits to captions

@@
 Solid blue lines = known mechansims.
-dahsed black lines = mechansims that may exacerbate or diminish this feedback loop but which we have a paucity of information on.
+dashed black lines = mechansims that may exacerbate or diminish this feedback loop but which we have a paucity of information on.
 Feedback loop based on new evidence obtained from experimental and field surveys conducted in this study (\textbf{b}).
@@
 \includegraphics{figures/fig2.pdf}
-\caption{\textbf{Mean \textit{Grevillea australis} seedling density along burning, fire severity (minimum twig diameter), adult density, altitude and Topographic Wetness Index (TWI) gradients.} Centered and standardized model coefficients (on the log scale) (\textbf{a}) and effects of burning (\textbf{b}), fire severity  (\textbf{c}), altitude (\textbf{d}), topographic wetness (\textbf{e}) \& adult density (\textbf{f}), in areas burnt by the 2003 fires.
-All errors bars are 95\% Bayesian Credible Intervals.
+\caption{\textbf{Mean \textit{Grevillea australis} seedling density along gradients of burning, fire severity (minimum twig diameter), adult density, altitude and Topographic Wetness Index (TWI).} Centered and standardized model coefficients (on the log scale) (\textbf{a}) and effects of burning (\textbf{b}), fire severity  (\textbf{c}), altitude (\textbf{d}), topographic wetness (\textbf{e}) \& adult density (\textbf{f}), in areas burnt by the 2003 fires.
+All errors are 95\% Bayesian Credible Intervals.
 Phi = overdispersion parameter for negative binomial distribution.}
@@
 Thin lines represent individual growth trajectories.
-Mean ($\pm$ 95\% Bayesian Credible Interval) projected height growth for each species over 70 years (\textbf{b}).
+Mean ($\pm$ 95\% Bayesian Credible Interval) projected growth trajectories (\textbf{b}).
 In all cases, red \& orange lines = seedlings growing in warmed (OTC) conditions and blue \& light blue lines = seedlings growing in control conditions.
 Projections were based on an logistic growth model using initial seedling size of 6 cm (the mean initial height observed in the OTC experiment) and mean maximum heights observed in long-unburnt sites (i.e.
-\Sexpr{hmax_values(hmax,'Asterolasia')}, \Sexpr{hmax_values(hmax,'Grevillea')}, \Sexpr{hmax_values(hmax,'Phebalium')} \& \Sexpr{hmax_values(hmax,'Prostanthera')} cm for \textit{Asterolasia}, \textit{Grevillea},\textit{Phebalium}, \& \textit{Prostanthera}, respectively).}
+\Sexpr{hmax_values(hmax,'Asterolasia')}, \Sexpr{hmax_values(hmax,'Grevillea')}, \Sexpr{hmax_values(hmax,'Phebalium')} \& \Sexpr{hmax_values(hmax,'Prostanthera')} cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Phebalium}, \& \textit{Prostanthera}, respectively).}
 \label{fig3}
@@
 \includegraphics{figures/fig4.pdf}
-\caption{\textbf{Mean ($\pm$ 95\% Bayesian Credible Interval) seedling probability of mortality for each species growing in control (blue) and warmed (red) conditions.}}
+\caption{\textbf{Mean ($\pm$ 95\% Bayesian Credible Interval) annual probability of mortality for each species growing in control (blue) and warmed (red) conditions.}}
 \label{fig4}
@@
 \caption{\textbf{Experimental warming and inter-tussock gap size effects on \textit{Grevillea australis} seedlings.} Top rows = Growth rate effects, Second row = Mortality effects.
-\textbf{a} \& \textbf{c} are the centered and standardized model coefficients; \textbf{b} \& \textbf{d} are growth and mortality rate response curves along an inter-tussock gap size gradient under warmed (red) and control (blue) conditions.
-All errors bars are 95\% Bayesian Credible Intervals.}
+Centered and standardized model coefficients (\textbf{a} \& \textbf{c}); growth and mortality rate response curves along an inter-tussock gap size gradient in warmed (red) and control (blue) conditions  (\textbf{b} \& \textbf{d}).
+All errors are 95\% Bayesian Credible Intervals.}
 \label{fig5}

### 2016-03-09 manuscript/supplementary.Rnw
More edits to text. Include updated priors for mortality models and changes to captions

@@
 The parameter $\beta_2$ is the effect of inter-tussock gap radius, which varied over time $t$, while $\beta_3$ is the interactive effect of OTC and time-varying inter-tussock gap radius.
-The inclusion of these parameters, coupled $Gap_{i,t}$ being centered and standardized (see Data inputs into growth and mortality models), alters the definition of the intercept, $\alpha$, which now represents the mean control plot growth rate at the average inter-tussock gap size (\Sexpr{mean_init_gap(tussock_otc_growth_data)} cm).
+The inclusion of these parameters, coupled with $Gap_{i,t}$ being centered and standardized (see Data inputs into growth and mortality models), alters the definition of the intercept, $\alpha$, which now represents the mean control plot growth rate at the average inter-tussock gap size (\Sexpr{mean_init_gap(tussock_otc_growth_data)} cm).
 The parameter  $\beta_1$ is the OTC effect and $\epsilon_{plot_k}$ and $\epsilon_i$ are residual errors at the plot and individual level, respectively.
@@
 
-As with growth rates, we modelled each individual's cumulative hazard per census, $Cumulative\;hazard_{i,t}$, separately for seedlings growing in bare ground patches to those in inter-tussock gaps.
+As with growth rates, we modelled each individual's cumulative hazard per census, \\
+$Cumulative\;hazard_{i,t}$, separately for seedlings growing in bare ground patches to those in inter-tussock gaps.
 In both cases, $Cumulative\;hazard_{i,t}$, was modelled on the complementary log-log scale as a function of an instantaneous hazard rate for each individual, $\lambda_i$, multiplied by the length (in years) between each census, $t$.
@@
 
-Like the growth models, the standard deviation parameters in both mortality models (i.e.
-$\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{ind}$) were each estimated from weakly informative positive half Cauchy priors with scale parameters of 25.
+Here the standard deviation parameters in both mortality models (i.e.
+$\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{ind}$) were each estimated from positive half Cauchy priors with scale parameters of 2.5.
 The intercept term $\alpha$ for inter-tussock seedling model and the mean hyper-parameter $\mu_{spp}^{CTL}$ in the multispecies models were estimated from cauchy priors with a scale parameter of 2.5.
@@
 In each model, OTC treatment, $OTC_k$, was a binary (0 = Control, 1 = OTC) plot level covariate.
-By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census t.
+By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census $t$.
 However, because some cardinal distances were recorded as censored data (i.e.
@@
 \subsection{\textit{Modelling censored Poa distances.}}
-Approximately \Sexpr{prop_censored(cleaned_poa_distances)}\% of the Poa cardinal distance data used to calculate the average inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, ≥ 26 cm.
+Approximately \Sexpr{prop_censored(cleaned_poa_distances)}\% of the $Poa$ cardinal distance data used to calculate the average inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, (i.e.
+$\geq$ 26 cm).
 We estimated these censored values by first estimating the distribution of distances in observed data, coupled with estimating covariate effects and then using these distributions to estimate censored observations.
@@
 gaps decreased over time) because it is theoretically possible for inter-tussock gaps to increase over time.
-Here $\alpha_i$ is a known constant that represents the initial inter-tussock gap radius for each gap $i$ at the beginning (i.e.
+Here $\alpha_i$ is a known constant that represents the initial inter-tussock gap radius for each gap $i$ (i.e.
 the intercept).
@@
 Half Cauchy priors with a scale of 25 were used for all standard deviation parameters and the overdispersion parameter $\phi$.
-Hyperparameter $\mu_site$ and coefficients $\beta_1$ to $\beta_5$ were estimated using normal priors centered on zero with a standard deviation of 2.5, encapsulating our prior knowledge that plots were unlikely to have more than 150 seedlings/m$^2$.
+Hyperparameter $\mu_{site}$ and coefficients $\beta_1$ to $\beta_5$ were estimated using normal priors centered on zero with a standard deviation of 2.5, encapsulating our prior knowledge that plots were unlikely to have more than 150 seedlings/m$^2$.
 
@@
 \begin{equation}\label{eq:maxht_obs} 
-maxht_i \sim LN(\widehat{\mu}_i, \sigma_{obs}),
+maxht_i \sim LN(\widehat{maxht}_i, \sigma_{obs}),
 \end{equation}
@@
 where the intercept, $\alpha_j$, varies by site and represents the average log maximum seedling height of each site, assuming all other parameters are at their mean.
-We estimated $\alpha_j$ from a normal distribution centred on $\mu_{site}$, the average log maximum height for the average site and a standard deviation of $sigma_{site}$ which defines site variability.
+We estimated $\alpha_j$ from a normal distribution centred on $\mu_{site}$, the average log maximum height for the average site and a standard deviation of $\sigma_{site}$ which defines site variability.
 The parameters $\beta_1$ and $\beta_2$ are the effects of altitude and topographic wetness, respectively and $\beta_3$ is the effect of fire severity.
@@
 \includegraphics{figures/figS1.pdf}
-\caption{\textbf{Mean \textit{Asterolasia trymalioides} seedling density along burning, fire severity (minimum twig diameter), altitude and Topographic Wetness Index (TWI) gradients.} Centered and standardized model coefficients (on the log scale) (\textbf{a}) and effects of burning (\textbf{b}), fire severity  (\textbf{c}), altitude (\textbf{d}) \& topographic wetness (\textbf{e}), in areas burnt by the 2003 fires.
+\caption{\textbf{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (minimum twig diameter), altitude and Topographic Wetness Index (TWI).} Centered and standardized model coefficients (on the log scale) (\textbf{a}) and effects of burning (\textbf{b}), fire severity  (\textbf{c}), altitude (\textbf{d}) \& topographic wetness (\textbf{e}), in areas burnt by the 2003 fires.
 All errors bars are 95\% Bayesian Credible Intervals.
@@
 Thin lines represent individual growth trajectories.
-Red or orange lines = seedlings growing in warmed (OTC) conditions and blue or light blue lines = seedlings growing in control conditions.}
+Red \& orange lines = seedlings growing in warmed (OTC) conditions and blue \& light blue lines = seedlings growing in control conditions.}
 \label{figs3}
@@
 Centered and standardized model coefficients (on the log scale) (\textbf{a} \& \textbf{b}) and effects of fire severity  (\textbf{c} \& \textbf{d}), altitude (\textbf{e} \& \textbf{f}) \& topographic wetness (\textbf{g} \& \textbf{h}), in areas burnt by the 2003 fires.
-All errors bars are 95\% Bayesian Credible Intervals.}
+All errors are 95\% Bayesian Credible Intervals.}
 \label{figs4}

### 2016-03-14 R/figures.R
Fixed few typos and removed phi from density coefficient plot

@@ -185,12 +185,12 @@ tussock_plots <- function(tussock_growth_model, tussock_mortality_model) {
 
 density_count_plots <- function(density_model, species, ylim=c(0,35)) {
   if(species=='Grevillea') {
-    params <- c('alpha_mu','b_unburnt','b_severity','b_altitude','b_twi','b_adult_density','phi')
-    y_axis_labels <- c('Intercept','unburnt','severity','altitude','twi','adult density','phi')
+    params <- c('alpha_mu','b_unburnt','b_severity','b_altitude','b_twi','b_adult_density')
+    y_axis_labels <- c('Intercept','Unburnt','Severity','Altitude','TWI','Adult density')
   }
   if(species=='Asterolasia'){
-    params <- c('alpha_mu','b_unburnt','b_severity','b_altitude','b_twi','phi')
-    y_axis_labels <- c('Intercept','unburnt','severity','altitude','twi','phi')
+    params <- c('alpha_mu','b_unburnt','b_severity','b_altitude','b_twi')
+    y_axis_labels <- c('Intercept','Unburnt','Severity','Altitude','TWI')
   }
   coeffs <- summarise_coefficients(density_model,params = params)
   predictions <- summarise_density_predictions(density_model,species)
@@ -205,7 +205,7 @@ density_count_plots <- function(density_model, species, ylim=c(0,35)) {
   
   if(species=='Grevillea') {
     p6 <- partial_plot_density_height(predictions$adult_density, x ='sim_adult_den', 
-                                      xlab = expression('Adult density/'~m^2), ylab = expression('Seedlings /'~m^2), ylim)
+                                      xlab = expression('Adults/'~m^2), ylab = expression('Seedlings /'~m^2), ylim)
     plot_grid(p1,p2,p3,p4,p5,p6, labels=letters[1:6], ncol = 2, label_size = 11)
   }
   

### 2016-03-14 R/models.R
Converted density models to poisson models as they were not overdispersed

@@ -757,7 +757,6 @@ parameters{ # Declare parameters the models must estimate
   real<lower=0> b_transect_sigma;
   real b_raw_transect[n_transects];
   %s
-  real<lower=0> phi;
 }
 transformed parameters { # Declare and define derived variables used in model
   real count[n_obs];
@@ -796,12 +795,11 @@ model { # Define priors and likelihood
   b_twi ~ normal(0,2.5);
   b_raw_transect ~ normal(0,1);
   b_transect_sigma ~ cauchy(0,25);
-  phi ~ cauchy(0,25);
   %s
 
   
   # Likelihood
-    y ~ neg_binomial_2(count, phi);
+    y ~ poisson(count);
 }
 generated quantities { # Calculate log likelihood, residuals or make predictions
   
@@ -840,12 +838,12 @@ generated quantities { # Calculate log likelihood, residuals or make predictions
   
   if(species =="Grevillea") {
     pars <- c("alpha_mu", "alpha_sigma","b_transect_sigma",
-              "b_unburnt","b_severity","b_altitude","b_twi","phi", "b_adult_density",
+              "b_unburnt","b_severity","b_altitude","b_twi", "b_adult_density",
               "pred_count_unburnt","pred_count_burnt","pred_count_severity","pred_count_altitude",
               "pred_count_twi","pred_count_adult_density")
   }else {
     pars <- c("alpha_mu", "alpha_sigma","b_transect_sigma",
-              "b_unburnt","b_severity","b_altitude","b_twi","phi",
+              "b_unburnt","b_severity","b_altitude","b_twi",
               "pred_count_unburnt","pred_count_burnt","pred_count_severity","pred_count_altitude",
               "pred_count_twi")}
   

### 2016-03-14 manuscript/manuscript.Rnw
Added JBaums edits

@@
 \title{Climatic warming strengthens a positive feedback between alpine shrubs and fire}
-\author{James S. Camac\textasteriskcentered$^{1,2,3}$, Richard J. Williams$^4$, Carl-Henrik Wahren$^3$, Ary A. Hoffmann$^5$ \\
+\author{James S. Camac$^{*,1,2,3}$, Richard J. Williams$^4$, Carl-Henrik Wahren$^3$, Ary A. Hoffmann$^5$ \\
 \& Peter A. Vesk$^2$}
@@
 \item Research Centre for Applied Alpine Ecology, La Trobe University, Melbourne 3086, Victoria, Australia.
-\item CSIRO Land and Water, tropical ecosystems research centre, PMB 44 Winnellie 0821, NT, Australia.
+\item CSIRO Land and Water, Tropical Ecosystems Research Centre, PMB 44 Winnellie 0821, NT, Australia.
 \item Bio21 Institute, School of BioSciences, The University of Melbourne, Vic. 3010, Australia.
@@
 \begin{abstract}
-Climate change is expected to increase fire activity and exacerbate woody plant encroachment in arctic and alpine landscapes.
-What is unknown is whether increases in woody plants and fire will interact to exacerbate changes in structure, function and composition of these ecosystems.
+Climate change is expected to increase fire activity and woody plant encroachment in arctic and alpine landscapes.
+However, it is unknown whether increases in woody species and fire will interact to exacerbate changes in structure, function and composition of these ecosystems.
 Here we use field surveys and experimental manipulations to examine warming and fire effects on recruitment, growth and survival of seedlings of evergreen obligate seeding alpine shrubs.
-We find that fire substantially increased shrub seedling establishment (up to \Sexpr{density_ratios(greaus_seedling_density_model, 'Grevillea')}) and that warming doubled tall shrub seedling growth rates and can increase survival.
-Warming also had weak or no effect on shrub recruitment, rates of grass post-fire gap infilling or grass competitive effects on shrub seedling growth and survival.
-These findings indicate that rising temperatures coupled with more frequent or severe fires will likely exacerbate increases in tall evergreen shrubs by increasing recruitment, doubling growth rates, and potentially increasing seedling survival.
+We find that fire substantially increased shrub seedling establishment (up to \Sexpr{density_ratios(greaus_seedling_density_model, 'Grevillea')}\textendash fold) and that warming doubled tall shrub seedling growth rates and could increase survival.
+Warming had limited or no effect on shrub recruitment, post-fire gap-infilling by grass, or competitive effects of grass on shrub seedling growth and survival.
+These findings indicate that rising temperatures coupled with more frequent or severe fires will likely exacerbate increases in tall evergreen shrubs by increasing recruitment, doubling growth rates, and potentially increasing shrub survival.
 \end{abstract}
@@
 Because disturbance strongly influences recruitment opportunities, and thus, composition and structure of plant communities and biomes worldwide\cite{Bond:2005eg, Turner:2010wo}, it has the capacity to exacerbate or diminish how climate affects vegetation.
-Consequently, it is imperative to understand how seedlings, the life stage most vulnerable to climate and disturbance and the life stage which determines the long-term persistence of a species and its capacity to establish in new areas, responds to both changing climatic and disturbance regimes\cite{Walck:2011dm, Briceno:2015iv}.
+Seedlings are the life stage most vulnerable to climate and disturbance, and the life stage which determines the long-term persistence of a species and its capacity to establish in new areas.
+Consequently, it is imperative to understand how seedlings respond to both changing climate and disturbance regimes\cite{Walck:2011dm, Briceno:2015iv}.
 
@@
 The frequency and extent of wildfires have also increased in these environments over recent decades, a trend expected to continue\cite{Westerling:2006ve, Flannigan:2009kl, Qiu:2009th, Bradstock:2014jp}.
-In alpine and tundra ecosystems, current evidence from long-term and experimental studies indicate that shrub recruitment and encroachment is highest in disturbed areas\cite{Williams:1987ua, Batllori:2009fe, Frost:2013tm}, that climatic warming is likely to improve woody species growth rates\cite{Arft:1999iz, Elmendorf:2012ba, MyersSmith:2011fp}, and that shrubs are the most flammable vegetation component of these ecosystems\cite{Williams:2006tl, Higuera:2008dw, Higuera:2009ga}.
-The combination of this evidence indicates that more frequent or severe fire may increase recruitment opportunities for woody species, which when coupled with higher growth and survival rates caused by rising temperatures, will exacerbate shrub expansion\cite{Racine:2004tn,Lantz:2013fh} and ultimately increase the likelihood of more fires\cite{Higuera:2008dw, Higuera:2009ga,MyersSmith:2011fp}.
-In effect, warming could strengthen an existing feedback loop that not only has the potential to cause rapid changes in the composition and structure of alpine and arctic vegetation, but also have serious social, biodiversity and carbon sequestration consequences\cite{Mack:2011ew}.
+In alpine and tundra ecosystems, current evidence from long-term and experimental studies indicate that shrub recruitment and encroachment is highest in disturbed areas\cite{Williams:1987ua, Batllori:2009fe, Frost:2013tm}, that climatic warming is likely to increase woody species growth rates\cite{Arft:1999iz, Elmendorf:2012ba, MyersSmith:2011fp}, and that shrubs are the most flammable component of these ecosystems\cite{Williams:2006tl, Higuera:2008dw, Higuera:2009ga}.
+The combination of this evidence indicates that more frequent or severe fire may increase recruitment opportunities for woody species, which when coupled with higher growth and survival rates caused by rising temperatures, will exacerbate shrub expansion\cite{Racine:2004tn,Lantz:2013fh} and ultimately increase the likelihood of fire\cite{Higuera:2008dw, Higuera:2009ga,MyersSmith:2011fp}.
+In effect, warming could strengthen an existing feedback loop that not only has the potential to cause rapid changes in the composition and structure of alpine and arctic vegetation, but also has serious social, biodiversity and carbon sequestration consequences\cite{Mack:2011ew}.
 
@@
 \ref{fig1}a).
-For example, we have little information on whether fire will stimulate shrub seedling recruitment in alpine or tundra ecosystems, nor do we know how rates of seedling growth and survival will be affected under warmer post-fire conditions.
-Furthermore, we do not know how tussock grasslands (the other dominant community in these ecosystems) will respond to warmer post-fire conditions, and whether grasses have competitive, facilitative or no effect on shrub seedling vital rates, or whether such effects are altered by warmer post-fire conditions.
+For example, we have little information on whether fire will stimulate shrub seedling recruitment in alpine or tundra ecosystems, nor do we know how rates of seedling growth and survival will be affected under warmer, more exposed, post-fire conditions.
+Furthermore, we do not know how tussock grasslands (the other dominant community in these ecosystems) will respond to warmer post-fire conditions, whether grasses have competitive, facilitative or no effect on shrub seedling vital rates, or whether such effects are altered by warmer post-fire conditions.
 
@@
 As with tundra ecosystems, the Australian Alps have experienced rapid changes in climate.
-Since 1979, mean growing season temperatures have risen by approximately 0.4\textdegree C and annual precipitation has fallen by 6\%\cite{Wahren:2013ji}, with a consequent decline in snow pack\cite{SanchezBayo:2013ci}.
-These climatic changes have been correlated with a 10 to 20\% increase in shrub cover and a 25\% decline in graminoids\cite{Wahren:2013ji}.
-Much of the Australian Alps has also been burnt by recent wildfires, the frequency and severity of which is expected to increase\cite{Bradstock:2014jp,Williams:2014ud}.
+Since 1979, mean growing season temperatures have risen by approximately 0.4\textdegree C and annual precipitation has fallen by 6\%\cite{Wahren:2013ji}, with a consequent decline in snow pack depth\cite{SanchezBayo:2013ci}.
+These climatic changes have been correlated with a 10 to 20\% increase in shrub cover and a 25\% decline in graminoids cover\cite{Wahren:2013ji}.
+Much of the Australian Alps has also been burnt by recent wildfires, the frequency and severity of which are expected to increase\cite{Bradstock:2014jp,Williams:2014ud}.
 
 We took advantage of recent fires in alpine open heathland, a plant community that occupies ca.
-25\% of the Australian alpine landscape.
-Under global warming, this plant community is likely to encroach grasslands and is itself susceptible to shrub thickening\cite{McDougall:2003gb}.
-To determine how warmer post-fire conditions affects shrub cover, we combined field observations with a warming experiment that used seedlings of four dominant Australian evergreen obligate seeding alpine shrubs \textemdash \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub),  \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub).
-This combination of data sources allowed us to examine: 1) landscape scale drivers of shrub seedling establishment; 2) how warmer post-fire conditions influence rates of both grass recovery and shrub seedling recruitment, growth and mortality; 3) how proximity to grass affect shrub seedling demographic rates, and 4) whether such grass effects change under warmer conditions post-fire conditions.
+25\% of the Australian alpine landscape\cite{Williams:2014ud}.
+Under global warming, this plant community is likely to encroach upon grasslands and is itself susceptible to shrub thickening\cite{McDougall:2003gb}.
+To determine drivers of shrub establishment and how warmer post-fire conditions affect shrub cover, we combined field observations with a warming experiment that used seedlings of four Australian evergreen obligate seeding shrubs: \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub),  \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub).
+This combination of data sources allowed us to examine: 1) landscape scale drivers of shrub seedling establishment; 2) how warmer post-fire conditions influence rates of grass recovery and shrub seedling recruitment, growth and mortality; 3) how proximity to grass affect shrub seedling demographic rates; and 4) whether such grass effects change under warmer post-fire conditions.
 
 \subsection{Drivers of shrub seedling establishment.}
-We first investigated how altitude, topographic wetness (TWI), adult density, fire and, if burnt, fire severity, influenced the density of \textit{Grevillea} and \textit{Asterolasia} seedlings (the two dominant shrubs of open heathland).
-In 2011-12, we estimated seedling density for both species across 40 alpine sites and found that the abundance of \textit{Grevillea} (Fig.
+We first investigated how altitude, Topographic Wetness Index (TWI), adult density, fire and fire severity (as measured by minimum twig diameters\textemdash see Supplementary Methods) influenced the density of \textit{Grevillea} and \textit{Asterolasia} seedlings (the two dominant shrubs of alpine open heathland).
+In 2011-12, we estimated seedling density for both species across 40 alpine sites and found that the abundances of \textit{Grevillea} (Fig.
 \ref{fig2}) and \textit{Asterolasia} (Supplementary Fig.
 S1) seedlings were strongly influenced by the occurrence of fire.
-Sites burnt by fire in 2003 had seedling densities between \Sexpr{density_ratios(asttry_seedling_density_model,'Asterolasia')} and \Sexpr{density_ratios(greaus_seedling_density_model,'Grevillea')} times higher than unburnt sites.
-In \textit{Grevillea}, mean seedling densities were \Sexpr{density_means(greaus_seedling_density_model,'Grevillea', TRUE)}/m$^2$ at burnt sites and \Sexpr{density_means(greaus_seedling_density_model,'Grevillea', FALSE)}/m$^2$ at unburnt sites.
+Sites burnt in 2003 had seedling densities between \Sexpr{density_ratios(asttry_seedling_density_model,'Asterolasia')} and \Sexpr{density_ratios(greaus_seedling_density_model,'Grevillea')} times higher than unburnt sites.
+The mean seedling density of \textit{Grevillea}, was \Sexpr{density_means(greaus_seedling_density_model,'Grevillea', TRUE)}/m$^2$ at burnt sites and \Sexpr{density_means(greaus_seedling_density_model,'Grevillea', FALSE)}/m$^2$ at unburnt sites.
 \textit{Asterolasia} had similar mean densities: \Sexpr{density_means(asttry_seedling_density_model,'Asterolasia', TRUE)} and \Sexpr{density_means(asttry_seedling_density_model, 'Asterolasia', FALSE)} seedlings/m$^2$ at burnt and unburnt sites, respectively.
-For both species, seedling density was also higher in severely burnt sites.
-As expected, pre-fire adult density positively influenced \textit{Grevillea} seedling density, with sites containing higher abundances of pre-fire \textit{Grevillea} adults containing more seedlings.
-For both shrub species, we detected no significant change in seedling density along a \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m altitudinal range (equivalent to a \Sexpr{round(0.77 * (200/100),1)}\textdegree C change in mean temperature\cite{Brown:1989vx}).
-The field warming experiment (see below) indicates that this lack of elevation response in \textit{Grevillea} is due to both rates of recruitment (Supplementary Fig.
+For both species, seedling density was also higher in severely burnt sites (i.e.
+sites with higher minimum twig diameters).
+As expected, pre-fire adult density positively influenced \textit{Grevillea} seedling density.
+For both shrub species, we detected no change in seedling density along a \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m altitudinal range (equivalent to a \Sexpr{round(0.77 * (200/100),1)}\textdegree C change in mean temperature\cite{Brown:1989vx}).
+The field warming experiment (see below) indicates that this lack of elevation response in \textit{Grevillea} is due to rates of recruitment (Supplementary Fig.
 S2) and mortality (see below) largely being insensitive to a \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C change in temperature.
-We detected no significant effect of topographic wetness for either species.
+We detected no significant effect of Topographic Wetness Index for either species.
 
@@
 As shrub seedlings were more abundant in burnt vegetation, we investigated how warmer post-fire conditions affected seedling growth and mortality rates.
-Seedlings of \textit{Grevillea}, \textit{Asterolasia}, \textit{Phebalium} and \textit{Prostanthera} emerging after a wildfire were transplanted into experimentally burnt plots that were either passively warmed by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C using open top chambers (OTCs) or subject to current ambient conditions (controls).
-After \Sexpr{n_days(chamber_dates, FALSE)} days (\Sexpr{n_days(chamber_dates, TRUE)} growing season days where OTC treatment was applied) or \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} years growth, mean seedling heights of the tall shrubs (\textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera}) growing in post-fire bare ground were taller in warmed plots by \Sexpr{ht_diff(non_tussock_otc_growth_data,'Grevillea')}, \Sexpr{ht_diff(non_tussock_otc_growth_data,'Phebalium')} and \Sexpr{ht_diff(non_tussock_otc_growth_data,'Prostanthera')} cm, respectively (Fig.
+Seedlings of \textit{Grevillea}, \textit{Asterolasia}, \textit{Phebalium} and \textit{Prostanthera} emerging after a wildfire were transplanted into experimentally burnt plots.
+These plots were either enclosed in Open Top Chambers (OTCs) whose temperature was \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C warmer than control plots subjected to current ambient conditions.
+After \Sexpr{n_days(chamber_dates, FALSE)} days (\Sexpr{n_days(chamber_dates, TRUE)} growing season days) or \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} years' growth, mean seedling heights of the tall shrubs (\textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera}) growing in post-fire bare ground were greater in warmed plots relative to controls by \Sexpr{ht_diff(non_tussock_otc_growth_data,'Grevillea')}, \Sexpr{ht_diff(non_tussock_otc_growth_data,'Phebalium')} and \Sexpr{ht_diff(non_tussock_otc_growth_data,'Prostanthera')} cm, respectively (Fig.
 \ref{fig3}).
-Warming increased heights of both \textit{Grevillea} and \textit{Prostanthera} seedlings in all years, while \textit{Phebalium} did not respond to warming treatment until the second growing season.
-By contrast, seedlings of the prostrate shrub, \textit{Asterolasia}, showed no significant difference in growth rate between warmed and control plots in any year.
+Warming increased heights of both \textit{Grevillea} and \textit{Prostanthera} seedlings in all years, while \textit{Phebalium} did not respond to the warming treatment until the second growing season.
+By contrast, seedlings of the prostrate shrub, \textit{Asterolasia}, showed no difference in growth rate between warmed and control plots in any year.
 For each species, we also observed similar treatment trends with stem diameter growth (Supplementary Fig.
 S3).
-Accounting for initial height and assuming logistic growth, mean annual predicted height growth rates of \textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera} were \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Grevillea')}, \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Phebalium')} and \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Prostanthera')} times that observed in control plots, respectively.
-Projecting these differences equates to an average 6 cm tall seedling (the mean initial size used in this experiment) growing in warmed conditions reaching maximum mature height in \Sexpr{hmax(non_tussock_otc_growth_model,'Grevillea')} fewer years for \textit{Grevillea} and \Sexpr{hmax(non_tussock_otc_growth_model,'Prostanthera')} fewer years for both \textit{Phebalium} and \textit{Prostanthera} (Fig.
+Accounting for initial height and assuming logistic growth, the rates of change in mean annual predicted height of \textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera} were \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Grevillea')}, \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Phebalium')} and \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Prostanthera')} times that observed in control plots, respectively.
+According to this model, a 6 cm seedling (the mean initial height of seedlings used in this experiment) attains maximum height 39 years sooner \Sexpr{hmax(non_tussock_otc_growth_model,'Grevillea')} (\textit{Grevillea}) or \Sexpr{hmax(non_tussock_otc_growth_model,'Prostanthera')} years sooner (\textit{Phebalium} and \textit{Prostanthera}) when warmed by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C (Fig.
 \ref{fig3}).
-\textit{Asterolasia} was predicted to reach a mature height in approximately \Sexpr{round(hmax(non_tussock_otc_growth_model,'Asterolasia', FALSE)[2])} to \Sexpr{round(hmax(non_tussock_otc_growth_model,'Asterolasia', FALSE)[1])} years.
+\textit{Asterolasia} was predicted to reach its maximum height in approximately \Sexpr{round(hmax(non_tussock_otc_growth_model,'Asterolasia', FALSE)[2])} to \Sexpr{round(hmax(non_tussock_otc_growth_model,'Asterolasia', FALSE)[1])} years, irrespective of warming treatment.
 
-We validated the \textit{Grevillea} and \textit{Asterolasia} growth responses observed in experimentally warmed plots by examining the effect of elevation on maximum height data collected from 2003 post-fire recruits across 40 open heathland sites.
-Here, a 200 m altitudinal range is equivalent to a mean ambient temperature difference of approximately \Sexpr{round(0.77 * (200/100),1)}\textdegree C\cite{Brown:1989vx}, which is comparable to that observed between experimental warmed and control plots (\Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C).
-In response to shifts in temperature, maximum heights of \textit{Grevillea} post-fire recruits were expected to decrease with elevation, whereas \textit{Asterolasia} seedlings were not expected to show this pattern.
+We validated the \textit{Grevillea} and \textit{Asterolasia} growth responses observed in experimentally warmed plots by examining the effect of elevation on maximum heights of post-fire recruits across 30 open heathland sites burnt in 2003.
+Here, a 200 m altitudinal range is equivalent to a mean ambient temperature difference of approximately \Sexpr{round(0.77 * (200/100),1)}\textdegree C\cite{Brown:1989vx}, which is comparable to that observed between experimentally warmed and control plots (\Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C).
+In response to shifts in temperature, the maximum height of \textit{Grevillea} post-fire recruits was expected to decrease with elevation, whereas \textit{Asterolasia} seedlings were not expected to show this pattern.
 Our experimental predictions were verified (Supplementary Fig.
 S4).
-Mean maximum heights of \textit{Grevillea} seedlings in burnt open heathland were \Sexpr{max_ht_alt(greaus_seedling_max_ht_model)} cm taller at 1670 m a.s.l compared to seedlings at 1860 m a.s.l.
+Mean maximum height of \textit{Grevillea} seedlings in burnt open heathland were \Sexpr{max_ht_alt(greaus_seedling_max_ht_model)} cm taller at 1670 m a.s.l compared to seedlings at 1860 m a.s.l.
 (\Sexpr{max_ht_alt(greaus_seedling_max_ht_model, FALSE)[1]} cm vs \Sexpr{max_ht_alt(greaus_seedling_max_ht_model, FALSE)[2]} cm; a difference comparable to our experimental findings).
-In contrast, mean maximum heights of \textit{Asterolasia} seedlings did not vary significantly with elevation.
-Topographic Wetness Index (TWI) and fire severity had no detectable influence on maximum seedling heights in either species.
+In contrast, mean maximum height of \textit{Asterolasia} seedlings did not vary significantly with elevation.
+Topographic Wetness Index and fire severity had no detectable influence on maximum seedling heights in either species.
 
-\subsection{Experimental warming effects on shrub seedling mortality.}
+\subsection{Experimental warming on shrub seedling mortality.}
 
@@
 Warming also reduced mean seedling mortality in \textit{Grevillea} and \textit{Phebalium} (Fig.
-\ref{fig4}); however, for both species, the effect was not statistically significant.
-By contrast, annual mortality rates in the prostrate shrub, \textit{Asterolasia}, were marginally higher in warmed plots, but again this effect was not significant (Fig.
+\ref{fig4}); however, for both species, the effect was highly uncertain (i.e.
+credible intervals overlap).
+By contrast, annual mortality rates in the prostrate shrub, \textit{Asterolasia}, were marginally higher in warmed plots, but again this effect was highly uncertain (Fig.
 \ref{fig4}).
 
-\subsection{Grass proximity effects on shrub seedling growth and mortality rates.}
+\subsection{Effects of tussock grass proximity on shrub seedling growth and mortality rates.}
 
 Climate change is expected to alter biotic interactions because their strength and direction depend strongly on climatic conditions, particularly in alpine and arctic ecosystems\cite{Callaway:2002vb, Klanderud:2005uw}.
-Here, we assess the interactive effects of warming and inter-tussock gap size (i.e.
-the proximity of shrub seedlings to grass tussocks) on the growth and survival of \textit{Grevillea} seedlings (Fig.
+Here, we assess the interactive effects of warming and grass proximity on the growth and survival of \textit{Grevillea} seedlings transplanted into various sized inter-tussock gaps (Fig.
 \ref{fig5}).
-As with the seedlings growing with no nearby \textit{Poa} tussock, we detected a strong positive effect of warming on growth rates and a marginal, yet not significant, decrease in mortality.
-We also found no significant effect of inter-tussock gap radius on growth or mortality rates.
-However, for growth, we did detect a marginally significant interaction between warming and gap size, such that in warmed plots growth rates slightly increased with gap size.
-This suggests that warmer conditions may increase competition between shrubs and grasses, a finding consistent with the abiotic stress hypothesis\cite{Callaway:2002vb}.
-However, this effect was small and unlikely to have significant biological consequence on shrub growth rates.
+We detected a strong positive effect of warming treatment on growth rates and a marginal, yet not significant, decrease in mortality.
+However, we did not detect significant inter-tussock gap size effects or an interaction between gap size and warming treatment on either growth or mortality rates (i.e.
+coefficient credible intervals overlap zero).
 
-\subsection{Experimental warming and rates of tussock grass gap infilling.}
+\subsection{Experimental warming and rates of gap infilling by tussock grass.}
 
-Despite having little impact on shrub seedling growth and mortality rates, tussock grass may still limit shrub recruitment, and thus, shrub expansion, by infilling post-fire bare ground gaps (whether by vegatative growth or seedlings) at a faster rate under warmer conditions.
-Using five years of post-fire inter-tussock gap size changes in warmed and unwarmed plots we found that gaps were being infilled by grasses (Fig.
+Despite having little impact on shrub seedling growth and mortality rates, tussock grass may still limit shrub recruitment, and thus, shrub expansion, by infilling post-fire bare ground gaps (whether by vegetative growth or seedlings) faster under warmer conditions.
+Using five years of post-fire inter-tussock gap size changes in warmed and unwarmed plots we found that gaps were being infilled by tussock grasses (Fig.
 \ref{fig6}).
 However, the rate at which this occurred was very slow, with a 10 cm radius gap predicted to decrease by approximately \Sexpr{gap_size(gap_dynamic_model, 'otc')} cm over a ten year period.
-We also detected no significant effect of a \Sexpr{micro_clim(overall_microclimate_summary,'CAmbient_Temp','mean','grand_mean')}\textdegree C temperature rise on the rate of grass infilling.
+We also detected no significant effect of a \Sexpr{micro_clim(overall_microclimate_summary,'CAmbient_Temp','mean','grand_mean')}\textdegree C temperature rise on the rate of infilling.
 
@@
 These demographic effects will ultimately result in shrub thickening and expansion into grasslands.
-These effects, coupled with field\cite{Williams:2006tl}, experimental\cite{Fraser:NHT5Kx_W} and paleoecological\cite{Higuera:2008dw} evidence indicate that shrubs are the most flammable vegetation component of alpine and tundra ecosystems, and suggest that flammable fuel loads will accumulate faster and cover a larger proportion of the alpine landscape under a warmer environment.
-This further strengthens the feedback\cite{Higuera:2008dw, Higuera:2009ga} by potentially increasing the frequency or severity of fires, which then create further recruitment opportunities for shrubs with little demographic impact of neighbouring tussock grass.
+Coupled with field\cite{Williams:2006tl}, experimental\cite{Fraser:NHT5Kx_W} and paleoecological\cite{Higuera:2008dw} evidence, which indicate that shrubs are the most flammable component of alpine and tundra ecosystems, our results suggest that flammable fuel loads will accumulate faster and cover a larger proportion of the alpine and arctic landscapes under a warmer environment.
+This further strengthens the feedback\cite{Higuera:2008dw, Higuera:2009ga} by potentially increasing the frequency or severity of fires, which then creates further recruitment opportunities for shrubs with little demographic impact of neighbouring tussock grass.
 
@@
 The most obvious mechanism that will break this cycle involves shorter fire intervals that prevent obligate seeding shrubs reaching reproductive age and thereby exhausting the seedbank.
-However, this scenario is unlikely for the majority of alpine (or tundra) landscapes including in Australia, where it would require current fire intervals of 50 to 100 years to decrease to less than 20 years \textemdash the time estimated for the species in this study to reach reproductive maturity\cite{Williams:2008ba}.
-Furthermore, if reproductive output is related to plant size\cite{Wenk:2015ko}, then climatic warming may allow obligate seeding species to reach reproductive maturity sooner, and consequently, increase species resilience to short fire intervals.
+However, this scenario is unlikely for the majority of alpine (or tundra) landscapes, including in Australia, where current fire intervals of 50 to 100 years would need to decrease to less than 20 years\textemdash the time estimated for the species in this study to reach reproductive maturity\cite{Williams:2008ba}.
+Furthermore, if reproductive output is related to plant size\cite{Wenk:2015ko}, then climatic warming may allow obligate seeding species to reach reproductive maturity sooner, and consequently, may increase species resilience to short fire intervals.
 Nevertheless, these factors and others such as changes in snow pack\cite{Wipf:2009jh}, soil moisture\cite{Engler:2011vq} or herbivory\cite{Post:2008wz} require further research because they are likely to be altered by the interactive effects of climate and disturbance in unpredictable ways\cite{MyersSmith:2011fp}.
@@
 But more importantly, our results provide evidence for underlying processes that could result in a warming-fire-shrub feedback that has been hypothesized in paleoecological studies\cite{Higuera:2008dw, Higuera:2009ga}.
-Based on current IPCC observations and predictions, average global temperature has already increased by 0.85\textdegree C since 1880 and is expected to rise by as much as 4.8\textdegree C by 2100\cite{IPCC:2013vv}.
+Based on current observations and predictions, average global temperature has already increased by 0.85\textdegree C since 1880 and is expected to rise by as much as 4.8\textdegree C by 2100\cite{IPCC:2013vv}.
 In alpine and tundra environments, temperatures\cite{ChapinIII:2005js}, shrub cover\cite{MyersSmith:2011fp} and the frequency and severity of fire\cite{Westerling:2006ve, Flannigan:2009kl, Qiu:2009th, Bradstock:2014jp} have all increased in the last few decades.
-These changes mean that the warming-shrub-fire feedback loop identified here has already strengthened, which could cause grasslands and other non-woody communities to transition to an alternative state of more shrubs and more fire, both of which are likely to have consequences for carbon sequestration, water supply and biodiversity.
+These changes mean that the warming-shrub-fire feedback loop identified here has already strengthened, which could cause grasslands and other non-woody communities to transition to an alternative state with more shrubs and more fire, both of which are likely to have consequences for carbon sequestration, water supply and biodiversity.
 
@@
 It is also an ecotone between closed heathland (\textgreater 70\% shrub cover) and tussock grassland consisting of \textit{Grevillea australis} shrubs interspersed among a \textit{Poa hiemata} sward.
-Aerial photography has revealed that the plant community has acted as an \textit{invasion front} of shrub expansion into grassland and has itself experienced shrub thickening by closed heathland dominants such as \textit{Prostanthera cuneata}.
+Aerial photography has revealed that the plant community has acted as an invasion front of shrub expansion into grassland and has itself experienced shrub thickening by closed heathland dominants such as \textit{Prostanthera cuneata}.
 
@@
 
-In March 2010, at 1750 m a.s.l, we burnt 32 randomly selected mature (60 cm tall and 1.5m$^2$) \textit{Grevillea australis} shrubs in open heathland to create patches of bare ground approximately 0.7 m$^2$ surrounded by burnt tussock grass, simulating disturbance in open heathland burnt by wildfire.
+In March 2010, at 1750 m a.s.l, we burnt 32 randomly selected mature (60 cm tall and 1.5 m$^2$) \textit{Grevillea australis} shrubs in open heathland to create patches of bare ground approximately 0.7 m$^2$ surrounded by burnt tussock grass, simulating disturbance in open heathland burnt by wildfire.
 After creating the bare ground patches we collected seedlings of dominant alpine shrub species from a nearby (\textless 2 km) site of similar altitude burnt by a late 2006 wildfire.
-We collected shrub seedlings of two dominant open heathland species (\textit{Grevillea australis} (Proteaceae; a tall shrub) and \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), a dominant closed heathland species (\textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) that typically grows on warmer aspects and a species common to both open and closed heathland (\textit{Phebalium squamulosum} (Rutaceae; a tall shrub).
-All four species are obligate seeders post-fire (i.e.
-fire killed)\cite{Camac:2013eb}.
+We collected shrub seedlings of two dominant open heathland species \textit{Grevillea australis} (Proteaceae; a tall shrub) and \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), a dominant closed heathland species \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) that typically grows on warmer aspects and a species common to both open and closed heathland \textit{Phebalium squamulosum} (Rutaceae; a tall shrub).
+All four species are killed by fire\cite{Camac:2013eb}.
 
-A total of 640 seedlings, 256 \textit{Grevillea} (half used in \textit{Poa} inter-tussock experiment \textemdash see below) and 128 for each of \textit{Asterolasia}, \textit{Prostanthera} and \textit{Phebalium} were used.
+A total of 640 seedlings, 256 \textit{Grevillea} (half used in \textit{Poa} inter-tussock experiment\textemdash see below) and 128 for each of \textit{Asterolasia}, \textit{Prostanthera} and \textit{Phebalium} were used.
 Four seedlings per species were randomly selected and transplanted into a 4\texttimes 4 square grid in the center of each burnt patch, with 14 cm between individuals and the edge of the patch dominated by resprouting tussock grass \textit{Poa hiemata}.
@@
 The experimental site was fenced to prevent grazing by deer and horses.
-Rabbits or invertebrates also did not graze on these seedlings.
+We detected no obvious signs of rabbit or invertebrate herbivory within our plots.
 
-To simulate near-term warmer conditions predicted under IPCC\cite{IPCC:2013vv}, we randomly assigned Open Top Chambers (OTCs) to half (16) the plots, with the remainder treated as unwarmed controls.
+To simulate near-term warmer conditions indicated by the IPCC\cite{IPCC:2013vv}, we randomly assigned Open Top Chambers (OTCs) to half (16) the plots, with the remainder treated as unwarmed controls.
 OTCs were placed over plots, ensuring all seedlings (including inter-tussock shrub seedlings) occurred within the 1.1 m$^2$ open top to minimise edge effects.
@@
 
-Microclimatic changes were measured hourly using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) at four control and four OTC plots.
-Across \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} growing seasons (\Sexpr{n_days(chamber_dates, TRUE)} chamber days), OTCs simulated warmer conditions at the lower end of IPCC predictions\cite{IPCC:2013vv} (Supplementary Methods and Supplementary Fig.
+Microclimatic conditions were measured hourly using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) at four control and four OTC plots.
+Across \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} growing seasons (\Sexpr{n_days(chamber_dates, TRUE)} growing season days), OTCs simulated warmer conditions at the lower end of IPCC projections\cite{IPCC:2013vv} (Supplementary Fig.
 S5-7).
@@
 Seedling survival, maximum height and stem diameter (nearest mm measured with Vernier calipers) were initially recorded in May 2010 and then subsequently re-measured at the end of each growing season (May-June).
-At the same time, we recorded the distance to nearest tussock or grass seedling in each of four cardinal directions for seedlings growing in inter-tussock gaps.
-Thus changes in inter-tussock gap size could be due to either vegetative growth or seedling establishment.
+At the same time, we recorded the distance to the nearest tussock or grass seedling in each of four cardinal directions for shrub seedlings growing in inter-tussock gaps.
+Thus, changes in inter-tussock gap size could be due to either vegetative growth or seedling establishment.
 We did not measure individual characteristics (e.g.
@@
 
-We used 40 long-term open heathland sites\cite{Camac:2013eb} established after the 2003 fires consisting of 30 burnt sites and 10 sites thought to be unburnt for over 70 years.
-In the summer of 2011-12, at each site, seedling density/m$^2$ was estimated using 40 1 m$^2$ quadrats, evenly distributed in groups of 10 along four 50 m transects, with 10 m between transect lines, subsampling an area of 2000 m$^2$.
+We used 40 long-term open heathland sites established after the 2003 fires\cite{Camac:2013eb} consisting of 30 burnt sites and 10 sites thought to be unburnt for over 70 years.
+In the summer of 2011-12, at each site, seedling density/m$^2$ was estimated using 40 quadrats, each 1 m$^2$, that were evenly distributed in groups of 10 along four 50 m transects, with 10 m between transect lines, subsampling an area of 2000 m$^2$.
 Within plots we recorded the number and maximum height of \textit{Grevillea} and \textit{Asterolasia} seedlings.
-For unburnt sites we counted the number of mature \textit{Grevillea} plants (\textgreater 0.5 m$^2$) within 5 m either side of each transect.
-In burnt sites, this required counting the number of skeletons (there were no living adults at any burnt site) that still persist post-fire.
-We were unable estimate numbers of adult \textit{Asterolasia} because this species does not have a persistent woody skeleton post-fire.
-Site level data, elevation and Topographic Wetness Index (TWI), were obtained from a 30 m resolution digital elevation model.
-Lastly, for burnt sites, fire severity was estimated using minimum twig diameters, collected immediately after the 2003 fires.
+For unburnt sites we counted the number of mature \textit{Grevillea} plants (\textgreater 0.5 m$^2$) within 5 m of each transect.
+In burnt sites, this required counting the number of skeletons (there were no living adults at any burnt site) that still persisted post-fire.
+We were unable to estimate numbers of adult \textit{Asterolasia} because this species does not have a persistent woody skeleton post-fire.
+Site level data, elevation and Topographic Wetness Index (TWI) were obtained from a 30 m resolution digital elevation model.
+Lastly, for burnt sites, fire severity was estimated by minimum twig diameters, collected immediately after the 2003 fires\cite{Camac:2013eb}.
 
@@
 
-In total we built 11 hierarchical models to examine factors influencing tussock grass post-fire recovery and shrub seedling establishment, growth and mortality.
-For each model (Supplementary Methods) we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package Rstan \Sexpr{packageVersion("rstan")}.
+In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality.
+For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package rstan \Sexpr{packageVersion("rstan")}.
 Data and source code for reproducing analysis and figures are available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.
-Additional information of experimental design and analysis can be found in Supplementary methods.
+Additional information about experimental design and analysis can be found in Supplementary Methods.
 \end{methods}
@@
 William Morris and John Baumgartner provided insightful chats about modelling approaches and Warwick Papst provided field logistic advice.
-Daniel Falster and Richard Fitzjohn for enabled these analyses to be readily reproducible.
-Special thanks to David Bowman, Anaïs Gibert, Brad Farmilo, John Morgan, Kimberley Millers, Dale Nimmo, Michaela Plein, Libby Rumpff and Inka Veltheim for their comments on this manuscript.
+Daniel Falster and Richard Fitzjohn for developing software that enabled these analyses to be readily reproducible.
+Special thanks to David Bowman, Ana{\"i}s Gibert, Brad Farmilo, John Morgan, Kimberley Millers, Dale Nimmo, Michaela Plein, Libby Rumpff and Inka Veltheim for their comments on this manuscript.
 
@@
 J.C.
-conceived, designed, performed the experiments, field surveys and analysis; C-H.W, R.W, A.H and P.V.
+conceived, designed and performed the experiments, field surveys and analysis; C-H.W, R.W, A.H and P.V.
 supervised the development of this work, aided in data collection and provided statistical advice.
@@
 \includegraphics{figures/fig1.pdf}
-\caption{\textbf{Conceptual diagram illustrating positive warming-shrub-fire feedback loop.} Hypothesised positive feedback loop between fire, climatic warming and landscape flammability (\textbf{a}).
-Solid blue lines = known mechansims.
-dashed black lines = mechansims that may exacerbate or diminish this feedback loop but which we have a paucity of information on.
-Feedback loop based on new evidence obtained from experimental and field surveys conducted in this study (\textbf{b}).
+\caption{\textbf{Conceptual diagram illustrating positive warming-shrub-fire feedback loop.} (\textbf{a}) Hypothesised positive feedback loop between fire, climatic warming and landscape flammability.
+Solid blue lines = known mechansims; Dashed black lines = mechansims that may exacerbate or diminish this feedback loop but which we have a paucity of information on.
+(\textbf{b}) Feedback loop based on new evidence obtained from experimental and field surveys conducted in this study.
 Also includes an example of a potential, but unlikely, effect (red dotted line) that could break this feedback.}
@@
 \includegraphics{figures/fig2.pdf}
-\caption{\textbf{Mean \textit{Grevillea australis} seedling density along gradients of burning, fire severity (minimum twig diameter), adult density, altitude and Topographic Wetness Index (TWI).} Centered and standardized model coefficients (on the log scale) (\textbf{a}) and effects of burning (\textbf{b}), fire severity  (\textbf{c}), altitude (\textbf{d}), topographic wetness (\textbf{e}) \& adult density (\textbf{f}), in areas burnt by the 2003 fires.
-All errors are 95\% Bayesian Credible Intervals.
-Phi = overdispersion parameter for negative binomial distribution.}
+\caption{\textbf{Mean \textit{Grevillea australis} seedling density along gradients of burning, fire severity (minimum twig diameter), altitude, Topographic Wetness Index (TWI) and adult density.} (\textbf{a}) Centered and standardized model coefficients (on the log scale); and effects of: (\textbf{b}) burning, (\textbf{c}) fire severity, (\textbf{d}) altitude, (\textbf{e}) topographic wetness and (\textbf{f}) adult density, in areas burnt by the 2003 fires.
+All bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
 \label{fig2}
@@
 \includegraphics{figures/fig3.pdf}
-\caption{\textbf{Observed and projected growth trajectories of four dominant Australian alpine shrubs.} Observed height growth (\textbf{a}): Thick lines represent mean ($\pm$ 95\% confidence intervals) observed heights at each May census.
+\caption{\textbf{Observed and projected growth trajectories of four dominant Australian alpine shrubs.} (\textbf{a}) Observed height growth: Thick lines represent mean ($\pm$ 95\% confidence intervals) observed heights at each May census.
 Thin lines represent individual growth trajectories.
-Mean ($\pm$ 95\% Bayesian Credible Interval) projected growth trajectories (\textbf{b}).
-In all cases, red \& orange lines = seedlings growing in warmed (OTC) conditions and blue \& light blue lines = seedlings growing in control conditions.
+(\textbf{b}) Mean ($\pm$ 95\% Bayesian Credible Interval) projected growth trajectories.
+In all cases, red \& orange lines = seedlings growing in warmed (OTC) conditions and blue and light blue lines = seedlings growing in control conditions.
 Projections were based on an logistic growth model using initial seedling size of 6 cm (the mean initial height observed in the OTC experiment) and mean maximum heights observed in long-unburnt sites (i.e.
@@
 \includegraphics{figures/fig4.pdf}
-\caption{\textbf{Mean ($\pm$ 95\% Bayesian Credible Interval) annual probability of mortality for each species growing in control (blue) and warmed (red) conditions.}}
+\caption{\textbf{Mean ($\pm$ 95\% Bayesian Credible Interval) annual probability of death for each species growing in control (blue) and warmed (red) conditions.}}
 \label{fig4}
@@
 \caption{\textbf{Experimental warming and inter-tussock gap size effects on \textit{Grevillea australis} seedlings.} Top rows = Growth rate effects, Second row = Mortality effects.
-Centered and standardized model coefficients (\textbf{a} \& \textbf{c}); growth and mortality rate response curves along an inter-tussock gap size gradient in warmed (red) and control (blue) conditions  (\textbf{b} \& \textbf{d}).
-All errors are 95\% Bayesian Credible Intervals.}
+(\textbf{a} \& \textbf{c}) Centered and standardized model coefficients; (\textbf{b} \& \textbf{d}) growth and mortality rate response curves along an inter-tussock gap size gradient in warmed (red) and control (blue) conditions.
+All error bars and shading are 95\% Bayesian Credible Intervals.}
 \label{fig5}
@@
 \includegraphics{figures/fig6.pdf}
-\caption{\textbf{Effects of warming on rates of tussock grass gap infilling.}
-Model coefficients (\textbf{a}) and projected temporal change in size for an average 10 cm inter-tussock gap in warmed (red) and unwarmed (blue) conditions (\textbf{b}).} 
+\caption{\textbf{Effects of warming on rates of gap infilling by tussock grass.} (\textbf{a}) Model coefficients and (\textbf{b}) projected temporal change in size for an average 10 cm inter-tussock gap in warmed (red) and unwarmed (blue) conditions.
+All error bars and shading are 95\% Bayesian Credible Intervals.}
 \label{fig6}

### 2016-03-14 manuscript/supplementary.Rnw
Added Jbaums edits

@@
 \title{Climatic warming strengthens a positive feedback between alpine shrubs and fire (Supplementary Material)}
-\author{James S. Camac\textasteriskcentered$^{1,2,3}$, Richard J. Williams$^4$, Carl-Henrik Wahren$^3$, Ary A. Hoffmann$^5$ \\
+\author{James S. Camac$^{*,1,2,3}$, Richard J. Williams$^4$, Carl-Henrik Wahren$^3$, Ary A. Hoffmann$^5$ \\
 \& Peter A. Vesk$^2$}
@@
 \section*{Open Top Chamber Experiment}
-We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of grass bare ground gap infilling and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affect shrub seedling demographic rates, and 3) whether such affects change under warmer conditions post-fire conditions.
+We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of post-fire gap-infilling by  tussock grass; and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affects shrub seedling demographic rates; and 3) whether such effects change under warmer conditions post-fire conditions.
 
 In alpine and tundra regions, shrub seedlings typically establish and grow in bare ground created by disturbance\cite{Williams:1987ua, Batllori:2009fe, Frost:2013tm}.
-Here, we used fire to create bare ground because 1) Australian alpine shrub seedlings occur in greatest abundance after fire (see seedling survey results) and 2) in many parts of the world, including Australia, wildfires are predicted to increase in frequency and severity\cite{Westerling:2006ve,Flannigan:2009kl}.
+Here, we used fire to create bare ground because Australian alpine shrub seedlings occur in greatest abundance after fire (see seedling survey results) and in many parts of the world, including Australia, wildfires are predicted to increase in frequency and severity\cite{Westerling:2006ve,Flannigan:2009kl}.
 Thus, to forecast likely changes in shrub abundance we need to know how warming affects growth and mortality rates of shrub seedlings growing in burnt sites.
 
-We collected seedlings from a site burnt by a late 2006 fire because it allowed us to collect sufficient seedling numbers from species occurring in both open and closed heathland.
+We collected seedlings from a site burnt by a late 2006 fire because it allowed us to collect sufficient numbers of seedlings from species occurring in both open and closed heathland.
 We focused on \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) because they are all: obligate seeders post-fire (i.e.
-fire killed), common in the Australian Alps, and under climatic warming, likely candidates to invade non-shrubby plant communities such as alpine grasslands and herbfields.
-For each species, we collected a range of seedling heights (\Sexpr{range_init_size(complete_otc_seedling_data)[1]} to \Sexpr{range_init_size(complete_otc_seedling_data)[2]} cm) using a forestry planter (12.5 cm deep by 5.5 cm wide).
+fire killed); common in the Australian Alps; and under climatic warming, are likely to invading non-shrubby plant communities such as alpine grasslands and herbfields.
+For each species, we collected seedlings with a range of heights (\Sexpr{range_init_size(complete_otc_seedling_data)[1]} to \Sexpr{range_init_size(complete_otc_seedling_data)[2]} cm) using a forestry planter (12.5 cm deep by 5.5 cm wide).
 The size of the planter was selected to minimize root damage across the range of seedling sizes and to allow for the extraction of a constant volume (378 cm$^3$) of intact soil and roots.
@@
 Four seedlings of each species were then transplanted into 70 cm$^2$ of burnt bare ground in the center of each experimental plot.
-An additional four \textit{Grevillea} seedlings were transplanted into various sized inter-tussock gaps surrounding this bare ground.
+An additional four \textit{Grevillea} seedlings were transplanted into naturally occuring inter-tussock gaps of varying size surrounding this bare ground.
 Only \textit{Grevillea} seedlings were transplanted into inter-tussock gaps because of space constraints but also because it is the dominant open heathland shrub species (i.e.
@@
 The combination of bare ground and inter-tussock seedlings meant that each experimental plot had a total of 20 seedlings per 1 m$^2$ (Fig.
-S8).
+S\ref{figs8}).
 However, the two sets of seedlings (i.e.
@@
 
-To simulate near-term warmer conditions predicted under IPCC\cite{IPCC:2013vv}, we used Open Top Chambers (OTCs)\cite{Molau:1996vc}.
+To simulate near-term warmer conditions indicated by the IPCC\cite{IPCC:2013vv}, we used Open Top Chambers (OTCs)\cite{Molau:1996vc}.
 OTCs were hexagonal, 58 cm tall with an open-top diameter of 110 cm and base diameter of 168 cm.
@@
 
-We measured changes in microclimate using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) randomly installed into four control plots and four OTC plots.
-Each Micro Station measured ambient temperature, relative humidity (both at 10 cm above ground), soil temperature (3 cm below ground) and volumetric soil moisture (between 3 and 10 cm) at hourly intervals for the entire duration of the experiment.
+We measured changes in microclimate using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) installed in four randomly chosen control plots and four OTC plots.
+Each Micro Station measured ambient temperature (10 cm above ground), relative humidity (10 cm above ground), soil temperature (3 cm below ground) and volumetric soil moisture (between 3 and 10 cm) at hourly intervals for the entire duration of the experiment.
 Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions\cite{IPCC:2013vv} (Supplementary Fig.
 S5-7).
-OTCs passively increased average ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C, and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively.
+OTCs passively increased average ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively.
 Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C).
@@
 This provided us with the opportunity to examine how warming directly affected recruitment.
-We marked (with paint on stems) and recorded the number of recruits that germinated in both 2011 and 2012 seasons.
+We marked (with paint on stems) recruits and recorded the number that germinated in each of the 2011 and 2012 seasons.
 
@@
 We measured seedling density and maximum heights in open heathland across the Bogong High Plains.
-The primary aim of this study was to evaluate how variables such as altitude, Topographic Wetness Index (a measure of plant available water), adult density, fire, and, if burnt, fire severity, influence the density of \textit{Asterolasia} and \textit{Grevillea} seedlings in open heathland.
-A secondary aim was to validate growth rates observed in the OTC experiment by assessing maximum seedling heights of \textit{Asterolasia} and \textit{Grevillea} along altitude, soil moisture and fire severity gradients in areas burnt by the 2003 wildfires.
+The primary aim of this study was to evaluate how variables such as altitude, Topographic Wetness Index (a measure of plant available water), adult density, fire, and fire severity influence the density of \textit{Asterolasia} and \textit{Grevillea} seedlings in open heathland.
+A secondary aim was to verify growth rates observed in the OTC experiment by assessing maximum seedling heights of \textit{Asterolasia} and \textit{Grevillea} along altitude, soil moisture and fire severity gradients in areas burnt by the 2003 wildfires.
 
 The study used 40 open heathland sites\cite{Camac:2013eb} established immediately after the 2003 fires\cite{Williams:2006tl}.
-All sites have not been grazed by cattle since the 2003 fires.
+All sites have been free of grazing by cattle since 2003.
 Thirty sites were burnt, while the remaining ten were unburnt, likely for more than 70 years\cite{Camac:2013eb}.
 This disparity in sampling effort between burnt and unburnt was due to very few open heathland remaining unburnt after both the 2003 and 2007 fires.
-These sites encompassed most of the Bogong High Plains (120 km$^2$) and ranged in altitudinal from 1660 m to 1860 m a.s.l.
+These sites encompassed most of the Bogong High Plains (120 km$^2$) and ranged in altitude from 1660 to 1860 m a.s.l.
 Sites were at least 0.25 ha and more than 500 m apart.
@@
 Transects were positioned in the center of each site and were parallel to each other with exactly 10 m between transect lines.
-This effectively created a grid of 40 plots (10 plots\texttimes4 transects) per site, subsampling an area of 2000 m$^2$.
+This effectively created a grid of 40 plots (10 plots~\texttimes~4 transects) per site, subsampling an area of 2000 m$^2$.
 
@@
 For each site, altitude and TWI were derived from a 30 m resolution digital elevation model of the Bogong High Plains.
-TWI, which approximates the relative plant-available water based on a site’s position in the landscape\cite{Moore:1993fk}, was derived for each site using $TWI= ln(\frac{a\alpha}{tan\theta})$, where, $\alpha$ is the upslope catchment area per unit width orthogonal to the flow direction and $\theta$ is the slope (in degrees).
+TWI, which approximates the relative plant-available water based on a site\textquotesingle s position in the landscape\cite{Moore:1993fk}, was derived for each site using $\textrm{TWI}= \textrm{ln}(\frac{\alpha}{\textrm{tan}\theta})$, where, $\alpha$ is the upslope catchment area per unit width orthogonal to the flow direction and $\theta$ is the slope (in radians).
 Across the 40 sites, TWI varied twofold (5 to 10 units; higher units = wetter).
@@
 
-The occurrence and severity of fire was initially determined immediately after the 2003 fire\cite{Williams:2006tl}.
-The severity of the fires in each burnt site was estimated using twig diameters, a proxy measure of fire severity\cite{Whight:1999vw}.
+The occurrence and severity of fire was initially determined immediately after the 2003 fires\cite{Williams:2006tl}.
+The severity of the fires in each burnt site was estimated using mean twig diameter, a proxy measure of fire severity\cite{Whight:1999vw}.
 These twig diameters were recorded between 1 and 10 months post-fire from 10 replicate \textit{Grevillea australis} individuals.
-Across the 30 burnt sites, the average twig diameter ranged from 2 to 13 mm, with smaller diameters assumed to have experienced lower fire intensity.
+Across the 30 burnt sites, the average site twig diameter across ranged from 2 to 13 mm, with smaller diameters assumed to correspond to lower fire intensity.
 This encompassed the full range of fire severity from lightly scorched sites containing shrubs with only the tips of the outer branches burnt, to severely burnt sites where the majority of large shrub branches were consumed.
@@
 \section*{Data analysis}
-In total we built 11 hierarchical models to examine factors influencing tussock grass post-fire recovery and shrub seedling establishment, growth and mortality.
+In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality.
 These models are well suited to datasets that have a hierarchical structure (e.g.
 plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset\cite{Gelman:2007te, Kery:2012wl}.
-For each model we used Bayesian inference and fitted models with sampling in R \Sexpr{paste(version$major, version$minor, sep='.')} using package Rstan \Sexpr{packageVersion("rstan")}.
-For each model, three independent chains were executed.
+For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package rstan \Sexpr{packageVersion("rstan")}.
+For each model, three independent chains were run.
+Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic\cite{Gelman:1992ts, Brooks:1998ju}.
 In all cases modelled parameters converged within 1000 iterations.
-Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic\cite{Gelman:1992ts, Brooks:1998ju}.
-After discarding the first 1000 iterations as ‘warm up’, a further 1000 iterations were taken from the joint posterior.
+After discarding the first 1000 iterations as ‘warm up’, a further 1000 samples were taken from the joint posterior.
 Below we describe the model structure, parameters and priors used in these models.
-To ensure all covariate effects were identifiable we selectively removed each covariate and visually inspected whether the direction and magnitude of effects were severely altered; none were.
+To ensure all covariate effects were identifiable we removed each covariate and visually inspected whether the direction and magnitude of effects were severely altered; none were.
 Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.
@@
 \subsection{\textit{Modelling shrub seedling growth rates.}}
-For the purpose of modelling seedling annual growth rates we used heights measured in May of each year from seedlings that survived throughout the passive warming experiment.
-This meant that our observations were of maximum height for each individual $i$ at each May census $t$.
-We modelled these heights, $height_{i,t}$, as a random realization from a normal distribution truncated at zero:
+We modelled the observed height of individual $i$ at May census $t$ as a random realization from a normal distribution truncated at zero:
 
 \begin{equation}\label{eq:growth_obs} 
-height_{i,t} \sim \mathcal{N}(\widehat{height}_{i,t}, \sigma_{obs}), 0 \leq height_{i,t} \leq \infty,
+height_{i,t} \sim \mathcal{N}(\widehat{height}_{i,t}, \sigma_{obs}), 0 \leq height_{i,t} < \infty,
 \end{equation}
@@
 That is, a curve whereby initial seedling growth is slow before increasing in speed and finally levelling off when adult height is reached.
-We used this particular parameterisation because it accounts for seedling size at the beginning of an experiment and contained biologically interpretable parameters.
-Here, $Hmax_{j[i]}$ is the asymptotic height: the average mature height an individual $i$ belonging to species $j$ can achieves.
-We treated Hmax as a constant, not to be estimated by the model, owing to our experimental observations being restricted to the first five years of growth.
-Based on maximum height data from 5 to 22 long-unburnt sites (depending on the species) we estimated the average mature height (and thus Hmax) to be approximately \Sexpr{hmax_values(hmax,'Asterolasia')}, \Sexpr{hmax_values(hmax,'Grevillea')}, \Sexpr{hmax_values(hmax,'Phebalium')} and \Sexpr{hmax_values(hmax,'Prostanthera')} cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera}, respectively.
+We used this particular parameterisation because it accounts for seedling size at the beginning of an experiment and contains biologically interpretable parameters.
+Here, $Hmax_{j[i]}$ is the asymptotic height: the average mature height that species $j$, to which individual $i$ belongs, can achieve.
+We did not use the model to estimate $Hmax$ and instead used species specific constants.
+We did this because our data spanned only the first five years of growth and thus individuals did not reach their maximum heights.
+We estimated these constants from maximum height data from 5 to 22 long-unburnt sites (depending on the species) to estimated the average maximum height to be approximately \Sexpr{hmax_values(hmax,'Asterolasia')}, \Sexpr{hmax_values(hmax,'Grevillea')}, \Sexpr{hmax_values(hmax,'Phebalium')} and \Sexpr{hmax_values(hmax,'Prostanthera')} cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera}, respectively.
 $Hinit_i$ is the height of individual $i$ at the start of the experiment (i.e.
@@
 The experiment ran for a total of \Sexpr{n_days(chamber_dates, FALSE)} days (or approximately \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} years).
-We divide the number of days at each census by 365.25 in order to estimate annual growth rates.
-The parameter of interest, $R_i$, is the mean growth rate constant for each individual $i$.
+We divided $Days_t$ by 365.25 in order to estimate annual growth rates.
+The parameter of interest, $R_i$, is the mean annual growth rate for each individual $i$.
 This parameter controls the slope at inflection (i.e.
@@
 When multiplied by the asymptotic height, $Hmax$, it yields the maximum annual growth rate at the inflection point.
-We modelled $R_i$ separately for seedlings growing in bare ground (equation \ref{eq:r_bg}) compared to those growing in inter-tussock gaps (equation \ref{eq:r_tussock}).
+We modelled $R_i$ separately for seedlings growing in bare ground (Equation \ref{eq:r_bg}) and those growing in inter-tussock gaps (Equation \ref{eq:r_tussock}).
 
-For individual seedlings of the four shrub species growing in the bare ground patch located in the centre of each plot we modelled $R_i$ using a varying intercept linear model:
+For individual seedlings of the four shrub species growing in the bare ground patch located in the centre of each plot, we modelled $R_i$ using a linear model with a varying intercept:
 
@@
 \epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
-\epsilon_i &\sim \mathcal{N}(0, \sigma_{obs_{ind}}).
+\epsilon_i &\sim \mathcal{N}(0, \sigma_{obs}).
 \end{aligned}
@@
 
-Here, the intercept term, $alpha_{j[i]}$, represents the expected control plot growth constant of species $j$ to which individual $i$ belongs.
-It is estimated from a normal distribution centred on $\mu_{spp}^{CTL}$ (the expected control plot growth rate for an average hypothetical species) with a standard deviation,$\sigma_{spp}^{CTL}$ which defines the variation between species’ growth rates at control plots.
-The parameter $\beta_{j[i]}$ is the OTC effect for each species $j$ to which individual $i$ belongs.
-$\beta_j$ is also estimated from a normal distribution, but unlike $\alpha_j$, it is centred on $\mu_{spp}^{OTC}$, (the expected OTC growth constant for an average hypothetical species) with a standard deviation, $\sigma_{spp}^{OTC}$ which defines the variation between species’ growth constants within OTCs.
+Here, the intercept term, $\alpha_{j[i]}$, represents the expected control plot growth rate of species $j$ to which individual $i$ belongs.
+It is estimated from a normal distribution centered on $\mu_{spp}^{CTL}$ (the expected control plot growth rate for an average hypothetical species) with a standard deviation, $\sigma_{spp}^{CTL}$ which defines the variation between species\textquotesingle growth rates at control plots.
+The parameter $\beta_{j[i]}$ is the OTC effect for species $j$ to which individual $i$ belongs.
+$\beta_j$ is also estimated from a normal distribution, but unlike $\alpha_j$, it is centered on $\mu_{spp}^{OTC}$, (the expected OTC growth rate for an average hypothetical species) with a standard deviation, $\sigma_{spp}^{OTC}$ which defines the variation between species\textquotesingle growth rate within OTCs.
 The terms $\epsilon_{plot_{k[i]}}$ and $\epsilon_i$ represent residual variation at the plot level (for each plot, $k$, within which individual $i$ is located), and the individual level (for each individual $i$), respectively.
-Both error terms were drawn from normal distributions centred on zero with standard deviations that defined the variation amongst plots, $\sigma_{plot}$, or amongst individuals, $\sigma_{obs_{ind}}$, respectively.
+These error terms were drawn from normal distributions centered on zero with standard deviations that defined the variation amongst plots, $\sigma_{plot}$, or amongst individuals, $\sigma_{obs}$, respectively.
 
-In contrast, for \textit{Grevillea australis} seedlings growing in inter-tussock gaps, the growth constant was modelled using a fixed intercept linear model:
+In contrast, for \textit{Grevillea australis} seedlings growing in inter-tussock gaps, growth rate was modelled using a fixed intercept linear model (i.e.
+no species random effect as there was only a single shrub species in inter-tussock gaps):
 
@@
 
-The structure and parameter definitions differ slightly to that in equation \ref{eq:r_bg}.
+The structure and parameter definitions differ slightly to that in Equation \ref{eq:r_bg}.
 Here, the model has two additional parameters, $\beta_2$ and $\beta_3$.
 The parameter $\beta_2$ is the effect of inter-tussock gap radius, which varied over time $t$, while $\beta_3$ is the interactive effect of OTC and time-varying inter-tussock gap radius.
-The inclusion of these parameters, coupled with $Gap_{i,t}$ being centered and standardized (see Data inputs into growth and mortality models), alters the definition of the intercept, $\alpha$, which now represents the mean control plot growth rate at the average inter-tussock gap size (\Sexpr{mean_init_gap(tussock_otc_growth_data)} cm).
+The inclusion of these parameters, coupled with $Gap_{i,t}$ being centered and standardized (see Data inputs for growth, mortality and recruitment models), alters the definition of the intercept, $\alpha$, which now represents the mean control plot growth rate at the average inter-tussock gap size (\Sexpr{mean_init_gap(tussock_otc_growth_data)} cm).
 The parameter $\beta_1$ is the OTC effect and $\epsilon_{plot_k}$ and $\epsilon_i$ are residual errors at the plot and individual level, respectively.
-Note that growth constant, $R$, now varies through time as a function of inter-tussock gap radius.
+Note that the growth rate, $R$, now varies through time as a function of inter-tussock gap radius.
 
@@
 In both growth models standard deviation parameters (i.e.
-$\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{ind}$) were all estimated from weakly informative positive half Cauchy priors with scale parameters of 25.
+$\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{obs}$) were all estimated from weakly informative positive half Cauchy priors with scale parameters of 25.
 We used these Cauchy priors in order to prevent over-inflated estimates of variances due to the small number of plots and species\cite{Gelman:2006uw}.
@@
 inter-tusscok model: $\beta_1$, $\beta_2$ and $\beta_3$) and hyperparameter (i.e.
-multispecies model) for the mean OTC effect $\mu_{spp}^{OTC}$ were estimated from normal priors centered on zero with a scale parameter of 0.5.
+in the multispecies model) for the mean OTC effect, $\mu_{spp}^{OTC}$, were estimated from normal priors centered on zero with a standard deviation of 0.5.
 We chose this prior as it spanned the range of likely growth trajectories from very slow to very fast growing.
@@
 \subsection{\textit{Modelling shrub seedling mortality rates.}}
-We were also interested in whether seedling instantaneous mortality rates varied amongst shrub species, OTC treatment and inter-tussock gap size.
+We were also interested in whether the instantaneous mortality rates of seedlings varied amongst shrub species, OTC treatment and inter-tussock gap size.
 Our observations, $y_{i,t}$, were binary (0 = alive or 1 = dead) for each individual $i$, in each census, $t$.
@@
 
-As with growth rates, we modelled each individual's cumulative hazard per census, \\
-$Cumulative\;hazard_{i,t}$, separately for seedlings growing in bare ground patches to those in inter-tussock gaps.
-In both cases, $Cumulative\;hazard_{i,t}$, was modelled on the complementary log-log scale as a function of an instantaneous hazard rate for each individual, $\lambda_i$, multiplied by the length (in years) between each census, $t$.
+As with growth rates, we modelled each individual's cumulative hazard, \\
+$Cumulative\;hazard_{i,t}$, separately for seedlings growing in bare ground patches and those in inter-tussock gaps.
+In both cases, $Cumulative\;hazard_{i,t}$, was modelled on the complementary log-log scale as a function of an instantaneous hazard rate for each individual, $\lambda_i$, multiplied by the time lapsed (in years) between censuses, $t$.
 By doing this our regression model becomes equivalent to a parametric survival analysis\cite{Kleinbaum:2012tn}.
@@
 \begin{equation}\label{eq:cum_hazard} 
-\log_e(-\log_e(1 - Cumulative\;hazard_{i,t}) = \lambda_i \times census\;length_t.
+\textrm{ln}(-\textrm{ln}(1 - Cumulative\;hazard_{i,t})) = \lambda_i \times census\;interval_t.
 \end{equation}
 
-We modelled individual instantaneous hazard rate, $\lambda_i$, in the same way we modelled $R_i$ in equations \ref{eq:r_bg} \& \ref{eq:r_tussock}. 
+We modelled the individual instantaneous hazard rate, $\lambda_i$, in the same way that we modelled $R_i$ in Equations \ref{eq:r_bg} \& \ref{eq:r_tussock}. 
 For individual seedlings of the four shrub species located in the bare ground patch of each plot, we modelled $\lambda_i$ using a varying intercept model:
@@
 
-Here, the intercept term, $\alpha_{j[i]}$, represents the expected instantaneous hazard rate in a control plot for species $j$ within which individual $i$ belongs.
-It is estimated from a normal distribution centred on $\mu_{spp}^{CTL}$  (the control plot expected instantaneous hazard rate for the average hypothetical species) with a standard deviation, $\sigma_{spp}^{CTL}$, which defines the variation between species’ instantaneous hazard rates in control plots.
+Here, the intercept term, $\alpha_{j[i]}$, represents the expected instantaneous hazard rate in a control plot for species $j$ to which individual $i$ belongs.
+It is estimated from a normal distribution centered on $\mu_{spp}^{CTL}$  (the control plot expected instantaneous hazard rate for the average hypothetical species) with a standard deviation, $\sigma_{spp}^{CTL}$, which defines the variation between species\textquotesingle instantaneous hazard rates in control plots.
 The parameter $\beta_{j[i]}$ is the OTC effect for each species $j$ to which individual $i$ belongs.
 $\beta_{j[i]}$ is also estimated from a normal distribution, but unlike $\alpha_j$, it is centered on $\mu_{spp}^{OTC}$, (the OTC expected probability of death for the average hypothetical species) with a standard deviation, $\sigma_{spp}^{OTC}$, which defines the variation between species’ instantaneous hazard rates in OTCs.
-The parameters $\epsilon_{plot_k}$ and $\epsilon_i$ are residual (unexplained) variation at the plot and individual level, respectively.
+The parameters $\epsilon_{plot_k}$ and $\epsilon_i$ are residuals at the plot and individual level, respectively.
 
@@
 
-Here, the structure and parameter definitions are similar to equation \ref{eq:r_tussock}.
-The intercept, $\alpha$, represents the expected instantaneous hazard rate in control plots in an average inter-tussock gap radius, $\beta_1$ is the effect of OTC, $\beta_2$ is the effect of inter-tussock gap radius, $\beta_3$ is the interactive effect of OTC and inter-tussock gap radius and $\epsilon_{plot_k}$ and $\epsilon_i$ are the residual variation at the plot and individual scales.
+Here, the structure and parameter definitions are similar to Equation \ref{eq:r_tussock}.
+The intercept, $\alpha$, represents the expected instantaneous hazard rate in control plots in an inter-tussock gap with average radius, $\beta_1$ is the effect of OTC, $\beta_2$ is the effect of inter-tussock gap radius, $\beta_3$ is the interactive effect of OTC and inter-tussock gap radius and $\epsilon_{plot_k}$ and $\epsilon_i$ are the residuals at the plot and individual scales.
 
 Here the standard deviation parameters in both mortality models (i.e.
-$\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{ind}$) were each estimated from positive half Cauchy priors with scale parameters of 2.5.
-The intercept term $\alpha$ for inter-tussock seedling model and the mean hyper-parameter $\mu_{spp}^{CTL}$ in the multispecies models were estimated from cauchy priors with a scale parameter of 2.5.
+$\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{obs}$) were each estimated from positive half Cauchy priors with standard deviation of 2.5.
+The intercept term $\alpha$ for inter-tussock seedling model and the mean hyper-parameter $\mu_{spp}^{CTL}$ in the multispecies models were estimated from cauchy priors with standard deviations of 2.5.
 We also used these Cauchy priors for inter-tussock model coefficients (i.e.
 $\beta_1$, $\beta_2$ and $\beta_3$) and the multi-species hyperparameter for the mean OTC effect $\mu_{spp}^{OTC}$.
-These Cauchy priors are the recommended default choice for logistic regression models\cite{Gelman:2008ua}.
+These cauchy priors are the recommended default choice for binomial regression models\cite{Gelman:2008ua}.
 
@@
 
-where $\mu_{k,t}$ is the log expected recruit count for each plot, and $\phi$ is the overdispersion parameter.
+where $\mu_{k,t}$ is the log expected recruit count for each plot $k$ at census $t$, and $\phi$ is the overdispersion parameter.
 We used a negative binomial distribution because count data was significantly overdispersed with $\phi$ being estimated as \Sexpr{phi_value(recruit_model,'mean')} (95\% BCI: \Sexpr{phi_value(recruit_model,'l_bci')}\textendash \Sexpr{phi_value(recruit_model,'u_bci')}).
@@
 \begin{aligned}
-\log_e(\mu_{k,t}) &=  \alpha + \beta_1 \times OTC_k + \beta_2 \times Census_t + \epsilon_{plot_k} \\
+\textrm{ln}(\mu_{k,t}) &=  \alpha + \beta_1 \times OTC_k + \beta_2 \times Census_t + \epsilon_{plot_k} \\
 \epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}),
@@
 
-where, the intercept, $\alpha$, represents the average log recruitment in control plots in 2011, assuming all other parameters are at their mean.
-The parameter $\beta_1$ is the effects of OTC and $\beta_2$ describes the change in recruit numbers from 2011 to 2012.
-$\epsilon_{plot_k}$ is the residual variation at the plot scale.
+where the intercept, $\alpha$, represents the average log recruitment in control plots in 2011.
+The parameter $\beta_1$ is the effects of OTC and $\beta_2$ describes the change in the log number of recruits from 2011 to 2012.
+$\epsilon_{plot_k}$ is the residual error at plot $k$.
 
@@
 \subsection{\textit{Data inputs for growth, mortality and recruitment models.}}
-In each model, OTC treatment, $OTC_k$, was a binary (0 = Control, 1 = OTC) plot level covariate.
+In each model, OTC treatment, $OTC$, was a binary (0 = Control, 1 = OTC) plot level covariate.
 By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census $t$.
-However, because some cardinal distances were recorded as censored data (i.e.
+However, because some distances were censored (i.e.
 \textgreater 25 cm), these data were incomplete.
 We estimated these censored data using a censored model (see below).
-Once censored data were estimated, we then estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each censuses, for each inter-tussock seedling.
+Once true distances were estimated for censored data, we estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each census, for each inter-tussock seedling.
 Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots.
 In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations\cite{Gelman:2007te}.
-The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as average responses and slope terms as partial dependencies conditional on other continuous variables at their mean.
-Furthermore, standardizing by two standard deviations allows the magnitude of effects to be directly compared between binary and continuous covariates\cite{Gelman:2007te}.
+The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as responses to average conditions and slope terms as partial dependencies conditional on other continuous variables being at their mean.
 
@@
 Approximately \Sexpr{prop_censored(cleaned_poa_distances)}\% of the $Poa$ cardinal distance data used to calculate the average inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, (i.e.
-$\geq$ 26 cm).
-We estimated these censored values by first estimating the distribution of distances in observed data, coupled with estimating covariate effects and then using these distributions to estimate censored observations.
+\textgreater  25 cm).
+We estimated these censored values by first estimating the distribution of distances in observed data, coupled with estimating covariate effects, and then using these distributions to estimate censored observations.
 
@@
 
-where $\mu_i$ is the expected log estimated distance for cardinal distance $i$, and  $\sigma_{obs}$ is the standard deviation of the observation error.
+where $\mu_i$ is the expected log distance for distance $i$, and $\sigma_{obs}$ is the standard deviation of the observation error.
 
@@
 Again, like other models, half Cauchy priors with a scale of 25 were used for all standard deviation parameters.
-Both the intercept, $\alpha$, and the effect of OTC, $\beta$ (which are on the log scale), were estimated from weakly informed normal distributions centered on zero with a standard deviation of 2.5.
-Specifically, this prior encapsulated our knowledge that distances were very unlikely to be \textgreater 150 cm.
+Both the intercept, $\alpha$, and the effect of OTC, $\beta$ (which are on the log scale), were estimated from weakly informative normal distributions centered on zero with a standard deviation of 2.5.
+Specifically, this prior encapsulated our knowledge that distances were very unlikely to be \textgreater 100 cm.
 
@@
 
-The mean posteriors from the censored data were then extracted from the model and used as our estimates for each censored observation.
-This complete dataset was then used subsequent analyses (e.g.
-inter-tussock seedling growth and mortality, changes in gap size \textemdash see below).
+The posterior means of censored data were then extracted from the model and used as our estimates for each censored observation.
+This complete dataset was then used in subsequent analyses (e.g.
+inter-tussock seedling growth and mortality, changes in gap size\textemdash see below).
 
 \subsection{\textit{Modelling rates of inter-tussock gap infilling by Poa.}}
-As well as modelling how OTC affected shrub seedling growth and mortality rates, we were also interested in examining how warming influenced rates of grass infilling of inter-tussock gaps.
-To examine this we first modelled the median gap radius, $y_{i,t}$, for each inter-tussock gap $i$ in each census $t$, as a random realisation from a lognormal distribution:
+As well as modelling how OTCs affected shrub seedling growth and mortality rates, we were also interested in examining how warming influenced rates of grass infilling of inter-tussock gaps.
+To examine this we first modelled the median observed gap radius, $y_{i,t}$, for inter-tussock gap $i$ in census $t$, as a random realisation from a lognormal distribution:
 
 \begin{equation}\label{eq:gap_obs} 
-y_{i,t} \sim LN(\mu_{i,t}, \sigma_{obs}),
+y_{i,t} \sim (\textrm{ln}(\mu_{i,t}), \sigma_{obs}),
 \end{equation}
 
-where $\mu_{i,t}$ is the log estimated median radius for each inter-tussock gap, at each census, and  $\sigma_{obs}$ is the standard deviation of the observation error.
+where $\textrm{ln}(\mu_{i,t})$ is the log estimated median radius for inter-tussock gap, at census $t$, and $\sigma_{obs}$ is the standard deviation of the observation error.
 
@@
 \begin{equation}\label{eq:log_gap_mod} 
-\mu_{i,t} = \log_e(\alpha_i) + \beta_i \times \frac{Days_t}{365.25},
+\textrm{ln}(\mu_{i,t}) = \textrm{ln}(\alpha_i) + \beta_i \times \frac{Days_t}{365.25},
 \end{equation}
@@
 We did not force this curve to be a negative exponential (i.e.
-gaps decreased over time) because it is theoretically possible for inter-tussock gaps to increase over time.
-Here $\alpha_i$ is a known constant that represents the initial inter-tussock gap radius for each gap $i$ (i.e.
+gaps decreasing over time) because it is theoretically possible for inter-tussock gaps to increase over time.
+Here $\alpha_i$ is the initial inter-tussock gap radius for gap $i$ (i.e.
 the intercept).
-$\beta_i$ represents the exponential rate for each gap and $Days_t$ is the number of days from the beginning of the experiment.
+$\beta_i$ represents the exponential rate of gap size change for gap $i$ and $Days_t$ is the number of days since the beginning of the experiment.
 
@@
 \begin{aligned}
-\beta_i &= \alpha_0 + \beta \times OTC_{k[i]} + \epsilon_{plot_{k[i]}} + \epsilon_i \\
+\beta_i &= \alpha + \beta \times OTC_{k[i]} + \epsilon_{plot_{k[i]}} + \epsilon_i \\
 \epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
@@
 
-Here, $\alpha_0$ is the average rate of gap radius change in control plots, $\beta$ is the effect of OTC and $\epsilon_{plot_k}$ and $\epsilon_i$ represent the plot and gap level residual errors. 
-Like all other models used in this study, we used half Cauchy priors with a scale of 25 for all standard deviation parameters.
-The intercept, $\alpha_0$, and the effect of OTC, $\beta$ were estimated from normal priors centered on zero with a standard deviation of 5.
-These priors encapsulated the full range of potential gap trajectories from increases or decreases or no change.
+Here, $\alpha$ is the average rate of gap radius change in control plots, $\beta$ is the effect of OTC and $\epsilon_{plot_k}$ and $\epsilon_i$ represent the plot and gap level residual errors. 
+Like most other models used in this study (the exception being mortality models), we used half Cauchy priors with a scale of 25 for all standard deviation parameters.
+The intercept, $\alpha$, and the effect of OTC, $\beta$ were estimated from normal priors centered on zero with a standard deviation of 5.
 
 \subsection{\textit{Modelling drivers of seedling density.}}
-We examined drivers of seedling density by modelling the number of seedlings per 1 m$^2$ as a random realization from a negative binomial distribution:
+We examined drivers of seedling density by modelling the number of seedlings square metre plot $i$ as a random realization from a Poisson distribution:
 
 \begin{equation}\label{eq:density_obs} 
-count_i \sim NegBinomial(\mu_i, \phi),
+count_i \sim Poisson(\lambda_i),
 \end{equation}
 
-where $\mu_i$ is the log expected seedling count for each plot, and $\phi$ is the overdispersion parameter.
-We used a negative binomial distribution because count data was significantly overdispersed for both species with $\phi$ being estimated as \Sexpr{phi_value(greaus_seedling_density_model$fit,'mean')} (95\% BCI: \Sexpr{phi_value(greaus_seedling_density_model$fit,'l_bci')}\textendash \Sexpr{phi_value(greaus_seedling_density_model$fit,'u_bci')}) and \Sexpr{phi_value(asttry_seedling_density_model$fit,'mean')} (95\% BCI: \Sexpr{phi_value(asttry_seedling_density_model$fit,'l_bci')}\textendash \Sexpr{phi_value(asttry_seedling_density_model$fit,'u_bci')}), for \textit{Grevillea} and \textit{Asterolasia}, respectively.
+where $\lambda_i$ is the log expected seedling count for each plot. We used a Poisson distribution because count data were not overdispersed for either species.
 
-$\mu_i$ was modelled separately for \textit{Grevillea australis} and \textit{Asterolasia trymalioides} because the latter contained no measurements of adult density.
+$\lambda_i$ was modelled separately for \textit{Grevillea australis} and \textit{Asterolasia trymalioides} because the latter contained no measurements of adult density.
 The two models are structurally similar and only differ in that parameter $\beta_5$ is not included for \textit{Asterolasia}.
-Below we outline the full model structure used to estimate $\mu_i$.
+Below we outline the full model structure used to estimate $\lambda_i$.
 
-We modeled $\mu_i$ on the log scale as a function of both site and transect covariates:
+We modeled $\lambda_i$ on the log scale as a function of both site and transect covariates:
 
 \begin{equation}\label{eq:density_mu_mod} 
-\mu_i = \delta_{j[i]} + \gamma_{k[i]}
+\lambda_i = \delta_{j[i]} + \gamma_{k[i]}
 \end{equation}
 
-where, $\delta_{j[i]}$ and $\gamma_{k[i]}$ represents the average log count for each site $j$ and transect $k$ within which plot $i$ is located, respectively.
+where, $\delta_{j[i]}$ and $\gamma_{k[i]}$ is the average log count for site $j$ and transect $k$ within which plot $i$ is located, respectively.
 
@@
 
-where the intercept, $\alpha_j$, varies by site and represents the average log count per plot at each site, assuming all other parameters are at their mean.
-We estimated $\alpha_j$ from a normal distribution centred on $\mu_{site}$, the average seedling count per 1 m$^2$ plot for the average site.
-The parameters $\beta_1$ and $\beta_2$ are the effects of altitude and wetness, respectively, while $\beta_3$ is the effect of being an unburnt site and $\beta_4$ is the effect of fire severity in burnt sites.
+where the intercept, $\alpha_j$, varies by site and is the average log count per plot at site $j$, assuming all other covariates are at their mean.
+We estimated $\alpha_j$ from a normal distribution centered on $\mu_{site}$, the average log seedling count per 1 m$^2$ plot.
+The parameters $\beta_1$ and $\beta_2$ are the effects of altitude ($Alt$) and Topographic Wetness Index ($TWI$), respectively, while $\beta_3$ is the effect of being an unburnt site and $\beta_4$ is the effect of fire severity in burnt sites.
 
@@
 
-where $\beta_5$ is the effect of mature \textit{Grevillea} counts and $\epsilon_{tran_k}$ is the random error associated with transect $k$.
-Parameter $\beta_5$ was not included when modelling \textit{Asterolasia} seedling counts because of a lack of mature individual counts.
+where $\beta_5$ is the effect of mature \textit{Grevillea} density ($ADen$) and $\epsilon_{tran_k}$ is the random error associated with transect $k$.
+Parameter $\beta_5$ was not included when modelling \textit{Asterolasia} seedling counts because of a lack of persistent adult skeletons.
 For this species $\gamma_k = \epsilon_{tran_k}$.
@@
 
-Half Cauchy priors with a scale of 25 were used for all standard deviation parameters and the overdispersion parameter $\phi$.
-Hyperparameter $\mu_{site}$ and coefficients $\beta_1$ to $\beta_5$ were estimated using normal priors centered on zero with a standard deviation of 2.5, encapsulating our prior knowledge that plots were unlikely to have more than 150 seedlings/m$^2$.
+Half Cauchy priors with a scale of 25 were used for all standard deviation parameters.
+Hyperparameter $\mu_{site}$ and coefficients $\beta_1$ to $\beta_5$ were estimated using normal priors centered on zero with a standard deviation of 2.5, encapsulating our prior belief that plots were unlikely to have more than 150 seedlings/m$^2$.
 
 \subsection{\textit{Modelling drivers of seedling maximum heights.}}
-In order to validate growth rate findings in the OTC experiment, we assessed how an elevation range of 190 m in elevation (equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C\cite{Brown:1989vx}) influenced \textit{Grevillea} and \textit{Asterolasia} seedlings heights in areas burnt by the 2003 fires.
+In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation (equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C\cite{Brown:1989vx}) influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires.
 We also assessed whether TWI and fire severity influenced these heights.
@@
 We used maximum heights because these were likely to represent individuals that established immediately post-fire, and would have had 9 years of post-fire growth.
-We modelled these heights, $maxht_i$, as a random realization from a log-normal distribution:
+We modelled these heights, $maxht_i$, as a random realizations from a log-normal distribution:
 
@@
 
-where $\widehat{maxht}_i$ is the model estimated log maximum height for each individual $i$ and $\sigma_{obs}$ is the standard deviation of the observation error.
-We modelled $\widehat{maxht}_i$ separately for \textit{Grevillea} and \textit{Asterolasia} using a varying intercept linear model:
+where $\widehat{maxht}_i$ is the model-estimated log maximum height for individual $i$ and $\sigma_{obs}$ is the standard deviation of the observation error.
+We modelled $\widehat{maxht}_i$ separately for \textit{Grevillea} and \textit{Asterolasia} using a linear model with a varying intercept:
 
@@
 
-where the intercept, $\alpha_j$, varies by site and represents the average log maximum seedling height of each site, assuming all other parameters are at their mean.
-We estimated $\alpha_j$ from a normal distribution centred on $\mu_{site}$, the average log maximum height for the average site and a standard deviation of $\sigma_{site}$ which defines site variability.
-The parameters $\beta_1$ and $\beta_2$ are the effects of altitude and topographic wetness, respectively and $\beta_3$ is the effect of fire severity.
+where the intercept, $\alpha_j$, varies by site and is the average log maximum seedling height of site $j$, assuming all other covariates are at their mean.
+We estimated $\alpha_j$ from a normal distribution centered on $\mu_{site}$, the average log maximum height, and a standard deviation of $\sigma_{site}$ which defines site variability.
+The parameters $\beta_1$ and $\beta_2$ are the effects of altitude and Topographic Wetness Index, respectively and $\beta_3$ is the effect of fire severity.
 
-Here, we estimated $\mu_{site}$ and coefficient parameter ($\beta_1$ to $\beta_3$) using weakly informative normal priors with a mean of 0 and a standard deviation of 2.5.
+Here, we estimated $\mu_{site}$ and coefficients ($\beta_1$ to $\beta_3$) using weakly informative normal priors with a mean of 0 and a standard deviation of 2.5.
 These priors encapsulated our prior knowledge that seedling maximum heights would be less than 150 cm.
@@
 Altitude ($Alt_j$), Topographic Wetness Index ($TWI_j$) and mature \textit{Grevillea} count ($ADen_k$) were also continuous covariates that affected all sites in both count and maximum height models.
-Collinearity among covariates was checked and all correlations were r \textless 0.16.
-The data from each continuous covariate was then centered on zero and standardized by two standard deviations\cite{Gelman:2007te}.
+Collinearity among covariates was checked and all correlations were \textit{r} < 0.16.
+The data from each continuous covariate was then centered on their mean and standardized by two standard deviations\cite{Gelman:2007te}.
 \pagebreak
@@
 \includegraphics{figures/figS1.pdf}
-\caption{\textbf{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (minimum twig diameter), altitude and Topographic Wetness Index (TWI).} Centered and standardized model coefficients (on the log scale) (\textbf{a}) and effects of burning (\textbf{b}), fire severity  (\textbf{c}), altitude (\textbf{d}) \& topographic wetness (\textbf{e}), in areas burnt by the 2003 fires.
-All errors bars are 95\% Bayesian Credible Intervals.
-Phi = overdispersion parameter for negative binomial distribution.}
+\caption{\textbf{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (minimum twig diameter), altitude and Topographic Wetness Index (TWI).} (\textbf{a}) Centered and standardized model coefficients (on the log scale); and effects of (\textbf{b}) burning, (\textbf{c}) fire severity, (\textbf{d}) altitude and (\textbf{e}) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires.
+All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
 \label{figs1}
@@
 \includegraphics{figures/figS2.pdf}
-\caption{\textbf{Number of \textit{Grevillea australis} recruits in warmed (red) and control (blue) conditions in both 2011 and 2012. Errors bars are 95\% Bayesian Credible Intervals.}}
+\caption{\textbf{Average number of \textit{Grevillea australis} recruits in warmed (red) and control (blue) conditions in 2011 and 2012. Errors bars are 95\% Bayesian Credible Intervals.}}
 \label{figs2}
@@
 \includegraphics{figures/figS3.pdf}
-\caption{\textbf{Observed and projected diameter growth trajectories of four dominant Australian alpine shrubs.} Thick lines represent mean ($\pm$ 95\% CI) observed heights at each May census.
-Thin lines represent individual growth trajectories.
-Red \& orange lines = seedlings growing in warmed (OTC) conditions and blue \& light blue lines = seedlings growing in control conditions.}
+\caption{\textbf{Observed seedling diameter growth trajectories of four dominant Australian alpine shrubs.} Red and Blue lines represent mean ($\pm$ 95\% CI) observed heights at each May census in warmed and control plots, respectively.
+orange and light blue lines represent individual growth trajectories in warmed and unwarmed plots, respectively.}
 \label{figs3}
@@
 \caption{\textbf{Mean maximum seedling height along fire severity (minimum twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires.} Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses.
-Centered and standardized model coefficients (on the log scale) (\textbf{a} \& \textbf{b}) and effects of fire severity  (\textbf{c} \& \textbf{d}), altitude (\textbf{e} \& \textbf{f}) \& topographic wetness (\textbf{g} \& \textbf{h}), in areas burnt by the 2003 fires.
-All errors are 95\% Bayesian Credible Intervals.}
+(\textbf{a} \& \textbf{b}) Centered and standardized model coefficients (on the log scale); and effects of (\textbf{c} \& \textbf{d}) fire severity, (\textbf{e} \& \textbf{f}) altitude and (\textbf{g} \& \textbf{h}) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires.
+All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
 \label{figs4}
@@
 \includegraphics{figures/figS5.pdf}
-\caption{\textbf{Mean daily microclimate difference (OTC-Control)}.
-Ambient temperature (\textbf{a}), soil temperature (\textbf{b}), volumetric water content (\textbf{c}) and relative humidity over experimental period (\textbf{d}).
-Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls.
+\caption{\textbf{Mean daily microclimate difference (OTC $-$ Control)}.
+(\textbf{a}) Ambient temperature, (\textbf{b}) soil temperature, (\textbf{c}) volumetric water content and (\textbf{d}) relative humidity over experimental period.
+Red lines = mean difference across all growing seasons (when chambers were applied); blue = no difference between OTCs and controls.
 Shaded areas are the winter snow periods when OTCs were not over plots.
-Note: amplitude declines in later years are due to shading effects caused by increased shrub growth.}
+Note: declines in amplitude in later years are due to shading effects caused by increased shrub size.}
 \label{figs5}
@@
 \includegraphics{figures/figS6.pdf}
-\caption{\textbf{Mean minimum daily microclimate difference (OTC-Control)}.
-Ambient temperature (\textbf{a}), soil temperature (\textbf{b}), volumetric water content (\textbf{c}) and relative humidity over experimental period (\textbf{d}).
-Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls.
+\caption{\textbf{Mean minimum daily microclimate difference (OTC $-$ Control)}.
+(\textbf{a}) Ambient temperature, (\textbf{b}) soil temperature, (\textbf{c}) volumetric water content and (\textbf{d}) relative humidity over experimental period.
+Red lines = mean difference across all growing seasons (when chambers were applied); blue = no difference between OTCs and controls.
 Shaded areas are the winter snow periods when OTCs were not over plots.
-Note: amplitude declines in later years are due to shading effects caused by increased shrub growth.}
+Note: declines in amplitude in later years are due to shading effects caused by increased shrub size.}
 \label{figs6}
@@
 \includegraphics{figures/figS7.pdf}
-\caption{\textbf{Mean maximum daily microclimate difference (OTC-Control)}.
+\caption{\textbf{Mean maximum daily microclimate difference (OTC $-$ Control)}.
 Ambient temperature (\textbf{a}), soil temperature (\textbf{b}), volumetric water content (\textbf{c}) and relative humidity over experimental period (\textbf{d}).
@@
 \includegraphics{figures/figS8.pdf}
-\caption{\textbf{Diagram outlining positions of transplanted seedling in an OTC plot.} The inner white square = the burnt bare ground patch used to assess heights of four dominant alpine shrub species.
+\caption{\textbf{Diagram outlining positions of transplanted seedling in an OTC plot.} The inner white square represents the burnt bare ground patch used to assess heights of four dominant alpine shrub species.
 The grey area contains \textit{Grevillea australis} seedlings randomly transplanted into burnt inter-tussock gaps.
@@
 
-\section*{References}
 \bibliography{refs.bib}

### 2016-03-14 R/figures.R
Simplified plots by removing 80% BCIs

@@ -33,7 +33,7 @@ plot_obs_growth <- function(observed_data, response ='ht', scales ='free', ncol=
   observed_data$otc <- factor(ifelse(observed_data$otc==1, 'otc','ctl'))
   observed_data$year <- as.numeric(year(observed_data$date))
   ggplot() + 
-  geom_path(data = observed_data, aes_string(x = 'year',y = response, colour='otc', group = 'ind'), alpha = 0.4) +
+  geom_path(data = observed_data, aes_string(x = 'year',y = response, colour='otc', group = 'ind'), alpha = 0.4, size=0.3) +
   geom_path(data = summary_obs, aes(x=year, y=mean, col=factor(paste0(otc, '_mean'))), alpha=0.6) +
   geom_pointrange(data= summary_obs,aes(x=year, y=mean,ymin = lower_95ci, ymax=upper_95ci, col=factor(paste0(otc, '_mean'))), size = 0.05) +
   scale_colour_manual('', values = c("ctl" ="sky blue","otc" ="orange", "otc_mean"="red", "ctl_mean"="blue")) +
@@ -50,8 +50,7 @@ coefficient_plot <- function(summarised_coefficients, y_axis_labels='',xlab ='Ef
   
   
   ggplot(summarised_coefficients, aes(x = mean,y = parameter)) + 
-    geom_segment(aes(x=`2.5%`,y=parameter, xend=`97.5%`, yend=parameter), size=0.3)+
-    geom_segment(aes(x=`10%`,y=parameter, xend=`90%`, yend=parameter), size = 0.6)+
+    geom_segment(aes(x=`2.5%`,y=parameter, xend=`97.5%`, yend=parameter))+
     geom_point(aes(x=mean, y=parameter)) +
     geom_vline(aes(xintercept=0), linetype=2) +
     scale_y_discrete(labels =rev(y_axis_labels)) +
@@ -138,7 +137,7 @@ partial_plot_density_height <- function(predictions, x, xlab=NULL,ylab=NULL, yli
 
 burnt_v_unburnt <- function(predictions, x, ylab=NULL) {
   ggplot(predictions, aes_string(x = x,y = 'mean')) + 
-    geom_pointrange(aes(ymin = `2.5%`, ymax=`97.5%`), size=0.3) +
+    geom_pointrange(aes(ymin = `2.5%`, ymax=`97.5%`)) +
     scale_x_discrete(labels =c('Unburnt','Burnt')) +
     xlab('') +
     ylab(ylab) +

### 2016-03-14 manuscript/manuscript.Rnw
More edits

@@
 \item Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia.
-\item The Centre of Excellence for Environmental Decisions, School of BioSciences, The University of Melbourne, Parkville 3010, Victoria, Australia.
-\item Research Centre for Applied Alpine Ecology, La Trobe University, Melbourne 3086, Victoria, Australia.
-\item CSIRO Land and Water, Tropical Ecosystems Research Centre, PMB 44 Winnellie 0821, NT, Australia.
-\item Bio21 Institute, School of BioSciences, The University of Melbourne, Vic. 3010, Australia.
+\item The Centre of Excellence for Environmental Decisions, School of BioSciences, The University of Melbourne, Melbourne, Victoria 3010, Australia.
+\item Research Centre for Applied Alpine Ecology, La Trobe University, Melbourne, Victoria 3086, Australia.
+\item CSIRO Land and Water, Tropical Ecosystems Research Centre, PMB 44 Winnellie, Northern Territory 0821, Australia.
+\item Bio21 Institute, School of BioSciences, The University of Melbourne, Victoria 3010, Australia.
 \textasteriskcentered Email for correspondence: \texttt{james.camac@gmail.com}
@@
 \subsection{Drivers of shrub seedling establishment.}
-We first investigated how altitude, Topographic Wetness Index (TWI), adult density, fire and fire severity (as measured by minimum twig diameters\textemdash see Supplementary Methods) influenced the density of \textit{Grevillea} and \textit{Asterolasia} seedlings (the two dominant shrubs of alpine open heathland).
+We first investigated how altitude, Topographic Wetness Index (TWI), adult density, fire and fire severity (as measured by twig diameters\textemdash see Supplementary Methods) influenced the density of \textit{Grevillea} and \textit{Asterolasia} seedlings (the two dominant shrubs of alpine open heathland).
 In 2011-12, we estimated seedling density for both species across 40 alpine sites and found that the abundances of \textit{Grevillea} (Fig.
@@
 For both species, seedling density was also higher in severely burnt sites (i.e.
-sites with higher minimum twig diameters).
+sites with higher twig diameters).
 As expected, pre-fire adult density positively influenced \textit{Grevillea} seedling density.
@@
 Site level data, elevation and Topographic Wetness Index (TWI) were obtained from a 30 m resolution digital elevation model.
-Lastly, for burnt sites, fire severity was estimated by minimum twig diameters, collected immediately after the 2003 fires\cite{Camac:2013eb}.
+Lastly, for burnt sites, fire severity was estimated by twig diameters, collected immediately after the 2003 fires\cite{Camac:2013eb}.
 
@@
 \includegraphics{figures/fig2.pdf}
-\caption{\textbf{Mean \textit{Grevillea australis} seedling density along gradients of burning, fire severity (minimum twig diameter), altitude, Topographic Wetness Index (TWI) and adult density.} (\textbf{a}) Centered and standardized model coefficients (on the log scale); and effects of: (\textbf{b}) burning, (\textbf{c}) fire severity, (\textbf{d}) altitude, (\textbf{e}) topographic wetness and (\textbf{f}) adult density, in areas burnt by the 2003 fires.
+\caption{\textbf{Mean \textit{Grevillea australis} seedling density along gradients of burning, fire severity (twig diameter), altitude, Topographic Wetness Index (TWI) and adult density.} (\textbf{a}) Centered and standardized model coefficients (on the log scale); and effects of: (\textbf{b}) burning, (\textbf{c}) fire severity, (\textbf{d}) altitude, (\textbf{e}) topographic wetness and (\textbf{f}) adult density, in areas burnt by the 2003 fires.
 All bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
@@
 \includegraphics{figures/fig3.pdf}
-\caption{\textbf{Observed and projected growth trajectories of four dominant Australian alpine shrubs.} (\textbf{a}) Observed height growth: Thick lines represent mean ($\pm$ 95\% confidence intervals) observed heights at each May census.
+\caption{\textbf{Observed and projected growth trajectories of four dominant Australian alpine shrubs.} (\textbf{a}) Observed height growth: Thick lines with error bars represent mean ($\pm$ 95\% confidence intervals) observed heights at each May census.
 Thin lines represent individual growth trajectories.
 (\textbf{b}) Mean ($\pm$ 95\% Bayesian Credible Interval) projected growth trajectories.
-In all cases, red \& orange lines = seedlings growing in warmed (OTC) conditions and blue and light blue lines = seedlings growing in control conditions.
+In all cases, red and orange lines = seedlings growing in warmed (OTC) conditions and blue and light blue lines = seedlings growing in control conditions.
 Projections were based on an logistic growth model using initial seedling size of 6 cm (the mean initial height observed in the OTC experiment) and mean maximum heights observed in long-unburnt sites (i.e.

### 2016-03-14 manuscript/supplementary.Rnw
More edits

@@
 \section*{Open Top Chamber Experiment}
-We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of post-fire gap-infilling by  tussock grass; and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affects shrub seedling demographic rates; and 3) whether such effects change under warmer conditions post-fire conditions.
+We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of post-fire gap-infilling by tussock grass and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affects shrub seedling demographic rates; and 3) whether such effects change under warmer conditions post-fire conditions.
 
@@
 S5-7).
-OTCs passively increased average ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively.
+OTCs passively increased average daily ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively.
 Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C).
@@
 We marked (with paint on stems) recruits and recorded the number that germinated in each of the 2011 and 2012 seasons.
+
+\subsection{\textit{Data inputs for growth, mortality and recruitment models.}}
+In each model, OTC treatment, $OTC$, was a binary (0 = Control, 1 = OTC) plot level covariate.
+By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census $t$.
+However, because some distances were censored (i.e.
+$\geq$ 26 cm), these data were incomplete.
+We estimated these censored data using a censored model (see Modelling censored Poa Distances).
+Once true distances were estimated for censored data, we estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each census, for each inter-tussock seedling.
+Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots.
+In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations\cite{Gelman:2007te}.
+The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as responses to average conditions and slope terms as partial dependencies conditional on other continuous variables being at their mean.
 
@@
 
-The occurrence and severity of fire was initially determined immediately after the 2003 fires\cite{Williams:2006tl}.
+The occurrence and severity of fire was determined immediately after the 2003 fires\cite{Williams:2006tl}.
 The severity of the fires in each burnt site was estimated using mean twig diameter, a proxy measure of fire severity\cite{Whight:1999vw}.
-These twig diameters were recorded between 1 and 10 months post-fire from 10 replicate \textit{Grevillea australis} individuals.
-Across the 30 burnt sites, the average site twig diameter across ranged from 2 to 13 mm, with smaller diameters assumed to correspond to lower fire intensity.
+These twig diameters were recorded between 1 and 10 months post-fire from 10 replicate \textit{Grevillea australis} individuals per site.
+Across the 30 burnt sites, the average site twig diameter ranged from 2 to 13 mm, with smaller diameters assumed to correspond to lower fire intensity.
 This encompassed the full range of fire severity from lightly scorched sites containing shrubs with only the tips of the outer branches burnt, to severely burnt sites where the majority of large shrub branches were consumed.
+
+\subsection{\textit{Data inputs into density and maximum height models.}}
+In all seedling survey models, $Burnt_j$ is a binary (0 = unburnt, 1 = burnt) site level covariate, which only affects seedling counts at unburnt sites.
+By contrast, fire severity ($Severity_j$), was a continuous site level covariate that only affected seedling counts and maximum heights at burnt sites.
+Altitude ($Alt_j$), Topographic Wetness Index ($TWI_j$) and mature \textit{Grevillea} count ($ADen_k$) were also continuous covariates that affected all sites in both count and maximum height models.
+Collinearity among covariates was checked and all correlations were \textit{r} \textless 0.16.
+The data from each continuous covariate was then centered on their mean and standardized by two standard deviations\cite{Gelman:2007te}.
 
@@
 \subsection{\textit{Modelling shrub seedling growth rates.}}
-We modelled the observed height of individual $i$ at May census $t$ as a random realization from a normal distribution truncated at zero:
+We modelled the observed height of individual $i$ at May census $t$ as a random realisation from a normal distribution truncated at zero:
 
@@
 
-where $\widehat{height}_{i,t}$ is the estimated maximum height for each individual $i$ at each May census $t$ and $\sigma_{obs}$ is the standard deviation of the observation error.
+where $\widehat{height}_{i,t}$ is the estimated maximum height for individual $i$ at May census $t$ and $\sigma_{obs}$ is the standard deviation of the observation error.
 The parameter $\widehat{height}_{i,t}$ was estimated using a logistic parameterisation of the unified Richards equation\cite{Tjorve:2010bk}:
@@
 We used a logistic equation because it can simulate sigmoidal growth.
-That is, a curve whereby initial seedling growth is slow before increasing in speed and finally levelling off when adult height is reached.
+That is, a curve whereby initial seedling growth is slow before increasing in speed and finally levelling off when maximum adult height is reached.
 We used this particular parameterisation because it accounts for seedling size at the beginning of an experiment and contains biologically interpretable parameters.
@@
 \epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
-\epsilon_i &\sim \mathcal{N}(0, \sigma_{obs}).
+\epsilon_i &\sim \mathcal{N}(0, \sigma_{ind}).
 \end{aligned}
@@
 Here, the intercept term, $\alpha_{j[i]}$, represents the expected control plot growth rate of species $j$ to which individual $i$ belongs.
-It is estimated from a normal distribution centered on $\mu_{spp}^{CTL}$ (the expected control plot growth rate for an average hypothetical species) with a standard deviation, $\sigma_{spp}^{CTL}$ which defines the variation between species\textquotesingle growth rates at control plots.
+It is estimated from a normal distribution centered on $\mu_{spp}^{CTL}$ (the expected control plot growth rate for an average hypothetical species) with a standard deviation, $\sigma_{spp}^{CTL}$ which defines the variation between species\textquotesingle{} growth rates at control plots.
 The parameter $\beta_{j[i]}$ is the OTC effect for species $j$ to which individual $i$ belongs.
-$\beta_j$ is also estimated from a normal distribution, but unlike $\alpha_j$, it is centered on $\mu_{spp}^{OTC}$, (the expected OTC growth rate for an average hypothetical species) with a standard deviation, $\sigma_{spp}^{OTC}$ which defines the variation between species\textquotesingle growth rate within OTCs.
+$\beta_j$ is also estimated from a normal distribution, but unlike $\alpha_j$, it is centered on $\mu_{spp}^{OTC}$, (the expected OTC growth rate for an average hypothetical species) with a standard deviation, $\sigma_{spp}^{OTC}$ which defines the variation between species\textquotesingle{} growth rate within OTCs.
 The terms $\epsilon_{plot_{k[i]}}$ and $\epsilon_i$ represent residual variation at the plot level (for each plot, $k$, within which individual $i$ is located), and the individual level (for each individual $i$), respectively.
-These error terms were drawn from normal distributions centered on zero with standard deviations that defined the variation amongst plots, $\sigma_{plot}$, or amongst individuals, $\sigma_{obs}$, respectively.
+These error terms were drawn from normal distributions centered on zero with standard deviations that defined the variation amongst plots, $\sigma_{plot}$, or amongst individuals, $\sigma_{ind}$, respectively.
 
@@
 In both growth models standard deviation parameters (i.e.
-$\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{obs}$) were all estimated from weakly informative positive half Cauchy priors with scale parameters of 25.
+$\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$, $\sigma_{ind}$ and $\sigma_{obs}$) were all estimated from weakly informative positive half Cauchy priors with scale parameters of 25.
 We used these Cauchy priors in order to prevent over-inflated estimates of variances due to the small number of plots and species\cite{Gelman:2006uw}.
@@
 Our observations, $y_{i,t}$, were binary (0 = alive or 1 = dead) for each individual $i$, in each census, $t$.
-We modelled these binary observations through time as random realizations from a Bernoulli distribution based on the model-estimated cumulative hazard rate for each individual at each census:
+We modelled these binary observations through time as random realisations from a Bernoulli distribution based on the model-estimated cumulative hazard rate for each individual at each census:
 
@@
 $Cumulative\;hazard_{i,t}$, separately for seedlings growing in bare ground patches and those in inter-tussock gaps.
-In both cases, $Cumulative\;hazard_{i,t}$, was modelled on the complementary log-log scale as a function of an instantaneous hazard rate for each individual, $\lambda_i$, multiplied by the time lapsed (in years) between censuses, $t$.
+In both cases, $Cumulative\;hazard_{i,t}$, was modelled on the complementary log-log scale as a function of an instantaneous hazard rate for each individual, $\lambda_i$, multiplied by the census interval (in years) between censuses, $t$.
 By doing this our regression model becomes equivalent to a parametric survival analysis\cite{Kleinbaum:2012tn}.
@@
 Here, the intercept term, $\alpha_{j[i]}$, represents the expected instantaneous hazard rate in a control plot for species $j$ to which individual $i$ belongs.
-It is estimated from a normal distribution centered on $\mu_{spp}^{CTL}$  (the control plot expected instantaneous hazard rate for the average hypothetical species) with a standard deviation, $\sigma_{spp}^{CTL}$, which defines the variation between species\textquotesingle instantaneous hazard rates in control plots.
+It is estimated from a normal distribution centered on $\mu_{spp}^{CTL}$  (the control plot expected instantaneous hazard rate for the average hypothetical species) with a standard deviation, $\sigma_{spp}^{CTL}$, which defines the variation between species\textquotesingle{} instantaneous hazard rates in control plots.
 The parameter $\beta_{j[i]}$ is the OTC effect for each species $j$ to which individual $i$ belongs.
@@
 \subsection{\textit{Modelling Grevillea recruitment rates.}}
-We examined whether warming influenced natural recruitment rates of \textit{Grevillea} by modelling the number of recruits per 1 m$^2$ plot $k$ in each census, $t$ as a random realization from a negative binomial distribution:
+We examined whether warming influenced natural recruitment rates of \textit{Grevillea} by modelling the number of recruits per 1 m$^2$ plot $k$ in each census, $t$ as a random realisation from a negative binomial distribution:
 
@@
 where the intercept, $\alpha$, represents the average log recruitment in control plots in 2011.
-The parameter $\beta_1$ is the effects of OTC and $\beta_2$ describes the change in the log number of recruits from 2011 to 2012.
+The parameter $\beta_1$ is the effects of OTC and $\beta_2$ describes the change in the log number of recruits from 2011 (i.e.
+$Census$ = 0) to 2012 (i.e.
+$Census$ = 1).
 $\epsilon_{plot_k}$ is the residual error at plot $k$.
@@
 
-\subsection{\textit{Data inputs for growth, mortality and recruitment models.}}
-In each model, OTC treatment, $OTC$, was a binary (0 = Control, 1 = OTC) plot level covariate.
-By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census $t$.
-However, because some distances were censored (i.e.
-\textgreater 25 cm), these data were incomplete.
-We estimated these censored data using a censored model (see below).
-Once true distances were estimated for censored data, we estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each census, for each inter-tussock seedling.
-Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots.
-In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations\cite{Gelman:2007te}.
-The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as responses to average conditions and slope terms as partial dependencies conditional on other continuous variables being at their mean.
-
 \subsection{\textit{Modelling censored Poa distances.}}
-Approximately \Sexpr{prop_censored(cleaned_poa_distances)}\% of the $Poa$ cardinal distance data used to calculate the average inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, (i.e.
-\textgreater  25 cm).
+Approximately \Sexpr{prop_censored(cleaned_poa_distances)}\% of the $Poa$ distance data used to calculate the median inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, (i.e.
+$\geq$  26 cm).
 We estimated these censored values by first estimating the distribution of distances in observed data, coupled with estimating covariate effects, and then using these distributions to estimate censored observations.
 
-Specifically, we first modelled the observed (non-censored) cardinal distances as random realizations from a lognormal distribution:
+Specifically, we first modelled the observed (non-censored) cardinal distances as random realisations from a lognormal distribution:
 
@@
 
-where $\textrm{ln}(\mu_{i,t})$ is the log estimated median radius for inter-tussock gap, at census $t$, and $\sigma_{obs}$ is the standard deviation of the observation error.
+where $\textrm{ln}(\mu_{i,t})$ is the log estimated median radius for inter-tussock gap $i$, at census $t$, and $\sigma_{obs}$ is the standard deviation of the observation error.
 
@@
 \subsection{\textit{Modelling drivers of seedling density.}}
-We examined drivers of seedling density by modelling the number of seedlings square metre plot $i$ as a random realization from a Poisson distribution:
+We examined drivers of seedling density by modelling the number of seedlings per square metre plot $i$ as a random realisation from a Poisson distribution:
 
@@
 We used maximum heights because these were likely to represent individuals that established immediately post-fire, and would have had 9 years of post-fire growth.
-We modelled these heights, $maxht_i$, as a random realizations from a log-normal distribution:
+We modelled these heights, $maxht_i$, as a random realisations from a log-normal distribution:
 
@@
 All standard deviation parameters were modelled using positive half Cauchy priors with a scale of 25.
-
-\subsection{\textit{Data inputs into density and maximum height models.}}
-In all seedling survey models, $Burnt_j$ is a binary (0 = unburnt, 1 = burnt) site level covariate, which only affects seedling counts at unburnt sites.
-By contrast, fire severity $Severity_j$, was a continuous site level covariate that only affected seedling counts and maximum heights at burnt sites.
-Altitude ($Alt_j$), Topographic Wetness Index ($TWI_j$) and mature \textit{Grevillea} count ($ADen_k$) were also continuous covariates that affected all sites in both count and maximum height models.
-Collinearity among covariates was checked and all correlations were \textit{r} < 0.16.
-The data from each continuous covariate was then centered on their mean and standardized by two standard deviations\cite{Gelman:2007te}.
 \pagebreak
@@
 \includegraphics{figures/figS1.pdf}
-\caption{\textbf{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (minimum twig diameter), altitude and Topographic Wetness Index (TWI).} (\textbf{a}) Centered and standardized model coefficients (on the log scale); and effects of (\textbf{b}) burning, (\textbf{c}) fire severity, (\textbf{d}) altitude and (\textbf{e}) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires.
+\caption{\textbf{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (twig diameter), altitude and Topographic Wetness Index (TWI).} (\textbf{a}) Centered and standardized model coefficients (on the log scale); and effects of (\textbf{b}) burning, (\textbf{c}) fire severity, (\textbf{d}) altitude and (\textbf{e}) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires.
 All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
@@
 \caption{\textbf{Observed seedling diameter growth trajectories of four dominant Australian alpine shrubs.} Red and Blue lines represent mean ($\pm$ 95\% CI) observed heights at each May census in warmed and control plots, respectively.
-orange and light blue lines represent individual growth trajectories in warmed and unwarmed plots, respectively.}
+Orange and light blue lines represent individual growth trajectories in warmed and unwarmed plots, respectively.}
 \label{figs3}
@@
 \includegraphics{figures/figS4.pdf}
-\caption{\textbf{Mean maximum seedling height along fire severity (minimum twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires.} Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses.
+\caption{\textbf{Mean maximum seedling height along fire severity (twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires.} Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses.
 (\textbf{a} \& \textbf{b}) Centered and standardized model coefficients (on the log scale); and effects of (\textbf{c} \& \textbf{d}) fire severity, (\textbf{e} \& \textbf{f}) altitude and (\textbf{g} \& \textbf{h}) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires.

### 2016-03-15 R/figures.R
Changed density ylim

@@ -123,7 +123,7 @@ recruit_plot <- function(recruit_model, ylab =expression('Recruits/'~m^2), xlab 
 
 
 # Plots covariate partial plots for density and height models
-partial_plot_density_height <- function(predictions, x, xlab=NULL,ylab=NULL, ylim =c(0,15)) {
+partial_plot_density_height <- function(predictions, x, xlab=NULL,ylab=NULL, ylim =c(0,40)) {
   ggplot(predictions, aes_string(x = x,y = 'mean'), axis.line = element_line()) + 
     geom_line() +
     scale_fill_manual("",values= 'grey60') +

### 2016-03-15 manuscript/supplementary.Rnw
Missed minus sign in logistic equation
Moved a few sections around to make text more coherent

@@
 plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset\cite{Gelman:2007te, Kery:2012wl}.
-For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package rstan \Sexpr{packageVersion("rstan")}.
+For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package rstan \Sexpr{packageVersion("rstan")}\cite{stan2015xx}.
 For each model, three independent chains were run.
@@
 \begin{equation}\label{eq:logistic} 
-\widehat{height}_{i,t} = \frac{Hmax_{j[i]}}{\left( 1 + \left(\frac{Hmax_{j[i]}}{Hinit_i} - 1\right)e^{R_i \frac{Days_t}{365.25}}\right)},
+\widehat{height}_{i,t} = \frac{Hmax_{j[i]}}{\left( 1 + \left(\frac{Hmax_{j[i]}}{Hinit_i} - 1\right)e^{-R_i \frac{Days_t}{365.25}}\right)},
 \end{equation}

### 2016-03-15 manuscript/manuscript.Rnw
Added font style for package names

@@
 In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality.
-For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package rstan \Sexpr{packageVersion("rstan")}.
+For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")}.
 Data and source code for reproducing analysis and figures are available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.

### 2016-03-15 manuscript/supplementary.Rnw
Added updated details about DEM

@@
 We measured seedling density and maximum heights in open heathland across the Bogong High Plains.
-The primary aim of this study was to evaluate how variables such as altitude, Topographic Wetness Index (a measure of plant available water), adult density, fire, and fire severity influence the density of \textit{Asterolasia} and \textit{Grevillea} seedlings in open heathland.
+The primary aim of this study was to evaluate how variables such as altitude, Topographic Wetness Index (TWI; a measure of plant available water), adult density, fire, and fire severity influence the density of \textit{Asterolasia} and \textit{Grevillea} seedlings in open heathland.
 A secondary aim was to verify growth rates observed in the OTC experiment by assessing maximum seedling heights of \textit{Asterolasia} and \textit{Grevillea} along altitude, soil moisture and fire severity gradients in areas burnt by the 2003 wildfires.
@@
 This disparity in sampling effort between burnt and unburnt was due to very few open heathland remaining unburnt after both the 2003 and 2007 fires.
-These sites encompassed most of the Bogong High Plains (120 km$^2$) and ranged in altitude from 1660 to 1860 m a.s.l.
+These sites encompassed most of the Bogong High Plains (120 km$^2$) and ranged in altitude from \Sexpr{round(min(complete_field_survey_data$altitude_m))} to \Sexpr{round(max(complete_field_survey_data$altitude_m))} m a.s.l.
 Sites were at least 0.25 ha and more than 500 m apart.
@@
 Different variables were recorded at the three sampling levels (site, transect and plot).
-For each site, altitude and TWI were derived from a 30 m resolution digital elevation model of the Bogong High Plains.
-TWI, which approximates the relative plant-available water based on a site\textquotesingle s position in the landscape\cite{Moore:1993fk}, was derived for each site using $\textrm{TWI}= \textrm{ln}(\frac{\alpha}{\textrm{tan}\theta})$, where, $\alpha$ is the upslope catchment area per unit width orthogonal to the flow direction and $\theta$ is the slope (in radians).
-Across the 40 sites, TWI varied twofold (5 to 10 units; higher units = wetter).
-Altitude ranged by \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m (i.e.
-from \Sexpr{round(min(complete_field_survey_data$altitude_m))} to \Sexpr{round(max(complete_field_survey_data$altitude_m))} m a.s.l), or by approximately \Sexpr{round(0.77 * (190/100),1)}\textdegree C from the lowest to the highest site (based on a lapse-rate of 0.77\textdegree C/100 m elevation\cite{Brown:1989vx}.
+For each site, altitude and TWI were derived from a 30 m resolution digital elevation model\cite{Gallant:2011er} of the Bogong High Plains.
+TWI, which approximates the relative plant-available water based on a site\textquotesingle s position in the landscape\cite{Moore:1993fk}, was derived for each site using R package \texttt{dynatopmodel}\cite{Metcalfe:2016ud}.
+Across the 40 sites, TWI varied twofold (\Sexpr{min(round(complete_field_survey_data$topo_wetness_index))} to \Sexpr{round(max(complete_field_survey_data$topo_wetness_index))} units; higher units = wetter).
+Altitude ranged by \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m, or by approximately \Sexpr{round(0.77 * (190/100),1)}\textdegree C from the lowest to the highest site (based on a lapse-rate of 0.77\textdegree C/100 m elevation\cite{Brown:1989vx}.
 This altitudinal temperature range is comparable to that observed in the OTC experiment.
@@
 plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset\cite{Gelman:2007te, Kery:2012wl}.
-For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package rstan \Sexpr{packageVersion("rstan")}\cite{stan2015xx}.
+For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")}\cite{stan:2015xx}.
 For each model, three independent chains were run.
@@
 Here the standard deviation parameters in both mortality models (i.e.
-$\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{obs}$) were each estimated from positive half Cauchy priors with standard deviation of 2.5.
+$\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{ind}$) were each estimated from positive half Cauchy priors with standard deviation of 2.5.
 The intercept term $\alpha$ for inter-tussock seedling model and the mean hyper-parameter $\mu_{spp}^{CTL}$ in the multispecies models were estimated from cauchy priors with standard deviations of 2.5.
@@
 Half Cauchy priors with a scale of 25 were used for both the overdispersion parameter, $\phi$ and the plot level standard deviation $\sigma_{plot}$.
-Coefficients $\beta_1$ and $\beta_2$ were estimated using normal priors centered on zero with a standard deviation of 2.5, encapsulating our prior knowledge that plots were unlikely to have more than 150 recruits/m$^2$.
+Coefficients $\beta_1$ and $\beta_2$ were estimated using normal priors centered on zero with a standard deviation of 2.5, encapsulating our prior knowledge that plots were unlikely to have more than 100 recruits/m$^2$.
 
@@
 Half Cauchy priors with a scale of 25 were used for all standard deviation parameters.
-Hyperparameter $\mu_{site}$ and coefficients $\beta_1$ to $\beta_5$ were estimated using normal priors centered on zero with a standard deviation of 2.5, encapsulating our prior belief that plots were unlikely to have more than 150 seedlings/m$^2$.
+Hyperparameter $\mu_{site}$ and coefficients $\beta_1$ to $\beta_5$ were estimated using normal priors centered on zero with a standard deviation of 2.5, encapsulating our prior belief that plots were unlikely to have more than 100 seedlings/m$^2$.
 
@@
 Here, we estimated $\mu_{site}$ and coefficients ($\beta_1$ to $\beta_3$) using weakly informative normal priors with a mean of 0 and a standard deviation of 2.5.
-These priors encapsulated our prior knowledge that seedling maximum heights would be less than 150 cm.
+These priors encapsulated our prior knowledge that seedling maximum heights would be less than 100 cm.
 All standard deviation parameters were modelled using positive half Cauchy priors with a scale of 25.

### 2016-03-15 manuscript/manuscript.Rnw
Final edits

@@
 
-\title{Climatic warming strengthens a positive feedback between alpine shrubs and fire}
+\title{Climatic warming strengthens a positive feedback\\ between alpine shrubs and fire}
 \author{James S. Camac$^{*,1,2,3}$, Richard J. Williams$^4$, Carl-Henrik Wahren$^3$, Ary A. Hoffmann$^5$ \\
@@
 \subsection{Drivers of shrub seedling establishment.}
-We first investigated how altitude, Topographic Wetness Index (TWI), adult density, fire and fire severity (as measured by twig diameters\textemdash see Supplementary Methods) influenced the density of \textit{Grevillea} and \textit{Asterolasia} seedlings (the two dominant shrubs of alpine open heathland).
+We first investigated how altitude, Topographic Wetness Index (TWI), adult density, fire and fire severity (as measured by post-fire twig diameters\textemdash see Supplementary Methods) influenced the density of \textit{Grevillea} and \textit{Asterolasia} seedlings (the two dominant shrubs of alpine open heathland).
 In 2011-12, we estimated seedling density for both species across 40 alpine sites and found that the abundances of \textit{Grevillea} (Fig.
@@
 For both species, seedling density was also higher in severely burnt sites (i.e.
-sites with higher twig diameters).
+sites with larger post-fire twig diameters).
 As expected, pre-fire adult density positively influenced \textit{Grevillea} seedling density.
@@
 S2) and mortality (see below) largely being insensitive to a \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C change in temperature.
-We detected no significant effect of Topographic Wetness Index for either species.
+We detected no strong effect of Topographic Wetness Index for either species.
 
@@
 After \Sexpr{n_days(chamber_dates, FALSE)} days (\Sexpr{n_days(chamber_dates, TRUE)} growing season days) or \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} years' growth, mean seedling heights of the tall shrubs (\textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera}) growing in post-fire bare ground were greater in warmed plots relative to controls by \Sexpr{ht_diff(non_tussock_otc_growth_data,'Grevillea')}, \Sexpr{ht_diff(non_tussock_otc_growth_data,'Phebalium')} and \Sexpr{ht_diff(non_tussock_otc_growth_data,'Prostanthera')} cm, respectively (Fig.
-\ref{fig3}).
+\ref{fig3}a).
 Warming increased heights of both \textit{Grevillea} and \textit{Prostanthera} seedlings in all years, while \textit{Phebalium} did not respond to the warming treatment until the second growing season.
@@
 According to this model, a 6 cm seedling (the mean initial height of seedlings used in this experiment) attains maximum height 39 years sooner \Sexpr{hmax(non_tussock_otc_growth_model,'Grevillea')} (\textit{Grevillea}) or \Sexpr{hmax(non_tussock_otc_growth_model,'Prostanthera')} years sooner (\textit{Phebalium} and \textit{Prostanthera}) when warmed by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C (Fig.
-\ref{fig3}).
+\ref{fig3}b).
 \textit{Asterolasia} was predicted to reach its maximum height in approximately \Sexpr{round(hmax(non_tussock_otc_growth_model,'Asterolasia', FALSE)[2])} to \Sexpr{round(hmax(non_tussock_otc_growth_model,'Asterolasia', FALSE)[1])} years, irrespective of warming treatment.
@@
 We validated the \textit{Grevillea} and \textit{Asterolasia} growth responses observed in experimentally warmed plots by examining the effect of elevation on maximum heights of post-fire recruits across 30 open heathland sites burnt in 2003.
-Here, a 200 m altitudinal range is equivalent to a mean ambient temperature difference of approximately \Sexpr{round(0.77 * (200/100),1)}\textdegree C\cite{Brown:1989vx}, which is comparable to that observed between experimentally warmed and control plots (\Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C).
+Here, a 190 m altitudinal range is equivalent to a mean ambient temperature difference of approximately \Sexpr{round(0.77 * (190/100),1)}\textdegree C\cite{Brown:1989vx}, which is comparable to that observed between experimentally warmed and control plots (\Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C).
 In response to shifts in temperature, the maximum height of \textit{Grevillea} post-fire recruits was expected to decrease with elevation, whereas \textit{Asterolasia} seedlings were not expected to show this pattern.
@@
 It is also an ecotone between closed heathland (\textgreater 70\% shrub cover) and tussock grassland consisting of \textit{Grevillea australis} shrubs interspersed among a \textit{Poa hiemata} sward.
-Aerial photography has revealed that the plant community has acted as an invasion front of shrub expansion into grassland and has itself experienced shrub thickening by closed heathland dominants such as \textit{Prostanthera cuneata}.
+Aerial photography has revealed that the plant community has acted as an invasion front of shrub expansion into grassland and has itself experienced shrub thickening by closed heathland dominants such as \textit{Prostanthera cuneata}\cite{McDougall:2003gb}.
 
@@
 Monica Camac, Brad Farmilo, Monica Hursburgh, Paulius Kviecinskas, Karen Stott, Imogen Fraser, Sarah Bartlett, Paul Smart, Steven Rumbold, Lachlan Yourn and Chris Jones provided assistance with field measurements.
-William Morris and John Baumgartner provided insightful chats about modelling approaches and Warwick Papst provided field logistic advice.
+William Morris and John Baumgartner provided coding advice and Warwick Papst provided field logistic advice.
 Daniel Falster and Richard Fitzjohn for developing software that enabled these analyses to be readily reproducible.
-Special thanks to David Bowman, Ana{\"i}s Gibert, Brad Farmilo, John Morgan, Kimberley Millers, Dale Nimmo, Michaela Plein, Libby Rumpff and Inka Veltheim for their comments on this manuscript.
+Special thanks to David Bowman, Ana{\"i}s Gibert, John Morgan, Kimberley Millers, Michaela Plein and Inka Veltheim for their comments on this manuscript.
 

### 2016-03-15 manuscript/supplementary.Rnw
Final edits

@@
 
-\title{Climatic warming strengthens a positive feedback between alpine shrubs and fire (Supplementary Material)}
+\title{Climatic warming strengthens a positive feedback \\ between alpine shrubs and fire \\ 
+(Supplementary Material)}
 \author{James S. Camac$^{*,1,2,3}$, Richard J. Williams$^4$, Carl-Henrik Wahren$^3$, Ary A. Hoffmann$^5$ \\
@@
 The occurrence and severity of fire was determined immediately after the 2003 fires\cite{Williams:2006tl}.
-The severity of the fires in each burnt site was estimated using mean twig diameter, a proxy measure of fire severity\cite{Whight:1999vw}.
+The severity of the fires in each burnt site was estimated using mean post-fire twig diameter, a proxy measure of fire severity\cite{Whight:1999vw}.
 These twig diameters were recorded between 1 and 10 months post-fire from 10 replicate \textit{Grevillea australis} individuals per site.

### 2016-04-11 R/build.R
usr/bin/ doesnt work with el captain

@@ -1,4 +1,4 @@
-#!/usr/bin/env Rscript
+#!/usr/local/bin/env Rscript
 
 latex_build <- function(filename, bibliography=NULL,
                         chdir=TRUE, interaction="nonstopmode",

### 2016-04-11 manuscript/manuscript.Rnw
Formatted manuscript for PNAS

@@
-\documentclass{nature}
-\bibliographystyle{naturemag.bst}
-\usepackage{lineno}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage[hidelinks]{hyperref}
+\documentclass[9pt,twocolumn,twoside, lineno]{pnas-new}
+% Use the lineno option to display guide line numbers if required.
+% Note that the use of elements such as single-column equations
+% may affect the guide line number alignment.
+
+\templatetype{pnasresearcharticle} % Choose template 
+% {pnasresearcharticle} = Template for a two-column research article
+% {pnasmathematics} = Template for a one-column mathematics article
+% {pnasinvited} = Template for a PNAS invited submission
 
 \title{Climatic warming strengthens a positive feedback\\ between alpine shrubs and fire}
-\author{James S. Camac$^{*,1,2,3}$, Richard J. Williams$^4$, Carl-Henrik Wahren$^3$, Ary A. Hoffmann$^5$ \\
-\& Peter A. Vesk$^2$}
 
-\begin{document}
-\linenumbers
-\maketitle
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\author[a,b,c,1]{James S Camac}
+\author[d,2]{Richard J Williams} 
+\author[c,3]{Carl-Henrik Wahren}
+\author[e,4]{Ary A Hoffmann}
+\author[b,5]{Peter A Vesk}
 
-\begin{affiliations}
-\item Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia.
-\item The Centre of Excellence for Environmental Decisions, School of BioSciences, The University of Melbourne, Melbourne, Victoria 3010, Australia.
-\item Research Centre for Applied Alpine Ecology, La Trobe University, Melbourne, Victoria 3086, Australia.
-\item CSIRO Land and Water, Tropical Ecosystems Research Centre, PMB 44 Winnellie, Northern Territory 0821, Australia.
-\item Bio21 Institute, School of BioSciences, The University of Melbourne, Victoria 3010, Australia.
-\textasteriskcentered Email for correspondence: \texttt{james.camac@gmail.com}
-\end{affiliations}
+\affil[a]{Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia}
+\affil[b]{The Centre of Excellence for Environmental Decisions, School of BioSciences, The University of Melbourne, Melbourne, Victoria 3010, Australia}
+\affil[c]{Research Centre for Applied Alpine Ecology, La Trobe University, Melbourne, Victoria 3086, Australia}
+\affil[d]{CSIRO Land and Water, Tropical Ecosystems Research Centre, PMB 44 Winnellie, Northern Territory 0821, Australia}
+\affil[e]{Bio21 Institute, School of BioSciences, The University of Melbourne, Victoria 3010, Australia}
 
-\pagebreak
+% Please give the surname of the lead author for the running footer
+\leadauthor{Camac}
+
+% Please add here a significance statement to explain the relevance of your work
+\significancestatement{Temperatures are rapidly increasing in alpine and tundra ecosystems worldwide.
+A consequence of this temperature increase has been significant increases in shrub cover and a predicted increase in the frequency or severity of fires.
+Here we use experiemental and observational data to explore whether there are antagonistic or synergistic feedbacks between temperature, fire and shrub demographic rates.
+We find that increases in temperatures, coupled with more frequent or severe fire is likely to exacerbate the rate of shrub cover increases, and consquently, increase landscape flammability.}
+
+% Please include corresponding author, author contribution and author declaration information
+\authorcontributions{J.C.
+conceived, designed and performed the experiments, field surveys and analysis; C-H.W, R.W, A.H and P.V.
+supervised the development of this work, aided in data collection and provided statistical advice.
+All authors contributed to the writing of this manuscript.}
+\authordeclaration{The authors declare that they have no competing financial interests.}
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: james.camac@gmail.com}
+
+% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
+\keywords{Bayesian models $|$ Biotic interactions $|$ Growth $|$ Recruitment $|$ Mortality}
 
@@
 
-Accurately forecasting the effects of climatic warming on vegetation dynamics requires an understanding of the mechanisms by which climate and vegetation interact.
+\dates{This manuscript was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
+
+\begin{document}
+% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
+% You should only change this length when you've finalised the article contents.
+\verticaladjustment{-2pt}
+
+\maketitle
+\thispagestyle{firststyle}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
+
+% If your first paragraph (i.e.
+with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation.
+If this is the case, add \parshape=0 to the end of the list environment.
+\dropcap{A}ccurately forecasting the effects of climatic warming on vegetation dynamics requires an understanding of the mechanisms by which climate and vegetation interact.
 Most forecasting models include the direct effects of climatic conditions on species distributions, but largely ignore biotic interactions and the type, frequency and severity of disturbances that are also likely to change\cite{Guisan:2005gz,Tucker:2012dr,Giannini:2013gi}.
@@
 Consequently, little is known about how climate change influences seedling demographic rates\cite{Briceno:2015iv} and thus, vegetation recovery in post-disturbance conditions.
-We consider these issues in Australian alpine heathland.
 
@@
 
+\begin{figure}
+\includegraphics{figures/fig1.pdf}
+\caption{Conceptual diagram illustrating positive warming-shrub-fire feedback loop.
+(A) Hypothesised positive feedback loop between fire, climatic warming and landscape flammability.
+Solid blue lines = known mechansims; Dashed black lines = mechansims that may exacerbate or diminish this feedback loop but which we have a paucity of information on.
+(B) Feedback loop based on new evidence obtained from experimental and field surveys conducted in this study.
+Also includes an example of a potential, but unlikely, effect (red dotted line) that could break this feedback.}
+\label{fig1}
+\end{figure}
+
 While paleoecological studies have indicated that such a feedback may exist\cite{Higuera:2008dw, Higuera:2009ga}, there is a paucity of information on multiple mechanisms that may strengthen, mitigate or break this feedback loop (Fig.
-\ref{fig1}a).
+\ref{fig1}A).
 For example, we have little information on whether fire will stimulate shrub seedling recruitment in alpine or tundra ecosystems, nor do we know how rates of seedling growth and survival will be affected under warmer, more exposed, post-fire conditions.
@@
 
-\subsection{Drivers of shrub seedling establishment.}
-We first investigated how altitude, Topographic Wetness Index (TWI), adult density, fire and fire severity (as measured by post-fire twig diameters\textemdash see Supplementary Methods) influenced the density of \textit{Grevillea} and \textit{Asterolasia} seedlings (the two dominant shrubs of alpine open heathland).
+\section*{Results and Discussion}
+
+\begin{figure}
+\centering
+\includegraphics{figures/fig2.pdf}
+\caption{Mean \textit{Grevillea australis} seedling density along gradients of burning, fire severity (twig diameter), altitude, Topographic Wetness Index (TWI) and adult density.
+(A) Centered and standardized model coefficients (on the log scale); and effects of: (B) burning, (C) fire severity, (D) altitude, (E) topographic wetness and (F) adult density, in areas burnt by the 2003 fires.
+All bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
+\label{fig2}
+\end{figure}
+
+\begin{figure*}
+\centering
+\includegraphics{figures/fig3.pdf}
+\caption{Observed and projected growth trajectories of four dominant Australian alpine shrubs.
+(A) Observed height growth: Thick lines with error bars represent mean ($\pm$ 95\% confidence intervals) observed heights at each May census.
+Thin lines represent individual growth trajectories.
+(B) Mean ($\pm$ 95\% Bayesian Credible Interval) projected growth trajectories.
+In all cases, red and orange lines = seedlings growing in warmed (OTC) conditions and blue and light blue lines = seedlings growing in control conditions.
+Projections were based on an logistic growth model using initial seedling size of 6 cm (the mean initial height observed in the OTC experiment) and mean maximum heights observed in long-unburnt sites (i.e.
+\Sexpr{hmax_values(hmax,'Asterolasia')}, \Sexpr{hmax_values(hmax,'Grevillea')}, \Sexpr{hmax_values(hmax,'Phebalium')} \& \Sexpr{hmax_values(hmax,'Prostanthera')} cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Phebalium}, \& \textit{Prostanthera}, respectively).}
+\label{fig3}
+\end{figure*}
+
+\subsection*{Drivers of shrub seedling establishment}
+We first investigated how altitude, Topographic Wetness Index (TWI), adult density, fire and fire severity (as measured by post-fire twig diameters\textemdash see SI Materials and Methods) influenced the density of \textit{Grevillea} and \textit{Asterolasia} seedlings (the two dominant shrubs of alpine open heathland).
 In 2011-12, we estimated seedling density for both species across 40 alpine sites and found that the abundances of \textit{Grevillea} (Fig.
@@
 
-\subsection{Experimental warming effects on shrub seedling growth.}
+\subsection*{Experimental warming effects on shrub seedling growth}
 
@@
 After \Sexpr{n_days(chamber_dates, FALSE)} days (\Sexpr{n_days(chamber_dates, TRUE)} growing season days) or \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} years' growth, mean seedling heights of the tall shrubs (\textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera}) growing in post-fire bare ground were greater in warmed plots relative to controls by \Sexpr{ht_diff(non_tussock_otc_growth_data,'Grevillea')}, \Sexpr{ht_diff(non_tussock_otc_growth_data,'Phebalium')} and \Sexpr{ht_diff(non_tussock_otc_growth_data,'Prostanthera')} cm, respectively (Fig.
-\ref{fig3}a).
+\ref{fig3}A).
 Warming increased heights of both \textit{Grevillea} and \textit{Prostanthera} seedlings in all years, while \textit{Phebalium} did not respond to the warming treatment until the second growing season.
@@
 According to this model, a 6 cm seedling (the mean initial height of seedlings used in this experiment) attains maximum height 39 years sooner \Sexpr{hmax(non_tussock_otc_growth_model,'Grevillea')} (\textit{Grevillea}) or \Sexpr{hmax(non_tussock_otc_growth_model,'Prostanthera')} years sooner (\textit{Phebalium} and \textit{Prostanthera}) when warmed by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C (Fig.
-\ref{fig3}b).
+\ref{fig3}B).
 \textit{Asterolasia} was predicted to reach its maximum height in approximately \Sexpr{round(hmax(non_tussock_otc_growth_model,'Asterolasia', FALSE)[2])} to \Sexpr{round(hmax(non_tussock_otc_growth_model,'Asterolasia', FALSE)[1])} years, irrespective of warming treatment.
@@
 
-\subsection{Experimental warming on shrub seedling mortality.}
+\subsection*{Experimental warming on shrub seedling mortality}
 
@@
 
-\subsection{Effects of tussock grass proximity on shrub seedling growth and mortality rates.}
+\subsection*{Effects of tussock grass proximity on shrub seedling growth and mortality rates}
 
@@
 
-\subsection{Experimental warming and rates of gap infilling by tussock grass.}
+\subsection*{Experimental warming and rates of gap infilling by tussock grass}
 
@@
 
-\subsection{Strengthening of the warming-shrub-fire feedback.}
+\subsection*{Strengthening of the warming-shrub-fire feedback}
 
 Here we quantified several unknown interactions between fire, shrub-grass relationships and climate to extend a conceptual model of alpine shrub dynamics under climate change (Fig.
-\ref{fig1}b).
+\ref{fig1}B).
 We showed how these interactions strengthen a hypothesized feedback loop that can rapidly increase shrub cover in alpine and tundra ecosystems\cite{Higuera:2008dw}.
@@
 This further strengthens the feedback\cite{Higuera:2008dw, Higuera:2009ga} by potentially increasing the frequency or severity of fires, which then creates further recruitment opportunities for shrubs with little demographic impact of neighbouring tussock grass.
+
+\begin{figure}
+\centering
+\includegraphics{figures/fig4.pdf}
+\caption{Mean ($\pm$ 95\% Bayesian Credible Interval) annual probability of death for each species growing in control (blue) and warmed (red) conditions.}
+\label{fig4}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{figures/fig5.pdf}
+\caption{Experimental warming and inter-tussock gap size effects on \textit{Grevillea australis} seedlings.
+Top rows = Growth rate effects, Second row = Mortality effects.
+(A \& C) Centered and standardized model coefficients; (B \& D) growth and mortality rate response curves along an inter-tussock gap size gradient in warmed (red) and control (blue) conditions.
+All error bars and shading are 95\% Bayesian Credible Intervals.}
+\label{fig5}
+\end{figure}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics{figures/fig6.pdf}
+\caption{Effects of warming on rates of gap infilling by tussock grass.
+(A) Model coefficients and (B) projected temporal change in size for an average 10 cm inter-tussock gap in warmed (red) and unwarmed (blue) conditions.
+All error bars and shading are 95\% Bayesian Credible Intervals.}
+\label{fig6}
+\end{figure}
+
 Our findings provide mechanistic understanding as to why shrub cover has increased, often at the expense of grasslands, in many alpine and arctic ecosystems\cite{Racine:2004tn,MyersSmith:2011fp,Frost:2013tm,Lantz:2013fh}.
@@
 
-\begin{methods}
-
+\matmethods{
 We investigated shrub dynamics in open heathland because it is a common and highly flammable plant community in the Australian Alps above 1600 m a.s.l\cite{Williams:2006tl}.
@@
 
-\subsection{Open Top Chamber Experiment.}
+\subsection*{Open Top Chamber Experiment}
 
@@
 
-\subsection{Seedling gradient study.}
+\subsection*{Seedling gradient study}
 
@@
 
-\subsection{Data analysis.}
+\subsection*{Data analysis}
 
@@
 Data and source code for reproducing analysis and figures are available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.
-Additional information about experimental design and analysis can be found in Supplementary Methods.
-\end{methods}
+Additional information about experimental design and analysis can be found in SI Materials and Methods.
+\\
+\\
+}
 
-\section*{Acknowledgements} 
+\showmatmethods % Display the Materials and Methods section
+
+\acknow{
 This research was funded through Australian Research Council Linkage Grants, partnered through the Department of Sustainability and Environment, Parks Victoria, Commonwealth Scientific and Industrial Research Organisation (CSIRO).
@@
 Daniel Falster and Richard Fitzjohn for developing software that enabled these analyses to be readily reproducible.
-Special thanks to David Bowman, Ana{\"i}s Gibert, John Morgan, Kimberley Millers, Michaela Plein and Inka Veltheim for their comments on this manuscript.
+Special thanks to David Bowman, Ana{\"i}s Gibert, John Morgan, Kimberley Millers, Michaela Plein and Inka Veltheim for their comments on this manuscript.}
 
-\section*{Author contributions}
-J.C.
-conceived, designed and performed the experiments, field surveys and analysis; C-H.W, R.W, A.H and P.V.
-supervised the development of this work, aided in data collection and provided statistical advice.
-All authors contributed to the writing of this manuscript.
+\showacknow % Display the acknowledgments section
 
-\section*{Competing financial interests} 
-The authors declare that they have no competing financial interests.
+% \pnasbreak splits and balances the columns before the references.
+% If you see unexpected formatting errors, try commenting out this line
+% as it can run into problems with floats and footnotes on the final page.
+%\pnasbreak
 
+% Bibliography
 \bibliography{refs.bib}
-
-\section*{Figures}
-
-\begin{figure}
-\begin{minipage}{\textwidth}
-\includegraphics{figures/fig1.pdf}
-\caption{\textbf{Conceptual diagram illustrating positive warming-shrub-fire feedback loop.} (\textbf{a}) Hypothesised positive feedback loop between fire, climatic warming and landscape flammability.
-Solid blue lines = known mechansims; Dashed black lines = mechansims that may exacerbate or diminish this feedback loop but which we have a paucity of information on.
-(\textbf{b}) Feedback loop based on new evidence obtained from experimental and field surveys conducted in this study.
-Also includes an example of a potential, but unlikely, effect (red dotted line) that could break this feedback.}
-\label{fig1}
-\end{minipage}
-\end{figure}
-
-\begin{figure}
-\begin{minipage}{\textwidth}
-\includegraphics{figures/fig2.pdf}
-\caption{\textbf{Mean \textit{Grevillea australis} seedling density along gradients of burning, fire severity (twig diameter), altitude, Topographic Wetness Index (TWI) and adult density.} (\textbf{a}) Centered and standardized model coefficients (on the log scale); and effects of: (\textbf{b}) burning, (\textbf{c}) fire severity, (\textbf{d}) altitude, (\textbf{e}) topographic wetness and (\textbf{f}) adult density, in areas burnt by the 2003 fires.
-All bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
-\label{fig2}
-\end{minipage}
-\end{figure}
-
-\begin{figure}
-\begin{minipage}{\textwidth}
-\includegraphics{figures/fig3.pdf}
-\caption{\textbf{Observed and projected growth trajectories of four dominant Australian alpine shrubs.} (\textbf{a}) Observed height growth: Thick lines with error bars represent mean ($\pm$ 95\% confidence intervals) observed heights at each May census.
-Thin lines represent individual growth trajectories.
-(\textbf{b}) Mean ($\pm$ 95\% Bayesian Credible Interval) projected growth trajectories.
-In all cases, red and orange lines = seedlings growing in warmed (OTC) conditions and blue and light blue lines = seedlings growing in control conditions.
-Projections were based on an logistic growth model using initial seedling size of 6 cm (the mean initial height observed in the OTC experiment) and mean maximum heights observed in long-unburnt sites (i.e.
-\Sexpr{hmax_values(hmax,'Asterolasia')}, \Sexpr{hmax_values(hmax,'Grevillea')}, \Sexpr{hmax_values(hmax,'Phebalium')} \& \Sexpr{hmax_values(hmax,'Prostanthera')} cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Phebalium}, \& \textit{Prostanthera}, respectively).}
-\label{fig3}
-\end{minipage}
-\end{figure}
-
-\begin{figure}
-\begin{minipage}{\textwidth}
-\includegraphics{figures/fig4.pdf}
-\caption{\textbf{Mean ($\pm$ 95\% Bayesian Credible Interval) annual probability of death for each species growing in control (blue) and warmed (red) conditions.}}
-\label{fig4}
-\end{minipage}
-\end{figure}
-\pagebreak
-
-\begin{figure}
-\begin{minipage}{\textwidth}
-\includegraphics{figures/fig5.pdf}
-\caption{\textbf{Experimental warming and inter-tussock gap size effects on \textit{Grevillea australis} seedlings.} Top rows = Growth rate effects, Second row = Mortality effects.
-(\textbf{a} \& \textbf{c}) Centered and standardized model coefficients; (\textbf{b} \& \textbf{d}) growth and mortality rate response curves along an inter-tussock gap size gradient in warmed (red) and control (blue) conditions.
-All error bars and shading are 95\% Bayesian Credible Intervals.}
-\label{fig5}
-\end{minipage}
-\end{figure}
-\pagebreak
-
-\begin{figure}
-\begin{minipage}{\textwidth}
-\includegraphics{figures/fig6.pdf}
-\caption{\textbf{Effects of warming on rates of gap infilling by tussock grass.} (\textbf{a}) Model coefficients and (\textbf{b}) projected temporal change in size for an average 10 cm inter-tussock gap in warmed (red) and unwarmed (blue) conditions.
-All error bars and shading are 95\% Bayesian Credible Intervals.}
-\label{fig6}
-\end{minipage}
-\end{figure}
-
+%\pnasbreak
 \end{document}

### 2016-04-11 R/figures.R
altered size of font to correctly size figures for publication

@@ -3,22 +3,22 @@
 partial_plot_theme <- function(legend.position = "none", strips = FALSE,...) {
   sb <- if(strips==TRUE) element_rect(fill='lightgrey') else element_blank()
   st <- if(strips==TRUE) element_text(face='italic') else element_blank()
-  theme_classic(base_size = 10) + theme(strip.text = st,
+  theme_classic(base_size = 7) + theme(strip.text = st,
                           legend.title = element_blank(),
                           strip.background = sb,
                           legend.position = legend.position,
                           axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
                           axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'),
-                          plot.margin = unit(c(5,5,5,5), "mm"))
+                          plot.margin = unit(c(3,3,3,3), "mm"))
 }
 
 # Theme for coefficient plots
 coefficent_plot_theme <- function() {
-  theme_classic(base_size = 10) + theme(
+  theme_classic(base_size = 7) + theme(
                           axis.title.y = element_blank(),
                           axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
                           axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'),
-                          plot.margin = unit(c(5,5,5,5), "mm"))
+                          plot.margin = unit(c(3,3,3,3), "mm"))
 }
 
 
@@ -55,7 +55,7 @@ coefficient_plot <- function(summarised_coefficients, y_axis_labels='',xlab ='Ef
     geom_vline(aes(xintercept=0), linetype=2) +
     scale_y_discrete(labels =rev(y_axis_labels)) +
     xlab(xlab) +
-    scale_x_continuous(breaks= scales::pretty_breaks(6)) +
+    scale_x_continuous(breaks= scales::pretty_breaks(4)) +
     labs(title=title) +
     coefficent_plot_theme()
 }
@@ -137,7 +137,7 @@ partial_plot_density_height <- function(predictions, x, xlab=NULL,ylab=NULL, yli
 
 burnt_v_unburnt <- function(predictions, x, ylab=NULL) {
[…]

### 2016-04-11 manuscript/supplementary.Rnw
Forgot to add supplementary methods title
formatted for PNAS

@@
-\documentclass{nature}
-\bibliographystyle{naturemag.bst}
-\usepackage{lineno}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{amsmath}
-\usepackage[hidelinks]{hyperref}
+\documentclass[9pt,twocolumn,twoside, lineno]{pnas-new}
+% Use the lineno option to display guide line numbers if required.
+% Note that the use of elements such as single-column equations
+% may affect the guide line number alignment.
 
-\title{Climatic warming strengthens a positive feedback \\ between alpine shrubs and fire \\ 
-(Supplementary Material)}
-\author{James S. Camac$^{*,1,2,3}$, Richard J. Williams$^4$, Carl-Henrik Wahren$^3$, Ary A. Hoffmann$^5$ \\
-\& Peter A. Vesk$^2$}
+\templatetype{pnasresearcharticle} % Choose template 
+% {pnasresearcharticle} = Template for a two-column research article
+% {pnasmathematics} = Template for a one-column mathematics article
+% {pnasinvited} = Template for a PNAS invited submission
+
+\title{Climatic warming strengthens a positive feedback\\ between alpine shrubs and fire}
+
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\author[a,b,c,1]{James S Camac}
+\author[d,2]{Richard J Williams} 
+\author[c,3]{Carl-Henrik Wahren}
+\author[e,4]{Ary A Hoffmann}
+\author[b,5]{Peter A Vesk}
+
+\affil[a]{Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia}
+\affil[b]{The Centre of Excellence for Environmental Decisions, School of BioSciences, The University of Melbourne, Melbourne, Victoria 3010, Australia}
+\affil[c]{Research Centre for Applied Alpine Ecology, La Trobe University, Melbourne, Victoria 3086, Australia}
+\affil[d]{CSIRO Land and Water, Tropical Ecosystems Research Centre, PMB 44 Winnellie, Northern Territory 0821, Australia}
+\affil[e]{Bio21 Institute, School of BioSciences, The University of Melbourne, Victoria 3010, Australia}
+
+% Please give the surname of the lead author for the running footer
+\leadauthor{Camac} 
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: james.camac@gmail.com}
+
+% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
+\keywords{Bayesian models $|$ Biotic interactions $|$ Growth $|$ Recruitment $|$ Mortality}
+
+\dates{This manuscript was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
 
 \begin{document}
-\linenumbers
 \maketitle
 
-\begin{affiliations}
-\item Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia.
-\item The Centre of Excellence for Environmental Decisions, School of BioSciences, The University of Melbourne, Melbourne, Victoria 3010, Australia.
-\item Research Centre for Applied Alpine Ecology, La Trobe University, Melbourne, Victoria 3086, Australia.
-\item CSIRO Land and Water, Tropical Ecosystems Research Centre, PMB 44 Winnellie, Northern Territory 0821, Australia.
-\item Bio21 Institute, School of BioSciences, The University of Melbourne, Victoria 3010, Australia.
-\textasteriskcentered Email for correspondence: \texttt{james.camac@gmail.com}
-\end{affiliations}
-
-\pagebreak
-
-\section*{Study site}
+\section*{SI Materials and Methods}
+\subsection*{Study site}
 Our research sites were on the Bogong High Plains, a series of alpine and high subalpine plateaux in the Alpine National Park, approximately 250 km northeast of Melbourne, Australia.
@@
 
-\section*{Open Top Chamber Experiment}
+\subsection*{Open Top Chamber Experiment}
 We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of post-fire gap-infilling by tussock grass and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affects shrub seedling demographic rates; and 3) whether such effects change under warmer conditions post-fire conditions.
@@
 
-\subsection{\textit{Data inputs for growth, mortality and recruitment models.}}
+\subsection*{Data inputs for growth, mortality and recruitment models.}
 In each model, OTC treatment, $OTC$, was a binary (0 = Control, 1 = OTC) plot level covariate.
@@
 
-\section*{Seedling survey}
+\subsection*{Seedling survey}
 We measured seedling density and maximum heights in open heathland across the Bogong High Plains.
@@
 
-\subsection{\textit{Modelling drivers of seedling maximum heights.}}
+\subsection*{\textit{Modelling drivers of seedling maximum heights.}}
 In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation (equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C\cite{Brown:1989vx}) influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires.
@@
 All standard deviation parameters were modelled using positive half Cauchy priors with a scale of 25.
-\pagebreak
-\section*{Supplementary figures}
-\begin{suppfigure}
-\begin{minipage}{\textwidth}
+
+\bibliography{refs.bib}
+
+\begin{figure*}
+\centering
 \includegraphics{figures/figS1.pdf}
-\caption{\textbf{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (twig diameter), altitude and Topographic Wetness Index (TWI).} (\textbf{a}) Centered and standardized model coefficients (on the log scale); and effects of (\textbf{b}) burning, (\textbf{c}) fire severity, (\textbf{d}) altitude and (\textbf{e}) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires.
+\caption{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (twig diameter), altitude and Topographic Wetness Index (TWI).
+(A) Centered and standardized model coefficients (on the log scale); and effects of (B) burning, (C) fire severity, (D) altitude and (E) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires.
 All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
 \label{figs1}
-\end{minipage}
-\end{suppfigure}
-\pagebreak
+\end{figure*}
+\clearpage
 
-\begin{suppfigure}
-\begin{minipage}{\textwidth}
+\begin{figure*}
+\centering
 \includegraphics{figures/figS2.pdf}
-\caption{\textbf{Average number of \textit{Grevillea australis} recruits in warmed (red) and control (blue) conditions in 2011 and 2012. Errors bars are 95\% Bayesian Credible Intervals.}}
+\caption{Average number of \textit{Grevillea australis} recruits in warmed (red) and control (blue) conditions in 2011 and 2012. Errors bars are 95\% Bayesian Credible Intervals.}
 \label{figs2}
-\end{minipage}
-\end{suppfigure}
-\pagebreak
+\end{figure*}
 
-\begin{suppfigure}
-\begin{minipage}{\textwidth}
+\clearpage
+\begin{figure*}
+\centering
 \includegraphics{figures/figS3.pdf}
-\caption{\textbf{Observed seedling diameter growth trajectories of four dominant Australian alpine shrubs.} Red and Blue lines represent mean ($\pm$ 95\% CI) observed heights at each May census in warmed and control plots, respectively.
+\caption{Observed seedling diameter growth trajectories of four dominant Australian alpine shrubs.
+Red and Blue lines represent mean ($\pm$ 95\% CI) observed heights at each May census in warmed and control plots, respectively.
 Orange and light blue lines represent individual growth trajectories in warmed and unwarmed plots, respectively.}
 \label{figs3}
-\end{minipage}
-\end{suppfigure}
-\pagebreak
+\end{figure*}
+\clearpage
 
-\begin{suppfigure}
-\begin{minipage}{\textwidth}
+\begin{figure*}
+\centering
 \includegraphics{figures/figS4.pdf}
-\caption{\textbf{Mean maximum seedling height along fire severity (twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires.} Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses.
-(\textbf{a} \& \textbf{b}) Centered and standardized model coefficients (on the log scale); and effects of (\textbf{c} \& \textbf{d}) fire severity, (\textbf{e} \& \textbf{f}) altitude and (\textbf{g} \& \textbf{h}) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires.
+\caption{Mean maximum seedling height along fire severity (twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires.
+Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses.
+(A \& B) Centered and standardized model coefficients (on the log scale); and effects of (C \& D) fire severity, (E \& F) altitude and (G \& H) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires.
 All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
 \label{figs4}
-\end{minipage}
-\end{suppfigure}
-\pagebreak
+\end{figure*}
+\clearpage
 
-\begin{suppfigure}
-\begin{minipage}{\textwidth}
+\begin{figure*}
+\centering
 \includegraphics{figures/figS5.pdf}
-\caption{\textbf{Mean daily microclimate difference (OTC $-$ Control)}.
-(\textbf{a}) Ambient temperature, (\textbf{b}) soil temperature, (\textbf{c}) volumetric water content and (\textbf{d}) relative humidity over experimental period.
+\caption{Mean daily microclimate difference (OTC $-$ Control).
+(A) Ambient temperature, (B) soil temperature, (C) volumetric water content and (D) relative humidity over experimental period.
 Red lines = mean difference across all growing seasons (when chambers were applied); blue = no difference between OTCs and controls.
@@
 \label{figs5}
-\end{minipage}
-\end{suppfigure}
-\pagebreak
+\end{figure*}
+\clearpage
 
-\begin{suppfigure}
-\begin{minipage}{\textwidth}
+\begin{figure*}
+\centering
 \includegraphics{figures/figS6.pdf}
-\caption{\textbf{Mean minimum daily microclimate difference (OTC $-$ Control)}.
-(\textbf{a}) Ambient temperature, (\textbf{b}) soil temperature, (\textbf{c}) volumetric water content and (\textbf{d}) relative humidity over experimental period.
+\caption{Mean minimum daily microclimate difference (OTC $-$ Control).
+(A) Ambient temperature, (B) soil temperature, (C) volumetric water content and (D) relative humidity over experimental period.
 Red lines = mean difference across all growing seasons (when chambers were applied); blue = no difference between OTCs and controls.
@@
 \label{figs6}
-\end{minipage}
-\end{suppfigure}
-\pagebreak
+\end{figure*}
+\clearpage
 
-\begin{suppfigure}
-\begin{minipage}{\textwidth}
+\begin{figure*}
+\centering
 \includegraphics{figures/figS7.pdf}
-\caption{\textbf{Mean maximum daily microclimate difference (OTC $-$ Control)}.
-Ambient temperature (\textbf{a}), soil temperature (\textbf{b}), volumetric water content (\textbf{c}) and relative humidity over experimental period (\textbf{d}).
+\caption{Mean maximum daily microclimate difference (OTC $-$ Control).
+Ambient temperature (A), soil temperature (B), volumetric water content (C) and relative humidity over experimental period (D).
 Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls.
@@
 \label{figs7}
-\end{minipage}
-\end{suppfigure}
-\pagebreak
+\end{figure*}
+\clearpage
 
-\begin{suppfigure}
-\begin{minipage}{\textwidth}
+\begin{figure*}
+\centering
 \includegraphics{figures/figS8.pdf}
-\caption{\textbf{Diagram outlining positions of transplanted seedling in an OTC plot.} The inner white square represents the burnt bare ground patch used to assess heights of four dominant alpine shrub species.
+\caption{Diagram outlining positions of transplanted seedling in an OTC plot.
+The inner white square represents the burnt bare ground patch used to assess heights of four dominant alpine shrub species.
 The grey area contains \textit{Grevillea australis} seedlings randomly transplanted into burnt inter-tussock gaps.
@@
 \label{figs8}
-\end{minipage}
-\end{suppfigure}
+\end{figure*}
+\clearpage
 
-\pagebreak
-
-\bibliography{refs.bib}
 \end{document}
-\pagebreak

### 2016-04-11 manuscript/manuscript.Rnw
edit to acknowledgements and to signifiance statement

@@
 \significancestatement{Temperatures are rapidly increasing in alpine and tundra ecosystems worldwide.
-A consequence of this temperature increase has been significant increases in shrub cover and a predicted increase in the frequency or severity of fires.
-Here we use experiemental and observational data to explore whether there are antagonistic or synergistic feedbacks between temperature, fire and shrub demographic rates.
-We find that increases in temperatures, coupled with more frequent or severe fire is likely to exacerbate the rate of shrub cover increases, and consquently, increase landscape flammability.}
+A consequence of this is an increase in shrub cover and in the frequency/ severity of fires.
+Here we use experimental and observational data to explore whether there are antagonistic or synergistic feedbacks between temperature, fire and shrub demographic rates.
+We find that feedback loops are likely to be synergistic; increases in temperature and fire will likely accelerate the rate of shrub cover increase, and this in turn will increase flammability, changing the composition of the vegetation in this landscape.}
 
@@
 Daniel Falster and Richard Fitzjohn for developing software that enabled these analyses to be readily reproducible.
-Special thanks to David Bowman, Ana{\"i}s Gibert, John Morgan, Kimberley Millers, Michaela Plein and Inka Veltheim for their comments on this manuscript.}
+Special thanks to John Baumgartner, David Bowman, Manuel Esper\'{o}n-Rodr\'{i}guez, Ana{\"i}s Gibert, John Morgan, Kimberley Millers, Michaela Plein and Inka Veltheim for their comments on this manuscript.}
 

### 2016-04-11 R/figures.R
panel labels capitalized

@@ -236,7 +236,7 @@ max_ht_plots <- function(greaus_max_ht_model,asttry_max_ht_model, ylim=c(0,35)) 
                                     xlab ='Topographic wetness index', ylab ='Maximum height (cm)',ylim)
   p8 <- partial_plot_density_height(predictions_greaus$pred_ht_twi, x ='sim_twi', 
                                     xlab ='Topographic wetness index', ylab ='Maximum height (cm)',ylim)
-  plot_grid(p1,p2,p3,p4,p5,p6,p7,p8, labels=letters[1:8], ncol = 2, label_size = 7)
+  plot_grid(p1,p2,p3,p4,p5,p6,p7,p8, labels=LETTERS[1:8], ncol = 2, label_size = 7)
 }
 
 gap_dynamics_plot <- function(gap_dynamic_model) {

### 2016-04-11 manuscript/manuscript.Rnw
additional change to significance statement from vesk
Further updates to sig statement

@@
 % Please add here a significance statement to explain the relevance of your work
-\significancestatement{Temperatures are rapidly increasing in alpine and tundra ecosystems worldwide.
-A consequence of this is an increase in shrub cover and in the frequency/ severity of fires.
-Here we use experimental and observational data to explore whether there are antagonistic or synergistic feedbacks between temperature, fire and shrub demographic rates.
-We find that feedback loops are likely to be synergistic; increases in temperature and fire will likely accelerate the rate of shrub cover increase, and this in turn will increase flammability, changing the composition of the vegetation in this landscape.}
+\significancestatement{Worldwide temperatures are rapidly increasing in alpine and tundra ecosystems.
+A consequence of this is increased cover of woody species and predicted increases in the frequency and severity of fires.
+Here we use experimental and observational data to explore feedbacks between temperature, fire and shrub demographics.
+We find that a 1\textdegree C warming doubles shrub seedling growth rates, and that feedback loops are likely to be synergistic.
+In effect increases in temperature and fire activity will likely accelerate increases in shrub cover.
+This, in turn, increases landscape flammability, ultimately changing disturbance regimes and vegetation composition in alpine and tundra ecosystems.}
 
@@
 In 2011-12, we estimated seedling density for both species across 40 alpine sites and found that the abundances of \textit{Grevillea} (Fig.
-\ref{fig2}) and \textit{Asterolasia} (Supplementary Fig.
+\ref{fig2}) and \textit{Asterolasia} (Fig.
 S1) seedlings were strongly influenced by the occurrence of fire.
@@
 For both shrub species, we detected no change in seedling density along a \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m altitudinal range (equivalent to a \Sexpr{round(0.77 * (200/100),1)}\textdegree C change in mean temperature\cite{Brown:1989vx}).
-The field warming experiment (see below) indicates that this lack of elevation response in \textit{Grevillea} is due to rates of recruitment (Supplementary Fig.
+The field warming experiment (see below) indicates that this lack of elevation response in \textit{Grevillea} is due to rates of recruitment (Fig.
 S2) and mortality (see below) largely being insensitive to a \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C change in temperature.
@@
 By contrast, seedlings of the prostrate shrub, \textit{Asterolasia}, showed no difference in growth rate between warmed and control plots in any year.
-For each species, we also observed similar treatment trends with stem diameter growth (Supplementary Fig.
+For each species, we also observed similar treatment trends with stem diameter growth (Fig.
 S3).
@@
 In response to shifts in temperature, the maximum height of \textit{Grevillea} post-fire recruits was expected to decrease with elevation, whereas \textit{Asterolasia} seedlings were not expected to show this pattern.
-Our experimental predictions were verified (Supplementary Fig.
+Our experimental predictions were verified (Fig.
 S4).
@@
 \ref{fig4}), with annual mortality rates estimated to be near \Sexpr{annual_mort(non_tussock_otc_mortality_model, 'Prostanthera','otc')}\% in warmed plots and \Sexpr{annual_mort(non_tussock_otc_mortality_model, 'Prostanthera','ctl')}\% in control plots.
-This significant decrease in mortality may be a consequence of OTCs reducing the severity of spring frosts by rising minimum ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary,'CAmbient_Temp','mean_min','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary,'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C, respectively (see Supplementary Fig S5-7).
+This significant decrease in mortality may be a consequence of OTCs reducing the severity of spring frosts by rising minimum ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary,'CAmbient_Temp','mean_min','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary,'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C, respectively (Fig S5-7).
 Warming also reduced mean seedling mortality in \textit{Grevillea} and \textit{Phebalium} (Fig.
@@
 Four seedlings per species were randomly selected and transplanted into a 4\texttimes 4 square grid in the center of each burnt patch, with 14 cm between individuals and the edge of the patch dominated by resprouting tussock grass \textit{Poa hiemata}.
-To examine interactions between tussock grass and shrub seedlings we also randomly transplanted four additional \textit{Grevillea australis} into various sized inter-tussock gaps between burnt \textit{Poa hiemata} immediately surrounding the bare ground (Supplementary Fig.
+To examine interactions between tussock grass and shrub seedlings we also randomly transplanted four additional \textit{Grevillea australis} into various sized inter-tussock gaps between burnt \textit{Poa hiemata} immediately surrounding the bare ground (Fig.
 S8).
@@
 Microclimatic conditions were measured hourly using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) at four control and four OTC plots.
-Across \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} growing seasons (\Sexpr{n_days(chamber_dates, TRUE)} growing season days), OTCs simulated warmer conditions at the lower end of IPCC projections\cite{IPCC:2013vv} (Supplementary Fig.
+Across \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} growing seasons (\Sexpr{n_days(chamber_dates, TRUE)} growing season days), OTCs simulated warmer conditions at the lower end of IPCC projections\cite{IPCC:2013vv} (Fig.
 S5-7).
@@
 \\
+\\
 }

### 2016-04-11 manuscript/supplementary.Rnw
Fixed formatting of figures and other small changes

@@
 
-\subsection*{Data inputs for growth, mortality and recruitment models.}
+\subsubsection*{Data inputs for growth, mortality and recruitment models}
 In each model, OTC treatment, $OTC$, was a binary (0 = Control, 1 = OTC) plot level covariate.
@@
 
-\subsection{\textit{Data inputs into density and maximum height models.}}
+\subsubsection*{Data inputs into density and maximum height models}
 In all seedling survey models, $Burnt_j$ is a binary (0 = unburnt, 1 = burnt) site level covariate, which only affects seedling counts at unburnt sites.
@@
 
-\section*{Data analysis}
+\subsection*{Data analysis}
 In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality.
@@
 
-\subsection{\textit{Modelling shrub seedling growth rates.}}
+\subsubsection*{Modelling shrub seedling growth rates}
 We modelled the observed height of individual $i$ at May census $t$ as a random realisation from a normal distribution truncated at zero:
@@
 \begin{aligned}
-R_{i,t} &=  \alpha + \beta_1 \times OTC_{k[i]} + \beta_2 \times Gap_{i,t} + \beta_3 \times (OTC_{k[i]} \times Gap_{i,t}) + \epsilon_{plot_{k[i]}} + \epsilon_i \\
+R_{i,t} &=  \alpha + \beta_1 \times OTC_{k[i]} + \beta_2 \times Gap_{i,t} \\
+& + \beta_3 \times (OTC_{k[i]} \times Gap_{i,t}) + \epsilon_{plot_{k[i]}} + \epsilon_i \\
 \epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
@@
 
-\subsection{\textit{Modelling shrub seedling mortality rates.}}
+\subsubsection*{Modelling shrub seedling mortality rates}
 We were also interested in whether the instantaneous mortality rates of seedlings varied amongst shrub species, OTC treatment and inter-tussock gap size.
@@
 \begin{aligned}
-\lambda_{i,t} &=  \alpha + \beta_1 \times OTC_{k[i]} + \beta_2 \times Gap_{i,t} + \beta_3 \times (OTC_{k[i]} \times Gap_{i,t}) + \epsilon_{plot_{k[i]}} + \epsilon_i \\
+\lambda_{i,t} &=  \alpha + \beta_1 \times OTC_{k[i]} + \beta_2 \times Gap_{i,t} \\
+& + \beta_3 \times (OTC_{k[i]} \times Gap_{i,t}) + \epsilon_{plot_{k[i]}} + \epsilon_i \\
 \epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
@@
 
-\subsection{\textit{Modelling Grevillea recruitment rates.}}
+\subsubsection*{Modelling Grevillea recruitment rates}
 We examined whether warming influenced natural recruitment rates of \textit{Grevillea} by modelling the number of recruits per 1 m$^2$ plot $k$ in each census, $t$ as a random realisation from a negative binomial distribution:
@@
 
-\subsection{\textit{Modelling censored Poa distances.}}
+\subsubsection*{Modelling censored Poa distances}
 Approximately \Sexpr{prop_censored(cleaned_poa_distances)}\% of the $Poa$ distance data used to calculate the median inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, (i.e.
@@
 \begin{equation}\label{eq:censored_obs} 
-censored\;distance_i \sim LN(\mu_i, \sigma_{obs}), 26 \leq censored\;distance_i \leq 100.
+censored_i \sim LN(\mu_i, \sigma_{obs}), 26 \leq censored_i \leq 100.
 \end{equation}
@@
 
-\subsection{\textit{Modelling rates of inter-tussock gap infilling by Poa.}}
+\subsubsection*{Modelling rates of inter-tussock gap infilling by Poa}
 As well as modelling how OTCs affected shrub seedling growth and mortality rates, we were also interested in examining how warming influenced rates of grass infilling of inter-tussock gaps.
@@
 
-\subsection{\textit{Modelling drivers of seedling density.}}
+\subsubsection*{Modelling drivers of seedling density}
 We examined drivers of seedling density by modelling the number of seedlings per square metre plot $i$ as a random realisation from a Poisson distribution:
@@
 \begin{aligned}
-\delta_j &= \alpha_j + \beta_1 \times Alt_j + \beta_2 \times TWI_j + \beta_3 \times (1 - Burnt_j) + \beta_4 \times (Severity_j \times Burnt_j) \\
+\delta_j &= \alpha_j + \beta_1 \times Alt_j + \beta_2 \times TWI_j + \beta_3 \times (1 - Burnt_j) \\
+& + \beta_4 \times (Severity_j \times Burnt_j) \\
 \alpha_j &\sim \mathcal{N}(\mu_{site}, \sigma_{site}),
@@
 
-\subsection*{\textit{Modelling drivers of seedling maximum heights.}}
+\subsubsection*{Modelling drivers of seedling maximum heights}
 In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation (equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C\cite{Brown:1989vx}) influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires.

### 2016-09-14 R/figures.R
Mortality was as a proportion not as a percentage

@@ -93,12 +93,12 @@ poadist_plot <- function(summarised_predictions, xlab=NULL,ylab=NULL) {
 }
 
 # Predicted mortality in OTC vs CTL for non tussock seedlings
-non_tussock_mortality_plot <- function(non_tussock_mortality_model, ylab ='Annual probability of death', xlab ='Species') {
+non_tussock_mortality_plot <- function(non_tussock_mortality_model, ylab ='Annual probability of death (%)', xlab ='Species') {
   mortality <- summarise_otc_model_predictions('non_tussock_mortality',non_tussock_mortality_model)
   
   
-  ggplot(mortality, aes(x = spp,y = mean, group= treatment, colour=treatment)) + 
-    geom_pointrange(aes(ymin = `2.5%`, ymax=`97.5%`), position=position_dodge(.1)) +
+  ggplot(mortality, aes(x = spp,y = mean*100, group= treatment, colour=treatment)) + 
+    geom_pointrange(aes(ymin = `2.5%`*100, ymax=`97.5%`*100), position=position_dodge(.1)) +
     scale_colour_manual('',values = c("ctl" ="blue","otc" ="red")) +
     ylab(ylab) + 
     xlab(xlab) +

### 2016-09-14 manuscript/manuscript.Rnw
Reformat for GCB

@@
-\documentclass[9pt,twocolumn,twoside, lineno]{pnas-new}
+\documentclass[9pt,twocolumn,twoside, lineno]{gcb}
 % Use the lineno option to display guide line numbers if required.
@@
 
-\templatetype{pnasresearcharticle} % Choose template 
+\templatetype{gcbresearcharticle} % Choose template 
 % {pnasresearcharticle} = Template for a two-column research article
@@
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,b,c,1]{James S Camac}
-\author[d,2]{Richard J Williams} 
-\author[c,3]{Carl-Henrik Wahren}
-\author[e,4]{Ary A Hoffmann}
-\author[b,5]{Peter A Vesk}
+\author[a,b,c]{James S Camac}
+\author[d]{Richard J Williams} 
+\author[c]{Carl-Henrik Wahren}
+\author[e]{Ary A Hoffmann}
+\author[b]{Peter A Vesk}
 
@@
 
-% Please add here a significance statement to explain the relevance of your work
-\significancestatement{Worldwide temperatures are rapidly increasing in alpine and tundra ecosystems.
-A consequence of this is increased cover of woody species and predicted increases in the frequency and severity of fires.
-Here we use experimental and observational data to explore feedbacks between temperature, fire and shrub demographics.
-We find that a 1\textdegree C warming doubles shrub seedling growth rates, and that feedback loops are likely to be synergistic.
-In effect increases in temperature and fire activity will likely accelerate increases in shrub cover.
-This, in turn, increases landscape flammability, ultimately changing disturbance regimes and vegetation composition in alpine and tundra ecosystems.}
-
 % Please include corresponding author, author contribution and author declaration information
@@
 % Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
-\keywords{Bayesian models $|$ Biotic interactions $|$ Growth $|$ Recruitment $|$ Mortality}
+\keywords{Bayesian models, Biotic interactions, Growth, Recruitment, Mortality}
 
@@
 \dates{This manuscript was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
+\doi{\url{doi:10.1111/GCB.XXXXX}}
 
@@
 If this is the case, add \parshape=0 to the end of the list environment.
-\dropcap{A}ccurately forecasting the effects of climatic warming on vegetation dynamics requires an understanding of the mechanisms by which climate and vegetation interact.
-Most forecasting models include the direct effects of climatic conditions on species distributions, but largely ignore biotic interactions and the type, frequency and severity of disturbances that are also likely to change\cite{Guisan:2005gz,Tucker:2012dr,Giannini:2013gi}.
-Because disturbance strongly influences recruitment opportunities, and thus, composition and structure of plant communities and biomes worldwide\cite{Bond:2005eg, Turner:2010wo}, it has the capacity to exacerbate or diminish how climate affects vegetation.
+\section*{Introduction}
+Accurately forecasting the effects of climatic warming on vegetation dynamics requires an understanding of the mechanisms by which climate and vegetation interact.
+Most forecasting models include the direct effects of climatic conditions on species distributions, but largely ignore biotic interactions and the type, frequency and severity of disturbances that are also likely to change \citep{Guisan:2005gz,Tucker:2012dr,Giannini:2013gi}.
+Because disturbance strongly influences recruitment opportunities, and thus, composition and structure of plant communities and biomes worldwide \citep{Bond:2005eg, Turner:2010wo}, it has the capacity to exacerbate or diminish how climate affects vegetation.
 Seedlings are the life stage most vulnerable to climate and disturbance, and the life stage which determines the long-term persistence of a species and its capacity to establish in new areas.
-Consequently, it is imperative to understand how seedlings respond to both changing climate and disturbance regimes\cite{Walck:2011dm, Briceno:2015iv}.
+Consequently, it is imperative to understand how seedlings respond to both changing climate and disturbance regimes \citep{Walck:2011dm, Briceno:2015iv}.
 
 Field manipulative experiments have been widely used to investigate the effects of climate change on plant communities.
-These experiments have provided invaluable information on how climate change directly influences plant phenology\cite{Hoffmann:2010kf, Dorji:2013tt}, reproduction\cite{Klady:2011dw}, morphology\cite{Hudson:2009cn}, growth\cite{Hollister:2005vf}, floristic composition\cite{Elmendorf:2012kz} and biotic interactions\cite{Klanderud:2005uw}.
-However, because disturbances vary both spatially and temporally, most have focused on mature plant responses in relatively undisturbed vegetation\cite{Briceno:2015iv}.
-As such, few field climate experiments have been conducted on seedlings or vegetation in post-disturbance conditions (but see\cite{Munier:2010cg,Graae:2010ir,Camac:2015hja}).
-Consequently, little is known about how climate change influences seedling demographic rates\cite{Briceno:2015iv} and thus, vegetation recovery in post-disturbance conditions.
+These experiments have provided invaluable information on how climate change directly influences plant phenology \citep{Hoffmann:2010kf, Dorji:2013tt}, reproduction \citep{Klady:2011dw}, morphology \citep{Hudson:2009cn}, growth \citep{Hollister:2005vf}, floristic composition \citep{Elmendorf:2012kz} and biotic interactions \citep{Klanderud:2005uw}.
+However, because disturbances vary both spatially and temporally, most have focused on mature plant responses in relatively undisturbed vegetation \citep{Briceno:2015iv}.
+As such, few field climate experiments have been conducted on seedlings or vegetation in post-disturbance conditions (but see \citep{Munier:2010cg,Graae:2010ir,Camac:2015hja}).
+Consequently, little is known about how climate change influences seedling demographic rates \citep{Briceno:2015iv} and thus, vegetation recovery in post-disturbance conditions.
 
-In alpine and arctic ecosystems, warming experiments and long-term monitoring have documented significant increases in the growth and cover of woody species\cite{Sturm:2001vd,Walker:2006ht,MyersSmith:2011fp}.
-The frequency and extent of wildfires have also increased in these environments over recent decades, a trend expected to continue\cite{Westerling:2006ve, Flannigan:2009kl, Qiu:2009th, Bradstock:2014jp}.
-In alpine and tundra ecosystems, current evidence from long-term and experimental studies indicate that shrub recruitment and encroachment is highest in disturbed areas\cite{Williams:1987ua, Batllori:2009fe, Frost:2013tm}, that climatic warming is likely to increase woody species growth rates\cite{Arft:1999iz, Elmendorf:2012ba, MyersSmith:2011fp}, and that shrubs are the most flammable component of these ecosystems\cite{Williams:2006tl, Higuera:2008dw, Higuera:2009ga}.
-The combination of this evidence indicates that more frequent or severe fire may increase recruitment opportunities for woody species, which when coupled with higher growth and survival rates caused by rising temperatures, will exacerbate shrub expansion\cite{Racine:2004tn,Lantz:2013fh} and ultimately increase the likelihood of fire\cite{Higuera:2008dw, Higuera:2009ga,MyersSmith:2011fp}.
-In effect, warming could strengthen an existing feedback loop that not only has the potential to cause rapid changes in the composition and structure of alpine and arctic vegetation, but also has serious social, biodiversity and carbon sequestration consequences\cite{Mack:2011ew}.
+In alpine and arctic ecosystems, warming experiments and long-term monitoring have documented significant increases in the growth and cover of woody species \citep{Sturm:2001vd,Walker:2006ht,MyersSmith:2011fp}.
+The frequency and extent of wildfires have also increased in these environments over recent decades, a trend expected to continue \citep{Westerling:2006ve, Flannigan:2009kl, Qiu:2009th, Bradstock:2014jp}.
+In alpine and tundra ecosystems, current evidence from long-term and experimental studies indicate that shrub recruitment and encroachment is highest in disturbed areas \citep{Williams:1987ua, Batllori:2009fe, Frost:2013tm}, that climatic warming is likely to increase woody species growth rates \citep{Arft:1999iz, Elmendorf:2012ba, MyersSmith:2011fp}, and that shrubs are the most flammable component of these ecosystems \citep{Williams:2006tl, Higuera:2008dw, Higuera:2009ga}.
+The combination of this evidence indicates that more frequent or severe fire may increase recruitment opportunities for woody species, which when coupled with higher growth and survival rates caused by rising temperatures, will exacerbate shrub expansion \citep{Racine:2004tn,Lantz:2013fh} and ultimately increase the likelihood of fire \citep{Higuera:2008dw, Higuera:2009ga,MyersSmith:2011fp}.
+In effect, warming could strengthen an existing feedback loop that not only has the potential to cause rapid changes in the composition and structure of alpine and arctic vegetation, but also has serious social, biodiversity and carbon sequestration consequences \citep{Mack:2011ew}.
 
-\begin{figure}
+\begin{figure}[!h]
 \includegraphics{figures/fig1.pdf}
@@
 
-While paleoecological studies have indicated that such a feedback may exist\cite{Higuera:2008dw, Higuera:2009ga}, there is a paucity of information on multiple mechanisms that may strengthen, mitigate or break this feedback loop (Fig.
+While paleoecological studies have indicated that such a feedback may exist \citep{Higuera:2008dw, Higuera:2009ga}, there is a paucity of information on multiple mechanisms that may strengthen, mitigate or break this feedback loop (Fig.
 \ref{fig1}A).
@@
 As with tundra ecosystems, the Australian Alps have experienced rapid changes in climate.
-Since 1979, mean growing season temperatures have risen by approximately 0.4\textdegree C and annual precipitation has fallen by 6\%\cite{Wahren:2013ji}, with a consequent decline in snow pack depth\cite{SanchezBayo:2013ci}.
-These climatic changes have been correlated with a 10 to 20\% increase in shrub cover and a 25\% decline in graminoids cover\cite{Wahren:2013ji}.
-Much of the Australian Alps has also been burnt by recent wildfires, the frequency and severity of which are expected to increase\cite{Bradstock:2014jp,Williams:2014ud}.
+Since 1979, mean growing season temperatures have risen by approximately 0.4\textdegree C and annual precipitation has fallen by 6\% \citep{Wahren:2013ji}, with a consequent decline in snow pack depth \citep{SanchezBayo:2013ci}.
+These climatic changes have been correlated with a 10 to 20\% increase in shrub cover and a 25\% decline in graminoids cover \citep{Wahren:2013ji}.
+Much of the Australian Alps has also been burnt by recent wildfires, the frequency and severity of which are expected to increase \citep{Bradstock:2014jp,Williams:2014ud}.
 
 We took advantage of recent fires in alpine open heathland, a plant community that occupies ca.
-25\% of the Australian alpine landscape\cite{Williams:2014ud}.
-Under global warming, this plant community is likely to encroach upon grasslands and is itself susceptible to shrub thickening\cite{McDougall:2003gb}.
+25\% of the Australian alpine landscape \citep{Williams:2014ud}.
+Under global warming, this plant community is likely to encroach upon grasslands and is itself susceptible to shrub thickening \citep{McDougall:2003gb}.
 To determine drivers of shrub establishment and how warmer post-fire conditions affect shrub cover, we combined field observations with a warming experiment that used seedlings of four Australian evergreen obligate seeding shrubs: \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub),  \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub).
@@
 
-\section*{Results and Discussion}
+\section*{Methods}
+We investigated shrub dynamics in open heathland because it is a common and highly flammable plant community in the Australian Alps above 1600 m a.s.l \citep{Williams:2006tl}.
+It is also an ecotone between closed heathland (\textgreater 70\% shrub cover) and tussock grassland consisting of \textit{Grevillea australis} shrubs interspersed among a \textit{Poa hiemata} sward.
+Aerial photography has revealed that the plant community has acted as an invasion front of shrub expansion into grassland and has itself experienced shrub thickening by closed heathland dominants such as \textit{Prostanthera cuneata} \citep{McDougall:2003gb}.
 
-\begin{figure}
+\subsection*{Open Top Chamber Experiment}
+
+In March 2010, at 1750 m a.s.l, we burnt 32 randomly selected mature (60 cm tall and 1.5 m$^2$) \textit{Grevillea australis} shrubs in open heathland to create patches of bare ground approximately 0.7 m$^2$ surrounded by burnt tussock grass, simulating disturbance in open heathland burnt by wildfire.
+After creating the bare ground patches we collected seedlings of dominant alpine shrub species from a nearby (\textless 2 km) site of similar altitude burnt by a late 2006 wildfire.
+We collected shrub seedlings of two dominant open heathland species \textit{Grevillea australis} (Proteaceae; a tall shrub) and \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), a dominant closed heathland species \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) that typically grows on warmer aspects and a species common to both open and closed heathland \textit{Phebalium squamulosum} (Rutaceae; a tall shrub).
+All four species are killed by fire \citep{Camac:2013eb}.
+
+A total of 640 seedlings, 256 \textit{Grevillea} (half used in \textit{Poa} inter-tussock experiment\textemdash see below) and 128 for each of \textit{Asterolasia}, \textit{Prostanthera} and \textit{Phebalium} were used.
+Four seedlings per species were randomly selected and transplanted into a 4\texttimes 4 square grid in the center of each burnt patch, with 14 cm between individuals and the edge of the patch dominated by resprouting tussock grass \textit{Poa hiemata}.
+To examine interactions between tussock grass and shrub seedlings we also randomly transplanted four additional \textit{Grevillea australis} into various sized inter-tussock gaps between burnt \textit{Poa hiemata} immediately surrounding the bare ground (Fig.
+S8).
+The experimental site was fenced to prevent grazing by deer and horses.
+We detected no obvious signs of rabbit or invertebrate herbivory within our plots.
+
+To simulate near-term warmer conditions indicated by the IPCC \citep{IPCC:2013vv}, we randomly assigned Open Top Chambers (OTCs) to half (16) the plots, with the remainder treated as unwarmed controls.
+OTCs were placed over plots, ensuring all seedlings (including inter-tussock shrub seedlings) occurred within the 1.1 m$^2$ open top to minimise edge effects.
+OTCs were placed out at the start of the growing season (October) where they remained until snowfall (early June).
+This procedure was repeated for five growing seasons from May 2010 to May 2015.
+
+Microclimatic conditions were measured hourly using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) at four control and four OTC plots.
+Across \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} growing seasons (\Sexpr{n_days(chamber_dates, TRUE)} growing season days), OTCs simulated warmer conditions at the lower end of IPCC projections \citep{IPCC:2013vv} (Fig.
+S5-7).
+OTCs passively increased average ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C, and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively.
+Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C).
+Chambers only marginally decreased soil moisture by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CMoisture3to10cmBG','mean','grand_mean'))}\% and relative humidity by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CRH','mean','grand_mean'))}\%.
+
+Seedling survival, maximum height and stem diameter (nearest mm measured with Vernier calipers) were initially recorded in May 2010 and then subsequently re-measured at the end of each growing season (May-June).
+At the same time, we recorded the distance to the nearest tussock or grass seedling in each of four cardinal directions for shrub seedlings growing in inter-tussock gaps.
+Thus, changes in inter-tussock gap size could be due to either vegetative growth or seedling establishment.
+We did not measure individual characteristics (e.g.
+height and basal diameter) of surrounding tussock grass because we could not distinguish individuals, and because height varied throughout the season.
+Numbers of natural \textit{Grevillea australis} recruits establishing within the plots were also recorded for the first two seasons.
+
+\subsection*{Seedling gradient study}
+
+We used 40 long-term open heathland sites established after the 2003 fires \citep{Camac:2013eb} consisting of 30 burnt sites and 10 sites thought to be unburnt for over 70 years.
+In the summer of 2011-12, at each site, seedling density/m$^2$ was estimated using 40 quadrats, each 1 m$^2$, that were evenly distributed in groups of 10 along four 50 m transects, with 10 m between transect lines, subsampling an area of 2000 m$^2$.
+Within plots we recorded the number and maximum height of \textit{Grevillea} and \textit{Asterolasia} seedlings.
+For unburnt sites we counted the number of mature \textit{Grevillea} plants (\textgreater 0.5 m$^2$) within 5 m of each transect.
+In burnt sites, this required counting the number of skeletons (there were no living adults at any burnt site) that still persisted post-fire.
+We were unable to estimate numbers of adult \textit{Asterolasia} because this species does not have a persistent woody skeleton post-fire.
+Site level data, elevation and Topographic Wetness Index (TWI) were obtained from a 30 m resolution digital elevation model.
+Lastly, for burnt sites, fire severity was estimated by twig diameters, collected immediately after the 2003 fires \citep{Camac:2013eb}.
+
+\subsection*{Data analysis}
+
+In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality.
+For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")}.
+Data and source code for reproducing analysis and figures are available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.
+Additional information about experimental design and analysis can be found in SI Materials and Methods.
+
+\section*{Results}
+\begin{figure}[ht]
 \centering
@@
 
-\begin{figure*}
+\begin{figure*}[!h]
 \centering
@@
 As expected, pre-fire adult density positively influenced \textit{Grevillea} seedling density.
-For both shrub species, we detected no change in seedling density along a \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m altitudinal range (equivalent to a \Sexpr{round(0.77 * (200/100),1)}\textdegree C change in mean temperature\cite{Brown:1989vx}).
+For both shrub species, we detected no change in seedling density along a \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m altitudinal range (equivalent to a \Sexpr{round(0.77 * (200/100),1)}\textdegree C change in mean temperature \citep{Brown:1989vx}).
 The field warming experiment (see below) indicates that this lack of elevation response in \textit{Grevillea} is due to rates of recruitment (Fig.
@@
 \subsection*{Experimental warming effects on shrub seedling growth}
-
 As shrub seedlings were more abundant in burnt vegetation, we investigated how warmer post-fire conditions affected seedling growth and mortality rates.
@@
 We validated the \textit{Grevillea} and \textit{Asterolasia} growth responses observed in experimentally warmed plots by examining the effect of elevation on maximum heights of post-fire recruits across 30 open heathland sites burnt in 2003.
-Here, a 190 m altitudinal range is equivalent to a mean ambient temperature difference of approximately \Sexpr{round(0.77 * (190/100),1)}\textdegree C\cite{Brown:1989vx}, which is comparable to that observed between experimentally warmed and control plots (\Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C).
+Here, a 190 m altitudinal range is equivalent to a mean ambient temperature difference of approximately \Sexpr{round(0.77 * (190/100),1)}\textdegree C \citep{Brown:1989vx}, which is comparable to that observed between experimentally warmed and control plots (\Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C).
 In response to shifts in temperature, the maximum height of \textit{Grevillea} post-fire recruits was expected to decrease with elevation, whereas \textit{Asterolasia} seedlings were not expected to show this pattern.
@@
 \subsection*{Experimental warming on shrub seedling mortality}
+\begin{figure}[!h]
+\centering
+\includegraphics{figures/fig4.pdf}
+\caption{Mean ($\pm$ 95\% Bayesian Credible Interval) annual probability of death for each species growing in control (blue) and warmed (red) conditions.}
+\label{fig4}
+\end{figure}
 
@@
 
-Climate change is expected to alter biotic interactions because their strength and direction depend strongly on climatic conditions, particularly in alpine and arctic ecosystems\cite{Callaway:2002vb, Klanderud:2005uw}.
+\begin{figure}[!h]
+\centering
+\includegraphics{figures/fig5.pdf}
+\caption{Experimental warming and inter-tussock gap size effects on \textit{Grevillea australis} seedlings.
+Top rows = Growth rate effects, Second row = Mortality effects.
+(A \& C) Centered and standardized model coefficients; (B \& D) growth and mortality rate response curves along an inter-tussock gap size gradient in warmed (red) and control (blue) conditions.
+All error bars and shading are 95\% Bayesian Credible Intervals.}
+\label{fig5}
+\end{figure}
+
+Climate change is expected to alter biotic interactions because their strength and direction depend strongly on climatic conditions, particularly in alpine and arctic ecosystems \citep{Callaway:2002vb, Klanderud:2005uw}.
 Here, we assess the interactive effects of warming and grass proximity on the growth and survival of \textit{Grevillea} seedlings transplanted into various sized inter-tussock gaps (Fig.
@@
 
-\subsection*{Strengthening of the warming-shrub-fire feedback}
-
-Here we quantified several unknown interactions between fire, shrub-grass relationships and climate to extend a conceptual model of alpine shrub dynamics under climate change (Fig.
-\ref{fig1}B).
-We showed how these interactions strengthen a hypothesized feedback loop that can rapidly increase shrub cover in alpine and tundra ecosystems\cite{Higuera:2008dw}.
-The combination of rising temperatures and more frequent or severe fire creates conditions that allow shrub seedlings to establish in greater densities, and for tall shrubs, double their growth rates and potentially increase their survival.
-These demographic effects will ultimately result in shrub thickening and expansion into grasslands.
-Coupled with field\cite{Williams:2006tl}, experimental\cite{Fraser:NHT5Kx_W} and paleoecological\cite{Higuera:2008dw} evidence, which indicate that shrubs are the most flammable component of alpine and tundra ecosystems, our results suggest that flammable fuel loads will accumulate faster and cover a larger proportion of the alpine and arctic landscapes under a warmer environment.
-This further strengthens the feedback\cite{Higuera:2008dw, Higuera:2009ga} by potentially increasing the frequency or severity of fires, which then creates further recruitment opportunities for shrubs with little demographic impact of neighbouring tussock grass.
-
-\begin{figure}
-\centering
-\includegraphics{figures/fig4.pdf}
-\caption{Mean ($\pm$ 95\% Bayesian Credible Interval) annual probability of death for each species growing in control (blue) and warmed (red) conditions.}
-\label{fig4}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{figures/fig5.pdf}
-\caption{Experimental warming and inter-tussock gap size effects on \textit{Grevillea australis} seedlings.
-Top rows = Growth rate effects, Second row = Mortality effects.
-(A \& C) Centered and standardized model coefficients; (B \& D) growth and mortality rate response curves along an inter-tussock gap size gradient in warmed (red) and control (blue) conditions.
-All error bars and shading are 95\% Bayesian Credible Intervals.}
-\label{fig5}
-\end{figure}
-
-While we have addressed several unknowns associated with this feedback between climatic warming, shrubs and fire, there are others we have not addressed that may also strengthen, weaken or break this feedback.
-The most obvious mechanism that will break this cycle involves shorter fire intervals that prevent obligate seeding shrubs reaching reproductive age and thereby exhausting the seedbank.
-However, this scenario is unlikely for the majority of alpine (or tundra) landscapes, including in Australia, where current fire intervals of 50 to 100 years would need to decrease to less than 20 years\textemdash the time estimated for the species in this study to reach reproductive maturity\cite{Williams:2008ba}.
-Furthermore, if reproductive output is related to plant size\cite{Wenk:2015ko}, then climatic warming may allow obligate seeding species to reach reproductive maturity sooner, and consequently, may increase species resilience to short fire intervals.
-Nevertheless, these factors and others such as changes in snow pack\cite{Wipf:2009jh}, soil moisture\cite{Engler:2011vq} or herbivory\cite{Post:2008wz} require further research because they are likely to be altered by the interactive effects of climate and disturbance in unpredictable ways\cite{MyersSmith:2011fp}.
-
-\begin{figure}
+\begin{figure}[!h]
 \centering
@@
 
-Our findings provide mechanistic understanding as to why shrub cover has increased, often at the expense of grasslands, in many alpine and arctic ecosystems\cite{Racine:2004tn,MyersSmith:2011fp,Frost:2013tm,Lantz:2013fh}.
-But more importantly, our results provide evidence for underlying processes that could result in a warming-fire-shrub feedback that has been hypothesized in paleoecological studies\cite{Higuera:2008dw, Higuera:2009ga}.
-Based on current observations and predictions, average global temperature has already increased by 0.85\textdegree C since 1880 and is expected to rise by as much as 4.8\textdegree C by 2100\cite{IPCC:2013vv}.
-In alpine and tundra environments, temperatures\cite{ChapinIII:2005js}, shrub cover\cite{MyersSmith:2011fp} and the frequency and severity of fire\cite{Westerling:2006ve, Flannigan:2009kl, Qiu:2009th, Bradstock:2014jp} have all increased in the last few decades.
+\section*{Discussion}
+\subsection*{Strengthening of the warming-shrub-fire feedback}
+
+Here we quantified several unknown interactions between fire, shrub-grass relationships and climate to extend a conceptual model of alpine shrub dynamics under climate change (Fig.
+\ref{fig1}B).
+We showed how these interactions strengthen a hypothesized feedback loop that can rapidly increase shrub cover in alpine and tundra ecosystems \citep{Higuera:2008dw}.
+The combination of rising temperatures and more frequent or severe fire creates conditions that allow shrub seedlings to establish in greater densities, and for tall shrubs, double their growth rates and potentially increase their survival.
+These demographic effects will ultimately result in shrub thickening and expansion into grasslands.
+Coupled with field \citep{Williams:2006tl}, experimental \citep{Fraser:NHT5Kx_W} and paleoecological \citep{Higuera:2008dw} evidence, which indicate that shrubs are the most flammable component of alpine and tundra ecosystems, our results suggest that flammable fuel loads will accumulate faster and cover a larger proportion of the alpine and arctic landscapes under a warmer environment.
+This further strengthens the feedback \citep{Higuera:2008dw, Higuera:2009ga} by potentially increasing the frequency or severity of fires, which then creates further recruitment opportunities for shrubs with little demographic impact of neighbouring tussock grass.
+
+While we have addressed several unknowns associated with this feedback between climatic warming, shrubs and fire, there are others we have not addressed that may also strengthen, weaken or break this feedback.
+The most obvious mechanism that will break this cycle involves shorter fire intervals that prevent obligate seeding shrubs reaching reproductive age and thereby exhausting the seedbank.
+However, this scenario is unlikely for the majority of alpine (or tundra) landscapes, including in Australia, where current fire intervals of 50 to 100 years would need to decrease to less than 20 years\textemdash the time estimated for the species in this study to reach reproductive maturity \citep{Williams:2008ba}.
+Furthermore, if reproductive output is related to plant size \citep{Wenk:2015ko}, then climatic warming may allow obligate seeding species to reach reproductive maturity sooner, and consequently, may increase species resilience to short fire intervals.
+Nevertheless, these factors and others such as changes in snow pack \citep{Wipf:2009jh}, soil moisture \citep{Engler:2011vq} or herbivory \citep{Post:2008wz} require further research because they are likely to be altered by the interactive effects of climate and disturbance in unpredictable ways \citep{MyersSmith:2011fp}.
+
+Our findings provide mechanistic understanding as to why shrub cover has increased, often at the expense of grasslands, in many alpine and arctic ecosystems \citep{Racine:2004tn,MyersSmith:2011fp,Frost:2013tm,Lantz:2013fh}.
+But more importantly, our results provide evidence for underlying processes that could result in a warming-fire-shrub feedback that has been hypothesized in paleoecological studies \citep{Higuera:2008dw, Higuera:2009ga}.
+Based on current observations and predictions, average global temperature has already increased by 0.85\textdegree C since 1880 and is expected to rise by as much as 4.8\textdegree C by 2100 \citep{IPCC:2013vv}.
+In alpine and tundra environments, temperatures \citep{ChapinIII:2005js}, shrub cover \citep{MyersSmith:2011fp} and the frequency and severity of fire \citep{Westerling:2006ve, Flannigan:2009kl, Qiu:2009th, Bradstock:2014jp} have all increased in the last few decades.
 These changes mean that the warming-shrub-fire feedback loop identified here has already strengthened, which could cause grasslands and other non-woody communities to transition to an alternative state with more shrubs and more fire, both of which are likely to have consequences for carbon sequestration, water supply and biodiversity.
-
-\matmethods{
-We investigated shrub dynamics in open heathland because it is a common and highly flammable plant community in the Australian Alps above 1600 m a.s.l\cite{Williams:2006tl}.
-It is also an ecotone between closed heathland (\textgreater 70\% shrub cover) and tussock grassland consisting of \textit{Grevillea australis} shrubs interspersed among a \textit{Poa hiemata} sward.
-Aerial photography has revealed that the plant community has acted as an invasion front of shrub expansion into grassland and has itself experienced shrub thickening by closed heathland dominants such as \textit{Prostanthera cuneata}\cite{McDougall:2003gb}.
-
-\subsection*{Open Top Chamber Experiment}
-
-In March 2010, at 1750 m a.s.l, we burnt 32 randomly selected mature (60 cm tall and 1.5 m$^2$) \textit{Grevillea australis} shrubs in open heathland to create patches of bare ground approximately 0.7 m$^2$ surrounded by burnt tussock grass, simulating disturbance in open heathland burnt by wildfire.
-After creating the bare ground patches we collected seedlings of dominant alpine shrub species from a nearby (\textless 2 km) site of similar altitude burnt by a late 2006 wildfire.
-We collected shrub seedlings of two dominant open heathland species \textit{Grevillea australis} (Proteaceae; a tall shrub) and \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), a dominant closed heathland species \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) that typically grows on warmer aspects and a species common to both open and closed heathland \textit{Phebalium squamulosum} (Rutaceae; a tall shrub).
-All four species are killed by fire\cite{Camac:2013eb}.
-
-A total of 640 seedlings, 256 \textit{Grevillea} (half used in \textit{Poa} inter-tussock experiment\textemdash see below) and 128 for each of \textit{Asterolasia}, \textit{Prostanthera} and \textit{Phebalium} were used.
-Four seedlings per species were randomly selected and transplanted into a 4\texttimes 4 square grid in the center of each burnt patch, with 14 cm between individuals and the edge of the patch dominated by resprouting tussock grass \textit{Poa hiemata}.
-To examine interactions between tussock grass and shrub seedlings we also randomly transplanted four additional \textit{Grevillea australis} into various sized inter-tussock gaps between burnt \textit{Poa hiemata} immediately surrounding the bare ground (Fig.
-S8).
-The experimental site was fenced to prevent grazing by deer and horses.
-We detected no obvious signs of rabbit or invertebrate herbivory within our plots.
-
-To simulate near-term warmer conditions indicated by the IPCC\cite{IPCC:2013vv}, we randomly assigned Open Top Chambers (OTCs) to half (16) the plots, with the remainder treated as unwarmed controls.
-OTCs were placed over plots, ensuring all seedlings (including inter-tussock shrub seedlings) occurred within the 1.1 m$^2$ open top to minimise edge effects.
-OTCs were placed out at the start of the growing season (October) where they remained until snowfall (early June).
-This procedure was repeated for five growing seasons from May 2010 to May 2015.
-
-Microclimatic conditions were measured hourly using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) at four control and four OTC plots.
-Across \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} growing seasons (\Sexpr{n_days(chamber_dates, TRUE)} growing season days), OTCs simulated warmer conditions at the lower end of IPCC projections\cite{IPCC:2013vv} (Fig.
-S5-7).
-OTCs passively increased average ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C, and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively.
-Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C).
-Chambers only marginally decreased soil moisture by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CMoisture3to10cmBG','mean','grand_mean'))}\% and relative humidity by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CRH','mean','grand_mean'))}\%.
-
-Seedling survival, maximum height and stem diameter (nearest mm measured with Vernier calipers) were initially recorded in May 2010 and then subsequently re-measured at the end of each growing season (May-June).
-At the same time, we recorded the distance to the nearest tussock or grass seedling in each of four cardinal directions for shrub seedlings growing in inter-tussock gaps.
-Thus, changes in inter-tussock gap size could be due to either vegetative growth or seedling establishment.
-We did not measure individual characteristics (e.g.
-height and basal diameter) of surrounding tussock grass because we could not distinguish individuals, and because height varied throughout the season.
-Numbers of natural \textit{Grevillea australis} recruits establishing within the plots were also recorded for the first two seasons.
-
-\subsection*{Seedling gradient study}
-
-We used 40 long-term open heathland sites established after the 2003 fires\cite{Camac:2013eb} consisting of 30 burnt sites and 10 sites thought to be unburnt for over 70 years.
-In the summer of 2011-12, at each site, seedling density/m$^2$ was estimated using 40 quadrats, each 1 m$^2$, that were evenly distributed in groups of 10 along four 50 m transects, with 10 m between transect lines, subsampling an area of 2000 m$^2$.
-Within plots we recorded the number and maximum height of \textit{Grevillea} and \textit{Asterolasia} seedlings.
-For unburnt sites we counted the number of mature \textit{Grevillea} plants (\textgreater 0.5 m$^2$) within 5 m of each transect.
-In burnt sites, this required counting the number of skeletons (there were no living adults at any burnt site) that still persisted post-fire.
-We were unable to estimate numbers of adult \textit{Asterolasia} because this species does not have a persistent woody skeleton post-fire.
-Site level data, elevation and Topographic Wetness Index (TWI) were obtained from a 30 m resolution digital elevation model.
-Lastly, for burnt sites, fire severity was estimated by twig diameters, collected immediately after the 2003 fires\cite{Camac:2013eb}.
-
-\subsection*{Data analysis}
-
-In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality.
-For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")}.
-Data and source code for reproducing analysis and figures are available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.
-Additional information about experimental design and analysis can be found in SI Materials and Methods.
-\\
-\\
-\\
-}
-
-\showmatmethods % Display the Materials and Methods section
 

### 2016-09-14 manuscript/supplementary.Rnw
Reformat for GCB

@@
 
-\templatetype{pnasresearcharticle} % Choose template 
-% {pnasresearcharticle} = Template for a two-column research article
-% {pnasmathematics} = Template for a one-column mathematics article
-% {pnasinvited} = Template for a PNAS invited submission
+\templatetype{gcbresearcharticle} % Choose template
+
+% Change label of supplementary figures
+\renewcommand{\thefigure}{S\arabic{figure}}
 
@@
 \correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: james.camac@gmail.com}
-
-% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
-\keywords{Bayesian models $|$ Biotic interactions $|$ Growth $|$ Recruitment $|$ Mortality}
-
 \dates{This manuscript was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
+\doi{\url{doi:10.1111/GCB.XXXXX}}
 
 \begin{document}
-\maketitle
+\suppmaketitle
 
@@
 Mean annual temperature is approximately 6\textdegree C, with persistent snow cover from June to September.
-Annual precipitation is 1200 mm, yet droughts occur at roughly decadal intervals\cite{Wahren:2013ji}.
+Annual precipitation is 1200 mm, yet droughts occur at roughly decadal intervals \citep{Wahren:2013ji}.
 The vegetation is a mosaic of Eucalypt-woodland and treeless vegetation, with the latter largely consisting of heathlands and tussock grasslands, but also herbfields and wetlands.
 The treeline is irregular (due to cold air drainage), occurring at altitudes ranging from 1600 to 1700 m a.s.l.
-Approximately 50\% of the alpine ‘treeless’ landscape consists of heathlands, most of which were extensively burned by a 400,000 ha wildfire in 2003\cite{Williams:2006tl,Camac:2013eb}.
-Large wildfires in Australian alpine landscapes have occurred once or twice per century and typically occur during periods of extended regional drought coupled with severe fire weather and multiple ignitions in the surrounding montane forests\cite{Williams:2008ba}.
+Approximately 50\% of the alpine ‘treeless’ landscape consists of heathlands, most of which were extensively burned by a 400,000 ha wildfire in 2003 \citep{Williams:2006tl,Camac:2013eb}.
+Large wildfires in Australian alpine landscapes have occurred once or twice per century and typically occur during periods of extended regional drought coupled with severe fire weather and multiple ignitions in the surrounding montane forests \citep{Williams:2008ba}.
 
@@
 
-In alpine and tundra regions, shrub seedlings typically establish and grow in bare ground created by disturbance\cite{Williams:1987ua, Batllori:2009fe, Frost:2013tm}.
-Here, we used fire to create bare ground because Australian alpine shrub seedlings occur in greatest abundance after fire (see seedling survey results) and in many parts of the world, including Australia, wildfires are predicted to increase in frequency and severity\cite{Westerling:2006ve,Flannigan:2009kl}.
+In alpine and tundra regions, shrub seedlings typically establish and grow in bare ground created by disturbance \citep{Williams:1987ua, Batllori:2009fe, Frost:2013tm}.
+Here, we used fire to create bare ground because Australian alpine shrub seedlings occur in greatest abundance after fire (see seedling survey results) and in many parts of the world, including Australia, wildfires are predicted to increase in frequency and severity \citep{Westerling:2006ve,Flannigan:2009kl}.
 Thus, to forecast likely changes in shrub abundance we need to know how warming affects growth and mortality rates of shrub seedlings growing in burnt sites.
@@
 
-To simulate near-term warmer conditions indicated by the IPCC\cite{IPCC:2013vv}, we used Open Top Chambers (OTCs)\cite{Molau:1996vc}.
+To simulate near-term warmer conditions indicated by the IPCC \citep{IPCC:2013vv}, we used Open Top Chambers (OTCs) \citep{Molau:1996vc}.
 OTCs were hexagonal, 58 cm tall with an open-top diameter of 110 cm and base diameter of 168 cm.
-The chambers were constructed following the International Tundra Experiment (ITEX) protocols\cite{Molau:1996vc} using sun-lite: a lightweight shatter-resistant fiberglass with high light transmittance (86\%) and low infrared transmittance (\textless 5\%; Solar Components Corporation, USA).
+The chambers were constructed following the International Tundra Experiment (ITEX) protocols \citep{Molau:1996vc} using sun-lite: a lightweight shatter-resistant fiberglass with high light transmittance (86\%) and low infrared transmittance (\textless 5\%; Solar Components Corporation, USA).
 
@@
 Each Micro Station measured ambient temperature (10 cm above ground), relative humidity (10 cm above ground), soil temperature (3 cm below ground) and volumetric soil moisture (between 3 and 10 cm) at hourly intervals for the entire duration of the experiment.
-Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions\cite{IPCC:2013vv} (Supplementary Fig.
+Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions \citep{IPCC:2013vv} (Supplementary Fig.
 S5-7).
@@
 Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots.
-In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations\cite{Gelman:2007te}.
+In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations \citep{Gelman:2007te}.
 The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as responses to average conditions and slope terms as partial dependencies conditional on other continuous variables being at their mean.
@@
 
-The study used 40 open heathland sites\cite{Camac:2013eb} established immediately after the 2003 fires\cite{Williams:2006tl}.
+The study used 40 open heathland sites \citep{Camac:2013eb} established immediately after the 2003 fires \citep{Williams:2006tl}.
 All sites have been free of grazing by cattle since 2003.
-Thirty sites were burnt, while the remaining ten were unburnt, likely for more than 70 years\cite{Camac:2013eb}.
+Thirty sites were burnt, while the remaining ten were unburnt, likely for more than 70 years \citep{Camac:2013eb}.
 This disparity in sampling effort between burnt and unburnt was due to very few open heathland remaining unburnt after both the 2003 and 2007 fires.
@@
 Different variables were recorded at the three sampling levels (site, transect and plot).
-For each site, altitude and TWI were derived from a 30 m resolution digital elevation model\cite{Gallant:2011er} of the Bogong High Plains.
-TWI, which approximates the relative plant-available water based on a site\textquotesingle s position in the landscape\cite{Moore:1993fk}, was derived for each site using R package \texttt{dynatopmodel}\cite{Metcalfe:2016ud}.
+For each site, altitude and TWI were derived from a 30 m resolution digital elevation model \citep{Gallant:2011er} of the Bogong High Plains.
+TWI, which approximates the relative plant-available water based on a site\textquotesingle s position in the landscape \citep{Moore:1993fk}, was derived for each site using R package \texttt{dynatopmodel} \citep{Metcalfe:2016ud}.
 Across the 40 sites, TWI varied twofold (\Sexpr{min(round(complete_field_survey_data$topo_wetness_index))} to \Sexpr{round(max(complete_field_survey_data$topo_wetness_index))} units; higher units = wetter).
-Altitude ranged by \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m, or by approximately \Sexpr{round(0.77 * (190/100),1)}\textdegree C from the lowest to the highest site (based on a lapse-rate of 0.77\textdegree C/100 m elevation\cite{Brown:1989vx}.
+Altitude ranged by \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m, or by approximately \Sexpr{round(0.77 * (190/100),1)}\textdegree C from the lowest to the highest site (based on a lapse-rate of 0.77\textdegree C/100 m elevation \citep{Brown:1989vx}.
 This altitudinal temperature range is comparable to that observed in the OTC experiment.
 
-The occurrence and severity of fire was determined immediately after the 2003 fires\cite{Williams:2006tl}.
-The severity of the fires in each burnt site was estimated using mean post-fire twig diameter, a proxy measure of fire severity\cite{Whight:1999vw}.
+The occurrence and severity of fire was determined immediately after the 2003 fires \citep{Williams:2006tl}.
+The severity of the fires in each burnt site was estimated using mean post-fire twig diameter, a proxy measure of fire severity \citep{Whight:1999vw}.
 These twig diameters were recorded between 1 and 10 months post-fire from 10 replicate \textit{Grevillea australis} individuals per site.
@@
 Collinearity among covariates was checked and all correlations were \textit{r} \textless 0.16.
-The data from each continuous covariate was then centered on their mean and standardized by two standard deviations\cite{Gelman:2007te}.
+The data from each continuous covariate was then centered on their mean and standardized by two standard deviations \citep{Gelman:2007te}.
 
@@
 These models are well suited to datasets that have a hierarchical structure (e.g.
-plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset\cite{Gelman:2007te, Kery:2012wl}.
-For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")}\cite{stan:2015xx}.
+plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset \citep{Gelman:2007te, Kery:2012wl}.
+For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")} \citep{stan:2015xx}.
 For each model, three independent chains were run.
-Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic\cite{Gelman:1992ts, Brooks:1998ju}.
+Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic \citep{Gelman:1992ts, Brooks:1998ju}.
 In all cases modelled parameters converged within 1000 iterations.
@@
 where $\widehat{height}_{i,t}$ is the estimated maximum height for individual $i$ at May census $t$ and $\sigma_{obs}$ is the standard deviation of the observation error.
-The parameter $\widehat{height}_{i,t}$ was estimated using a logistic parameterisation of the unified Richards equation\cite{Tjorve:2010bk}:
+The parameter $\widehat{height}_{i,t}$ was estimated using a logistic parameterisation of the unified Richards equation \citep{Tjorve:2010bk}:
 
@@
 $\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$, $\sigma_{ind}$ and $\sigma_{obs}$) were all estimated from weakly informative positive half Cauchy priors with scale parameters of 25.
-We used these Cauchy priors in order to prevent over-inflated estimates of variances due to the small number of plots and species\cite{Gelman:2006uw}.
+We used these Cauchy priors in order to prevent over-inflated estimates of variances due to the small number of plots and species \citep{Gelman:2006uw}.
 In both sets of models the intercept terms ($\mu_{spp}^{CTL}$, $\alpha$), coefficients (i.e.
@@
 In both cases, $Cumulative\;hazard_{i,t}$, was modelled on the complementary log-log scale as a function of an instantaneous hazard rate for each individual, $\lambda_i$, multiplied by the census interval (in years) between censuses, $t$.
-By doing this our regression model becomes equivalent to a parametric survival analysis\cite{Kleinbaum:2012tn}.
+By doing this our regression model becomes equivalent to a parametric survival analysis \citep{Kleinbaum:2012tn}.
 
@@
 $\beta_1$, $\beta_2$ and $\beta_3$) and the multi-species hyperparameter for the mean OTC effect $\mu_{spp}^{OTC}$.
-These cauchy priors are the recommended default choice for binomial regression models\cite{Gelman:2008ua}.
+These cauchy priors are the recommended default choice for binomial regression models \citep{Gelman:2008ua}.
 
@@
 \subsubsection*{Modelling drivers of seedling maximum heights}
-In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation (equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C\cite{Brown:1989vx}) influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires.
+In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation (equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C \citep{Brown:1989vx}) influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires.
 We also assessed whether TWI and fire severity influenced these heights.

### 2016-09-15 manuscript/manuscript.Rnw
More edits for GCB

@@
 % may affect the guide line number alignment. 
-
+% Enable cross referencing to Evidence document
+\usepackage{xr}
+\externaldocument{supplementary}
+\pdfminorversion=4
 \templatetype{gcbresearcharticle} % Choose template 
@@
 % Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
-\keywords{Bayesian models, Biotic interactions, Growth, Recruitment, Mortality}
+\keywords{Bayesian models, Biotic interactions, Growth, Recruitment, Mortality, Gap dynamics}
 
@@
 However, because disturbances vary both spatially and temporally, most have focused on mature plant responses in relatively undisturbed vegetation \citep{Briceno:2015iv}.
-As such, few field climate experiments have been conducted on seedlings or vegetation in post-disturbance conditions (but see \citep{Munier:2010cg,Graae:2010ir,Camac:2015hja}).
+As such, few field climate experiments have been conducted on seedlings or vegetation in post-disturbance conditions \citep[but see:][]{Munier:2010cg,Graae:2010ir,Camac:2015hja}.
 Consequently, little is known about how climate change influences seedling demographic rates \citep{Briceno:2015iv} and thus, vegetation recovery in post-disturbance conditions.
@@
 The frequency and extent of wildfires have also increased in these environments over recent decades, a trend expected to continue \citep{Westerling:2006ve, Flannigan:2009kl, Qiu:2009th, Bradstock:2014jp}.
-In alpine and tundra ecosystems, current evidence from long-term and experimental studies indicate that shrub recruitment and encroachment is highest in disturbed areas \citep{Williams:1987ua, Batllori:2009fe, Frost:2013tm}, that climatic warming is likely to increase woody species growth rates \citep{Arft:1999iz, Elmendorf:2012ba, MyersSmith:2011fp}, and that shrubs are the most flammable component of these ecosystems \citep{Williams:2006tl, Higuera:2008dw, Higuera:2009ga}.
+In alpine and tundra ecosystems, current evidence from long-term and experimental studies indicate that shrub recruitment and encroachment is highest in disturbed areas \citep{Williams:1987ua, Batllori:2009fe, Frost:2013tm}, that climatic warming is likely to increase woody species growth rates \citep{Arft:1999iz, Elmendorf:2012ba, MyersSmith:2011fp}, and that shrubs are the most flammable component of these ecosystems \citep{Williams:2006tl, Higuera:2009ga, Fraser:NHT5Kx_W}.
 The combination of this evidence indicates that more frequent or severe fire may increase recruitment opportunities for woody species, which when coupled with higher growth and survival rates caused by rising temperatures, will exacerbate shrub expansion \citep{Racine:2004tn,Lantz:2013fh} and ultimately increase the likelihood of fire \citep{Higuera:2008dw, Higuera:2009ga,MyersSmith:2011fp}.
@@
 \caption{Conceptual diagram illustrating positive warming-shrub-fire feedback loop.
-(A) Hypothesised positive feedback loop between fire, climatic warming and landscape flammability.
+(a) Hypothesised positive feedback loop between fire, climatic warming and landscape flammability.
 Solid blue lines = known mechansims; Dashed black lines = mechansims that may exacerbate or diminish this feedback loop but which we have a paucity of information on.
-(B) Feedback loop based on new evidence obtained from experimental and field surveys conducted in this study.
+(b) Feedback loop based on new evidence obtained from experimental and field surveys conducted in this study.
 Also includes an example of a potential, but unlikely, effect (red dotted line) that could break this feedback.}
@@
 While paleoecological studies have indicated that such a feedback may exist \citep{Higuera:2008dw, Higuera:2009ga}, there is a paucity of information on multiple mechanisms that may strengthen, mitigate or break this feedback loop (Fig.
-\ref{fig1}A).
+\ref{fig1}a).
 For example, we have little information on whether fire will stimulate shrub seedling recruitment in alpine or tundra ecosystems, nor do we know how rates of seedling growth and survival will be affected under warmer, more exposed, post-fire conditions.
@@
 To examine interactions between tussock grass and shrub seedlings we also randomly transplanted four additional \textit{Grevillea australis} into various sized inter-tussock gaps between burnt \textit{Poa hiemata} immediately surrounding the bare ground (Fig.
-S8).
+\ref{figs8}).
 The experimental site was fenced to prevent grazing by deer and horses.
@@
 
-To simulate near-term warmer conditions indicated by the IPCC \citep{IPCC:2013vv}, we randomly assigned Open Top Chambers (OTCs) to half (16) the plots, with the remainder treated as unwarmed controls.
+To simulate near-term warmer conditions indicated by the \citet*{IPCC:2013vv}, we randomly assigned Open Top Chambers (OTCs) to half (16) the plots, with the remainder treated as unwarmed controls.
 OTCs were placed over plots, ensuring all seedlings (including inter-tussock shrub seedlings) occurred within the 1.1 m$^2$ open top to minimise edge effects.
@@
 Microclimatic conditions were measured hourly using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) at four control and four OTC plots.
-Across \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} growing seasons (\Sexpr{n_days(chamber_dates, TRUE)} growing season days), OTCs simulated warmer conditions at the lower end of IPCC projections \citep{IPCC:2013vv} (Fig.
-S5-7).
-OTCs passively increased average ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C, and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively.
+Across \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} growing seasons (\Sexpr{n_days(chamber_dates, TRUE)} growing season days), OTCs simulated warmer conditions at the lower end of \citet*{IPCC:2013vv} projections (Fig.
+\ref{figs5}-\ref{figs7}).
+OTCs passively increased average ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C.
 Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C).
@@
 We used 40 long-term open heathland sites established after the 2003 fires \citep{Camac:2013eb} consisting of 30 burnt sites and 10 sites thought to be unburnt for over 70 years.
+The disparity in sampling effort between burnt and unburnt was due to very few open heathland remaining unburnt after both the 2003 and 2007 fires.
 In the summer of 2011-12, at each site, seedling density/m$^2$ was estimated using 40 quadrats, each 1 m$^2$, that were evenly distributed in groups of 10 along four 50 m transects, with 10 m between transect lines, subsampling an area of 2000 m$^2$.
@@
 We were unable to estimate numbers of adult \textit{Asterolasia} because this species does not have a persistent woody skeleton post-fire.
-Site level data, elevation and Topographic Wetness Index (TWI) were obtained from a 30 m resolution digital elevation model.
+Site level data, elevation and Topographic Wetness Index \citep[TWI; a measure of plant available water,][]{Moore:1993fk} were obtained from a 30 m resolution digital elevation model.
 Lastly, for burnt sites, fire severity was estimated by twig diameters, collected immediately after the 2003 fires \citep{Camac:2013eb}.
@@
 Data and source code for reproducing analysis and figures are available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.
-Additional information about experimental design and analysis can be found in SI Materials and Methods.
+Additional information about experimental design and analysis can be found in Supporting Information.
 
@@
 \caption{Mean \textit{Grevillea australis} seedling density along gradients of burning, fire severity (twig diameter), altitude, Topographic Wetness Index (TWI) and adult density.
-(A) Centered and standardized model coefficients (on the log scale); and effects of: (B) burning, (C) fire severity, (D) altitude, (E) topographic wetness and (F) adult density, in areas burnt by the 2003 fires.
+(a) Centered and standardized model coefficients (on the log scale); and effects of: (b) burning, (c) fire severity, (d) altitude, (e) topographic wetness and (F) adult density, in areas burnt by the 2003 fires.
 All bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
@@
 \caption{Observed and projected growth trajectories of four dominant Australian alpine shrubs.
-(A) Observed height growth: Thick lines with error bars represent mean ($\pm$ 95\% confidence intervals) observed heights at each May census.
+(a) Observed height growth: Thick lines with error bars represent mean ($\pm$ 95\% confidence intervals) observed heights at each May census.
 Thin lines represent individual growth trajectories.
-(B) Mean ($\pm$ 95\% Bayesian Credible Interval) projected growth trajectories.
+(b) Mean ($\pm$ 95\% Bayesian Credible Interval) projected growth trajectories.
 In all cases, red and orange lines = seedlings growing in warmed (OTC) conditions and blue and light blue lines = seedlings growing in control conditions.
 Projections were based on an logistic growth model using initial seedling size of 6 cm (the mean initial height observed in the OTC experiment) and mean maximum heights observed in long-unburnt sites (i.e.
-\Sexpr{hmax_values(hmax,'Asterolasia')}, \Sexpr{hmax_values(hmax,'Grevillea')}, \Sexpr{hmax_values(hmax,'Phebalium')} \& \Sexpr{hmax_values(hmax,'Prostanthera')} cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Phebalium}, \& \textit{Prostanthera}, respectively).}
+\Sexpr{hmax_values(hmax,'Asterolasia')}, \Sexpr{hmax_values(hmax,'Grevillea')}, \Sexpr{hmax_values(hmax,'Phebalium')} \& \Sexpr{hmax_values(hmax,'Prostanthera')} cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Phebalium}, \& \textit{Prostanthera}, respectively).
+Vertical lines delimit year in which maximum height is obtained}
 \label{fig3}
@@
 \subsection*{Drivers of shrub seedling establishment}
-We first investigated how altitude, Topographic Wetness Index (TWI), adult density, fire and fire severity (as measured by post-fire twig diameters\textemdash see SI Materials and Methods) influenced the density of \textit{Grevillea} and \textit{Asterolasia} seedlings (the two dominant shrubs of alpine open heathland).
+We first investigated how altitude, Topographic Wetness Index (TWI), adult density, fire and fire severity (as measured by post-fire twig diameters\textemdash see Supporting Information) influenced the density of \textit{Grevillea} and \textit{Asterolasia} seedlings (the two dominant shrubs of alpine open heathland).
 In 2011-12, we estimated seedling density for both species across 40 alpine sites and found that the abundances of \textit{Grevillea} (Fig.
 \ref{fig2}) and \textit{Asterolasia} (Fig.
-S1) seedlings were strongly influenced by the occurrence of fire.
+\ref{figs1}) seedlings were strongly influenced by the occurrence of fire.
 Sites burnt in 2003 had seedling densities between \Sexpr{density_ratios(asttry_seedling_density_model,'Asterolasia')} and \Sexpr{density_ratios(greaus_seedling_density_model,'Grevillea')} times higher than unburnt sites.
@@
 As expected, pre-fire adult density positively influenced \textit{Grevillea} seedling density.
-For both shrub species, we detected no change in seedling density along a \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m altitudinal range (equivalent to a \Sexpr{round(0.77 * (200/100),1)}\textdegree C change in mean temperature \citep{Brown:1989vx}).
+For both shrub species, we detected no change in seedling density along a \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m altitudinal range \citep[equivalent to a \Sexpr{round(0.77 * (200/100),1)}\textdegree C change in mean temperature;][]{Brown:1989vx}.
 The field warming experiment (see below) indicates that this lack of elevation response in \textit{Grevillea} is due to rates of recruitment (Fig.
-S2) and mortality (see below) largely being insensitive to a \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C change in temperature.
+\ref{figs2}) and mortality (see below) largely being insensitive to a \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C change in temperature.
 We detected no strong effect of Topographic Wetness Index for either species.
@@
 Seedlings of \textit{Grevillea}, \textit{Asterolasia}, \textit{Phebalium} and \textit{Prostanthera} emerging after a wildfire were transplanted into experimentally burnt plots.
-These plots were either enclosed in Open Top Chambers (OTCs) whose temperature was \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C warmer than control plots subjected to current ambient conditions.
+These plots were either subjected to ambient conditions (i.e.
+controls), or enclosed in Open Top Chambers (OTCs) which rose temperature by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C.
 After \Sexpr{n_days(chamber_dates, FALSE)} days (\Sexpr{n_days(chamber_dates, TRUE)} growing season days) or \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} years' growth, mean seedling heights of the tall shrubs (\textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera}) growing in post-fire bare ground were greater in warmed plots relative to controls by \Sexpr{ht_diff(non_tussock_otc_growth_data,'Grevillea')}, \Sexpr{ht_diff(non_tussock_otc_growth_data,'Phebalium')} and \Sexpr{ht_diff(non_tussock_otc_growth_data,'Prostanthera')} cm, respectively (Fig.
-\ref{fig3}A).
+\ref{fig3}a).
 Warming increased heights of both \textit{Grevillea} and \textit{Prostanthera} seedlings in all years, while \textit{Phebalium} did not respond to the warming treatment until the second growing season.
@@
 For each species, we also observed similar treatment trends with stem diameter growth (Fig.
-S3).
+\ref{figs3}).
 Accounting for initial height and assuming logistic growth, the rates of change in mean annual predicted height of \textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera} were \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Grevillea')}, \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Phebalium')} and \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Prostanthera')} times that observed in control plots, respectively.
-According to this model, a 6 cm seedling (the mean initial height of seedlings used in this experiment) attains maximum height 39 years sooner \Sexpr{hmax(non_tussock_otc_growth_model,'Grevillea')} (\textit{Grevillea}) or \Sexpr{hmax(non_tussock_otc_growth_model,'Prostanthera')} years sooner (\textit{Phebalium} and \textit{Prostanthera}) when warmed by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C (Fig.
-\ref{fig3}B).
+According to this model, a 6 cm seedling (the mean initial height of seedlings used in this experiment) attains maximum height \Sexpr{hmax(non_tussock_otc_growth_model,'Grevillea')} (\textit{Grevillea}) or \Sexpr{hmax(non_tussock_otc_growth_model,'Prostanthera')} years sooner (\textit{Phebalium} and \textit{Prostanthera}) when warmed by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C (Fig.
+\ref{fig3}b).
 \textit{Asterolasia} was predicted to reach its maximum height in approximately \Sexpr{round(hmax(non_tussock_otc_growth_model,'Asterolasia', FALSE)[2])} to \Sexpr{round(hmax(non_tussock_otc_growth_model,'Asterolasia', FALSE)[1])} years, irrespective of warming treatment.
@@
 Our experimental predictions were verified (Fig.
-S4).
+\ref{figs4}).
 Mean maximum height of \textit{Grevillea} seedlings in burnt open heathland were \Sexpr{max_ht_alt(greaus_seedling_max_ht_model)} cm taller at 1670 m a.s.l compared to seedlings at 1860 m a.s.l.
@@
 After five years and across all plots, \Sexpr{total_deaths(non_tussock_otc_mortality_data,TRUE)}\% (\Sexpr{total_deaths(non_tussock_otc_mortality_data,FALSE)} out of 511) of all seedlings growing in burnt bare ground patches had died.
-Most deaths occurred in \textit{Asterolasia} (\Sexpr{death_counts(non_tussock_otc_mortality_data,'Asterolasia')}) followed by \textit{Phebalium} (\Sexpr{death_counts(non_tussock_otc_mortality_data, 'Phebalium')}), \textit{Grevillea} (\Sexpr{death_counts(non_tussock_otc_mortality_data,'Grevillea')}) and \textit{Prostanthera} (\Sexpr{death_counts(non_tussock_otc_mortality_data,'Prostanthera')}).
+Most deaths occurred in \textit{Asterolasia} (\Sexpr{death_counts(non_tussock_otc_mortality_data,'Asterolasia')}) and \textit{Phebalium} (\Sexpr{death_counts(non_tussock_otc_mortality_data, 'Phebalium')}), followed by \textit{Grevillea} (\Sexpr{death_counts(non_tussock_otc_mortality_data,'Grevillea')}) and \textit{Prostanthera} (\Sexpr{death_counts(non_tussock_otc_mortality_data,'Prostanthera')}).
 \textit{Prostanthera} showed the largest treatment effect (Fig.
 \ref{fig4}), with annual mortality rates estimated to be near \Sexpr{annual_mort(non_tussock_otc_mortality_model, 'Prostanthera','otc')}\% in warmed plots and \Sexpr{annual_mort(non_tussock_otc_mortality_model, 'Prostanthera','ctl')}\% in control plots.
-This significant decrease in mortality may be a consequence of OTCs reducing the severity of spring frosts by rising minimum ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary,'CAmbient_Temp','mean_min','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary,'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C, respectively (Fig S5-7).
+This significant decrease in mortality may be a consequence of OTCs reducing the severity of spring frosts by rising minimum ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary,'CAmbient_Temp','mean_min','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary,'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C, respectively (Fig.
+\ref{figs5}-\ref{figs7}).
 Warming also reduced mean seedling mortality in \textit{Grevillea} and \textit{Phebalium} (Fig.
@@
 Top rows = Growth rate effects, Second row = Mortality effects.
-(A \& C) Centered and standardized model coefficients; (B \& D) growth and mortality rate response curves along an inter-tussock gap size gradient in warmed (red) and control (blue) conditions.
+(a \& c) Centered and standardized model coefficients; (b \& d) growth and mortality rate response curves along an inter-tussock gap size gradient in warmed (red) and control (blue) conditions.
 All error bars and shading are 95\% Bayesian Credible Intervals.}
@@
 \caption{Effects of warming on rates of gap infilling by tussock grass.
-(A) Model coefficients and (B) projected temporal change in size for an average 10 cm inter-tussock gap in warmed (red) and unwarmed (blue) conditions.
+(a) Model coefficients and (b) projected temporal change in size for an average 10 cm inter-tussock gap in warmed (red) and unwarmed (blue) conditions.
 All error bars and shading are 95\% Bayesian Credible Intervals.}
@@
 Here we quantified several unknown interactions between fire, shrub-grass relationships and climate to extend a conceptual model of alpine shrub dynamics under climate change (Fig.
-\ref{fig1}B).
+\ref{fig1}b).
 We showed how these interactions strengthen a hypothesized feedback loop that can rapidly increase shrub cover in alpine and tundra ecosystems \citep{Higuera:2008dw}.

### 2016-09-15 manuscript/supplementary.Rnw
More edits for GCB

@@
 % may affect the guide line number alignment. 
-
+\pdfminorversion=4
 \templatetype{gcbresearcharticle} % Choose template
@@
 The combination of bare ground and inter-tussock seedlings meant that each experimental plot had a total of 20 seedlings per 1 m$^2$ (Fig.
-S\ref{figs8}).
+\ref{figs8}).
 However, the two sets of seedlings (i.e.
@@
 Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions \citep{IPCC:2013vv} (Supplementary Fig.
-S5-7).
+\ref{figs5}-\ref{figs7}).
 OTCs passively increased average daily ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively.
@@
 \subsubsection*{Modelling drivers of seedling maximum heights}
-In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation (equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C \citep{Brown:1989vx}) influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires.
+In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation \citep[equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C;][]{Brown:1989vx} influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires.
 We also assessed whether TWI and fire severity influenced these heights.

### 2016-09-15 R/figures.R

@@ -98,7 +98,7 @@ non_tussock_mortality_plot <- function(non_tussock_mortality_model, ylab ='Annua
   
   
   ggplot(mortality, aes(x = spp,y = mean*100, group= treatment, colour=treatment)) + 
-    geom_pointrange(aes(ymin = `2.5%`*100, ymax=`97.5%`*100), position=position_dodge(.1)) +
+    geom_pointrange(aes(ymin = `2.5%`*100, ymax=`97.5%`*100), position=position_dodge(.2), size= 0.3) +
     scale_colour_manual('',values = c("ctl" ="blue","otc" ="red")) +
     ylab(ylab) + 
     xlab(xlab) +
@@ -162,7 +162,7 @@ gap_dynamics_curve <- function(summarised_predictions, xlab='Year',ylab='Inter-t
 obs_pred_ht_growth <- function(observed_data, non_tussock_otc_growth_model) {
   p1 <- plot_obs_growth(observed_data,'ht')
   p2 <- plot_growth_curves(non_tussock_otc_growth_model)
-  print(plot_grid(p1,p2, labels=LETTERS[1:2], ncol = 1, label_size = 7),vp=viewport(layout.pos.row = 1, layout.pos.col = 1))
+  print(plot_grid(p1,p2, labels=letters[1:2], ncol = 1, label_size = 7),vp=viewport(layout.pos.row = 1, layout.pos.col = 1))
 }
   
 # Partial plots for growth and mortality against poa distance size
@@ -178,7 +178,7 @@ tussock_plots <- function(tussock_growth_model, tussock_mortality_model) {
                          xlab = 'Mortality coefficents')
   
   p4 <- poadist_plot(mortality_preds, xlab = 'Inter-tussock gap radius (cm)',ylab = 'Annual probability of death')
-  plot_grid(p1,p2,p3,p4, labels=LETTERS[1:4], ncol = 2, label_size = 7)
+  plot_grid(p1,p2,p3,p4, labels=letters[1:4], ncol = 2, label_size = 7)
 }
 
 density_count_plots <- function(density_model, species, ylim=c(0,35)) {
@@ -204,11 +204,11 @@ density_count_plots <- function(density_model, species, ylim=c(0,35)) {
   if(species=='Grevillea') {
     p6 <- partial_plot_density_height(predictions$adult_density, x ='sim_adult_den', 
[…]

### 2016-10-05 manuscript/manuscript.Rnw
Updates to manuscript

@@
 However, it is unknown whether increases in woody species and fire will interact to exacerbate changes in structure, function and composition of these ecosystems.
-Here we use field surveys and experimental manipulations to examine warming and fire effects on recruitment, growth and survival of seedlings of evergreen obligate seeding alpine shrubs.
+Here we use field surveys and experimental manipulations to examine warming and fire effects on recruitment, growth and survival of seedlings of evergreen fire-killed alpine shrubs.
 We find that fire substantially increased shrub seedling establishment (up to \Sexpr{density_ratios(greaus_seedling_density_model, 'Grevillea')}\textendash fold) and that warming doubled tall shrub seedling growth rates and could increase survival.
@@
 
-Field manipulative experiments have been widely used to investigate the effects of climate change on plant communities.
-These experiments have provided invaluable information on how climate change directly influences plant phenology \citep{Hoffmann:2010kf, Dorji:2013tt}, reproduction \citep{Klady:2011dw}, morphology \citep{Hudson:2009cn}, growth \citep{Hollister:2005vf}, floristic composition \citep{Elmendorf:2012kz} and biotic interactions \citep{Klanderud:2005uw}.
-However, because disturbances vary both spatially and temporally, most have focused on mature plant responses in relatively undisturbed vegetation \citep{Briceno:2015iv}.
+Field manipulative experiments have been widely used to investigate the effects of climate change in alpine and tundra plant communities.
+These experiments have provided invaluable information on how climate directly influences plant phenology \citep{Hoffmann:2010kf, Dorji:2013tt}, reproduction \citep{Klady:2011dw}, morphology \citep{Hudson:2009cn}, growth \citep{Hollister:2005vf}, floristic composition \citep{Elmendorf:2012kz} and biotic interactions \citep{Klanderud:2005uw}.
+However, because the general perception is that alpine and tundra vegetation is predominately clonal \citep{Billings:1968bn}, despite contradictory evidence \citep[e.g.][]{Venn:2009bw,Briceno:2015iv}, coupled with the fact that disturbances vary both spatially and temporally, most studies have focused on mature plant responses in relatively undisturbed vegetation \citep{Briceno:2015iv}.
 As such, few field climate experiments have been conducted on seedlings or vegetation in post-disturbance conditions \citep[but see:][]{Munier:2010cg,Graae:2010ir,Camac:2015hja}.
@@
 The frequency and extent of wildfires have also increased in these environments over recent decades, a trend expected to continue \citep{Westerling:2006ve, Flannigan:2009kl, Qiu:2009th, Bradstock:2014jp}.
-In alpine and tundra ecosystems, current evidence from long-term and experimental studies indicate that shrub recruitment and encroachment is highest in disturbed areas \citep{Williams:1987ua, Batllori:2009fe, Frost:2013tm}, that climatic warming is likely to increase woody species growth rates \citep{Arft:1999iz, Elmendorf:2012ba, MyersSmith:2011fp}, and that shrubs are the most flammable component of these ecosystems \citep{Williams:2006tl, Higuera:2009ga, Fraser:NHT5Kx_W}.
+In alpine and tundra ecosystems, current evidence from long-term and experimental studies indicate that shrub recruitment and encroachment is highest in disturbed areas \citep{Williams:1987ua, Batllori:2009fe, Frost:2013tm} and lowest in areas with minimal bare ground cover \citep[e.g.
+undisturbed alpine tussock grassland;][]{Williams:1987ua}.
+Evidence also indicates that shrubs can have strong supressive effects on other life forms \citep{MyersSmith:2011fp}, that climatic warming is likely to increase woody species growth rates \citep{Arft:1999iz, Elmendorf:2012ba, MyersSmith:2011fp}, and that shrubs are the most flammable component of these ecosystems \citep{Williams:2006tl, Higuera:2009ga, Fraser:NHT5Kx_W}.
 The combination of this evidence indicates that more frequent or severe fire may increase recruitment opportunities for woody species, which when coupled with higher growth and survival rates caused by rising temperatures, will exacerbate shrub expansion \citep{Racine:2004tn,Lantz:2013fh} and ultimately increase the likelihood of fire \citep{Higuera:2008dw, Higuera:2009ga,MyersSmith:2011fp}.
@@
 (a) Hypothesised positive feedback loop between fire, climatic warming and landscape flammability.
-Solid blue lines = known mechansims; Dashed black lines = mechansims that may exacerbate or diminish this feedback loop but which we have a paucity of information on.
+Solid lines = known mechansims; Dashed lines = mechansims that may exacerbate or diminish this feedback loop but which we have a paucity of information on.
 (b) Feedback loop based on new evidence obtained from experimental and field surveys conducted in this study.
-Also includes an example of a potential, but unlikely, effect (red dotted line) that could break this feedback.}
+Also includes an example of a potential, but unlikely, effect (red dashed line) that could break this feedback.
+Blue = positive relationships, Red = negative relationships.}
 \label{fig1}
@@
 25\% of the Australian alpine landscape \citep{Williams:2014ud}.
-Under global warming, this plant community is likely to encroach upon grasslands and is itself susceptible to shrub thickening \citep{McDougall:2003gb}.
-To determine drivers of shrub establishment and how warmer post-fire conditions affect shrub cover, we combined field observations with a warming experiment that used seedlings of four Australian evergreen obligate seeding shrubs: \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub),  \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub).
+Under global warming, this plant community is likely to encroach upon grasslands, wetlands and herbfields \citep[including the endangered snow patch herbfields;][]{Williams:2015gj}, and is itself susceptible to shrub thickening \citep{McDougall:2003gb}.
+To determine drivers of shrub establishment and how warmer post-fire conditions affect shrub vital rates, we combined field observations with a warming experiment that used seedlings of four dominant Australian evergreen fire-killed shrubs: \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub),  \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub).
 This combination of data sources allowed us to examine: 1) landscape scale drivers of shrub seedling establishment; 2) how warmer post-fire conditions influence rates of grass recovery and shrub seedling recruitment, growth and mortality; 3) how proximity to grass affect shrub seedling demographic rates; and 4) whether such grass effects change under warmer post-fire conditions.
@@
 We collected shrub seedlings of two dominant open heathland species \textit{Grevillea australis} (Proteaceae; a tall shrub) and \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), a dominant closed heathland species \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) that typically grows on warmer aspects and a species common to both open and closed heathland \textit{Phebalium squamulosum} (Rutaceae; a tall shrub).
-All four species are killed by fire \citep{Camac:2013eb}.
+We focused on these four species because they are all fire-killed \citep[the dominant shrub post-fire strategy in the Australian mainland Alps;][]{Walsh:2004vk}; common in the Australian Alps; and under climatic warming, are likely candidates to invading non-shrubby plant communities such as alpine grasslands and herbfields.
 
@@
 
-\section*{Results}
-\begin{figure}[ht]
+\begin{figure}[!htb]
 \centering
@@
 
-\begin{figure*}[!h]
+\begin{figure*}[!htb]
 \centering
@@
 
+\section*{Results}
 \subsection*{Drivers of shrub seedling establishment}
+
 We first investigated how altitude, Topographic Wetness Index (TWI), adult density, fire and fire severity (as measured by post-fire twig diameters\textemdash see Supporting Information) influenced the density of \textit{Grevillea} and \textit{Asterolasia} seedlings (the two dominant shrubs of alpine open heathland).
@@
 \subsection*{Experimental warming on shrub seedling mortality}
-\begin{figure}[!h]
-\centering
-\includegraphics{figures/fig4.pdf}
-\caption{Mean ($\pm$ 95\% Bayesian Credible Interval) annual probability of death for each species growing in control (blue) and warmed (red) conditions.}
-\label{fig4}
-\end{figure}
 
@@
 
+\begin{figure}[!h]
+\centering
+\includegraphics{figures/fig4.pdf}
+\caption{Mean ($\pm$ 95\% Bayesian Credible Interval) annual probability of death for each species growing in control (blue) and warmed (red) conditions.}
+\label{fig4}
+\end{figure}
+
 \subsection*{Effects of tussock grass proximity on shrub seedling growth and mortality rates}
+
+Climate change is expected to alter biotic interactions because their strength and direction depend strongly on climatic conditions, particularly in alpine and arctic ecosystems \citep{Callaway:2002vb, Klanderud:2005uw}.
+Here, we assess the interactive effects of warming and grass proximity on the growth and survival of \textit{Grevillea} seedlings transplanted into various sized inter-tussock gaps (Fig.
+\ref{fig5}).
+We detected a strong positive effect of warming treatment on growth rates and a marginal, yet not significant, decrease in mortality.
+However, we did not detect significant inter-tussock gap size effects or an interaction between gap size and warming treatment on either growth or mortality rates (i.e.
+coefficient credible intervals overlap zero).
 
@@
 \end{figure}
-
-Climate change is expected to alter biotic interactions because their strength and direction depend strongly on climatic conditions, particularly in alpine and arctic ecosystems \citep{Callaway:2002vb, Klanderud:2005uw}.
-Here, we assess the interactive effects of warming and grass proximity on the growth and survival of \textit{Grevillea} seedlings transplanted into various sized inter-tussock gaps (Fig.
-\ref{fig5}).
-We detected a strong positive effect of warming treatment on growth rates and a marginal, yet not significant, decrease in mortality.
-However, we did not detect significant inter-tussock gap size effects or an interaction between gap size and warming treatment on either growth or mortality rates (i.e.
-coefficient credible intervals overlap zero).
 
@@
 We showed how these interactions strengthen a hypothesized feedback loop that can rapidly increase shrub cover in alpine and tundra ecosystems \citep{Higuera:2008dw}.
-The combination of rising temperatures and more frequent or severe fire creates conditions that allow shrub seedlings to establish in greater densities, and for tall shrubs, double their growth rates and potentially increase their survival.
-These demographic effects will ultimately result in shrub thickening and expansion into grasslands.
-Coupled with field \citep{Williams:2006tl}, experimental \citep{Fraser:NHT5Kx_W} and paleoecological \citep{Higuera:2008dw} evidence, which indicate that shrubs are the most flammable component of alpine and tundra ecosystems, our results suggest that flammable fuel loads will accumulate faster and cover a larger proportion of the alpine and arctic landscapes under a warmer environment.
-This further strengthens the feedback \citep{Higuera:2008dw, Higuera:2009ga} by potentially increasing the frequency or severity of fires, which then creates further recruitment opportunities for shrubs with little demographic impact of neighbouring tussock grass.
+The combination of rising temperatures and more frequent or severe fire creates bare ground required for shrub seedlings to establish in greater densities, and for tall shrubs, double their growth rates and potentially increase their survival.
+These demographic effects will ultimately manifest in shrub thickening within and at heathland boundaries, which in turn, will facilitate shrub establishment into shrub-free communities by decreasing dispersal distances.
+Coupled with slow tussock grass gap infilling and, field \citep{Williams:2006tl}, experimental \citep{Fraser:NHT5Kx_W} and paleoecological \citep{Higuera:2008dw} evidence which indicate that shrubs are the most flammable component of alpine and tundra ecosystems  \citep[particularly moreso then tussock grassland; ][]{Fraser:NHT5Kx_W}, our results suggest that flammable fuel loads will accumulate faster and cover a larger proportion of the alpine and arctic landscapes under a warmer environment.
+This further strengthens the feedback \citep{Higuera:2008dw, Higuera:2009ga} by increasing the frequency or severity of fires, which then creates more shrub recruitment opportunities both within and beyond heathland boundaries (i.e.
+via more fire-induced tussock death or increased exposure of inter-tussock gaps) with little demographic impact of neighbouring tussock grass.
 
 While we have addressed several unknowns associated with this feedback between climatic warming, shrubs and fire, there are others we have not addressed that may also strengthen, weaken or break this feedback.
-The most obvious mechanism that will break this cycle involves shorter fire intervals that prevent obligate seeding shrubs reaching reproductive age and thereby exhausting the seedbank.
+The most obvious mechanism that will break this cycle involves shorter fire intervals that prevent fire-killed shrubs reaching reproductive age and thereby exhausting the seedbank.
 However, this scenario is unlikely for the majority of alpine (or tundra) landscapes, including in Australia, where current fire intervals of 50 to 100 years would need to decrease to less than 20 years\textemdash the time estimated for the species in this study to reach reproductive maturity \citep{Williams:2008ba}.
-Furthermore, if reproductive output is related to plant size \citep{Wenk:2015ko}, then climatic warming may allow obligate seeding species to reach reproductive maturity sooner, and consequently, may increase species resilience to short fire intervals.
+Furthermore, if reproductive output is related to plant size \citep{Wenk:2015ko}, then climatic warming may allow fire-killed species to reach reproductive maturity sooner, and consequently, may increase species resilience to short fire intervals.
 Nevertheless, these factors and others such as changes in snow pack \citep{Wipf:2009jh}, soil moisture \citep{Engler:2011vq} or herbivory \citep{Post:2008wz} require further research because they are likely to be altered by the interactive effects of climate and disturbance in unpredictable ways \citep{MyersSmith:2011fp}.
 
-Our findings provide mechanistic understanding as to why shrub cover has increased, often at the expense of grasslands, in many alpine and arctic ecosystems \citep{Racine:2004tn,MyersSmith:2011fp,Frost:2013tm,Lantz:2013fh}.
+By focusing on seedlings, the life stage most vulnerable to climate and disturbance, and which determines the long-term persistence of a species and its capacity to establish in new areas \citep{Walck:2011dm, Briceno:2015iv}, our findings provide mechanistic understanding as to why shrub cover has increased, often at the expense of grasslands, in many alpine and arctic ecosystems \citep{Racine:2004tn,MyersSmith:2011fp,Frost:2013tm,Lantz:2013fh}.
 But more importantly, our results provide evidence for underlying processes that could result in a warming-fire-shrub feedback that has been hypothesized in paleoecological studies \citep{Higuera:2008dw, Higuera:2009ga}.

### 2016-10-05 manuscript/supplementary.Rnw
Updates to manuscript

@@
 We collected seedlings from a site burnt by a late 2006 fire because it allowed us to collect sufficient numbers of seedlings from species occurring in both open and closed heathland.
-We focused on \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) because they are all: obligate seeders post-fire (i.e.
-fire killed); common in the Australian Alps; and under climatic warming, are likely to invading non-shrubby plant communities such as alpine grasslands and herbfields.
+We focused on \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) because they are all fire-killed (the dominant shrub post-fire strategy in the Australian mainland Alps; \citet{Walsh:2004vk}); common in the Australian Alps; and under climatic warming, are likely candidates to invading non-shrubby plant communities such as alpine grasslands and herbfields.
+Resprouter shrub species were not examined in this study because we were unable to locate sufficient numbers of seedlings; and are unlikely to be the primary contributors to shrub expansion in these landscapes.
 For each species, we collected seedlings with a range of heights (\Sexpr{range_init_size(complete_otc_seedling_data)[1]} to \Sexpr{range_init_size(complete_otc_seedling_data)[2]} cm) using a forestry planter (12.5 cm deep by 5.5 cm wide).
@@
 Four seedlings of each species were then transplanted into 70 cm$^2$ of burnt bare ground in the center of each experimental plot.
-An additional four \textit{Grevillea} seedlings were transplanted into naturally occuring inter-tussock gaps of varying size surrounding this bare ground.
+An additional four \textit{Grevillea} seedlings were transplanted into occuring post-fire inter-tussock gaps of varying size surrounding this bare ground.
 Only \textit{Grevillea} seedlings were transplanted into inter-tussock gaps because of space constraints but also because it is the dominant open heathland shrub species (i.e.

### 2016-10-18 manuscript/manuscript.Rnw
Updates to manuscript

@@
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,b,c]{James S Camac}
-\author[d]{Richard J Williams} 
-\author[c]{Carl-Henrik Wahren}
-\author[e]{Ary A Hoffmann}
-\author[b]{Peter A Vesk}
+\author[a,b]{James S Camac}
+\author[c]{Richard J Williams} 
+\author[b]{Carl-Henrik Wahren}
+\author[d]{Ary A Hoffmann}
+\author[a]{Peter A Vesk}
 
-\affil[a]{Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia}
-\affil[b]{The Centre of Excellence for Environmental Decisions, School of BioSciences, The University of Melbourne, Melbourne, Victoria 3010, Australia}
-\affil[c]{Research Centre for Applied Alpine Ecology, La Trobe University, Melbourne, Victoria 3086, Australia}
-\affil[d]{CSIRO Land and Water, Tropical Ecosystems Research Centre, PMB 44 Winnellie, Northern Territory 0821, Australia}
-\affil[e]{Bio21 Institute, School of BioSciences, The University of Melbourne, Victoria 3010, Australia}
+\affil[a]{The Centre of Excellence for Environmental Decisions, School of BioSciences, The University of Melbourne, Melbourne, Victoria 3010, Australia}
+\affil[b]{Research Centre for Applied Alpine Ecology, La Trobe University, Melbourne, Victoria 3086, Australia}
+\affil[c]{CSIRO Land and Water, Tropical Ecosystems Research Centre, PMB 44 Winnellie, Northern Territory 0821, Australia}
+\affil[d]{Bio21 Institute, School of BioSciences, The University of Melbourne, Victoria 3010, Australia}
 
@@
 Climate change is expected to increase fire activity and woody plant encroachment in arctic and alpine landscapes.
-However, it is unknown whether increases in woody species and fire will interact to exacerbate changes in structure, function and composition of these ecosystems.
-Here we use field surveys and experimental manipulations to examine warming and fire effects on recruitment, growth and survival of seedlings of evergreen fire-killed alpine shrubs.
-We find that fire substantially increased shrub seedling establishment (up to \Sexpr{density_ratios(greaus_seedling_density_model, 'Grevillea')}\textendash fold) and that warming doubled tall shrub seedling growth rates and could increase survival.
-Warming had limited or no effect on shrub recruitment, post-fire gap-infilling by grass, or competitive effects of grass on shrub seedling growth and survival.
-These findings indicate that rising temperatures coupled with more frequent or severe fires will likely exacerbate increases in tall evergreen shrubs by increasing recruitment, doubling growth rates, and potentially increasing shrub survival.
+However, the extent to which these increases interact to affect changes in structure, function and composition of alpine ecosystems is unknown.
+Here we use field surveys and experimental manipulations to examine warming and fire effects on recruitment, growth and survival of seedlings of four dominant, evergreen fire-killed alpine shrubs.
+We find that fire substantially increased shrub seedling establishment (up to \Sexpr{density_ratios(greaus_seedling_density_model, 'Grevillea')}\textendash fold) and that experimental warming doubled tall shrub seedling growth rates and could potentially increase survival.
+Warming had no effect on shrub recruitment, post-fire infilling of gaps by a dominant alpine grass, or competitive effects of grass on shrub seedling growth and survival.
+These findings indicate that warming, coupled with more frequent or severe fires, will likely exacerbate increases in tall evergreen shrubs by increasing recruitment, doubling growth rates, and potentially increasing shrub survival.
+Thus, warming is likely to enhance a positive feedback between woody species abundance and fire activity in alpine and tundra environments.
 \end{abstract}
@@
 Accurately forecasting the effects of climatic warming on vegetation dynamics requires an understanding of the mechanisms by which climate and vegetation interact.
-Most forecasting models include the direct effects of climatic conditions on species distributions, but largely ignore biotic interactions and the type, frequency and severity of disturbances that are also likely to change \citep{Guisan:2005gz,Tucker:2012dr,Giannini:2013gi}.
-Because disturbance strongly influences recruitment opportunities, and thus, composition and structure of plant communities and biomes worldwide \citep{Bond:2005eg, Turner:2010wo}, it has the capacity to exacerbate or diminish how climate affects vegetation.
-Seedlings are the life stage most vulnerable to climate and disturbance, and the life stage which determines the long-term persistence of a species and its capacity to establish in new areas.
-Consequently, it is imperative to understand how seedlings respond to both changing climate and disturbance regimes \citep{Walck:2011dm, Briceno:2015iv}.
+Most forecasting models include the direct effects of climatic conditions on species distributions, but largely ignore disturbances, the type, frequency and severity of which are also likely to change \citep{Guisan:2005gz,Tucker:2012dr}.
+Recurrent disturbance is integral to all ecosystems, strongly influencing fundamental demographic processes such as recruitment and mortality, and thus the composition and structure of plant communities and biomes worldwide \citep{Bond:2005eg, Turner:2010wo}.
+Disturbance regimes are likely to change as a consequence of climate change \citep{Turner:2010wo, Westerling:2011jp,Bradstock:2014jp}.
+Moreover, the effects of disturbance will interact with climate change, and thus may amplify or diminish how changes to temperature and moisture affect vegetation \citep{Post:2008wz,Camac:2015hja,Enright:2015em}
 
-Field manipulative experiments have been widely used to investigate the effects of climate change in alpine and tundra plant communities.
+Feedbacks between climate change, vegetation and fire will be particularly important determinants of the trajectory of ecosystem change in coming decades \citep{Bowman:2009kp}.
+For example, in addition to the direct effects of warming on fire weather, changing climate may also alter fire regimes by altering vegetation productivity, structure and plant trait composition \citep{Keeley:2012wo,Matthews:2012gia,Bowman:2014dr}.
+Feedbacks may be positive or negative, and have been documented in tropical savanna \citep{Hoffmann:2003br}, boreal forests \citep{Krawchuk:2011wr}, Mediterranean ecosystems \citep{Keeley:2012wo} and grasslands \citep{Flannigan:2009kl}.
+Change in both vegetation and fire regimes in alpine and arctic environments are also likely to be subject to feedbacks \citep{Goetz:2007gj,Wookey:2009wo}.
+
+Alpine and arctic vegetation are particularly vulnerable to the effects of changing climate \citep{Engler:2011vq,Dullinger:2012cg, Elmendorf:2012kz}.
+Field manipulative experiments have been widely used to investigate the potential effects of climate change in alpine and tundra plant communities.
 These experiments have provided invaluable information on how climate directly influences plant phenology \citep{Hoffmann:2010kf, Dorji:2013tt}, reproduction \citep{Klady:2011dw}, morphology \citep{Hudson:2009cn}, growth \citep{Hollister:2005vf}, floristic composition \citep{Elmendorf:2012kz} and biotic interactions \citep{Klanderud:2005uw}.
-However, because the general perception is that alpine and tundra vegetation is predominately clonal \citep{Billings:1968bn}, despite contradictory evidence \citep[e.g.][]{Venn:2009bw,Briceno:2015iv}, coupled with the fact that disturbances vary both spatially and temporally, most studies have focused on mature plant responses in relatively undisturbed vegetation \citep{Briceno:2015iv}.
-As such, few field climate experiments have been conducted on seedlings or vegetation in post-disturbance conditions \citep[but see:][]{Munier:2010cg,Graae:2010ir,Camac:2015hja}.
-Consequently, little is known about how climate change influences seedling demographic rates \citep{Briceno:2015iv} and thus, vegetation recovery in post-disturbance conditions.
+
+However, most studies have focused on plant responses in mature, undisturbed vegetation \citep{Briceno:2015iv}.
+Few climate-change field experiments have included disturbance as a factor \citep[but see:][]{Munier:2010cg,Graae:2010ir,Camac:2015hja}, or examined the influence of climate change on seedlings in post-disturbance environments.
+Although alpine and tundra vegetation has a strong capacity to regenerate vegetatively \citep{Billings:1968bn} there is mounting evidence that seedling regeneration is important in these ecosystems \citep{Venn:2009bw, Briceno:2015iv}, particularly for wood species \citep{Camac:2013eb,Williams:2014ud}.
+Although seedlings are highly vulnerable to climate and disturbance, they are the life stage which determines the long-term persistence of a species and its capacity to establish in new areas.
+However, little is known about how climate change affects seedling demography \citep{Briceno:2015iv}, particularly in post-disturbance conditions in alpine and tundra ecosystems.
+It is therefore imperative we understand how seedlings respond to both changing climate and disturbance regimes \citep{Walck:2011dm, Briceno:2015iv} if we wish to accurately predict future trajectories of vegetation change.
 
@@
 The frequency and extent of wildfires have also increased in these environments over recent decades, a trend expected to continue \citep{Westerling:2006ve, Flannigan:2009kl, Qiu:2009th, Bradstock:2014jp}.
-In alpine and tundra ecosystems, current evidence from long-term and experimental studies indicate that shrub recruitment and encroachment is highest in disturbed areas \citep{Williams:1987ua, Batllori:2009fe, Frost:2013tm} and lowest in areas with minimal bare ground cover \citep[e.g.
-undisturbed alpine tussock grassland;][]{Williams:1987ua}.
-Evidence also indicates that shrubs can have strong supressive effects on other life forms \citep{MyersSmith:2011fp}, that climatic warming is likely to increase woody species growth rates \citep{Arft:1999iz, Elmendorf:2012ba, MyersSmith:2011fp}, and that shrubs are the most flammable component of these ecosystems \citep{Williams:2006tl, Higuera:2009ga, Fraser:NHT5Kx_W}.
-The combination of this evidence indicates that more frequent or severe fire may increase recruitment opportunities for woody species, which when coupled with higher growth and survival rates caused by rising temperatures, will exacerbate shrub expansion \citep{Racine:2004tn,Lantz:2013fh} and ultimately increase the likelihood of fire \citep{Higuera:2008dw, Higuera:2009ga,MyersSmith:2011fp}.
-In effect, warming could strengthen an existing feedback loop that not only has the potential to cause rapid changes in the composition and structure of alpine and arctic vegetation, but also has serious social, biodiversity and carbon sequestration consequences \citep{Mack:2011ew}.
+In alpine and tundra ecosystems, current evidence from long-term and experimental studies indicate that shrub recruitment and encroachment is highest in disturbed areas \citep{Williams:1987ws, Batllori:2009fe, Frost:2013tm} and lowest in areas with minimal bare ground cover \citep[e.g.
+undisturbed alpine tussock grassland;][]{Williams:1987ws}.
+Current evidence indicates that climatic warming is likely to increase woody species growth rates \citep{Arft:1999iz, Elmendorf:2012ba, MyersSmith:2011fp} and that shrubs are the most flammable vegetation component of these ecosystems \citep{Williams:2006tl, Higuera:2009ga, Fraser:NHT5Kx_W}.
+Thus, increases in the cover of woody species in alpine and arctic environments is likely to increase the flammability of these ecosystems.
+
+The combination of this evidence indicates a potential positive feedback between warmer temperatures, woody species dynamics, and fire activity in alpine and tundra environments (Fig.
+\ref{fig1}a).
+Warmer temperatures are likely to lead to more frequent or severe fire.
+This is likely to increase woody species recruitment opportunities (i.e.
+more bare ground) and exacerbate both shrub thickening within shrub communities and shrub expansion into shrub-free communities such as grassland and herbfields \citep{Racine:2004tn,Lantz:2013fh}.
+If this effect is coupled with increased shurb seedling growth and mortality, highly flammable fuels will accumulate faster and have higher landscape connectivity, ultimately leading to further increases in the likelihood of fire \citep{Higuera:2008dw, Higuera:2009ga,MyersSmith:2011fp}.
+Thus, warming could strengthen an existing climate-disturbance feedback, that not only has the potential to cause rapid changes in the composition and structure of alpine and arctic vegetation, but also has serious social, biodiversity and carbon sequestration consequences \citep{Mack:2011ew}.
 
@@
 \includegraphics{figures/fig1.pdf}
-\caption{Conceptual diagram illustrating positive warming-shrub-fire feedback loop.
+\caption{Conceptual diagram illustrating positive warming-shrub-fire feedback.
 (a) Hypothesised positive feedback loop between fire, climatic warming and landscape flammability.
-Solid lines = known mechansims; Dashed lines = mechansims that may exacerbate or diminish this feedback loop but which we have a paucity of information on.
+Solid lines = known mechansims; Dashed lines = mechansims that may exacerbate or diminish this feedback but which we have a paucity of information on.
 (b) Feedback loop based on new evidence obtained from experimental and field surveys conducted in this study.
@@
 
-While paleoecological studies have indicated that such a feedback may exist \citep{Higuera:2008dw, Higuera:2009ga}, there is a paucity of information on multiple mechanisms that may strengthen, mitigate or break this feedback loop (Fig.
-\ref{fig1}a).
+While paleoecological studies have indicated that such a feedback may exist in the arctic \citep{Higuera:2008dw, Higuera:2009ga}, there is a paucity of information on what may strengthen, mitigate or break this feedback.
 For example, we have little information on whether fire will stimulate shrub seedling recruitment in alpine or tundra ecosystems, nor do we know how rates of seedling growth and survival will be affected under warmer, more exposed, post-fire conditions.
-Furthermore, we do not know how tussock grasslands (the other dominant community in these ecosystems) will respond to warmer post-fire conditions, whether grasses have competitive, facilitative or no effect on shrub seedling vital rates, or whether such effects are altered by warmer post-fire conditions.
+Furthermore, we do not know how tussock grasslands (a dominant structural vegetation type in arctic and alpine ecosystems) will respond to warmer post-fire conditions, whether grasses have competitive, facilitative or no effect on shrub seedling vital rates, or whether such effects are altered by warmer post-fire conditions.
 
-We examine these unknowns and their impacts on the feedback between climatic warming, shrubs and fire in Australian alpine vegetation.
-As with tundra ecosystems, the Australian Alps have experienced rapid changes in climate.
+Here, we examine these unknowns and their impacts on this proposed feedback between climatic warming, shrubs and fire in Australian alpine vegetation.
+As with other tundra ecosystems, Australian alpine landscapes consist of a matrix of grass-herb dominanted vegetation types, and shrub-dominanted types \citep{Williams:2014ud}.
+They have also experienced significant changes in climate.
 Since 1979, mean growing season temperatures have risen by approximately 0.4\textdegree C and annual precipitation has fallen by 6\% \citep{Wahren:2013ji}, with a consequent decline in snow pack depth \citep{SanchezBayo:2013ci}.
 These climatic changes have been correlated with a 10 to 20\% increase in shrub cover and a 25\% decline in graminoids cover \citep{Wahren:2013ji}.
-Much of the Australian Alps has also been burnt by recent wildfires, the frequency and severity of which are expected to increase \citep{Bradstock:2014jp,Williams:2014ud}.
+Much of the Australian Alps has also been burnt by recent (2003 and 2006) wildfires, the frequency and/or severity of which are expected to increase as a consequence of climate change \citep{Bradstock:2014jp,Williams:2014ud}.
 
@@
 25\% of the Australian alpine landscape \citep{Williams:2014ud}.
-Under global warming, this plant community is likely to encroach upon grasslands, wetlands and herbfields \citep[including the endangered snow patch herbfields;][]{Williams:2015gj}, and is itself susceptible to shrub thickening \citep{McDougall:2003gb}.
-To determine drivers of shrub establishment and how warmer post-fire conditions affect shrub vital rates, we combined field observations with a warming experiment that used seedlings of four dominant Australian evergreen fire-killed shrubs: \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub),  \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub).
-This combination of data sources allowed us to examine: 1) landscape scale drivers of shrub seedling establishment; 2) how warmer post-fire conditions influence rates of grass recovery and shrub seedling recruitment, growth and mortality; 3) how proximity to grass affect shrub seedling demographic rates; and 4) whether such grass effects change under warmer post-fire conditions.
+Under global warming, this plant community is likely to encroach upon grasslands, wetlands and herbfields \citep[including the nationally endangered snow patch herbfields;][]{Williams:2015gj}, and is itself susceptible to shrub thickening \citep{McDougall:2003gb}.
+To determine drivers of shrub establishment and how warmer post-fire conditions affect shrub vital rates, we combined field observations of post-fire shrub seedling density with a warming experiment that used seedlings of four dominant Australian evergreen fire-killed shrubs: \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub),  \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub).
+
+This combination of data sources allowed us to quantify some of the outstanding unknowns in the hypothesised climate-fire-shrub feedback loop (Fig. \ref{fig1}a). Specifically, we examined: 
+\begin{enumerate}\itemsep 0pt
+  \item Landscape scale drivers of shrub seedling establishment.
+  \item How warmer post-fire conditions influence rates of grass recovery and shrub seedling recruitment, growth and mortality.
+  \item How proximity to grass affects shrub seedling demographic rates; and whether such grass effects change under warmer post-fire conditions.
+\end{enumerate}
 
 \section*{Methods}
-We investigated shrub dynamics in open heathland because it is a common and highly flammable plant community in the Australian Alps above 1600 m a.s.l \citep{Williams:2006tl}.
-It is also an ecotone between closed heathland (\textgreater 70\% shrub cover) and tussock grassland consisting of \textit{Grevillea australis} shrubs interspersed among a \textit{Poa hiemata} sward.
-Aerial photography has revealed that the plant community has acted as an invasion front of shrub expansion into grassland and has itself experienced shrub thickening by closed heathland dominants such as \textit{Prostanthera cuneata} \citep{McDougall:2003gb}.
+
+A summary of study sites, experimental design and approach to data analysis is given below. Full details are presented in the Supporting Information.
+
+\subsection*{Study Sites}
+
+The Australian Alps are an ideal place to examine potential fire-climate-vegetation feedbacks in high mountain environments because Australian alpine ecosystems have all the elements needed to demonstrate such potential feedbacks.
+Here, the Alps are subject to recurrent landscape fire, and were burnt extensively in 2003 \citep{Williams:2006tl,Williams:2014ud}.
+The vegetation consists of a range of life forms (graminoids, other herbs, shrubs), and variation in life-form dominance between and within vegetation types is a strong determinant of landscape flammability \citep{Williams:2006tl,Fraser:NHT5Kx_W}.
+There is also evidence that both fire regimes and the relative abundance of lifeforms in the Australian Alps have changed as a consequence of recent climate change \citep{Bradstock:2014jp,Wahren:2013ji,Camac:2015hja}.
+
+We investigated shrub-grass dynamics in open heathland vegetation on the Bogong High Plains, Australia.
+Open heathland is a common Australian alpine vegetation type, occupying approximately 25\% of the treeless landscapes above 1600m in the Australian Alps.
+It is dominated by the shrub \textit{Grevillea australis} (shrub cover generally 20\textendash 60\%) with inter-shrub regions occupied by snow grasses (\textit{Poa hiemata} and \textit{Poa costiniana}) and other herbaceous species.
+It is an ecotone between closed heathland (\textgreater 70\% shrub cover) and tussock grassland (shrub cover < 20\%).
+\textit{Grevillea australis} is an obligate seeding shrub, and establishment of seedlings is dependent on disturbance creating bare ground in the grass\textendash herb sward \citep{Williams:1987ws}.
+Relative to tussock grassland, open heathland is a highly flammable plant community \citep{Williams:2006tl,Fraser:NHT5Kx_W}.
+A consequence of this differential flammability lead to the 2003 wildfires on the Bogong High Plains burnting approximately 60\% of open heathland and only 13\% of tussock grasslands.
+Following these fires, seedling regeneration of \textit{Grevillea australis} and other shrubs was prolific following the 2003 fires \citep{Williams:2014ud}.
 
@@
 
-In March 2010, at 1750 m a.s.l, we burnt 32 randomly selected mature (60 cm tall and 1.5 m$^2$) \textit{Grevillea australis} shrubs in open heathland to create patches of bare ground approximately 0.7 m$^2$ surrounded by burnt tussock grass, simulating disturbance in open heathland burnt by wildfire.
-After creating the bare ground patches we collected seedlings of dominant alpine shrub species from a nearby (\textless 2 km) site of similar altitude burnt by a late 2006 wildfire.
-We collected shrub seedlings of two dominant open heathland species \textit{Grevillea australis} (Proteaceae; a tall shrub) and \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), a dominant closed heathland species \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) that typically grows on warmer aspects and a species common to both open and closed heathland \textit{Phebalium squamulosum} (Rutaceae; a tall shrub).
-We focused on these four species because they are all fire-killed \citep[the dominant shrub post-fire strategy in the Australian mainland Alps;][]{Walsh:2004vk}; common in the Australian Alps; and under climatic warming, are likely candidates to invading non-shrubby plant communities such as alpine grasslands and herbfields.
+In March 2010, at 1750 m a.s.l, we burnt 32 randomly selected mature (60 cm tall and 1.5 m$^2$) \textit{Grevillea australis} shrubs in an open heathland site that was not burnt by wildfire in 2003 or 2006.
+This created patches of bare ground approximately 0.7 m$^2$ surrounded by burnt tussock grass, simulating disturbance of individual shrubs in open heathland burnt by wildfire.
+We collected seedlings of dominant alpine shrub species from a nearby (\textless 2 km) site of similar altitude burnt by 2006 wildfires.
+We collected seedlings of two dominant open heathland shrub species \textit{Grevillea australis} (Proteaceae; a tall shrub) and \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), a dominant closed heathland species \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) that typically grows on warmer aspects and a species common to both open and closed heathland \textit{Phebalium squamulosum} (Rutaceae; a tall shrub).
+We focused on these four species because they are common in the Australian Alps; are fire killed and thereby re-establish via seed \citep[the dominant shrub post-fire strategy in the Australian mainland Alps;][]{Walsh:2004vk}; and under climatic warming, have the potential to increase in cover and height within heathlands, and invade non-shrubby plant communities such as alpine grasslands and herbfields.
 
 A total of 640 seedlings, 256 \textit{Grevillea} (half used in \textit{Poa} inter-tussock experiment\textemdash see below) and 128 for each of \textit{Asterolasia}, \textit{Prostanthera} and \textit{Phebalium} were used.
-Four seedlings per species were randomly selected and transplanted into a 4\texttimes 4 square grid in the center of each burnt patch, with 14 cm between individuals and the edge of the patch dominated by resprouting tussock grass \textit{Poa hiemata}.
-To examine interactions between tussock grass and shrub seedlings we also randomly transplanted four additional \textit{Grevillea australis} into various sized inter-tussock gaps between burnt \textit{Poa hiemata} immediately surrounding the bare ground (Fig.
-\ref{figs8}).
+Four seedlings per species were randomly selected and transplanted into a 4\texttimes 4 square grid in the center of each burnt patch, with 14 cm between individuals and the edge of the patch, which was dominated by resprouting tussock grass \textit{Poa hiemata}.
+To examine interactions between tussock grass and shrub seedlings we also randomly transplanted four additional \textit{Grevillea australis} seedlings into various sized bare gaps between burnt \textit{Poa hiemata} tussocks that were immediately surrounding the 4\texttimes 4 bare ground square grid (Fig.
+\ref{figs1}).
 The experimental site was fenced to prevent grazing by deer and horses.
-We detected no obvious signs of rabbit or invertebrate herbivory within our plots.
+We detected no obvious signs of herbivory by invertebrates or hares within our plots.
 
 To simulate near-term warmer conditions indicated by the \citet*{IPCC:2013vv}, we randomly assigned Open Top Chambers (OTCs) to half (16) the plots, with the remainder treated as unwarmed controls.
+The chambers were constructed following the International Tundra Experiment (ITEX) protocols \citep{Molau:1996vc}.
 OTCs were placed over plots, ensuring all seedlings (including inter-tussock shrub seedlings) occurred within the 1.1 m$^2$ open top to minimise edge effects.
-OTCs were placed out at the start of the growing season (October) where they remained until snowfall (early June).
-This procedure was repeated for five growing seasons from May 2010 to May 2015.
+OTCs were placed out at the start of the growing season (October) where they remained until first snowfall (early June).
+This procedure was repeated for six growing seasons from May 2010 to May 2016.
 
@@
 Across \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} growing seasons (\Sexpr{n_days(chamber_dates, TRUE)} growing season days), OTCs simulated warmer conditions at the lower end of \citet*{IPCC:2013vv} projections (Fig.
-\ref{figs5}-\ref{figs7}).
+\ref{figs6}-\ref{figs8}).
 OTCs passively increased average ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C.
@@
 Seedling survival, maximum height and stem diameter (nearest mm measured with Vernier calipers) were initially recorded in May 2010 and then subsequently re-measured at the end of each growing season (May-June).
-At the same time, we recorded the distance to the nearest tussock or grass seedling in each of four cardinal directions for shrub seedlings growing in inter-tussock gaps.
-Thus, changes in inter-tussock gap size could be due to either vegetative growth or seedling establishment.
+At the same time, we recorded the distance to the nearest tussock or grass seedling in each of four cardinal directions for \textit{Grevillea} seedlings growing in inter-tussock gaps.
+Thus, temporal changes in inter-tussock gap size could be due to either vegetative growth of resprouting tussocks, or the establishment of grass seedlings.
 We did not measure individual characteristics (e.g.
-height and basal diameter) of surrounding tussock grass because we could not distinguish individuals, and because height varied throughout the season.
-Numbers of natural \textit{Grevillea australis} recruits establishing within the plots were also recorded for the first two seasons.
+height and basal diameter) of surrounding tussock grass because we could not distinguish individuals, and because height varied substantially throughout the season.
+Numbers of natural \textit{Grevillea australis} recruits establishing within the plots were also recorded and identified for the first two years of the experiment.
 
@@
 
-We used 40 long-term open heathland sites established after the 2003 fires \citep{Camac:2013eb} consisting of 30 burnt sites and 10 sites thought to be unburnt for over 70 years.
-The disparity in sampling effort between burnt and unburnt was due to very few open heathland remaining unburnt after both the 2003 and 2007 fires.
+We used 40 open heathland monitoring sites established after the 2003 fires \citep{Williams:2006tl,Camac:2013eb}.
+These sites consisted of 30 burnt sites and 10 sites known to be unburnt for over 70 years.
+The disparity in sampling effort between burnt and unburnt was due to very few open heathland sites of sufficient size (0.25 ha) remaining unburnt after both the 2003 and 2006 fires.
 In the summer of 2011-12, at each site, seedling density/m$^2$ was estimated using 40 quadrats, each 1 m$^2$, that were evenly distributed in groups of 10 along four 50 m transects, with 10 m between transect lines, subsampling an area of 2000 m$^2$.
@@
 Site level data, elevation and Topographic Wetness Index \citep[TWI; a measure of plant available water,][]{Moore:1993fk} were obtained from a 30 m resolution digital elevation model.
-Lastly, for burnt sites, fire severity was estimated by twig diameters, collected immediately after the 2003 fires \citep{Camac:2013eb}.
+Lastly, for burnt sites, fire severity was estimated by twig diameters \citep{Whight:1999vw}, collected immediately after the 2003 fires \citep{Williams:2006tl}.
 
@@
 
-In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality.
-For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")}.
-Data and source code for reproducing analysis and figures are available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.
-Additional information about experimental design and analysis can be found in Supporting Information.
+We built multiple hierarchical models to examine how increased temperature and other factors influenced shrub seedling recruitment, growth and mortality, as well as tussock-grass gap dynamics.
+For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} \citep{Rcran:2016xx} using package \texttt{rstan} \Sexpr{packageVersion("rstan")} \citep{stan:2015xx}.
+Detailed information about experimental design and analysis is provided in the Supporting Information.
+Data and source code are available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.
+In order to aid in the reproducibility of this work, our code was written using a \texttt{remake} framework \citep{fitzjohn2015}, such that others can readily reproduce our entire workflow from data processing, through to producing a pdf of this manuscript by calling \texttt{remake()} in R.
 
-\begin{figure}[!htb]
+\begin{figure}[!h]
 \centering
 \includegraphics{figures/fig2.pdf}
-\caption{Mean \textit{Grevillea australis} seedling density along gradients of burning, fire severity (twig diameter), altitude, Topographic Wetness Index (TWI) and adult density.
+\caption{Mean \textit{Grevillea australis} seedling density as a function of burning (burnt/unburnt), fire severity (twig diameter), altitude, Topographic Wetness Index (TWI) and adult density.
 (a) Centered and standardized model coefficients (on the log scale); and effects of: (b) burning, (c) fire severity, (d) altitude, (e) topographic wetness and (F) adult density, in areas burnt by the 2003 fires.
@@
 We first investigated how altitude, Topographic Wetness Index (TWI), adult density, fire and fire severity (as measured by post-fire twig diameters\textemdash see Supporting Information) influenced the density of \textit{Grevillea} and \textit{Asterolasia} seedlings (the two dominant shrubs of alpine open heathland).
-In 2011-12, we estimated seedling density for both species across 40 alpine sites and found that the abundances of \textit{Grevillea} (Fig.
+Across the 40 alpine sites surveyed in 2011-12 we found that the abundance of \textit{Grevillea} (Fig.
 \ref{fig2}) and \textit{Asterolasia} (Fig.
-\ref{figs1}) seedlings were strongly influenced by the occurrence of fire.
+\ref{figs2}) seedlings was strongly influenced by the occurrence of fire.
 Sites burnt in 2003 had seedling densities between \Sexpr{density_ratios(asttry_seedling_density_model,'Asterolasia')} and \Sexpr{density_ratios(greaus_seedling_density_model,'Grevillea')} times higher than unburnt sites.
@@
 \textit{Asterolasia} had similar mean densities: \Sexpr{density_means(asttry_seedling_density_model,'Asterolasia', TRUE)} and \Sexpr{density_means(asttry_seedling_density_model, 'Asterolasia', FALSE)} seedlings/m$^2$ at burnt and unburnt sites, respectively.
-For both species, seedling density was also higher in severely burnt sites (i.e.
+For both species, seedling density increased with increasing fire severity (i.e.
 sites with larger post-fire twig diameters).
-As expected, pre-fire adult density positively influenced \textit{Grevillea} seedling density.
-For both shrub species, we detected no change in seedling density along a \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m altitudinal range \citep[equivalent to a \Sexpr{round(0.77 * (200/100),1)}\textdegree C change in mean temperature;][]{Brown:1989vx}.
-The field warming experiment (see below) indicates that this lack of elevation response in \textit{Grevillea} is due to rates of recruitment (Fig.
-\ref{figs2}) and mortality (see below) largely being insensitive to a \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C change in temperature.
+As hypothesised, pre-fire adult density also positively influenced \textit{Grevillea} seedling density.
+For both shrub species, we detected no change in seedling density along a \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m elevational gradient \citep[equivalent to a \Sexpr{round(0.77 * (200/100),1)}\textdegree C change in mean temperature;][]{Brown:1989vx}.
+This is consistent with the field warming experiment (see below) which indicated rates of recruitment and mortality in \textit{Grevillea} (Fig.
+\ref{figs3}) is largely insensitive to a \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C change in temperature.
 We detected no strong effect of Topographic Wetness Index for either species.
 
-\subsection*{Experimental warming effects on shrub seedling growth}
-As shrub seedlings were more abundant in burnt vegetation, we investigated how warmer post-fire conditions affected seedling growth and mortality rates.
+\subsection*{Experimental warming and shrub seedling growth}
+Because seedlings were more abundant in burnt vegetation, we investigated how warmer post-fire conditions affected seedling growth and mortality rates.
 Seedlings of \textit{Grevillea}, \textit{Asterolasia}, \textit{Phebalium} and \textit{Prostanthera} emerging after a wildfire were transplanted into experimentally burnt plots.
@@
 controls), or enclosed in Open Top Chambers (OTCs) which rose temperature by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C.
+
 After \Sexpr{n_days(chamber_dates, FALSE)} days (\Sexpr{n_days(chamber_dates, TRUE)} growing season days) or \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} years' growth, mean seedling heights of the tall shrubs (\textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera}) growing in post-fire bare ground were greater in warmed plots relative to controls by \Sexpr{ht_diff(non_tussock_otc_growth_data,'Grevillea')}, \Sexpr{ht_diff(non_tussock_otc_growth_data,'Phebalium')} and \Sexpr{ht_diff(non_tussock_otc_growth_data,'Prostanthera')} cm, respectively (Fig.
@@
 By contrast, seedlings of the prostrate shrub, \textit{Asterolasia}, showed no difference in growth rate between warmed and control plots in any year.
-For each species, we also observed similar treatment trends with stem diameter growth (Fig.
-\ref{figs3}).
+For each species, similar treatment effects were observed for stem diameter growth (Fig.
+\ref{figs4}).
 Accounting for initial height and assuming logistic growth, the rates of change in mean annual predicted height of \textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera} were \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Grevillea')}, \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Phebalium')} and \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Prostanthera')} times that observed in control plots, respectively.
@@
 
-We validated the \textit{Grevillea} and \textit{Asterolasia} growth responses observed in experimentally warmed plots by examining the effect of elevation on maximum heights of post-fire recruits across 30 open heathland sites burnt in 2003.
+The growth responses of both \textit{Grevillea} and \textit{Asterolasia} observed in experimentally warmed plots were validated by the maximum heights of post-fire recruits across 30 open heathland sites burnt in 2003.
 Here, a 190 m altitudinal range is equivalent to a mean ambient temperature difference of approximately \Sexpr{round(0.77 * (190/100),1)}\textdegree C \citep{Brown:1989vx}, which is comparable to that observed between experimentally warmed and control plots (\Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C).
@@
 Our experimental predictions were verified (Fig.
-\ref{figs4}).
+\ref{figs5}).
 Mean maximum height of \textit{Grevillea} seedlings in burnt open heathland were \Sexpr{max_ht_alt(greaus_seedling_max_ht_model)} cm taller at 1670 m a.s.l compared to seedlings at 1860 m a.s.l.
@@
 
-\subsection*{Experimental warming on shrub seedling mortality}
+\subsection*{Experimental warming and shrub seedling mortality}
 
-After five years and across all plots, \Sexpr{total_deaths(non_tussock_otc_mortality_data,TRUE)}\% (\Sexpr{total_deaths(non_tussock_otc_mortality_data,FALSE)} out of 511) of all seedlings growing in burnt bare ground patches had died.
+After six years and across all plots, \Sexpr{total_deaths(non_tussock_otc_mortality_data,TRUE)}\% (\Sexpr{total_deaths(non_tussock_otc_mortality_data,FALSE)} out of 511) of all seedlings transplanted into the 4\texttimes 4 bare ground square grid had died.
 Most deaths occurred in \textit{Asterolasia} (\Sexpr{death_counts(non_tussock_otc_mortality_data,'Asterolasia')}) and \textit{Phebalium} (\Sexpr{death_counts(non_tussock_otc_mortality_data, 'Phebalium')}), followed by \textit{Grevillea} (\Sexpr{death_counts(non_tussock_otc_mortality_data,'Grevillea')}) and \textit{Prostanthera} (\Sexpr{death_counts(non_tussock_otc_mortality_data,'Prostanthera')}).
@@
 This significant decrease in mortality may be a consequence of OTCs reducing the severity of spring frosts by rising minimum ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary,'CAmbient_Temp','mean_min','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary,'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C, respectively (Fig.
-\ref{figs5}-\ref{figs7}).
+\ref{figs6}-\ref{figs8}).
 Warming also reduced mean seedling mortality in \textit{Grevillea} and \textit{Phebalium} (Fig.
@@
 Climate change is expected to alter biotic interactions because their strength and direction depend strongly on climatic conditions, particularly in alpine and arctic ecosystems \citep{Callaway:2002vb, Klanderud:2005uw}.
-Here, we assess the interactive effects of warming and grass proximity on the growth and survival of \textit{Grevillea} seedlings transplanted into various sized inter-tussock gaps (Fig.
+Here, we assess the interactive effects of warming and grass proximity on the growth and survival of \textit{Grevillea} seedlings transplanted into various sized inter-tussock gaps.
+We detected a strong positive effect of warming treatment on growth rates and a marginal, yet not significant, decrease in mortality (Fig.
 \ref{fig5}).
-We detected a strong positive effect of warming treatment on growth rates and a marginal, yet not significant, decrease in mortality.
 However, we did not detect significant inter-tussock gap size effects or an interaction between gap size and warming treatment on either growth or mortality rates (i.e.
@@
 Despite having little impact on shrub seedling growth and mortality rates, tussock grass may still limit shrub recruitment, and thus, shrub expansion, by infilling post-fire bare ground gaps (whether by vegetative growth or seedlings) faster under warmer conditions.
-Using five years of post-fire inter-tussock gap size changes in warmed and unwarmed plots we found that gaps were being infilled by tussock grasses (Fig.
+Using six years of post-fire inter-tussock gap size changes in warmed and unwarmed plots we found that gaps were being infilled by tussock grasses (Fig.
 \ref{fig6}).
@@
 
-Here we quantified several unknown interactions between fire, shrub-grass relationships and climate to extend a conceptual model of alpine shrub dynamics under climate change (Fig.
+In this study we have quantified several unknown interactions between fire, shrub-grass relationships and climate to extend a conceptual model of alpine shrub dynamics under climate change (Fig.
 \ref{fig1}b).
-We showed how these interactions strengthen a hypothesized feedback loop that can rapidly increase shrub cover in alpine and tundra ecosystems \citep{Higuera:2008dw}.
-The combination of rising temperatures and more frequent or severe fire creates bare ground required for shrub seedlings to establish in greater densities, and for tall shrubs, double their growth rates and potentially increase their survival.
-These demographic effects will ultimately manifest in shrub thickening within and at heathland boundaries, which in turn, will facilitate shrub establishment into shrub-free communities by decreasing dispersal distances.
-Coupled with slow tussock grass gap infilling and, field \citep{Williams:2006tl}, experimental \citep{Fraser:NHT5Kx_W} and paleoecological \citep{Higuera:2008dw} evidence which indicate that shrubs are the most flammable component of alpine and tundra ecosystems  \citep[particularly moreso then tussock grassland; ][]{Fraser:NHT5Kx_W}, our results suggest that flammable fuel loads will accumulate faster and cover a larger proportion of the alpine and arctic landscapes under a warmer environment.
-This further strengthens the feedback \citep{Higuera:2008dw, Higuera:2009ga} by increasing the frequency or severity of fires, which then creates more shrub recruitment opportunities both within and beyond heathland boundaries (i.e.
-via more fire-induced tussock death or increased exposure of inter-tussock gaps) with little demographic impact of neighbouring tussock grass.
+In particular, we have shown how these interactions strengthen a hypothesized feedback loop that can rapidly increase shrub cover in the Australian Alps, and potentially other alpine and tundra ecosystems \citep{Higuera:2008dw}.
+While high cover of tussock grass is known to inhibit the establishment of shrub seedlings \citep{Williams:1987ws}, more frequent and severe fire caused by changing climate, will create bare ground required for shrub seedlings to establish.
+Our results, highlight that given a shrub seedling establishes in bare ground, its growth and survival rates are not affected by proximity to tussock grass, irrespective or warming.
+More importantly, however, our results indicate that for tall shrubs, a 1\textdegree C increase in mean growing season temperature resulted in a doubling of growth rates and a potential increase in survival.
+These demographic effects will ultimately manifest in shrub thickening within and at shrub-grassland/herbfield boundaries.
+It will also facilitate shrub establishment into non-shrubby communities by increasing seed sources at the boundary (i.e.
+more mature plants) and decreasing dispersal distances.
+
+Evidence from field studies of burning patterns \citep{Williams:2006tl}, quantification of fuel mass and architecture \citep{Fraser:NHT5Kx_W} and paleoecological studies \citep{Higuera:2008dw} indicate that shrubs are the most flammable component of alpine and tundra ecosystems \citep[particularly moreso then tussock grassland; ][]{Fraser:NHT5Kx_W}.
+Consequently, our results indicate that flammable fuel loads will accumulate twice as fast under a warmer environment.
+This is likely to further strengthen the feedback \citep{Higuera:2008dw, Higuera:2009ga} by increasing the frequency and severity of fires, which then creates more shrub recruitment opportunities both within and beyond shrub-grassland boundaries that may persist for decades.
 
 While we have addressed several unknowns associated with this feedback between climatic warming, shrubs and fire, there are others we have not addressed that may also strengthen, weaken or break this feedback.
-The most obvious mechanism that will break this cycle involves shorter fire intervals that prevent fire-killed shrubs reaching reproductive age and thereby exhausting the seedbank.
-However, this scenario is unlikely for the majority of alpine (or tundra) landscapes, including in Australia, where current fire intervals of 50 to 100 years would need to decrease to less than 20 years\textemdash the time estimated for the species in this study to reach reproductive maturity \citep{Williams:2008ba}.
-Furthermore, if reproductive output is related to plant size \citep{Wenk:2015ko}, then climatic warming may allow fire-killed species to reach reproductive maturity sooner, and consequently, may increase species resilience to short fire intervals.
+The most obvious mechanism that will break this cycle involves shorter fire intervals that prevent fire-killed shrubs reaching reproductive age and thereby exhausting the seedbank \citep{Enright:2015em}.
+However, this scenario is unlikely for the majority of alpine (or tundra) landscapes, including those in Australia, where current fire intervals of 50 to 100 years would need to decrease to less than 20 years\textemdash the time estimated for the species in this study to reach reproductive maturity \citep{Williams:2008ba}.
+Furthermore, if, as hypothesised, reproductive output is related to plant size \citep{Wenk:2015ko}, then climatic warming may allow fire-killed species to reach reproductive maturity sooner, and consequently, may increase species resilience to short fire intervals.
 Nevertheless, these factors and others such as changes in snow pack \citep{Wipf:2009jh}, soil moisture \citep{Engler:2011vq} or herbivory \citep{Post:2008wz} require further research because they are likely to be altered by the interactive effects of climate and disturbance in unpredictable ways \citep{MyersSmith:2011fp}.
@@
 By focusing on seedlings, the life stage most vulnerable to climate and disturbance, and which determines the long-term persistence of a species and its capacity to establish in new areas \citep{Walck:2011dm, Briceno:2015iv}, our findings provide mechanistic understanding as to why shrub cover has increased, often at the expense of grasslands, in many alpine and arctic ecosystems \citep{Racine:2004tn,MyersSmith:2011fp,Frost:2013tm,Lantz:2013fh}.
-But more importantly, our results provide evidence for underlying processes that could result in a warming-fire-shrub feedback that has been hypothesized in paleoecological studies \citep{Higuera:2008dw, Higuera:2009ga}.
+But more importantly, our results provide evidence for the underlying processes that could result in a warming-fire-shrub feedback that has been hypothesized in paleoecological studies \citep{Higuera:2008dw, Higuera:2009ga}.
 Based on current observations and predictions, average global temperature has already increased by 0.85\textdegree C since 1880 and is expected to rise by as much as 4.8\textdegree C by 2100 \citep{IPCC:2013vv}.
-In alpine and tundra environments, temperatures \citep{ChapinIII:2005js}, shrub cover \citep{MyersSmith:2011fp} and the frequency and severity of fire \citep{Westerling:2006ve, Flannigan:2009kl, Qiu:2009th, Bradstock:2014jp} have all increased in the last few decades.
-These changes mean that the warming-shrub-fire feedback loop identified here has already strengthened, which could cause grasslands and other non-woody communities to transition to an alternative state with more shrubs and more fire, both of which are likely to have consequences for carbon sequestration, water supply and biodiversity.
+In alpine and tundra environments, temperatures \citep{ChapinIII:2005js}, shrub cover \citep{MyersSmith:2011fp} and the frequency and severity of fire \citep{Westerling:2006ve, Flannigan:2009kl, Qiu:2009th, Bradstock:2014jp} have all increased across the globe in the last few decades as a consequence of rising temperatures.
+These changes mean that the warming-shrub-fire feedback loop identified here is likely to have already strengthened, which could cause grasslands and other non-woody communities to transition to an alternative state with more shrubs and more fire, both of which are likely to have consequences for carbon sequestration, water supply and biodiversity.
 

### 2016-10-18 manuscript/supplementary.Rnw
Updates to manuscript

@@
-\documentclass[9pt,twocolumn,twoside, lineno]{gcb}
+\documentclass[9pt,onecolumn,twoside, lineno]{gcb}
 % Use the lineno option to display guide line numbers if required.
@@
 \templatetype{gcbresearcharticle} % Choose template 
+\usepackage{float}
 
@@
 \renewcommand{\thefigure}{S\arabic{figure}}
-
-\title{Climatic warming strengthens a positive feedback\\ between alpine shrubs and fire}
+\renewcommand{\bibsection}{\section{\refname}}
+\title{Climatic warming strengthens a positive feedback\\ between alpine shrubs and fire \\[1cm]
+\huge Supplementary Material}
 
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,b,c]{James S Camac}
-\author[d]{Richard J Williams} 
-\author[c]{Carl-Henrik Wahren}
-\author[e]{Ary A Hoffmann}
-\author[b]{Peter A Vesk}
+\author[a,b]{James S Camac}
+\author[c]{Richard J Williams} 
+\author[b]{Carl-Henrik Wahren}
+\author[d]{Ary A Hoffmann}
+\author[a]{Peter A Vesk}
 
-\affil[a]{Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia}
-\affil[b]{The Centre of Excellence for Environmental Decisions, School of BioSciences, The University of Melbourne, Melbourne, Victoria 3010, Australia}
-\affil[c]{Research Centre for Applied Alpine Ecology, La Trobe University, Melbourne, Victoria 3086, Australia}
-\affil[d]{CSIRO Land and Water, Tropical Ecosystems Research Centre, PMB 44 Winnellie, Northern Territory 0821, Australia}
-\affil[e]{Bio21 Institute, School of BioSciences, The University of Melbourne, Victoria 3010, Australia}
+\affil[a]{The Centre of Excellence for Environmental Decisions, School of BioSciences, The University of Melbourne, Melbourne, Victoria 3010, Australia}
+\affil[b]{Research Centre for Applied Alpine Ecology, La Trobe University, Melbourne, Victoria 3086, Australia}
+\affil[c]{CSIRO Land and Water, Tropical Ecosystems Research Centre, PMB 44 Winnellie, Northern Territory 0821, Australia}
+\affil[d]{Bio21 Institute, School of BioSciences, The University of Melbourne, Victoria 3010, Australia}
 
@@
 \suppmaketitle
+\tableofcontents
+\listoffigures
 
-\section*{SI Materials and Methods}
-\subsection*{Study site}
+\section{Study site}
 Our research sites were on the Bogong High Plains, a series of alpine and high subalpine plateaux in the Alpine National Park, approximately 250 km northeast of Melbourne, Australia.
@@
 
-\subsection*{Open Top Chamber Experiment}
+\section{Open Top Chamber Experiment}
 We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of post-fire gap-infilling by tussock grass and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affects shrub seedling demographic rates; and 3) whether such effects change under warmer conditions post-fire conditions.
@@
 We collected seedlings from a site burnt by a late 2006 fire because it allowed us to collect sufficient numbers of seedlings from species occurring in both open and closed heathland.
-We focused on \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) because they are all fire-killed (the dominant shrub post-fire strategy in the Australian mainland Alps; \citet{Walsh:2004vk}); common in the Australian Alps; and under climatic warming, are likely candidates to invading non-shrubby plant communities such as alpine grasslands and herbfields.
+We focused on \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) because they are all fire-killed \citep[the dominant shrub post-fire strategy in the Australian mainland Alps;][]{Walsh:2004vk}; common in the Australian Alps; and under climatic warming, are likely candidates to invading non-shrubby plant communities such as alpine grasslands and herbfields.
 Resprouter shrub species were not examined in this study because we were unable to locate sufficient numbers of seedlings; and are unlikely to be the primary contributors to shrub expansion in these landscapes.
@@
 The combination of bare ground and inter-tussock seedlings meant that each experimental plot had a total of 20 seedlings per 1 m$^2$ (Fig.
-\ref{figs8}).
+\ref{figs1}).
 However, the two sets of seedlings (i.e.
@@
 Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions \citep{IPCC:2013vv} (Supplementary Fig.
-\ref{figs5}-\ref{figs7}).
+\ref{figs6}-\ref{figs8}).
 OTCs passively increased average daily ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively.
@@
 
-\subsubsection*{Data inputs for growth, mortality and recruitment models}
+\subsection{Data inputs for growth, mortality and recruitment models}
 In each model, OTC treatment, $OTC$, was a binary (0 = Control, 1 = OTC) plot level covariate.
@@
 
-\subsection*{Seedling survey}
+\section{Seedling survey}
 We measured seedling density and maximum heights in open heathland across the Bogong High Plains.
@@
 
-\subsubsection*{Data inputs into density and maximum height models}
+\subsection{Data inputs into density and maximum height models}
 In all seedling survey models, $Burnt_j$ is a binary (0 = unburnt, 1 = burnt) site level covariate, which only affects seedling counts at unburnt sites.
@@
 
-\subsection*{Data analysis}
+\section{Data analysis}
 In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality.
@@
 
-\subsubsection*{Modelling shrub seedling growth rates}
+\subsection{Modelling shrub seedling growth rates}
 We modelled the observed height of individual $i$ at May census $t$ as a random realisation from a normal distribution truncated at zero:
@@
 
-\subsubsection*{Modelling shrub seedling mortality rates}
+\subsection{Modelling shrub seedling mortality rates}
 We were also interested in whether the instantaneous mortality rates of seedlings varied amongst shrub species, OTC treatment and inter-tussock gap size.
@@
 
-\subsubsection*{Modelling Grevillea recruitment rates}
+\subsection{Modelling Grevillea recruitment rates}
 We examined whether warming influenced natural recruitment rates of \textit{Grevillea} by modelling the number of recruits per 1 m$^2$ plot $k$ in each census, $t$ as a random realisation from a negative binomial distribution:
@@
 
-\subsubsection*{Modelling censored Poa distances}
+\subsection{Modelling censored Poa distances}
 Approximately \Sexpr{prop_censored(cleaned_poa_distances)}\% of the $Poa$ distance data used to calculate the median inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, (i.e.
@@
 
-\subsubsection*{Modelling rates of inter-tussock gap infilling by Poa}
+\subsection{Modelling rates of inter-tussock gap infilling by Poa}
 As well as modelling how OTCs affected shrub seedling growth and mortality rates, we were also interested in examining how warming influenced rates of grass infilling of inter-tussock gaps.
@@
 
-\subsubsection*{Modelling drivers of seedling density}
+\subsection{Modelling drivers of seedling density}
 We examined drivers of seedling density by modelling the number of seedlings per square metre plot $i$ as a random realisation from a Poisson distribution:
@@
 
-\subsubsection*{Modelling drivers of seedling maximum heights}
+\subsection{Modelling drivers of seedling maximum heights}
 In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation \citep[equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C;][]{Brown:1989vx} influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires.
@@
 \bibliography{refs.bib}
+\pagebreak
 
-\begin{figure*}
+\section{Supplementary Figures}
+
+\begin{figure}[h]
 \centering
 \includegraphics{figures/figS1.pdf}
-\caption{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (twig diameter), altitude and Topographic Wetness Index (TWI).
-(A) Centered and standardized model coefficients (on the log scale); and effects of (B) burning, (C) fire severity, (D) altitude and (E) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires.
-All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
-\label{figs1}
-\end{figure*}
-\clearpage
-
-\begin{figure*}
-\centering
-\includegraphics{figures/figS2.pdf}
-\caption{Average number of \textit{Grevillea australis} recruits in warmed (red) and control (blue) conditions in 2011 and 2012. Errors bars are 95\% Bayesian Credible Intervals.}
-\label{figs2}
-\end{figure*}
-
-\clearpage
-\begin{figure*}
-\centering
-\includegraphics{figures/figS3.pdf}
-\caption{Observed seedling diameter growth trajectories of four dominant Australian alpine shrubs.
-Red and Blue lines represent mean ($\pm$ 95\% CI) observed heights at each May census in warmed and control plots, respectively.
-Orange and light blue lines represent individual growth trajectories in warmed and unwarmed plots, respectively.}
-\label{figs3}
-\end{figure*}
-\clearpage
-
-\begin{figure*}
-\centering
-\includegraphics{figures/figS4.pdf}
-\caption{Mean maximum seedling height along fire severity (twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires.
-Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses.
-(A \& B) Centered and standardized model coefficients (on the log scale); and effects of (C \& D) fire severity, (E \& F) altitude and (G \& H) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires.
-All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
-\label{figs4}
-\end{figure*}
-\clearpage
-
-\begin{figure*}
-\centering
-\includegraphics{figures/figS5.pdf}
-\caption{Mean daily microclimate difference (OTC $-$ Control).
-(A) Ambient temperature, (B) soil temperature, (C) volumetric water content and (D) relative humidity over experimental period.
-Red lines = mean difference across all growing seasons (when chambers were applied); blue = no difference between OTCs and controls.
-Shaded areas are the winter snow periods when OTCs were not over plots.
-Note: declines in amplitude in later years are due to shading effects caused by increased shrub size.}
-\label{figs5}
-\end{figure*}
-\clearpage
-
-\begin{figure*}
-\centering
-\includegraphics{figures/figS6.pdf}
-\caption{Mean minimum daily microclimate difference (OTC $-$ Control).
-(A) Ambient temperature, (B) soil temperature, (C) volumetric water content and (D) relative humidity over experimental period.
-Red lines = mean difference across all growing seasons (when chambers were applied); blue = no difference between OTCs and controls.
-Shaded areas are the winter snow periods when OTCs were not over plots.
-Note: declines in amplitude in later years are due to shading effects caused by increased shrub size.}
-\label{figs6}
-\end{figure*}
-\clearpage
-
-\begin{figure*}
-\centering
-\includegraphics{figures/figS7.pdf}
-\caption{Mean maximum daily microclimate difference (OTC $-$ Control).
-Ambient temperature (A), soil temperature (B), volumetric water content (C) and relative humidity over experimental period (D).
-Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls.
-Shaded areas are the winter snow periods when OTCs were not over plots.
-Note: amplitude declines in later years are due to shading effects caused by increased shrub growth.}
-\label{figs7}
-\end{figure*}
-\clearpage
-
-\begin{figure*}
-\centering
-\includegraphics{figures/figS8.pdf}
 \caption{Diagram outlining positions of transplanted seedling in an OTC plot.
@@
 it was not within the 110 cm opening of the chamber) and therefore was not used in the experiment.}
-\label{figs8}
-\end{figure*}
+\label{figs1}
+\end{figure}
 \clearpage
 
+\begin{figure}[!hb]
+\centering
+\includegraphics{figures/figS2.pdf}
+\caption{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (twig diameter), altitude and Topographic Wetness Index (TWI).
+(a) Centered and standardized model coefficients (on the log scale); and effects of (b) burning, (c) fire severity, (d) altitude and (e) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires.
+All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
+\label{figs2}
+\end{figure}
+\clearpage
+
+\begin{figure}[!hb]
+\centering
+\includegraphics{figures/figS3.pdf}
+\caption{Average number of \textit{Grevillea australis} recruits in warmed (red) and control (blue) conditions in 2011 and 2012. Errors bars are 95\% Bayesian Credible Intervals.}
+\label{figs3}
+\end{figure}
+\clearpage
+
+\begin{figure}[!hb]
+\centering
+\includegraphics{figures/figS4.pdf}
+\caption{Observed seedling diameter growth trajectories of four dominant Australian alpine shrubs.
+Red and Blue lines represent mean ($\pm$ 95\% CI) observed heights at each May census in warmed and control plots, respectively.
+Orange and light blue lines represent individual growth trajectories in warmed and unwarmed plots, respectively.}
+\label{figs4}
+\end{figure}
+\clearpage
+
+\begin{figure}[!hb]
+\centering
+\includegraphics{figures/figS5.pdf}
+\caption{Mean maximum seedling height along fire severity (twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires.
+Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses.
+(a \& b) Centered and standardized model coefficients (on the log scale); and effects of (c \& d) fire severity, (e \& f) altitude and (g \& h) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires.
+All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
+\label{figs5}
+\end{figure}
+\clearpage
+
+\begin{figure}[!hb]
+\centering
+\includegraphics{figures/figS6.pdf}
+\caption{Mean daily microclimate difference (OTC $-$ Control).
+(a) Ambient temperature, (b) soil temperature, (c) volumetric water content and (d) relative humidity over experimental period.
+Red lines = mean difference across all growing seasons (when chambers were applied); blue = no difference between OTCs and controls.
+Shaded areas are the winter snow periods when OTCs were not over plots.
+Note: declines in amplitude in later years are due to shading effects caused by increased shrub size.}
+\label{figs6}
+\end{figure}
+\clearpage
+
+\begin{figure}[!hb]
+\centering
+\includegraphics{figures/figS7.pdf}
+\caption{Mean minimum daily microclimate difference (OTC $-$ Control).
+(a) Ambient temperature, (b) soil temperature, (c) volumetric water content and (d) relative humidity over experimental period.
+Red lines = mean difference across all growing seasons (when chambers were applied); blue = no difference between OTCs and controls.
+Shaded areas are the winter snow periods when OTCs were not over plots.
+Note: declines in amplitude in later years are due to shading effects caused by increased shrub size.}
+\label{figs7}
+\end{figure}
+\clearpage
+
+\begin{figure}[!hb]
+\centering
+\includegraphics{figures/figS8.pdf}
+\caption{Mean maximum daily microclimate difference (OTC $-$ Control).
+Ambient temperature (a), soil temperature (b), volumetric water content (c) and relative humidity over experimental period (d).
+Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls.
+Shaded areas are the winter snow periods when OTCs were not over plots.
+Note: amplitude declines in later years are due to shading effects caused by increased shrub growth.}
+\label{figs8}
+\end{figure}
+
 \end{document}

### 2016-10-18 manuscript/manuscript.Rnw
Fixing typos and adding subheading for introduction

@@
 
-However, most studies have focused on plant responses in mature, undisturbed vegetation \citep{Briceno:2015iv}.
+However, most studies in arctic and alpine areas have focused on plant responses in mature, undisturbed vegetation \citep{Briceno:2015iv}.
 Few climate-change field experiments have included disturbance as a factor \citep[but see:][]{Munier:2010cg,Graae:2010ir,Camac:2015hja}, or examined the influence of climate change on seedlings in post-disturbance environments.
 Although alpine and tundra vegetation has a strong capacity to regenerate vegetatively \citep{Billings:1968bn} there is mounting evidence that seedling regeneration is important in these ecosystems \citep{Venn:2009bw, Briceno:2015iv}, particularly for wood species \citep{Camac:2013eb,Williams:2014ud}.
-Although seedlings are highly vulnerable to climate and disturbance, they are the life stage which determines the long-term persistence of a species and its capacity to establish in new areas.
+Although seedlings are highly vulnerable to climate and disturbance, they are the life stage that determines the long-term persistence of a species and its capacity to establish in new areas.
 However, little is known about how climate change affects seedling demography \citep{Briceno:2015iv}, particularly in post-disturbance conditions in alpine and tundra ecosystems.
 It is therefore imperative we understand how seedlings respond to both changing climate and disturbance regimes \citep{Walck:2011dm, Briceno:2015iv} if we wish to accurately predict future trajectories of vegetation change.
+
+\section*{Conceptual model}
 
@@
 more bare ground) and exacerbate both shrub thickening within shrub communities and shrub expansion into shrub-free communities such as grassland and herbfields \citep{Racine:2004tn,Lantz:2013fh}.
-If this effect is coupled with increased shurb seedling growth and mortality, highly flammable fuels will accumulate faster and have higher landscape connectivity, ultimately leading to further increases in the likelihood of fire \citep{Higuera:2008dw, Higuera:2009ga,MyersSmith:2011fp}.
+If this effect is coupled with increased shrub seedling growth and mortality, highly flammable fuels will accumulate faster and have higher landscape connectivity, ultimately leading to further increases in the likelihood of fire \citep{Higuera:2008dw, Higuera:2009ga,MyersSmith:2011fp}.
 Thus, warming could strengthen an existing climate-disturbance feedback, that not only has the potential to cause rapid changes in the composition and structure of alpine and arctic vegetation, but also has serious social, biodiversity and carbon sequestration consequences \citep{Mack:2011ew}.
@@
 Solid lines = known mechansims; Dashed lines = mechansims that may exacerbate or diminish this feedback but which we have a paucity of information on.
-(b) Feedback loop based on new evidence obtained from experimental and field surveys conducted in this study.
+(b) Updated feedback based on new evidence obtained from experimental and field surveys conducted in this study.
 Also includes an example of a potential, but unlikely, effect (red dashed line) that could break this feedback.
@@
 Relative to tussock grassland, open heathland is a highly flammable plant community \citep{Williams:2006tl,Fraser:NHT5Kx_W}.
-A consequence of this differential flammability lead to the 2003 wildfires on the Bogong High Plains burnting approximately 60\% of open heathland and only 13\% of tussock grasslands.
-Following these fires, seedling regeneration of \textit{Grevillea australis} and other shrubs was prolific following the 2003 fires \citep{Williams:2014ud}.
+A consequence of this differential flammability lead to the 2003 wildfires on the Bogong High Plains burning approximately 60\% of open heathland and only 13\% of tussock grasslands.
+Following these fires, seedling regeneration of \textit{Grevillea australis} and other shrubs was prolific \citep{Williams:2014ud}.
 
@@
 We collected seedlings of two dominant open heathland shrub species \textit{Grevillea australis} (Proteaceae; a tall shrub) and \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), a dominant closed heathland species \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) that typically grows on warmer aspects and a species common to both open and closed heathland \textit{Phebalium squamulosum} (Rutaceae; a tall shrub).
-We focused on these four species because they are common in the Australian Alps; are fire killed and thereby re-establish via seed \citep[the dominant shrub post-fire strategy in the Australian mainland Alps;][]{Walsh:2004vk}; and under climatic warming, have the potential to increase in cover and height within heathlands, and invade non-shrubby plant communities such as alpine grasslands and herbfields.
+We focused on these four species because they are common in the Australian Alps; are fire-killed and thereby re-establish via seed \citep[the dominant shrub post-fire strategy in the Australian mainland Alps;][]{Walsh:2004vk}; and under climatic warming, have the potential to increase in cover and height within heathlands, and invade non-shrubby plant communities such as alpine grasslands and herbfields.
 
@@
 Microclimatic conditions were measured hourly using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) at four control and four OTC plots.
-Across \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} growing seasons (\Sexpr{n_days(chamber_dates, TRUE)} growing season days), OTCs simulated warmer conditions at the lower end of \citet*{IPCC:2013vv} projections (Fig.
+Across six growing seasons (\Sexpr{n_days(chamber_dates, TRUE)} growing season days), OTCs simulated warmer conditions at the lower end of \citet*{IPCC:2013vv} projections (Fig.
 \ref{figs6}-\ref{figs8}).
@@
 These plots were either subjected to ambient conditions (i.e.
-controls), or enclosed in Open Top Chambers (OTCs) which rose temperature by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C.
+controls), or enclosed in Open Top Chambers (OTCs) which increased temperature by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C.
 

### 2016-10-18 R/figures.R

@@ -241,7 +241,7 @@ max_ht_plots <- function(greaus_max_ht_model,asttry_max_ht_model, ylim=c(0,35)) 
 gap_dynamics_plot <- function(gap_dynamic_model) {
   coeffs <- summarise_coefficients(gap_dynamic_model, params = c('alpha','b_otc'))
   predictions <- summarise_otc_model_predictions('gap_dynamics',gap_dynamic_model)
-  p1 <- coefficient_plot(coeffs,y_axis_labels = c('Intercept', 'otc'), xlab = 'log coefficients')
+  p1 <- coefficient_plot(coeffs,y_axis_labels = c('Intercept', 'warmed'), xlab = 'log coefficients')
   p2 <- gap_dynamics_curve(predictions)
   plot_grid(p1,p2, labels=letters[1:2], ncol = 2, label_size = 7)
 }

### 2016-10-19 manuscript/manuscript.Rnw
More edits to manuscript

@@
 Here we use field surveys and experimental manipulations to examine warming and fire effects on recruitment, growth and survival of seedlings of four dominant, evergreen fire-killed alpine shrubs.
-We find that fire substantially increased shrub seedling establishment (up to \Sexpr{density_ratios(greaus_seedling_density_model, 'Grevillea')}\textendash fold) and that experimental warming doubled tall shrub seedling growth rates and could potentially increase survival.
+We found that fire substantially increased shrub seedling establishment (up to \Sexpr{density_ratios(greaus_seedling_density_model, 'Grevillea')}\textendash fold) and that experimental warming doubled tall shrub seedling growth rates and could potentially increase survival.
 Warming had no effect on shrub recruitment, post-fire infilling of gaps by a dominant alpine grass, or competitive effects of grass on shrub seedling growth and survival.
-These findings indicate that warming, coupled with more frequent or severe fires, will likely exacerbate increases in tall evergreen shrubs by increasing recruitment, doubling growth rates, and potentially increasing shrub survival.
-Thus, warming is likely to enhance a positive feedback between woody species abundance and fire activity in alpine and tundra environments.
+These findings indicate that warming, coupled with more frequent or severe fires, will likely result in an increase in the cover and abundance of evergreen shrubs by increasing recruitment, doubling growth rates, and potentially increasing shrub survival.
+Thus, warming is likely to strengthen an existing feedback between woody species abundance and fire activity in alpine and tundra environments.
 \end{abstract}
@@
 Accurately forecasting the effects of climatic warming on vegetation dynamics requires an understanding of the mechanisms by which climate and vegetation interact.
-Most forecasting models include the direct effects of climatic conditions on species distributions, but largely ignore disturbances, the type, frequency and severity of which are also likely to change \citep{Guisan:2005gz,Tucker:2012dr}.
+Most forecasting models include the direct effects of climatic conditions on species distributions, but largely ignore disturbances, particularly their type, frequency and severity \citep{Guisan:2005gz,Tucker:2012dr}.
 Recurrent disturbance is integral to all ecosystems, strongly influencing fundamental demographic processes such as recruitment and mortality, and thus the composition and structure of plant communities and biomes worldwide \citep{Bond:2005eg, Turner:2010wo}.
-Disturbance regimes are likely to change as a consequence of climate change \citep{Turner:2010wo, Westerling:2011jp,Bradstock:2014jp}.
+Disturbance regimes are changing as a consequence of climate change \citep{Turner:2010wo, Westerling:2011jp,Bradstock:2014jp}.
 Moreover, the effects of disturbance will interact with climate change, and thus may amplify or diminish how changes to temperature and moisture affect vegetation \citep{Post:2008wz,Camac:2015hja,Enright:2015em}
@@
 For example, in addition to the direct effects of warming on fire weather, changing climate may also alter fire regimes by altering vegetation productivity, structure and plant trait composition \citep{Keeley:2012wo,Matthews:2012gia,Bowman:2014dr}.
-Feedbacks may be positive or negative, and have been documented in tropical savanna \citep{Hoffmann:2003br}, boreal forests \citep{Krawchuk:2011wr}, Mediterranean ecosystems \citep{Keeley:2012wo} and grasslands \citep{Flannigan:2009kl}.
-Change in both vegetation and fire regimes in alpine and arctic environments are also likely to be subject to feedbacks \citep{Goetz:2007gj,Wookey:2009wo}.
+Feedbacks, which may be positive or negative, have been documented in tropical savanna \citep{Hoffmann:2003br}, boreal forests \citep{Krawchuk:2011wr}, Mediterranean ecosystems \citep{Keeley:2012wo}, grasslands \citep{Flannigan:2009kl} and alpine and arctic environments \citep{Goetz:2007gj,Wookey:2009wo}.
 
@@
 Few climate-change field experiments have included disturbance as a factor \citep[but see:][]{Munier:2010cg,Graae:2010ir,Camac:2015hja}, or examined the influence of climate change on seedlings in post-disturbance environments.
-Although alpine and tundra vegetation has a strong capacity to regenerate vegetatively \citep{Billings:1968bn} there is mounting evidence that seedling regeneration is important in these ecosystems \citep{Venn:2009bw, Briceno:2015iv}, particularly for wood species \citep{Camac:2013eb,Williams:2014ud}.
-Although seedlings are highly vulnerable to climate and disturbance, they are the life stage that determines the long-term persistence of a species and its capacity to establish in new areas.
-However, little is known about how climate change affects seedling demography \citep{Briceno:2015iv}, particularly in post-disturbance conditions in alpine and tundra ecosystems.
-It is therefore imperative we understand how seedlings respond to both changing climate and disturbance regimes \citep{Walck:2011dm, Briceno:2015iv} if we wish to accurately predict future trajectories of vegetation change.
+Although alpine and tundra vegetation has a strong capacity to regenerate vegetatively \citep{Billings:1968bn} there is mounting evidence that seedling regeneration is important in these ecosystems \citep{Venn:2009bw, Briceno:2015iv}, particularly for woody species \citep{Camac:2013eb,Williams:2014ud}.
+Despite seedlings being highly vulnerable to climate and disturbance, they are the life stage that determines the long-term persistence of a species and its capacity to establish in new areas \citep{Walck:2011dm}.
+Little is known about how climate change affects seedling demography \citep{Briceno:2015iv}, particularly in alpine and tundra post-disturbance conditions.
+This is unfortunate, because if we wish to accurately predict future trajectories of vegetation change in these ecosystems, it is imperative we understand how seedlings respond to both changing climate and disturbance regimes \citep{Walck:2011dm, Briceno:2015iv}.
 
-\section*{Conceptual model}
+\section*{Conceptual feedback model}
 
@@
 The frequency and extent of wildfires have also increased in these environments over recent decades, a trend expected to continue \citep{Westerling:2006ve, Flannigan:2009kl, Qiu:2009th, Bradstock:2014jp}.
-In alpine and tundra ecosystems, current evidence from long-term and experimental studies indicate that shrub recruitment and encroachment is highest in disturbed areas \citep{Williams:1987ws, Batllori:2009fe, Frost:2013tm} and lowest in areas with minimal bare ground cover \citep[e.g.
-undisturbed alpine tussock grassland;][]{Williams:1987ws}.
-Current evidence indicates that climatic warming is likely to increase woody species growth rates \citep{Arft:1999iz, Elmendorf:2012ba, MyersSmith:2011fp} and that shrubs are the most flammable vegetation component of these ecosystems \citep{Williams:2006tl, Higuera:2009ga, Fraser:NHT5Kx_W}.
+In alpine and tundra ecosystems, current evidence from long-term and experimental studies indicate that shrub recruitment and encroachment is highest in disturbed areas \citep{Williams:1987ws, Batllori:2009fe, Frost:2013tm} and lowest in areas with minimal bare ground cover, such as undisturbed alpine tussock grassland \citep{Williams:1987ws}.
+Current evidence indicates that climatic warming is likely to increase the growth rates of woody species \citep{Arft:1999iz, Elmendorf:2012ba, MyersSmith:2011fp} and that shrubs are the most flammable vegetation component of these ecosystems \citep{Williams:2006tl, Higuera:2009ga, Fraser:NHT5Kx_W}.
 Thus, increases in the cover of woody species in alpine and arctic environments is likely to increase the flammability of these ecosystems.
 
-The combination of this evidence indicates a potential positive feedback between warmer temperatures, woody species dynamics, and fire activity in alpine and tundra environments (Fig.
+The results of these studies suggest a positive feedback between warmer temperatures, woody species and fire in alpine and tundra environments (Fig.
 \ref{fig1}a).
 Warmer temperatures are likely to lead to more frequent or severe fire.
-This is likely to increase woody species recruitment opportunities (i.e.
-more bare ground) and exacerbate both shrub thickening within shrub communities and shrub expansion into shrub-free communities such as grassland and herbfields \citep{Racine:2004tn,Lantz:2013fh}.
-If this effect is coupled with increased shrub seedling growth and mortality, highly flammable fuels will accumulate faster and have higher landscape connectivity, ultimately leading to further increases in the likelihood of fire \citep{Higuera:2008dw, Higuera:2009ga,MyersSmith:2011fp}.
+This is likely to increase recruitment opportunities for woody species and increase shrub thickening both within and beyond shrub boundaries \citep{Racine:2004tn,Lantz:2013fh}.
+If this effect is coupled with an increase in the growth and mortality of shrub seedlings, highly flammable fuels will accumulate faster, have higher landscape connectivity, and ultimately lead to increases in the likelihood of fire \citep{Higuera:2008dw, Higuera:2009ga,MyersSmith:2011fp}.
 Thus, warming could strengthen an existing climate-disturbance feedback, that not only has the potential to cause rapid changes in the composition and structure of alpine and arctic vegetation, but also has serious social, biodiversity and carbon sequestration consequences \citep{Mack:2011ew}.
@@
 Solid lines = known mechansims; Dashed lines = mechansims that may exacerbate or diminish this feedback but which we have a paucity of information on.
-(b) Updated feedback based on new evidence obtained from experimental and field surveys conducted in this study.
+(b) Feedback updated based on new evidence obtained from experimental and field surveys conducted in this study.
 Also includes an example of a potential, but unlikely, effect (red dashed line) that could break this feedback.
@@
 For example, we have little information on whether fire will stimulate shrub seedling recruitment in alpine or tundra ecosystems, nor do we know how rates of seedling growth and survival will be affected under warmer, more exposed, post-fire conditions.
-Furthermore, we do not know how tussock grasslands (a dominant structural vegetation type in arctic and alpine ecosystems) will respond to warmer post-fire conditions, whether grasses have competitive, facilitative or no effect on shrub seedling vital rates, or whether such effects are altered by warmer post-fire conditions.
+Furthermore, we do not know how tussock grasslands will respond to warmer post-fire conditions, how grasses affect vital rates of shrub seedlings, or whether such effects are altered by warmer post-fire conditions.
 
 Here, we examine these unknowns and their impacts on this proposed feedback between climatic warming, shrubs and fire in Australian alpine vegetation.
-As with other tundra ecosystems, Australian alpine landscapes consist of a matrix of grass-herb dominanted vegetation types, and shrub-dominanted types \citep{Williams:2014ud}.
+As with other tundra ecosystems, Australian alpine landscapes consist of a mosaic of vegetation types dominated by either herbaceous or woody species \citep{Williams:2014ud}.
 They have also experienced significant changes in climate.
@@
 These climatic changes have been correlated with a 10 to 20\% increase in shrub cover and a 25\% decline in graminoids cover \citep{Wahren:2013ji}.
-Much of the Australian Alps has also been burnt by recent (2003 and 2006) wildfires, the frequency and/or severity of which are expected to increase as a consequence of climate change \citep{Bradstock:2014jp,Williams:2014ud}.
+Much of the Australian Alps has also been burnt by recent (2003 and 2006) wildfires, the frequency and/or severity of which are expected to increase as a consequence of further climate change \citep{Bradstock:2014jp,Williams:2014ud}.
 
@@
 Under global warming, this plant community is likely to encroach upon grasslands, wetlands and herbfields \citep[including the nationally endangered snow patch herbfields;][]{Williams:2015gj}, and is itself susceptible to shrub thickening \citep{McDougall:2003gb}.
-To determine drivers of shrub establishment and how warmer post-fire conditions affect shrub vital rates, we combined field observations of post-fire shrub seedling density with a warming experiment that used seedlings of four dominant Australian evergreen fire-killed shrubs: \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub),  \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub).
+To identify some of the biotic and abiotic factors that affect shrub establishment and how warmer post-fire conditions affect shrub seedling vital rates, we combined field observations of post-fire seedling densities with a warming experiment that used seedlings of four dominant Australian evergreen fire-killed shrubs: \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub).
 
-This combination of data sources allowed us to quantify some of the outstanding unknowns in the hypothesised climate-fire-shrub feedback loop (Fig. \ref{fig1}a). Specifically, we examined: 
+This combination of data allowed us to quantify the following unknowns in the hypothesised climate-fire-shrub feedback loop (Fig. \ref{fig1}a): 
 \begin{enumerate}\itemsep 0pt
-  \item Landscape scale drivers of shrub seedling establishment.
-  \item How warmer post-fire conditions influence rates of grass recovery and shrub seedling recruitment, growth and mortality.
-  \item How proximity to grass affects shrub seedling demographic rates; and whether such grass effects change under warmer post-fire conditions.
+  \item Does fire stimulate shrub recruitment?
+  \item Does warming alter the recruitment of shrubs post-fire?
+  \item Does warming alter the growth and survival of shrub seedlings?
+  \item Does proximity to tussock grass affect the growth and survival of shrub seedlings?
+  \item Does warming alter the post-fire regeneration of tussock grass?
 \end{enumerate}
@@
 
-A summary of study sites, experimental design and approach to data analysis is given below. Full details are presented in the Supporting Information.
+Below, we provide a summary of the experimental design and analysis. Full details are presented in the Supporting Information.
 
@@
 
-The Australian Alps are an ideal place to examine potential fire-climate-vegetation feedbacks in high mountain environments because Australian alpine ecosystems have all the elements needed to demonstrate such potential feedbacks.
-Here, the Alps are subject to recurrent landscape fire, and were burnt extensively in 2003 \citep{Williams:2006tl,Williams:2014ud}.
-The vegetation consists of a range of life forms (graminoids, other herbs, shrubs), and variation in life-form dominance between and within vegetation types is a strong determinant of landscape flammability \citep{Williams:2006tl,Fraser:NHT5Kx_W}.
-There is also evidence that both fire regimes and the relative abundance of lifeforms in the Australian Alps have changed as a consequence of recent climate change \citep{Bradstock:2014jp,Wahren:2013ji,Camac:2015hja}.
+The Australian Alps are an ideal place to examine potential fire-climate-vegetation feedbacks in high mountain environments because they have all the elements needed to demonstrate such potential feedbacks.
+The Alps are subject to recurrent landscape fire, and were burnt extensively in 2003 \citep{Williams:2006tl,Williams:2014ud}.
+The vegetation consists of a range of life forms (graminoids, other herbs, shrubs), the abundance of which strongly determines landscape flammability \citep{Williams:2006tl,Fraser:NHT5Kx_W}.
+There is also evidence that both fire regimes and the relative abundance of life forms in the Australian Alps have changed as a consequence of recent climate change \citep{Bradstock:2014jp,Wahren:2013ji,Camac:2015hja}.
 
@@
 Relative to tussock grassland, open heathland is a highly flammable plant community \citep{Williams:2006tl,Fraser:NHT5Kx_W}.
-A consequence of this differential flammability lead to the 2003 wildfires on the Bogong High Plains burning approximately 60\% of open heathland and only 13\% of tussock grasslands.
+A consequence of this differential flammability led to the 2003 wildfires on the Bogong High Plains burning approximately 60\% of open heathland and only 13\% of tussock grasslands.
 Following these fires, seedling regeneration of \textit{Grevillea australis} and other shrubs was prolific \citep{Williams:2014ud}.
@@
 These sites consisted of 30 burnt sites and 10 sites known to be unburnt for over 70 years.
-The disparity in sampling effort between burnt and unburnt was due to very few open heathland sites of sufficient size (0.25 ha) remaining unburnt after both the 2003 and 2006 fires.
+The disparity in sample number between burnt and unburnt treatments was due to very few open heathland sites of sufficient size (0.25 ha) remaining unburnt after both the 2003 and 2006 fires.
 In the summer of 2011-12, at each site, seedling density/m$^2$ was estimated using 40 quadrats, each 1 m$^2$, that were evenly distributed in groups of 10 along four 50 m transects, with 10 m between transect lines, subsampling an area of 2000 m$^2$.
@@
 Here, we assess the interactive effects of warming and grass proximity on the growth and survival of \textit{Grevillea} seedlings transplanted into various sized inter-tussock gaps.
-We detected a strong positive effect of warming treatment on growth rates and a marginal, yet not significant, decrease in mortality (Fig.
+We detected a strong positive effect of warming treatment on growth rates and a marginally non-significant decrease in mortality (Fig.
 \ref{fig5}).
@@
 Despite having little impact on shrub seedling growth and mortality rates, tussock grass may still limit shrub recruitment, and thus, shrub expansion, by infilling post-fire bare ground gaps (whether by vegetative growth or seedlings) faster under warmer conditions.
-Using six years of post-fire inter-tussock gap size changes in warmed and unwarmed plots we found that gaps were being infilled by tussock grasses (Fig.
+Using six years of post-fire inter-tussock gap size changes in warmed and unwarmed plots, we found that gaps were being infilled by tussock grasses (Fig.
 \ref{fig6}).
@@
 While high cover of tussock grass is known to inhibit the establishment of shrub seedlings \citep{Williams:1987ws}, more frequent and severe fire caused by changing climate, will create bare ground required for shrub seedlings to establish.
-Our results, highlight that given a shrub seedling establishes in bare ground, its growth and survival rates are not affected by proximity to tussock grass, irrespective or warming.
-More importantly, however, our results indicate that for tall shrubs, a 1\textdegree C increase in mean growing season temperature resulted in a doubling of growth rates and a potential increase in survival.
-These demographic effects will ultimately manifest in shrub thickening within and at shrub-grassland/herbfield boundaries.
-It will also facilitate shrub establishment into non-shrubby communities by increasing seed sources at the boundary (i.e.
-more mature plants) and decreasing dispersal distances.
+Our results highlight that if a shrub seedling establishes in bare ground, its growth and survival rates are not affected by proximity to tussock grass, irrespective or warming.
+More importantly, our results indicate that for tall shrubs, a 1\textdegree C increase in mean growing season temperature resulted in a doubling of growth rates and a potential increase in survival.
+These demographic effects will manifest in shrub thickening within and at shrub community boundaries, and ultimately, facilitate shrub establishment into non-shrubby communities.
 
-Evidence from field studies of burning patterns \citep{Williams:2006tl}, quantification of fuel mass and architecture \citep{Fraser:NHT5Kx_W} and paleoecological studies \citep{Higuera:2008dw} indicate that shrubs are the most flammable component of alpine and tundra ecosystems \citep[particularly moreso then tussock grassland; ][]{Fraser:NHT5Kx_W}.
-Consequently, our results indicate that flammable fuel loads will accumulate twice as fast under a warmer environment.
-This is likely to further strengthen the feedback \citep{Higuera:2008dw, Higuera:2009ga} by increasing the frequency and severity of fires, which then creates more shrub recruitment opportunities both within and beyond shrub-grassland boundaries that may persist for decades.
+Evidence from field studies of burning patterns \citep{Williams:2006tl}, quantification of fuel mass and architecture \citep{Fraser:NHT5Kx_W} and paleoecological studies \citep{Higuera:2008dw} indicate that shrubs are the most flammable component of alpine and tundra ecosystems \citep[significantly more than tussock grassland; ][]{Fraser:NHT5Kx_W}.
+Consequently, our results indicate that flammable fuel loads will accumulate twice as fast under a warmer climate.
+This is likely to further strengthen the feedback \citep{Higuera:2008dw, Higuera:2009ga} by increasing the frequency and severity of fires, which then creates more shrub recruitment opportunities both within and beyond shrub-grassland boundaries.
+Such changes may persist for decades.
 
@@
 Furthermore, if, as hypothesised, reproductive output is related to plant size \citep{Wenk:2015ko}, then climatic warming may allow fire-killed species to reach reproductive maturity sooner, and consequently, may increase species resilience to short fire intervals.
-Nevertheless, these factors and others such as changes in snow pack \citep{Wipf:2009jh}, soil moisture \citep{Engler:2011vq} or herbivory \citep{Post:2008wz} require further research because they are likely to be altered by the interactive effects of climate and disturbance in unpredictable ways \citep{MyersSmith:2011fp}.
+Nevertheless, these factors and others such as changes in snow pack \citep{Wipf:2009jh}, soil moisture \citep{Engler:2011vq}, herbivory \citep{Post:2008wz} and adaptation \citep{Byars:2009fd, Hoffmann:2011fp} require further research because they are likely to be altered by the interactive effects of climate and disturbance in unpredictable ways \citep{MyersSmith:2011fp}.
 
-By focusing on seedlings, the life stage most vulnerable to climate and disturbance, and which determines the long-term persistence of a species and its capacity to establish in new areas \citep{Walck:2011dm, Briceno:2015iv}, our findings provide mechanistic understanding as to why shrub cover has increased, often at the expense of grasslands, in many alpine and arctic ecosystems \citep{Racine:2004tn,MyersSmith:2011fp,Frost:2013tm,Lantz:2013fh}.
+By focusing on seedlings, the life stage most vulnerable to climate and disturbance, and which determines the long-term persistence of shrub species and its capacity to establish in new areas \citep{Walck:2011dm, Briceno:2015iv}, our findings provide a possible explanation as to why shrub cover has increased in the Australian Alps, often at the expense of grasslands \citep{McDougall:2003gb,Wahren:2013ji}.
+An explanation that may also apply to other arctic and alpine ecosystems \citep{Racine:2004tn,MyersSmith:2011fp,Frost:2013tm,Lantz:2013fh}.
 But more importantly, our results provide evidence for the underlying processes that could result in a warming-fire-shrub feedback that has been hypothesized in paleoecological studies \citep{Higuera:2008dw, Higuera:2009ga}.
 Based on current observations and predictions, average global temperature has already increased by 0.85\textdegree C since 1880 and is expected to rise by as much as 4.8\textdegree C by 2100 \citep{IPCC:2013vv}.
-In alpine and tundra environments, temperatures \citep{ChapinIII:2005js}, shrub cover \citep{MyersSmith:2011fp} and the frequency and severity of fire \citep{Westerling:2006ve, Flannigan:2009kl, Qiu:2009th, Bradstock:2014jp} have all increased across the globe in the last few decades as a consequence of rising temperatures.
-These changes mean that the warming-shrub-fire feedback loop identified here is likely to have already strengthened, which could cause grasslands and other non-woody communities to transition to an alternative state with more shrubs and more fire, both of which are likely to have consequences for carbon sequestration, water supply and biodiversity.
+In alpine and tundra environments, temperatures \citep{ChapinIII:2005js}, shrub cover \citep{MyersSmith:2011fp} and the frequency and severity of fire \citep{Westerling:2006ve, Flannigan:2009kl,Qiu:2009th, Bradstock:2014jp} have all increased across the globe in the last few decades as a consequence of rising temperatures.
+These changes mean that the warming-shrub-fire feedback loop identified here is likely to have already strengthened.
+Thus, other non-woody communities will become shrubbier and more prone to fire, the effects of which, will have significant consequences for carbon sequestration, water supply and biodiversity.
 
@@
 This research was funded through Australian Research Council Linkage Grants, partnered through the Department of Sustainability and Environment, Parks Victoria, Commonwealth Scientific and Industrial Research Organisation (CSIRO).
+Plot infrastructure was supported by the Long Term Ecological Research Network.
 Funding was also received from the Centre of Excellence for Environmental Decisions (CEED) and Holsworth Research Committee.

### 2016-10-20 manuscript/manuscript.Rnw
Proof read edits

@@
 \begin{abstract}
-Climate change is expected to increase fire activity and woody plant encroachment in arctic and alpine landscapes.
-However, the extent to which these increases interact to affect changes in structure, function and composition of alpine ecosystems is unknown.
-Here we use field surveys and experimental manipulations to examine warming and fire effects on recruitment, growth and survival of seedlings of four dominant, evergreen fire-killed alpine shrubs.
-We found that fire substantially increased shrub seedling establishment (up to \Sexpr{density_ratios(greaus_seedling_density_model, 'Grevillea')}\textendash fold) and that experimental warming doubled tall shrub seedling growth rates and could potentially increase survival.
-Warming had no effect on shrub recruitment, post-fire infilling of gaps by a dominant alpine grass, or competitive effects of grass on shrub seedling growth and survival.
-These findings indicate that warming, coupled with more frequent or severe fires, will likely result in an increase in the cover and abundance of evergreen shrubs by increasing recruitment, doubling growth rates, and potentially increasing shrub survival.
-Thus, warming is likely to strengthen an existing feedback between woody species abundance and fire activity in alpine and tundra environments.
+Climate change is expected to increase fire and woody plant encroachment in arctic and alpine landscapes.
+However, the extent to which these increases interact to affect the structure, function and composition of alpine ecosystems is largely unknown.
+Here we use field surveys and experimental manipulations to examine how warming and fire affect recruitment, seedling growth and seedling survival in four dominant Australian alpine shrubs.
+We found that fire increased establishment of shrub seedlings by as much as \Sexpr{density_ratios(greaus_seedling_density_model, 'Grevillea')}\textendash fold.
+Experimental warming also doubled growth rates of tall shrub seedlings and could potentially increase their survival.
+By contrast, warming had no effect on shrub recruitment, post-fire tussock regeneration, or how tussock grass affected shrub seedling growth and survival.
+These findings indicate that warming, coupled with more frequent or severe fires, will likely result in an increase in the cover and abundance of evergreen shrubs.
+Given shrubs are one of the most flammable components in alpine and tundra environments, warming is likely to strengthen an existing feedback between woody species abundance and fire in these ecosystems.
 \end{abstract}
@@
 Recurrent disturbance is integral to all ecosystems, strongly influencing fundamental demographic processes such as recruitment and mortality, and thus the composition and structure of plant communities and biomes worldwide \citep{Bond:2005eg, Turner:2010wo}.
-Disturbance regimes are changing as a consequence of climate change \citep{Turner:2010wo, Westerling:2011jp,Bradstock:2014jp}.
-Moreover, the effects of disturbance will interact with climate change, and thus may amplify or diminish how changes to temperature and moisture affect vegetation \citep{Post:2008wz,Camac:2015hja,Enright:2015em}
+Disturbance regimes are already changing as a consequence of climate change \citep{Turner:2010wo, Westerling:2011jp,Bradstock:2014jp}.
+It is therefore imperative we understand how the effects of disturbance will interact with climate change, and whether such effects amplify or diminish how vegetation responds to changing temperature and moisture \citep{Post:2008wz,Camac:2015hja,Enright:2015em}
 
-Feedbacks between climate change, vegetation and fire will be particularly important determinants of the trajectory of ecosystem change in coming decades \citep{Bowman:2009kp}.
-For example, in addition to the direct effects of warming on fire weather, changing climate may also alter fire regimes by altering vegetation productivity, structure and plant trait composition \citep{Keeley:2012wo,Matthews:2012gia,Bowman:2014dr}.
-Feedbacks, which may be positive or negative, have been documented in tropical savanna \citep{Hoffmann:2003br}, boreal forests \citep{Krawchuk:2011wr}, Mediterranean ecosystems \citep{Keeley:2012wo}, grasslands \citep{Flannigan:2009kl} and alpine and arctic environments \citep{Goetz:2007gj,Wookey:2009wo}.
+Understanding the feedback between climate, vegetation and fire will be particularly important in order to accurately predict the trajectory of ecosystem change in coming decades \citep{Bowman:2009kp}.
+In addition to the direct effects of warming on fire weather, changing climate may also indirectly alter fire regimes by altering vegetation productivity, structure and composition \citep{Keeley:2012wo,Matthews:2012gia,Bowman:2014dr}.
+Such feedbacks may be positive or negative, and have been documented in tropical savanna \citep{Hoffmann:2003br}, boreal forests \citep{Krawchuk:2011wr}, grasslands \citep{Flannigan:2009kl} and alpine and arctic environments \citep{Goetz:2007gj,Wookey:2009wo}.
 
-Alpine and arctic vegetation are particularly vulnerable to the effects of changing climate \citep{Engler:2011vq,Dullinger:2012cg, Elmendorf:2012kz}.
-Field manipulative experiments have been widely used to investigate the potential effects of climate change in alpine and tundra plant communities.
-These experiments have provided invaluable information on how climate directly influences plant phenology \citep{Hoffmann:2010kf, Dorji:2013tt}, reproduction \citep{Klady:2011dw}, morphology \citep{Hudson:2009cn}, growth \citep{Hollister:2005vf}, floristic composition \citep{Elmendorf:2012kz} and biotic interactions \citep{Klanderud:2005uw}.
+Alpine and arctic vegetation are considered to be particularly vulnerable to the effects of changing climate \citep{Engler:2011vq,Dullinger:2012cg, Elmendorf:2012kz}.
+In these ecosystems, field manipulative experiments have shown that climate directly influences plant phenology \citep{Hoffmann:2010kf, Dorji:2013tt}, reproduction \citep{Klady:2011dw}, morphology \citep{Hudson:2009cn}, growth \citep{Hollister:2005vf}, floristic composition \citep{Elmendorf:2012kz} and biotic interactions \citep{Klanderud:2005uw}.
+However, most studies in these ecosystems have focused on mature plant responses in undisturbed vegetation \citep{Briceno:2015iv}.
+Few have included disturbance as a factor \citep[but see:][]{Munier:2010cg,Graae:2010ir,Camac:2015hja}, or examined the influence of climate change on seedlings in post-disturbance environments.
+As a consequence, little is known about how climate affect seedling demographic rates \citep{Briceno:2015iv}, particularly in alpine and tundra post-disturbance conditions.
+This is despite mounting evidence that seedling regeneration is important in alpine and arctic ecosystems \citep{Venn:2009bw, Briceno:2015iv}, particularly for woody species \citep{Camac:2013eb,Williams:2014ud}.
+Seedlings are the life stage that determines the long-term persistence of a species as well as its capacity to establish in new areas \citep{Walck:2011dm}.
+As such, in order to accurately predict future trajectories of vegetation change in alpine and arctic ecosystems, it is impertative we understand how seedlings respond to both changing climate and disturbance regimes \citep{Walck:2011dm, Briceno:2015iv}.
 
-However, most studies in arctic and alpine areas have focused on plant responses in mature, undisturbed vegetation \citep{Briceno:2015iv}.
-Few climate-change field experiments have included disturbance as a factor \citep[but see:][]{Munier:2010cg,Graae:2010ir,Camac:2015hja}, or examined the influence of climate change on seedlings in post-disturbance environments.
-Although alpine and tundra vegetation has a strong capacity to regenerate vegetatively \citep{Billings:1968bn} there is mounting evidence that seedling regeneration is important in these ecosystems \citep{Venn:2009bw, Briceno:2015iv}, particularly for woody species \citep{Camac:2013eb,Williams:2014ud}.
-Despite seedlings being highly vulnerable to climate and disturbance, they are the life stage that determines the long-term persistence of a species and its capacity to establish in new areas \citep{Walck:2011dm}.
-Little is known about how climate change affects seedling demography \citep{Briceno:2015iv}, particularly in alpine and tundra post-disturbance conditions.
-This is unfortunate, because if we wish to accurately predict future trajectories of vegetation change in these ecosystems, it is imperative we understand how seedlings respond to both changing climate and disturbance regimes \citep{Walck:2011dm, Briceno:2015iv}.
-
-\section*{Conceptual feedback model}
+\section*{Feedback between climatic warming, fire and shrubs}
 
 In alpine and arctic ecosystems, warming experiments and long-term monitoring have documented significant increases in the growth and cover of woody species \citep{Sturm:2001vd,Walker:2006ht,MyersSmith:2011fp}.
-The frequency and extent of wildfires have also increased in these environments over recent decades, a trend expected to continue \citep{Westerling:2006ve, Flannigan:2009kl, Qiu:2009th, Bradstock:2014jp}.
-In alpine and tundra ecosystems, current evidence from long-term and experimental studies indicate that shrub recruitment and encroachment is highest in disturbed areas \citep{Williams:1987ws, Batllori:2009fe, Frost:2013tm} and lowest in areas with minimal bare ground cover, such as undisturbed alpine tussock grassland \citep{Williams:1987ws}.
-Current evidence indicates that climatic warming is likely to increase the growth rates of woody species \citep{Arft:1999iz, Elmendorf:2012ba, MyersSmith:2011fp} and that shrubs are the most flammable vegetation component of these ecosystems \citep{Williams:2006tl, Higuera:2009ga, Fraser:NHT5Kx_W}.
-Thus, increases in the cover of woody species in alpine and arctic environments is likely to increase the flammability of these ecosystems.
+The frequency and extent of wildfires in these environments have also increased over recent decades, a trend expected to continue \citep{Westerling:2006ve, Flannigan:2009kl, Qiu:2009th, Bradstock:2014jp}.
+Current evidence from these ecosystems indicate that shrub recruitment and encroachment is highest in disturbed areas and lowest in areas with minimal bare ground cover \citep{Williams:1987ws, Batllori:2009fe, Frost:2013tm}.
+Evidence also suggests that climatic warming is likely to increase growth rates of woody species \citep{Arft:1999iz, Elmendorf:2012ba, MyersSmith:2011fp} and that shrubs are potentially the most flammable vegetation component in these ecosystems \citep{Williams:2006tl, Higuera:2009ga, Fraser:NHT5Kx_W}.
+Thus, increases in the cover of woody species in alpine and arctic environments may likely to increase the flammability of these ecosystems.
 
-The results of these studies suggest a positive feedback between warmer temperatures, woody species and fire in alpine and tundra environments (Fig.
+The results of these studies suggest a positive feedback could exist between warmer temperatures, woody species and fire in alpine and tundra environments (Fig.
 \ref{fig1}a).
-Warmer temperatures are likely to lead to more frequent or severe fire.
-This is likely to increase recruitment opportunities for woody species and increase shrub thickening both within and beyond shrub boundaries \citep{Racine:2004tn,Lantz:2013fh}.
-If this effect is coupled with an increase in the growth and mortality of shrub seedlings, highly flammable fuels will accumulate faster, have higher landscape connectivity, and ultimately lead to increases in the likelihood of fire \citep{Higuera:2008dw, Higuera:2009ga,MyersSmith:2011fp}.
+Specifically, warmer temperatures may lead to more frequent and severe fire, which in turn, may increase recruitment opportunities (i.e.
+more bare ground) for woody species and increase shrub thickening both within and beyond shrub boundaries \citep{Racine:2004tn,Lantz:2013fh}.
+If this effect is coupled with an increase in the growth and survival of shrub seedlings, highly flammable fuels will accumulate at a faster rate, have higher landscape connectivity, and ultimately lead to increases in the likelihood of fire.
+The consequence of which will further increase shrub recruitment opportunities.
 Thus, warming could strengthen an existing climate-disturbance feedback, that not only has the potential to cause rapid changes in the composition and structure of alpine and arctic vegetation, but also has serious social, biodiversity and carbon sequestration consequences \citep{Mack:2011ew}.
@@
 These climatic changes have been correlated with a 10 to 20\% increase in shrub cover and a 25\% decline in graminoids cover \citep{Wahren:2013ji}.
-Much of the Australian Alps has also been burnt by recent (2003 and 2006) wildfires, the frequency and/or severity of which are expected to increase as a consequence of further climate change \citep{Bradstock:2014jp,Williams:2014ud}.
+Much of the Australian Alps has also been burnt by recent (2003 and 2006) wildfires, the frequency and severity of which are expected to increase as a consequence of further climate change \citep{Bradstock:2014jp,Williams:2014ud}.
 
-We took advantage of recent fires in alpine open heathland, a plant community that occupies ca.
+In this study we took advantage of recent fires in alpine open heathland, a plant community that occupies ca.
 25\% of the Australian alpine landscape \citep{Williams:2014ud}.
-Under global warming, this plant community is likely to encroach upon grasslands, wetlands and herbfields \citep[including the nationally endangered snow patch herbfields;][]{Williams:2015gj}, and is itself susceptible to shrub thickening \citep{McDougall:2003gb}.
+Under global warming, open heathland is likely to encroach upon grasslands, wetlands and herbfields \citep[including the nationally endangered snow patch herbfields;][]{Williams:2015gj}, and is itself susceptible to shrub thickening \citep{McDougall:2003gb}.
 To identify some of the biotic and abiotic factors that affect shrub establishment and how warmer post-fire conditions affect shrub seedling vital rates, we combined field observations of post-fire seedling densities with a warming experiment that used seedlings of four dominant Australian evergreen fire-killed shrubs: \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub).
@@
 
-The Australian Alps are an ideal place to examine potential fire-climate-vegetation feedbacks in high mountain environments because they have all the elements needed to demonstrate such potential feedbacks.
-The Alps are subject to recurrent landscape fire, and were burnt extensively in 2003 \citep{Williams:2006tl,Williams:2014ud}.
-The vegetation consists of a range of life forms (graminoids, other herbs, shrubs), the abundance of which strongly determines landscape flammability \citep{Williams:2006tl,Fraser:NHT5Kx_W}.
-There is also evidence that both fire regimes and the relative abundance of life forms in the Australian Alps have changed as a consequence of recent climate change \citep{Bradstock:2014jp,Wahren:2013ji,Camac:2015hja}.
+The Australian Alps are an ideal place to examine the proposed fire-climate-vegetation feedback because they have all the elements needed to demonstrate such a feedback.
+The Alps are subject to recurrent landscape fire (approximately every 50 to 100 years), and were burnt extensively in 2003 \citep{Williams:2006tl,Williams:2014ud}.
+The vegetation consists of a range of life forms (graminoids, forbs and shrubs), of which shrub abundance strongly determines landscape flammability \citep{Williams:2006tl,Fraser:NHT5Kx_W}.
+There is also evidence that both fire regimes \citep{Bradstock:2014jp} and the relative abundance of life forms in the Australian Alps have changed as a consequence of recent climate change \citep{Wahren:2013ji,Camac:2015hja}.
 
-We investigated shrub-grass dynamics in open heathland vegetation on the Bogong High Plains, Australia.
+We investigated shrub-grass dynamics in open heathland on the Bogong High Plains, Australia.
 Open heathland is a common Australian alpine vegetation type, occupying approximately 25\% of the treeless landscapes above 1600m in the Australian Alps.
-It is dominated by the shrub \textit{Grevillea australis} (shrub cover generally 20\textendash 60\%) with inter-shrub regions occupied by snow grasses (\textit{Poa hiemata} and \textit{Poa costiniana}) and other herbaceous species.
 It is an ecotone between closed heathland (\textgreater 70\% shrub cover) and tussock grassland (shrub cover < 20\%).
+Open heathland is dominated by the shrub \textit{Grevillea australis} (shrub cover generally 20\textendash 60\%) with inter-shrub regions occupied by snow grasses (\textit{Poa hiemata} and \textit{Poa costiniana}) and other herbaceous species.
 \textit{Grevillea australis} is an obligate seeding shrub, and establishment of seedlings is dependent on disturbance creating bare ground in the grass\textendash herb sward \citep{Williams:1987ws}.
@@
 \section*{Discussion}
-\subsection*{Strengthening of the warming-shrub-fire feedback}
 
@@
 \ref{fig1}b).
-In particular, we have shown how these interactions strengthen a hypothesized feedback loop that can rapidly increase shrub cover in the Australian Alps, and potentially other alpine and tundra ecosystems \citep{Higuera:2008dw}.
-While high cover of tussock grass is known to inhibit the establishment of shrub seedlings \citep{Williams:1987ws}, more frequent and severe fire caused by changing climate, will create bare ground required for shrub seedlings to establish.
-Our results highlight that if a shrub seedling establishes in bare ground, its growth and survival rates are not affected by proximity to tussock grass, irrespective or warming.
-More importantly, our results indicate that for tall shrubs, a 1\textdegree C increase in mean growing season temperature resulted in a doubling of growth rates and a potential increase in survival.
-These demographic effects will manifest in shrub thickening within and at shrub community boundaries, and ultimately, facilitate shrub establishment into non-shrubby communities.
+In particular, we have shown how these interactions strengthen a hypothesized feedback that can rapidly increase shrub cover in the Australian Alps, and potentially, other alpine and tundra ecosystems \citep{Higuera:2008dw}.
+While high cover of tussock grass is known to inhibit the establishment of shrub seedlings \citep{Williams:1987ws}, fire can create bare ground neccesary for shrub seedling establishment.
+Our results highlight that fire can increase shrub seedling establishment by as much \Sexpr{density_ratios(greaus_seedling_density_model, 'Grevillea')}\textendash fold.
+Our results also indicate that if a shrub seedling manages to establish in a bare ground patch, its rate of growth and survival are not affected by tussock grass proximity, irrespective or warming.
+More importantly, our findings indicate that for tall shrubs, a 1\textdegree C increase in mean growing season temperature will result in a doubling of growth rates and a potential increase in survival.
+The consequence of these demographic effects will likely manifest as shrub thickening within and at shrub community boundaries.
+This in turn, will facilitate shrub establishment, via increased seed pools, into non-shrubby communities.
 
-Evidence from field studies of burning patterns \citep{Williams:2006tl}, quantification of fuel mass and architecture \citep{Fraser:NHT5Kx_W} and paleoecological studies \citep{Higuera:2008dw} indicate that shrubs are the most flammable component of alpine and tundra ecosystems \citep[significantly more than tussock grassland; ][]{Fraser:NHT5Kx_W}.
+Landscape flammability, and thus fire, are also likely to increase as a result of warming effects on shrub seedling vital rates.
+Evidence from field studies of burning patterns \citep{Williams:2006tl}, quantification of fuel mass and architecture \citep{Fraser:NHT5Kx_W} and paleoecological studies \citep{Higuera:2008dw, Higuera:2009ga} indicate that shrubs are the most flammable component of alpine and tundra ecosystems \citep[significantly more than tussock grassland; ][]{Fraser:NHT5Kx_W}.
+In the Australian Alps, this differential flammability resulted in approximately 60\% of open heathland and only 13\% of tussock grasslands being burnt by the extensive 2003 wildfires.
 Consequently, our results indicate that flammable fuel loads will accumulate twice as fast under a warmer climate.
-This is likely to further strengthen the feedback \citep{Higuera:2008dw, Higuera:2009ga} by increasing the frequency and severity of fires, which then creates more shrub recruitment opportunities both within and beyond shrub-grassland boundaries.
-Such changes may persist for decades.
+Ultimately this will further strengthen the feedback between shrubs and fire by increasing the frequency and severity of fires, which in turn, will create more bare ground, and thus more shrub recruitment opportunities both within and beyond current shrub boundaries.
+Recruitment opportunities that could persist for decades \citep{Williams:2014ud}.
 
 While we have addressed several unknowns associated with this feedback between climatic warming, shrubs and fire, there are others we have not addressed that may also strengthen, weaken or break this feedback.
-The most obvious mechanism that will break this cycle involves shorter fire intervals that prevent fire-killed shrubs reaching reproductive age and thereby exhausting the seedbank \citep{Enright:2015em}.
-However, this scenario is unlikely for the majority of alpine (or tundra) landscapes, including those in Australia, where current fire intervals of 50 to 100 years would need to decrease to less than 20 years\textemdash the time estimated for the species in this study to reach reproductive maturity \citep{Williams:2008ba}.
-Furthermore, if, as hypothesised, reproductive output is related to plant size \citep{Wenk:2015ko}, then climatic warming may allow fire-killed species to reach reproductive maturity sooner, and consequently, may increase species resilience to short fire intervals.
-Nevertheless, these factors and others such as changes in snow pack \citep{Wipf:2009jh}, soil moisture \citep{Engler:2011vq}, herbivory \citep{Post:2008wz} and adaptation \citep{Byars:2009fd, Hoffmann:2011fp} require further research because they are likely to be altered by the interactive effects of climate and disturbance in unpredictable ways \citep{MyersSmith:2011fp}.
+The most obvious mechanism that will break this cycle involves short fire intervals that prevent fire-killed shrubs reaching reproductive age and thereby exhausting the seedbank \citep{Enright:2015em}.
+However, this scenario is unlikely for the majority of alpine (or tundra) landscapes, including those in Australia.
+For example, in the Australian Alps, current fire intervals of 50 to 100 years would need to decrease to less than 20 years\textemdash the time estimated for the species in this study to reach reproductive maturity \citep{Williams:2008ba}.
+Furthermore, if reproductive output is related to plant size \citep{Wenk:2015ko}, then climatic warming may allow fire-killed species to reach reproductive maturity sooner, and consequently, may increase their resilience to short fire intervals.
+Nevertheless, this and other factors such as changes in snow pack \citep{Wipf:2009jh}, soil moisture \citep{Engler:2011vq}, herbivory \citep{Post:2008wz} and adaptation \citep{Byars:2009fd, Hoffmann:2011fp} require further research because they are all likely to be altered by the interactive effects of climate and disturbance in unpredictable ways.
 
-By focusing on seedlings, the life stage most vulnerable to climate and disturbance, and which determines the long-term persistence of shrub species and its capacity to establish in new areas \citep{Walck:2011dm, Briceno:2015iv}, our findings provide a possible explanation as to why shrub cover has increased in the Australian Alps, often at the expense of grasslands \citep{McDougall:2003gb,Wahren:2013ji}.
-An explanation that may also apply to other arctic and alpine ecosystems \citep{Racine:2004tn,MyersSmith:2011fp,Frost:2013tm,Lantz:2013fh}.
-But more importantly, our results provide evidence for the underlying processes that could result in a warming-fire-shrub feedback that has been hypothesized in paleoecological studies \citep{Higuera:2008dw, Higuera:2009ga}.
-Based on current observations and predictions, average global temperature has already increased by 0.85\textdegree C since 1880 and is expected to rise by as much as 4.8\textdegree C by 2100 \citep{IPCC:2013vv}.
-In alpine and tundra environments, temperatures \citep{ChapinIII:2005js}, shrub cover \citep{MyersSmith:2011fp} and the frequency and severity of fire \citep{Westerling:2006ve, Flannigan:2009kl,Qiu:2009th, Bradstock:2014jp} have all increased across the globe in the last few decades as a consequence of rising temperatures.
-These changes mean that the warming-shrub-fire feedback loop identified here is likely to have already strengthened.
-Thus, other non-woody communities will become shrubbier and more prone to fire, the effects of which, will have significant consequences for carbon sequestration, water supply and biodiversity.
+By focusing on the life stage most vulnerable to climate and disturbance, and which determines a species capacity to establish in new areas \citep{Walck:2011dm, Briceno:2015iv}, our analyses provide a possible explanations as to why shrub cover has increased in the Australian Alps, often at the expense of grasslands \citep{McDougall:2003gb,Wahren:2013ji}.
+Explanations that may also apply to other arctic and alpine ecosystems \citep{Racine:2004tn,MyersSmith:2011fp,Frost:2013tm,Lantz:2013fh}.
+But more importantly, our results provide evidence for the underlying processes that could result in a warming-fire-shrub feedback that has been hypothesized in arctic paleoecological studies \citep{Higuera:2008dw, Higuera:2009ga}.
+Based on current observations, average global temperature has already increased by 0.85\textdegree C since 1880 and is expected to rise by as much as 4.8\textdegree C by 2100 \citep{IPCC:2013vv}.
+In alpine and tundra environments, temperatures \citep{ChapinIII:2005js}, shrub cover \citep{MyersSmith:2011fp} and the frequency and severity of fire \citep{Westerling:2006ve, Flannigan:2009kl,Qiu:2009th, Bradstock:2014jp} have all increased in the last few decades.
+These changes mean that the warming-shrub-fire feedback identified here is likely to have already strengthened.
+If this is the case, other non-woody communities will become shrubbier and more flammable, the effects of which, will have significant consequences for carbon sequestration, water supply and biodiversity.
 

### 2016-10-23 R/build.R
Removed callr dependency

@@ -57,11 +57,11 @@ run_latex <- function(filename, interaction="nonstopmode", engine="pdflatex") {
   args <- c(paste0("-interaction=", interaction),
             "-halt-on-error",
             filename)
-  callr::call_system(Sys_which(engine), args)
+  call_system(Sys_which(engine), args)
 }
 
 run_bibtex <- function(filename) {
-  callr::call_system(Sys_which("bibtex"), sub(".tex$", "", filename))
+  call_system(Sys_which("bibtex"), sub(".tex$", "", filename))
 }
 
 Sys_which <- function(x) {
@@ -70,4 +70,86 @@ Sys_which <- function(x) {
     stop(sprintf("%s not found in $PATH", x))
   }
   ret
+}
+
+##' Function imported from callr package;  makes it easy to call a
+##' system command from R and have it behave.
+##'
+##' This function uses \code{system2} to call a system command fairly
+##' portably.  What it adds is a particular way of dealing with
+##' errors.  \code{call_system} runs the command \code{command} with
+##' arguments \code{args} (and with optionally set environment
+##' variables \code{env}) and hides \emph{all} produced output to
+##' stdout and stderr.  If the command fails (currently any nonzero
+##' exit code is counted as a failure) then \code{call_system} will
+##' throw an R error giving
+##' \itemize{
+##' \item the full string of the command run
+##' \item the exit code of the command
+##' \item any \code{errmsg} attribute that might have been returned
+##' \item all output that the program produced to either stdout and
+##' stderr
+##' }
+##'
+##' This means that a successful invocation of a program produces no
+##' output while the unsuccessful invocation throws an error and
+##' prints all information to the screen (though this is delayed until
+##' failure happens).
+##'
+##'
+##' \code{call_system} also returns the contents of both stderr and
+##' stdout \emph{invisibly} so that it can be inspected if needed.
+##'
+##' The function \code{run_system} does the same thing and will be
+##' removed as soon as code that depends on it is out of use.
+##'
+##' @title Run a system command, stopping on error
+##' @param command The system command to be invoked, as a character
+##' string.  \code{\link{Sys.which}} is useful here.
+##' @param args A character vector of arguments to \code{command}
+##' @param env A character vector of name=value pairs to be set as
[…]

### 2016-10-24 manuscript/manuscript.Rnw
Final proof check

@@
 % Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
-\keywords{Bayesian models, Biotic interactions, Growth, Recruitment, Mortality, Gap dynamics}
+\keywords{Bayesian models, Biotic interactions, Gap dynamics, Growth,  Mortality, Recruitment}
 
 \begin{abstract}
-Climate change is expected to increase fire and woody plant encroachment in arctic and alpine landscapes.
+Climate change is expected to increase fire activity and woody plant encroachment in arctic and alpine landscapes.
 However, the extent to which these increases interact to affect the structure, function and composition of alpine ecosystems is largely unknown.
@@
 Evidence also suggests that climatic warming is likely to increase growth rates of woody species \citep{Arft:1999iz, Elmendorf:2012ba, MyersSmith:2011fp} and that shrubs are potentially the most flammable vegetation component in these ecosystems \citep{Williams:2006tl, Higuera:2009ga, Fraser:NHT5Kx_W}.
-Thus, increases in the cover of woody species in alpine and arctic environments may likely to increase the flammability of these ecosystems.
+Thus, increases in the cover of woody species in alpine and arctic environments may increase the flammability of these ecosystems.
 
@@
 For both shrub species, we detected no change in seedling density along a \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m elevational gradient \citep[equivalent to a \Sexpr{round(0.77 * (200/100),1)}\textdegree C change in mean temperature;][]{Brown:1989vx}.
-This is consistent with the field warming experiment (see below) which indicated rates of recruitment and mortality in \textit{Grevillea} (Fig.
-\ref{figs3}) is largely insensitive to a \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C change in temperature.
+This is consistent with the field warming experiment (see below) which indicated rates of recruitment (Fig.
+\ref{figs3}) and mortality in \textit{Grevillea} is largely insensitive to a \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C change in temperature.
 We detected no strong effect of Topographic Wetness Index for either species.
@@
 \subsection*{Experimental warming and shrub seedling growth}
-Because seedlings were more abundant in burnt vegetation, we investigated how warmer post-fire conditions affected seedling growth and mortality rates.
+Given that seedlings were more abundant in burnt vegetation, we investigated how warmer post-fire conditions affected seedling growth and mortality rates.
 Seedlings of \textit{Grevillea}, \textit{Asterolasia}, \textit{Phebalium} and \textit{Prostanthera} emerging after a wildfire were transplanted into experimentally burnt plots.
@@
 Consequently, our results indicate that flammable fuel loads will accumulate twice as fast under a warmer climate.
-Ultimately this will further strengthen the feedback between shrubs and fire by increasing the frequency and severity of fires, which in turn, will create more bare ground, and thus more shrub recruitment opportunities both within and beyond current shrub boundaries.
-Recruitment opportunities that could persist for decades \citep{Williams:2014ud}.
+Ultimately this will further strengthen the feedback between shrubs and fire by increasing the frequency and severity of fires, which in turn, will create more bare ground, and thus more shrub recruitment opportunities both within and beyond current shrub boundaries \textemdash Recruitment opportunities that could persist for decades \citep{Williams:2014ud}.
 
-While we have addressed several unknowns associated with this feedback between climatic warming, shrubs and fire, there are others we have not addressed that may also strengthen, weaken or break this feedback.
+While we have addressed several unknowns between climatic warming, shrubs and fire, there are others we have not addressed that may also strengthen, weaken or break this feedback.
 The most obvious mechanism that will break this cycle involves short fire intervals that prevent fire-killed shrubs reaching reproductive age and thereby exhausting the seedbank \citep{Enright:2015em}.
@@
 
-By focusing on the life stage most vulnerable to climate and disturbance, and which determines a species capacity to establish in new areas \citep{Walck:2011dm, Briceno:2015iv}, our analyses provide a possible explanations as to why shrub cover has increased in the Australian Alps, often at the expense of grasslands \citep{McDougall:2003gb,Wahren:2013ji}.
-Explanations that may also apply to other arctic and alpine ecosystems \citep{Racine:2004tn,MyersSmith:2011fp,Frost:2013tm,Lantz:2013fh}.
+By focusing on the life stage most vulnerable to climate and disturbance, and which determines a species capacity to establish in new areas \citep{Walck:2011dm, Briceno:2015iv}, our analyses provide a possible explanation as to why shrub cover is increasing in the Australian Alps, often at the expense of grasslands \citep{McDougall:2003gb,Wahren:2013ji}.
+An explanation that may also apply to other arctic and alpine ecosystems \citep{Racine:2004tn,MyersSmith:2011fp,Frost:2013tm,Lantz:2013fh}.
 But more importantly, our results provide evidence for the underlying processes that could result in a warming-fire-shrub feedback that has been hypothesized in arctic paleoecological studies \citep{Higuera:2008dw, Higuera:2009ga}.

### 2017-01-03 manuscript/manuscript.Rnw
Problem with compiling pdf with line numbers using older versions of TeX Live (see #3 and annakrystalli/OpenConBerlin_ReproHack#2). Temp solution is to remove line numbers
Final edits for GCB

@@
-\documentclass[9pt,twocolumn,twoside, lineno]{gcb}
+\documentclass[9pt,twocolumn,twoside]{gcb}
 % Use the lineno option to display guide line numbers if required.
@@
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,b]{James S Camac}
+\author[a,b,*]{James S Camac}
 \author[c]{Richard J Williams} 
@@
 \affil[b]{Research Centre for Applied Alpine Ecology, La Trobe University, Melbourne, Victoria 3086, Australia}
-\affil[c]{CSIRO Land and Water, Tropical Ecosystems Research Centre, PMB 44 Winnellie, Northern Territory 0821, Australia}
+\affil[c]{Research Institute for the Environment and Livelihoods, Charles Darwin University, Darwin, Northern Territory 0909, Australia}
 \affil[d]{Bio21 Institute, School of BioSciences, The University of Melbourne, Victoria 3010, Australia}
+\affil[*]{Current address: The Centre of Excellence for Biosecurity Risk Analysis, School of BioSciences, The University of Melbourne, Melbourne, Victoria 3010, Australia}
 
@@
 Few have included disturbance as a factor \citep[but see:][]{Munier:2010cg,Graae:2010ir,Camac:2015hja}, or examined the influence of climate change on seedlings in post-disturbance environments.
-As a consequence, little is known about how climate affect seedling demographic rates \citep{Briceno:2015iv}, particularly in alpine and tundra post-disturbance conditions.
+As a consequence, little is known about how climate affects seedling demographic rates \citep{Briceno:2015iv}, particularly in alpine and tundra post-disturbance conditions.
 This is despite mounting evidence that seedling regeneration is important in alpine and arctic ecosystems \citep{Venn:2009bw, Briceno:2015iv}, particularly for woody species \citep{Camac:2013eb,Williams:2014ud}.
 Seedlings are the life stage that determines the long-term persistence of a species as well as its capacity to establish in new areas \citep{Walck:2011dm}.
-As such, in order to accurately predict future trajectories of vegetation change in alpine and arctic ecosystems, it is impertative we understand how seedlings respond to both changing climate and disturbance regimes \citep{Walck:2011dm, Briceno:2015iv}.
+As such, in order to accurately predict future trajectories of vegetation change in alpine and arctic ecosystems, it is imperative we understand how seedlings respond to both changing climate and disturbance regimes \citep{Walck:2011dm, Briceno:2015iv}.
 
@@
 \caption{Conceptual diagram illustrating positive warming-shrub-fire feedback.
-(a) Hypothesised positive feedback loop between fire, climatic warming and landscape flammability.
+(a) Hypothesised positive feedback loop between fire, climatic warming and shrub demographics.
 Solid lines = known mechansims; Dashed lines = mechansims that may exacerbate or diminish this feedback but which we have a paucity of information on.
@@
 Since 1979, mean growing season temperatures have risen by approximately 0.4\textdegree C and annual precipitation has fallen by 6\% \citep{Wahren:2013ji}, with a consequent decline in snow pack depth \citep{SanchezBayo:2013ci}.
-These climatic changes have been correlated with a 10 to 20\% increase in shrub cover and a 25\% decline in graminoids cover \citep{Wahren:2013ji}.
+These climatic changes have been correlated with a 10 to 20\% increase in shrub cover and a 25\% decline in graminoid cover \citep{Wahren:2013ji}.
 Much of the Australian Alps has also been burnt by recent (2003 and 2006) wildfires, the frequency and severity of which are expected to increase as a consequence of further climate change \citep{Bradstock:2014jp,Williams:2014ud}.
@@
 The Australian Alps are an ideal place to examine the proposed fire-climate-vegetation feedback because they have all the elements needed to demonstrate such a feedback.
-The Alps are subject to recurrent landscape fire (approximately every 50 to 100 years), and were burnt extensively in 2003 \citep{Williams:2006tl,Williams:2014ud}.
+The Alps are subject to recurrent landscape fire (approximately every 50 to 100 years), and were burnt extensively in 2003 \citep[Fig.
+S1;][]{Williams:2006tl,Williams:2014ud}.
 The vegetation consists of a range of life forms (graminoids, forbs and shrubs), of which shrub abundance strongly determines landscape flammability \citep{Williams:2006tl,Fraser:NHT5Kx_W}.
@@
 We investigated shrub-grass dynamics in open heathland on the Bogong High Plains, Australia.
-Open heathland is a common Australian alpine vegetation type, occupying approximately 25\% of the treeless landscapes above 1600m in the Australian Alps.
-It is an ecotone between closed heathland (\textgreater 70\% shrub cover) and tussock grassland (shrub cover < 20\%).
+Open heathland is a common Australian alpine vegetation type, occupying approximately 25\% of the treeless landscapes above 1600 m in the Australian Alps (Fig.
+S2).
+It is an ecotone between closed heathland (\textgreater 70\% shrub cover) and tussock grassland (shrub cover < 20\%) that typically occurs on relatively flat topography.
 Open heathland is dominated by the shrub \textit{Grevillea australis} (shrub cover generally 20\textendash 60\%) with inter-shrub regions occupied by snow grasses (\textit{Poa hiemata} and \textit{Poa costiniana}) and other herbaceous species.
 \textit{Grevillea australis} is an obligate seeding shrub, and establishment of seedlings is dependent on disturbance creating bare ground in the grass\textendash herb sward \citep{Williams:1987ws}.
-Relative to tussock grassland, open heathland is a highly flammable plant community \citep{Williams:2006tl,Fraser:NHT5Kx_W}.
+Relative to tussock grassland, open heathland is a highly flammable plant community \citep[Fig.
+S3;][]{Williams:2006tl,Fraser:NHT5Kx_W}.
 A consequence of this differential flammability led to the 2003 wildfires on the Bogong High Plains burning approximately 60\% of open heathland and only 13\% of tussock grasslands.
-Following these fires, seedling regeneration of \textit{Grevillea australis} and other shrubs was prolific \citep{Williams:2014ud}.
 
@@
 In March 2010, at 1750 m a.s.l, we burnt 32 randomly selected mature (60 cm tall and 1.5 m$^2$) \textit{Grevillea australis} shrubs in an open heathland site that was not burnt by wildfire in 2003 or 2006.
-This created patches of bare ground approximately 0.7 m$^2$ surrounded by burnt tussock grass, simulating disturbance of individual shrubs in open heathland burnt by wildfire.
+This created patches of bare ground approximately 0.7 m$^2$ surrounded by burnt tussock grass \textemdash simulating disturbance of individual shrubs in open heathland burnt by wildfire.
 We collected seedlings of dominant alpine shrub species from a nearby (\textless 2 km) site of similar altitude burnt by 2006 wildfires.
@@
 A total of 640 seedlings, 256 \textit{Grevillea} (half used in \textit{Poa} inter-tussock experiment\textemdash see below) and 128 for each of \textit{Asterolasia}, \textit{Prostanthera} and \textit{Phebalium} were used.
-Four seedlings per species were randomly selected and transplanted into a 4\texttimes 4 square grid in the center of each burnt patch, with 14 cm between individuals and the edge of the patch, which was dominated by resprouting tussock grass \textit{Poa hiemata}.
-To examine interactions between tussock grass and shrub seedlings we also randomly transplanted four additional \textit{Grevillea australis} seedlings into various sized bare gaps between burnt \textit{Poa hiemata} tussocks that were immediately surrounding the 4\texttimes 4 bare ground square grid (Fig.
-\ref{figs1}).
+Four seedlings per species were randomly selected and transplanted into a 4 by 4 square grid in the center of each burnt patch, with 14 cm between individuals and the edge of the patch, which was dominated by resprouting tussock grass \textit{Poa hiemata}.
+To examine interactions between tussock grass and shrub seedlings we also randomly transplanted four additional \textit{Grevillea australis} seedlings into various sized bare gaps between burnt \textit{Poa hiemata} tussocks that were immediately surrounding the 4 by 4 bare ground square grid (Fig.
+S4 \& S5).
 The experimental site was fenced to prevent grazing by deer and horses.
@@
 Across six growing seasons (\Sexpr{n_days(chamber_dates, TRUE)} growing season days), OTCs simulated warmer conditions at the lower end of \citet*{IPCC:2013vv} projections (Fig.
-\ref{figs6}-\ref{figs8}).
+S6-S8).
 OTCs passively increased average ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C.
@@
 
-We used 40 open heathland monitoring sites established after the 2003 fires \citep{Williams:2006tl,Camac:2013eb}.
+We used 40 open heathland monitoring sites established after the 2003 fires \citep[Fig.
+S9;][]{Williams:2006tl,Camac:2013eb}.
 These sites consisted of 30 burnt sites and 10 sites known to be unburnt for over 70 years.
 The disparity in sample number between burnt and unburnt treatments was due to very few open heathland sites of sufficient size (0.25 ha) remaining unburnt after both the 2003 and 2006 fires.
-In the summer of 2011-12, at each site, seedling density/m$^2$ was estimated using 40 quadrats, each 1 m$^2$, that were evenly distributed in groups of 10 along four 50 m transects, with 10 m between transect lines, subsampling an area of 2000 m$^2$.
-Within plots we recorded the number and maximum height of \textit{Grevillea} and \textit{Asterolasia} seedlings.
-For unburnt sites we counted the number of mature \textit{Grevillea} plants (\textgreater 0.5 m$^2$) within 5 m of each transect.
+In the summer of 2011-12, at each site, seedling density/m$^2$ was estimated using 40 quadrats, each 1 m$^2$.
+Quadrats were evenly distributed in groups of 10 along four 50 m transects, with 10 m between transect lines, subsampling an area of 2000 m$^2$.
+Within quadrats we recorded the number and maximum height of \textit{Grevillea} and \textit{Asterolasia} seedlings.
+In unburnt sites, we counted the number of mature \textit{Grevillea} plants (\textgreater 0.5 m$^2$) within 5 m of each transect.
 In burnt sites, this required counting the number of skeletons (there were no living adults at any burnt site) that still persisted post-fire.
@@
 Site level data, elevation and Topographic Wetness Index \citep[TWI; a measure of plant available water,][]{Moore:1993fk} were obtained from a 30 m resolution digital elevation model.
-Lastly, for burnt sites, fire severity was estimated by twig diameters \citep{Whight:1999vw}, collected immediately after the 2003 fires \citep{Williams:2006tl}.
+Lastly, for burnt sites, fire severity was estimated by twig diameters \citep{Whight:1999vw}, collected immediately after the 2003 fires \citep[Fig S10;][]{Williams:2006tl}.
 
@@
 We built multiple hierarchical models to examine how increased temperature and other factors influenced shrub seedling recruitment, growth and mortality, as well as tussock-grass gap dynamics.
-For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} \citep{Rcran:2016xx} using package \texttt{rstan} \Sexpr{packageVersion("rstan")} \citep{stan:2015xx}.
+For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} \citep{Rcran:2016xx} using package \texttt{rstan} \Sexpr{packageVersion("rstan")} \citep{Stan:2016xx}.
 Detailed information about experimental design and analysis is provided in the Supporting Information.
 Data and source code are available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.
-In order to aid in the reproducibility of this work, our code was written using a \texttt{remake} framework \citep{fitzjohn2015}, such that others can readily reproduce our entire workflow from data processing, through to producing a pdf of this manuscript by calling \texttt{remake()} in R.
+In order to aid in the reproducibility of this work, our code was written using a \texttt{remake} framework \citep{fitzjohn2015}.
+This allows others to readily reproduce our entire workflow from data processing, through to producing a pdf of this manuscript by calling \texttt{remake()} in R.
+To safeguard against cross-platform issues and future software changes, we've embedded this framework within a Docker image (\url{https://hub.docker.com/r/jscamac/alpine_shrub_experiment/}).
 
@@
 \includegraphics{figures/fig2.pdf}
-\caption{Mean \textit{Grevillea australis} seedling density as a function of burning (burnt/unburnt), fire severity (twig diameter), altitude, Topographic Wetness Index (TWI) and adult density.
-(a) Centered and standardized model coefficients (on the log scale); and effects of: (b) burning, (c) fire severity, (d) altitude, (e) topographic wetness and (F) adult density, in areas burnt by the 2003 fires.
-All bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
+\caption{Mean \textit{Grevillea australis} seedling density as a function of burning (burnt/unburnt), fire severity (MTD), altitude, Topographic Wetness Index (TWI) and adult density.
+(a) Centered and standardized model coefficients (on the log scale); and effects of: (b) burning, (c) fire severity; as measured by minimum twig diameter (MTD), (d) altitude, (e) topographic wetness and (f) adult density, in areas burnt by the 2003 fires.
+All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
 \label{fig2}
@@
 \caption{Observed and projected growth trajectories of four dominant Australian alpine shrubs.
-(a) Observed height growth: Thick lines with error bars represent mean ($\pm$ 95\% confidence intervals) observed heights at each May census.
-Thin lines represent individual growth trajectories.
-(b) Mean ($\pm$ 95\% Bayesian Credible Interval) projected growth trajectories.
-In all cases, red and orange lines = seedlings growing in warmed (OTC) conditions and blue and light blue lines = seedlings growing in control conditions.
-Projections were based on an logistic growth model using initial seedling size of 6 cm (the mean initial height observed in the OTC experiment) and mean maximum heights observed in long-unburnt sites (i.e.
+(a) Observed height growth: Thick lines with error bars represent mean ($\pm$ 95\% confidence intervals) observed heights in warmed (red solid line/closed circles) and control (blue dashed line/open circles) conditions at each May census from 2010 to 2016.
+Thin lines represent individual growth trajectories in warmed (orange) and control (light blue) conditions.
+(b) Mean ($\pm$ 95\% Bayesian Credible Interval) projected growth trajectories in warmed (red solid line) and control (blue dashed line) conditions.
+Projections were based on a logistic growth model using initial seedling size of 6 cm (the mean initial height observed in the OTC experiment) and mean maximum heights observed in long-unburnt sites (i.e.
 \Sexpr{hmax_values(hmax,'Asterolasia')}, \Sexpr{hmax_values(hmax,'Grevillea')}, \Sexpr{hmax_values(hmax,'Phebalium')} \& \Sexpr{hmax_values(hmax,'Prostanthera')} cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Phebalium}, \& \textit{Prostanthera}, respectively).
-Vertical lines delimit year in which maximum height is obtained}
+Vertical dotted lines delimit year in which maximum height is obtained}
 \label{fig3}
@@
 \ref{fig2}) and \textit{Asterolasia} (Fig.
-\ref{figs2}) seedlings was strongly influenced by the occurrence of fire.
+S11) seedlings was strongly influenced by the occurrence of fire.
 Sites burnt in 2003 had seedling densities between \Sexpr{density_ratios(asttry_seedling_density_model,'Asterolasia')} and \Sexpr{density_ratios(greaus_seedling_density_model,'Grevillea')} times higher than unburnt sites.
@@
 This is consistent with the field warming experiment (see below) which indicated rates of recruitment (Fig.
-\ref{figs3}) and mortality in \textit{Grevillea} is largely insensitive to a \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C change in temperature.
+S12) and mortality in \textit{Grevillea} is largely insensitive to a \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C change in temperature.
 We detected no strong effect of Topographic Wetness Index for either species.
@@
 Given that seedlings were more abundant in burnt vegetation, we investigated how warmer post-fire conditions affected seedling growth and mortality rates.
-Seedlings of \textit{Grevillea}, \textit{Asterolasia}, \textit{Phebalium} and \textit{Prostanthera} emerging after a wildfire were transplanted into experimentally burnt plots.
+Seedlings of \textit{Grevillea}, \textit{Asterolasia}, \textit{Phebalium} and \textit{Prostanthera} emerging after a 2006 wildfire were transplanted into experimentally burnt plots.
 These plots were either subjected to ambient conditions (i.e.
@@
 For each species, similar treatment effects were observed for stem diameter growth (Fig.
-\ref{figs4}).
+S13).
 Accounting for initial height and assuming logistic growth, the rates of change in mean annual predicted height of \textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera} were \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Grevillea')}, \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Phebalium')} and \Sexpr{growth_rate_ratio(non_tussock_otc_growth_model,'Prostanthera')} times that observed in control plots, respectively.
-According to this model, a 6 cm seedling (the mean initial height of seedlings used in this experiment) attains maximum height \Sexpr{hmax(non_tussock_otc_growth_model,'Grevillea')} (\textit{Grevillea}) or \Sexpr{hmax(non_tussock_otc_growth_model,'Prostanthera')} years sooner (\textit{Phebalium} and \textit{Prostanthera}) when warmed by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C (Fig.
+According to this model, a 6 cm seedling (the mean initial height of seedlings used in this experiment) attains maximum height \Sexpr{hmax(non_tussock_otc_growth_model,'Grevillea')} (\textit{Grevillea}) or \Sexpr{hmax(non_tussock_otc_growth_model,'Prostanthera')} years sooner (\textit{Phebalium} and \textit{Prostanthera}) when growing in conditions \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C warmer (Fig.
 \ref{fig3}b).
@@
 Our experimental predictions were verified (Fig.
-\ref{figs5}).
+S14).
 Mean maximum height of \textit{Grevillea} seedlings in burnt open heathland were \Sexpr{max_ht_alt(greaus_seedling_max_ht_model)} cm taller at 1670 m a.s.l compared to seedlings at 1860 m a.s.l.
@@
 
-After six years and across all plots, \Sexpr{total_deaths(non_tussock_otc_mortality_data,TRUE)}\% (\Sexpr{total_deaths(non_tussock_otc_mortality_data,FALSE)} out of 511) of all seedlings transplanted into the 4\texttimes 4 bare ground square grid had died.
+After six years and across all plots, \Sexpr{total_deaths(non_tussock_otc_mortality_data,TRUE)}\% (\Sexpr{total_deaths(non_tussock_otc_mortality_data,FALSE)} out of 511) of all seedlings transplanted into the 4 by 4 bare ground square grid had died.
 Most deaths occurred in \textit{Asterolasia} (\Sexpr{death_counts(non_tussock_otc_mortality_data,'Asterolasia')}) and \textit{Phebalium} (\Sexpr{death_counts(non_tussock_otc_mortality_data, 'Phebalium')}), followed by \textit{Grevillea} (\Sexpr{death_counts(non_tussock_otc_mortality_data,'Grevillea')}) and \textit{Prostanthera} (\Sexpr{death_counts(non_tussock_otc_mortality_data,'Prostanthera')}).
@@
 This significant decrease in mortality may be a consequence of OTCs reducing the severity of spring frosts by rising minimum ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary,'CAmbient_Temp','mean_min','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary,'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C, respectively (Fig.
-\ref{figs6}-\ref{figs8}).
+S7).
 Warming also reduced mean seedling mortality in \textit{Grevillea} and \textit{Phebalium} (Fig.
@@
 \includegraphics{figures/fig4.pdf}
-\caption{Mean ($\pm$ 95\% Bayesian Credible Interval) annual probability of death for each species growing in control (blue) and warmed (red) conditions.}
+\caption{Mean ($\pm$ 95\% Bayesian Credible Interval) annual probability of death for each species growing in warmed (red/closed circle) and control (blue/open circle) conditions.}
 \label{fig4}
@@
 \caption{Experimental warming and inter-tussock gap size effects on \textit{Grevillea australis} seedlings.
-Top rows = Growth rate effects, Second row = Mortality effects.
-(a \& c) Centered and standardized model coefficients; (b \& d) growth and mortality rate response curves along an inter-tussock gap size gradient in warmed (red) and control (blue) conditions.
+Top rows = Growth rate effects, bottom row = Mortality effects.
+(a \& c) Centered and standardized model coefficients; (b \& d) growth and mortality rate response curves along an inter-tussock gap size gradient in warmed (red solid line) and control (blue dashed line) conditions.
 All error bars and shading are 95\% Bayesian Credible Intervals.}
@@
 \ref{fig6}).
-However, the rate at which this occurred was very slow, with a 10 cm radius gap predicted to decrease by approximately \Sexpr{gap_size(gap_dynamic_model, 'otc')} cm over a ten year period.
+However, the rate at which this occurred was very slow, with a 10 cm radius gap predicted to decrease by approximately \Sexpr{gap_size(gap_dynamic_model, 'otc')} cm over a ten\textendash year period.
 We also detected no significant effect of a \Sexpr{micro_clim(overall_microclimate_summary,'CAmbient_Temp','mean','grand_mean')}\textdegree C temperature rise on the rate of infilling.
@@
 \caption{Effects of warming on rates of gap infilling by tussock grass.
-(a) Model coefficients and (b) projected temporal change in size for an average 10 cm inter-tussock gap in warmed (red) and unwarmed (blue) conditions.
+(a) Model coefficients and (b) projected temporal change in size for an average 10 cm inter-tussock gap in warmed (red solid line) and control (blue dashed line) conditions over a 10-year period.
 All error bars and shading are 95\% Bayesian Credible Intervals.}
@@
 In particular, we have shown how these interactions strengthen a hypothesized feedback that can rapidly increase shrub cover in the Australian Alps, and potentially, other alpine and tundra ecosystems \citep{Higuera:2008dw}.
-While high cover of tussock grass is known to inhibit the establishment of shrub seedlings \citep{Williams:1987ws}, fire can create bare ground neccesary for shrub seedling establishment.
+While high cover of tussock grass is known to inhibit the establishment of shrub seedlings \citep{Williams:1987ws}, fire can create bare ground necessary for shrub seedling establishment.
 Our results highlight that fire can increase shrub seedling establishment by as much \Sexpr{density_ratios(greaus_seedling_density_model, 'Grevillea')}\textendash fold.
@@
 Landscape flammability, and thus fire, are also likely to increase as a result of warming effects on shrub seedling vital rates.
-Evidence from field studies of burning patterns \citep{Williams:2006tl}, quantification of fuel mass and architecture \citep{Fraser:NHT5Kx_W} and paleoecological studies \citep{Higuera:2008dw, Higuera:2009ga} indicate that shrubs are the most flammable component of alpine and tundra ecosystems \citep[significantly more than tussock grassland; ][]{Fraser:NHT5Kx_W}.
+Evidence from field studies of burning patterns \citep{Williams:2006tl}, quantification of fuel mass and architecture \citep{Fraser:NHT5Kx_W} and paleoecological studies \citep{Higuera:2008dw, Higuera:2009ga} indicate that shrubs are the most flammable component of alpine and tundra ecosystems \citep[significantly moreso than tussock grassland; ][]{Fraser:NHT5Kx_W}.
 In the Australian Alps, this differential flammability resulted in approximately 60\% of open heathland and only 13\% of tussock grasslands being burnt by the extensive 2003 wildfires.
@@
 Daniel Falster and Richard Fitzjohn for developing software that enabled these analyses to be readily reproducible.
-Special thanks to John Baumgartner, David Bowman, Manuel Esper\'{o}n-Rodr\'{i}guez, Ana{\"i}s Gibert, John Morgan, Kimberley Millers, Michaela Plein and Inka Veltheim for their comments on this manuscript.}
+Special thanks to John Baumgartner, David Bowman, Manuel Esper\'{o}n-Rodr\'{i}guez, Ana{\"i}s Gibert, John Morgan, Kimberley Millers, Michaela Plein and Inka Veltheim for their comments on this manuscript.
+This research was conducted under Parks Victoria permits 10005212 \& 10006611}
 
@@
 
-% \pnasbreak splits and balances the columns before the references.
+%\gcbbreak
+% \gcbbreak splits and balances the columns before the references.
 % If you see unexpected formatting errors, try commenting out this line
@@
 \bibliography{refs.bib}
-%\pnasbreak
+%\gcbbreak
 \end{document}

### 2017-01-03 manuscript/supplementary.Rnw
Problem with compiling pdf with line numbers using older versions of TeX Live (see #3 and annakrystalli/OpenConBerlin_ReproHack#2). Temp solution is to remove line numbers
Final edits for GCB

@@
-\documentclass[9pt,onecolumn,twoside, lineno]{gcb}
+\documentclass[9pt,onecolumn,twoside]{gcb}
 % Use the lineno option to display guide line numbers if required.
@@
 \tableofcontents
+\clearpage
 \listoffigures
-
+\clearpage
 \section{Study site}
@@
 The undulating terrain ranges in altitude from 1400 to 2000 m a.s.l and contains a well-developed mantle of soil.
+The most widespread soil type is ``alpine humus soil'', a friable, organic loam up to 1 m deep \citep{Williams:2014ud}.
 Mean annual temperature is approximately 6\textdegree C, with persistent snow cover from June to September.
@@
 The treeline is irregular (due to cold air drainage), occurring at altitudes ranging from 1600 to 1700 m a.s.l.
-Approximately 50\% of the alpine ‘treeless’ landscape consists of heathlands, most of which were extensively burned by a 400,000 ha wildfire in 2003 \citep{Williams:2006tl,Camac:2013eb}.
+Approximately 50\% of the alpine ‘treeless’ landscape consists of heathlands, most of which were extensively burned by a 400,000 ha wildfire in 2003 \citep[Fig.
+\ref{figS1};][]{Williams:2006tl,Camac:2013eb}.
 Large wildfires in Australian alpine landscapes have occurred once or twice per century and typically occur during periods of extended regional drought coupled with severe fire weather and multiple ignitions in the surrounding montane forests \citep{Williams:2008ba}.
@@
 \section{Open Top Chamber Experiment}
-We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of post-fire gap-infilling by tussock grass and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affects shrub seedling demographic rates; and 3) whether such effects change under warmer conditions post-fire conditions.
+We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of post-fire gap-infilling by tussock grass as well as shrub seedling recruitment, growth and mortality; 2) how proximity to tussock grass affects shrub seedling demographic rates; and 3) whether such effects change under warmer conditions post-fire conditions.
 
@@
 The combination of bare ground and inter-tussock seedlings meant that each experimental plot had a total of 20 seedlings per 1 m$^2$ (Fig.
-\ref{figs1}).
+\ref{figs5}).
 However, the two sets of seedlings (i.e.
@@
 
-To simulate near-term warmer conditions indicated by the IPCC \citep{IPCC:2013vv}, we used Open Top Chambers (OTCs) \citep{Molau:1996vc}.
+To simulate near-term warmer conditions indicated by the IPCC \citep{IPCC:2013vv}, we used Open Top Chambers (OTCs; Fig.
+\ref{figS2}) \citep{Molau:1996vc}.
 OTCs were hexagonal, 58 cm tall with an open-top diameter of 110 cm and base diameter of 168 cm.
@@
 \ref{figs6}-\ref{figs8}).
-OTCs passively increased average daily ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively.
+OTCs passively increased average daily ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C.
 Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C).
@@
 $\geq$ 26 cm), these data were incomplete.
-We estimated these censored data using a censored model (see Modelling censored Poa Distances).
+We estimated these censored data using a censored model (see: Modelling censored Poa distances).
 Once true distances were estimated for censored data, we estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each census, for each inter-tussock seedling.
@@
 
-The study used 40 open heathland sites \citep{Camac:2013eb} established immediately after the 2003 fires \citep{Williams:2006tl}.
+The study used 40 open heathland sites \citep[Fig.
+\ref{figS9};][]{Camac:2013eb} established immediately after the 2003 fires \citep{Williams:2006tl}.
 All sites have been free of grazing by cattle since 2003.
@@
 Transects were positioned in the center of each site and were parallel to each other with exactly 10 m between transect lines.
-This effectively created a grid of 40 plots (10 plots~\texttimes~4 transects) per site, subsampling an area of 2000 m$^2$.
+This effectively created a grid of 40 plots (10 plots by 4 transects) per site, subsampling an area of 2000 m$^2$.
 
@@
 plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset \citep{Gelman:2007te, Kery:2012wl}.
-For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")} \citep{stan:2015xx}.
+For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")} \citep{stan:2016xx}.
 For each model, three independent chains were run.
@@
 \bibliography{refs.bib}
-\pagebreak
 
+\clearpage
 \section{Supplementary Figures}
 
-\begin{figure}[h]
+\begin{figure}[!h]
 \centering
-\includegraphics{figures/figS1.pdf}
+\includegraphics{figures/figS1.jpg}
+\caption{Approximately 400,000 ha of the Victorian Alpine National Park was burned by the wildfires in 2003.} 
+\label{figs1}
+\end{figure}
+\clearpage
+
+\begin{figure}[!h]
+\centering
+\includegraphics{figures/figS2.jpg}
+\caption{Open-Top-Chambers in \textit{Grevillea australis} and \textit{Poa hiemata} dominated open heathland.} 
+\label{figs2}
+\end{figure}
+\clearpage
+
+\begin{figure}[!h]
+\centering
+\includegraphics{figures/figS3.jpg}
+\caption{Shrubs are the most flammable component in Australian alpine 'treeless' vegetation \citep{Williams:2006tl}. Burned patches were shrubs surrounded by lightly scorched tussock grass.} 
+\label{figs3}
+\end{figure}
+\clearpage
+
+\begin{figure}[!h]
+\centering
+\includegraphics{figures/figS4.jpg}
+\caption{Setting up of experimental plots.
+(a): unburned \textit{Grevillea australis}; (b): \textit{Grevillea} after being burned; (c) \textit{Grevillea} remains removed leaving burned bare ground patch.
+Transplant holes were then created using hamilton planter (note: photo does not show holes for inter-tussock seedlings; (d) all seedlings planted within 1 m$^2$.}
+\label{figs4}
+\end{figure}
+\clearpage
+
+\begin{figure}[!h]
+\centering
+\includegraphics{figures/figS5.pdf}
 \caption{Diagram outlining positions of transplanted seedling in an OTC plot.
@@
 it was not within the 110 cm opening of the chamber) and therefore was not used in the experiment.}
-\label{figs1}
-\end{figure}
-\clearpage
-
-\begin{figure}[!hb]
-\centering
-\includegraphics{figures/figS2.pdf}
-\caption{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (twig diameter), altitude and Topographic Wetness Index (TWI).
-(a) Centered and standardized model coefficients (on the log scale); and effects of (b) burning, (c) fire severity, (d) altitude and (e) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires.
-All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
-\label{figs2}
-\end{figure}
-\clearpage
-
-\begin{figure}[!hb]
-\centering
-\includegraphics{figures/figS3.pdf}
-\caption{Average number of \textit{Grevillea australis} recruits in warmed (red) and control (blue) conditions in 2011 and 2012. Errors bars are 95\% Bayesian Credible Intervals.}
-\label{figs3}
-\end{figure}
-\clearpage
-
-\begin{figure}[!hb]
-\centering
-\includegraphics{figures/figS4.pdf}
-\caption{Observed seedling diameter growth trajectories of four dominant Australian alpine shrubs.
-Red and Blue lines represent mean ($\pm$ 95\% CI) observed heights at each May census in warmed and control plots, respectively.
-Orange and light blue lines represent individual growth trajectories in warmed and unwarmed plots, respectively.}
-\label{figs4}
-\end{figure}
-\clearpage
-
-\begin{figure}[!hb]
-\centering
-\includegraphics{figures/figS5.pdf}
-\caption{Mean maximum seedling height along fire severity (twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires.
-Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses.
-(a \& b) Centered and standardized model coefficients (on the log scale); and effects of (c \& d) fire severity, (e \& f) altitude and (g \& h) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires.
-All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
 \label{figs5}
@@
 
-\begin{figure}[!hb]
+\begin{figure}[!h]
 \centering
@@
 \end{figure}
+\clearpage
+
+\begin{figure}[!h]
+\centering
+\includegraphics{figures/figS9.png}
+\caption{Distribution of open heathland sites across the Bogong high plains.
+Red points: sites burned by the 2003 wildfires; Blue points: unburned sites.
+White areas signify alpine treeless vegetation.}
+\label{figs9}
+\end{figure}
+\clearpage
+
+\begin{figure}[!h]
+\centering
+\includegraphics{figures/figS10.jpg}
+\caption{Example of a shrub severely burned by the 2003 wildfires.}
+\label{figs10}
+\end{figure}
+\clearpage
+
+\begin{figure}[!h]
+\centering
+\includegraphics{figures/figS11.pdf}
+\caption{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (MTD), altitude and Topographic Wetness Index (TWI).
+(a) Centered and standardized model coefficients (on the log scale); and effects of (b) burning, (c) fire severity; as measured by minimum twig diameter (MTD), (d) altitude and (e) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires.
+All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
+\label{figs11}
+\end{figure}
+\clearpage
+
+\begin{figure}[!h]
+\centering
+\includegraphics{figures/figS12.pdf}
+\caption{Average number of \textit{Grevillea australis} recruits in warmed (red closed circle) and control (blue open circle) conditions in 2011 and 2012.
+Errors bars are 95\% Bayesian Credible Intervals.}
+\label{figs12}
+\end{figure}
+\clearpage
+
+\begin{figure}[!h]
+\centering
+\includegraphics{figures/figS13.pdf}
+\caption{Observed seedling diameter growth trajectories of four dominant Australian alpine shrubs.
+Thick lines represent mean ($\pm$ 95\% CI) observed diameters in warmed (red solid line/closed circles) and control (blue dashed line/open circles) at each May census in warmed and control plots, respectively.
+Thin lines represent individual growth trajectories in warmed (orange) and control (light blue) conditions.}
+\label{figs13}
+\end{figure}
+\clearpage
+
+\begin{figure}[!h]
+\centering
+\includegraphics{figures/figS14.pdf}
+\caption{2012 mean maximum seedling height along fire severity (MTD), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires.
+Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses.
+(a \& b) Centered and standardized model coefficients (on the log scale); and effects of (c \& d) fire severity; as measured by minimum twig diameter (MTD), (e \& f) altitude and (g \& h) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires.
+All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
+\label{figs14}
+\end{figure}
+\clearpage
 

### 2017-01-03 R/figures.R

@@ -1,24 +1,35 @@
 
 # Theme for partial plots
 partial_plot_theme <- function(legend.position = "none", strips = FALSE,...) {
-  sb <- if(strips==TRUE) element_rect(fill='lightgrey') else element_blank()
+  sb <- if(strips==TRUE) element_rect(fill='lightgrey', size=0.7) else element_blank()
   st <- if(strips==TRUE) element_text(face='italic') else element_blank()
   theme_classic(base_size = 7) + theme(strip.text = st,
                           legend.title = element_blank(),
                           strip.background = sb,
                           legend.position = legend.position,
-                          axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
-                          axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'),
-                          plot.margin = unit(c(3,3,3,3), "mm"))
+                          plot.margin = unit(c(3,3,3,3), "mm"),
+                          axis.text = element_text(colour="black"),
+                          axis.line = element_blank(),
+                          panel.border = element_rect(fill=NA, colour ="black"),
+                          axis.ticks = element_line(size =0.3, colour="black"),
+                          axis.ticks.length=unit(-0.7, "mm"),
+                          axis.text.x = element_text(margin=unit(c(2,0,0,0), "mm")),
+                          axis.text.y = element_text(margin=unit(c(0,2,0,0), "mm")))
 }
 
 # Theme for coefficient plots
 coefficent_plot_theme <- function() {
   theme_classic(base_size = 7) + theme(
                           axis.title.y = element_blank(),
-                          axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
-                          axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'),
-                          plot.margin = unit(c(3,3,3,3), "mm"))
+                          plot.margin = unit(c(3,3,3,3), "mm"),
+                          axis.line = element_blank(),
[…]

### 2017-01-03 manuscript/supplementary.Rnw
Fixed figure references
Remove DOI, issue and volume placemarkers

@@
 \leadauthor{Camac} 
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: james.camac@gmail.com}
-\dates{This manuscript was compiled on \today}
-\doi{\url{doi:10.1111/GCB.XXXXX}}
+\correspondingauthor{To whom correspondence should be addressed. E-mail: james.camac@gmail.com}
+%\dates{This manuscript was compiled on \today}
+%\doi{\url{doi:10.1111/GCB.XXXXX}}
 
@@
 Approximately 50\% of the alpine ‘treeless’ landscape consists of heathlands, most of which were extensively burned by a 400,000 ha wildfire in 2003 \citep[Fig.
-\ref{figS1};][]{Williams:2006tl,Camac:2013eb}.
+\ref{figs1};][]{Williams:2006tl,Camac:2013eb}.
 Large wildfires in Australian alpine landscapes have occurred once or twice per century and typically occur during periods of extended regional drought coupled with severe fire weather and multiple ignitions in the surrounding montane forests \citep{Williams:2008ba}.
@@
 To simulate near-term warmer conditions indicated by the IPCC \citep{IPCC:2013vv}, we used Open Top Chambers (OTCs; Fig.
-\ref{figS2}) \citep{Molau:1996vc}.
+\ref{figs2}) \citep{Molau:1996vc}.
 OTCs were hexagonal, 58 cm tall with an open-top diameter of 110 cm and base diameter of 168 cm.
@@
 The study used 40 open heathland sites \citep[Fig.
-\ref{figS9};][]{Camac:2013eb} established immediately after the 2003 fires \citep{Williams:2006tl}.
+\ref{figs9};][]{Camac:2013eb} established immediately after the 2003 fires \citep{Williams:2006tl}.
 All sites have been free of grazing by cattle since 2003.
